### 2013-10-05 paper/appendix.tex
first commit

\section{Analytic equilibrium for the polytropic disk}\label{appen1}
For a polytropic disk with $P = K\rho^2$ the dimensional equilibrium equation to
be solved is
\begin{align}
  0=\csmid^2\frac{d^2}{dz^2}\left(\frac{\rho}{\rho_0}\right) +
  \Omega_z^2 + \frac{\Omega^2}{Q}\left(\frac{\rho}{\rho_0}\right), 
\end{align}
which is obtained by combining Eq. \ref{eqm_eqns1} and
\ref{eqm_eqns2} with the above equation of state. The solution is
\begin{align}  
  \frac{\rho}{\rho_0} = \left(1 + \frac{\Omega_z^2}{\Omega^2}Q\right)\cos{\left(a z\right)} - 
\frac{\Omega_z^2}{\Omega^2}Q,  
\end{align}
where  
\begin{align}
  a^2 \equiv \frac{\Omega^2}{Q\csmid^2}. 
\end{align}
The polytropic disk thickness is
\begin{align}
  H =
  \frac{\csmid}{\Omega}\sqrt{Q}\arccos\left(\frac{\Omega_z^2Q}{\Omega^2+ \Omega_z^2Q}\right).  
\end{align}

\section{Reduction to linear hydrodynamics}\label{reduction} 
Our task here is to remove the magnetic field and vertical velocity
perturbations from the non-dimensional linearized equations. 
Let us first define operators
\begin{align}
  D_0 = 1, \quad D_1 = \frac{\rho^\prime}{\rho} + \frac{d}{dz}, \quad
  D_2 = \frac{\rho^{\prime\prime}}{\rho} +
  \frac{2\rho^\prime}{\rho}\frac{d}{dz} + \frac{d^2}{d z^2},
\end{align}
and 
\begin{align}
  \hat{D}_0 = \eta D_0, \quad \hat{D}_1 = \eta^\prime D_0 + \eta
  D_1,\quad \hat{D}_2 = \eta^{\prime\prime} D_0 + 2\eta^\prime D_1 +
  \eta D_2. 
\end{align}
Denoting the $n^\mathrm{th}$ vertical derivative as $^{(n)}$, the
equations of motion give
\begin{align}
  &\dbx^{(n)} = \frac{\beta\rho}{f}D_{n-1}\left(\imgi\sigma\dvx - 2\dvy +
  \imgi f k_x\w\right) + \imgi k_x \dbz^{(n-1)},\label{bx_eq}\\
  &\dby^{(n)} = \frac{\beta\rho}{f}D_{n-1}\left(\imgi\sigma\dvy +
  \frac{\kappa^2}{2}\dvx\right),\label{by_eq}
\end{align}
for $n\geq1$. Differentiating the divergence-free condition for the magnetic
field gives
\begin{align}
  \imgi k_x \dbx^{\prime} + \dbz^{\prime\prime} = 0.   
\end{align}
We insert the expression for $\dbx^\prime$ from Eq. \ref{bx_eq} and the
expressoin for $\dbz^{\prime\prime}$ from the $z$ component of the
linearized induction equation (Eq. \ref{induct_vert}) to obtain
\begin{align}\label{bz_eq}
  -\sigma\dbz^{(n)} = f k_x \dvx^{(n)} + \frac{k_x\beta\rho}{f}\hat{D}_n\left(\imgi\sigma\dvx - 2\dvy +
  \imgi f k_x\w \right). 
\end{align}

\subsection{Eliminating $\dbx$}
Inserting the above expressions for $\dbx^{\prime\prime}$,
$\dbx^\prime$ (Eq. \ref{bx_eq}) and $\dbz^\prime$  (Eq. \ref{bz_eq})
into the right-hand-side of the $x$-induction equation (Eq. \ref{induct_x} ) gives   
\begin{align}
  \imgi\sigma \frac{f}{\beta\rho}\dbx =
  \frac{f^2}{\beta\rho}\dvx^\prime + \hat{D}_1\left(\imgi\sigma\dvx -
  2\dvy + \imgi f k_x\w\right). \label{bx_expression}
\end{align}
($\bar{\sigma}\neq0$ has been assumed to obtain this.) 
We multiply this expression by $\rho$, differentiate with respect to
$z$ and eliminate the resulting $\dbx^\prime$ using
Eq. \ref{bx_eq}, to obtain
\begin{align}
  0 = \frac{f^2}{\beta\rho}\dvx^{\prime\prime} -
  \frac{f^2}{\beta\rho}k_x^2\dvx + \left(\hat{D}_2 - k_x^2 \hat{D}_0
    - \imgi\sigma D_0\right)\left( \imgi\sigma\dvx -
  2\dvy + \imgi f k_x\w\right). \label{final_vx} 
\end{align}

\subsection{Eliminating $\dby$}
We follow a similar procedure as above to remove $\dby$. We first use
the divergence-free condition to eliminate $\dbx$ from the
right-hand-side of the $y$-induction equation (Eq. \ref{induct_y}),
\begin{align}
  &\imgi\bar{\sigma}\dby = f\dvy^\prime - \imgi
  S\frac{\dbz^\prime}{k_x} +
  \eta\dby^{\prime\prime}+\eta^\prime\dby^\prime.  
\end{align}
We can now insert expressions for the magnetic field derivatives using
Eq. \ref{by_eq} and \ref{bz_eq} to obtain an expression for $\dby$, 
\begin{align}
  \imgi\bar{\sigma}\dby = f \dvy^\prime + \frac{\imgi S}{\sigma}\left[
    f\dvx^\prime + \frac{\beta\rho}{f}\hat{D}_1\left(\imgi\sigma\dvx -
  2\dvy + \imgi f k_x\w \right)\right] +
  \frac{\beta\rho}{f}\hat{D}_1\left(\imgi\sigma\dvy +
  \frac{\kappa^2}{2}\dvx\right). \label{by_expression}
\end{align}
We differentiate this expression with respect to $z$, then eliminate 
$\dby$ and  $\dby^\prime$ from the left-hand-side of the resulting
expression using Eq. \ref{by_expression} and 
Eq. \ref{by_eq}, respectively. We obtain
\begin{align}
  0 =& \frac{f^2}{\beta\rho}\left(\dvy^{\prime\prime} -
  \frac{\bar{\sigma}^\prime}{\bar{\sigma}}\dvy^\prime\right) 
  + \sbar\sigma D_0\dvy  + \frac{\imgi 
    S}{\sigma}\frac{f^2}{\beta\rho}\left(\dvx^{\prime\prime} -
    \frac{\sbar^\prime}{\sbar}\dvx^\prime\right) -
    \frac{\imgi\sbar\kappa^2}{2}D_0\dvx \notag\\
    & + \left(\hat{D}_2 -
    \frac{\sbar^\prime}{\sbar}\hat{D}_1\right)\left[\imgi\left(\sigma
      - \frac{2S}{\sigma}\right)\dvy + \left(\frac{\kappa^2}{2} -
      S\right)\dvx - \frac{Sfk_x}{\sigma}\w\right]. \label{final_vy}
\end{align}

\subsection{Elminating $\dvz$}
The vertical velocity perturbation is
\begin{align}
  \dvz = \frac{\imgi f}{\sigma}\w^\prime. 
\end{align}
Inserting this into the linearized continuity equation
(Eq. \ref{lin_cont}) and using the Poisson equation, we obtain
\begin{align}
0=  W^{\prime\prime} + \left(\ln{\rho}\right)^\prime W^\prime +
  \frac{1}{c_s^2f^2}\left(\frac{\rho}{Q} + \sigma^2\right)W +
  \left(\ln{\rho}\right)^\prime\dphi^\prime + k_x^2\dphi + \frac{\sigma
  k_x}{f}\dvx.\label{final_w}
\end{align}

### 2013-10-05 paper/cover.tex
first commit

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=4in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Min-Kai Lin}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{The Editor, The Astrophysical Journal}

\begin{flushleft}
{\large\bf Min-Kai Lin}
\end{flushleft}
\medskip\hrule height 1pt
\begin{flushright}
\hfill Canadian Institute for Theoretical Astrophysics, \\
\hfill 60 St. George Street, Toronto, ON \\
\hfill M5S 3H8, Canada\\
\hfill +1 647 896 4314
\end{flushright} 
\vfill % forces letterhead to top of page

\opening{Dear Editor,}

I am writing to submit the second revised manuscript for `Non-barotropic linear Rossby wave instability
in three-dimensional disks'. I include a list of revisions made in light of the reviewer's comments.

\closing{Yours sincerely,}

%\encl{}  				% Enclosures

\end{letter}

\end{document}

### 2013-10-05 paper/iso.tex
first commit

\section{Isothermal limit}\label{isothermal}
We now examine the limit $\Gamma\to 1$, where the unperturbed disk becomes isothermal
, but perturbations are evolved with an adiabatic index
$\gamma=1.4$. We consider a nearly-isothermal polytropic background
and strictly isothermal backgrounds. These cases are treated
separately because the equilibrium structures have different
functional forms. A comparison between them provide another
check on our numerical results.

\subsection{Large polytropic index}
We first consider setting $n=10$ to produce an almost radially
isothermal equilibrium with $p\propto \rho^{1.1}$. This allows us to use the
numerical code as set up for polytropic equilibria without
modification. We also adopt $A=2.5$ and $h=0.25$ for reasons given
in \S\ref{varpolyn}. The relatively large aspect-ratio does not
violate the thin-disk approximation as large $n$ implies the 
density decays rapidly away from the midplane. Also because of this,
we set the upper disk boundary at $Z_s=0.6$ to avoid very low 
densities. %All other parameters are the same as before.

%3d-ness of core is 0.55
%3d-ness averaged over 0.9, 1.1 is 0.37
%%  Despite the larger
%% value of $\gamma/\Gamma$ than case 8 (see Table \ref{linsims}), the flow
%% three-dimensinality is the same.

For this setup we obtained $\omega/m\Omega_0=0.9883$,
$\nu/\Omega_0=0.1375$ and $\avg{\theta_m}=0.35$. The top panel of
Fig. \ref{polyn10} shows the meridional flow at the vortex core. The
vortical motion is distinct and more apparent than case 3a, despite
the smaller value of $\gamma/\Gamma$ in the present case. 
However, apart from this difference, the solution is qualitatively
similar to case 3a.

\begin{figure}[!t]
  \centering
  %% \includegraphics[scale=.425,clip=true,trim=0cm 1.cm 0cm
  %%   0cm]{figures/n10.0_gmma1.4_wq_3d_real_zslices.ps}
  \includegraphics[scale=.425,clip=true,trim=0cm 1.83cm 0cm
    0.25cm]{figures/n10.0_gmma1.4_vel3d_rz_dvz_nofill.ps}
   \includegraphics[scale=.425,clip=true,trim=0cm 0cm 0cm
    0.25cm]{figures/iso_vel3d_rz_dvz_nofill.ps}
  \caption{Perturbed meridional flow at $\phi=\phi_0$ for a $n=10$
    polytropic disk equilibrium (top) and a strictly isothermal
    equilibrium (bottom). 
    \label{polyn10}}
\end{figure}

\subsection{Strictly isothermal equilibrium}
%Because we have formulated the linear problem in terms of entropy and
%pressure length-scales, in order to calculate the RWI for strictly
%isothermal disks, we can just replace 
Modifications to our standard setup are required to treat  
disk equilibria with $p=\ciso^2\rho$ ($\Gamma\equiv 1$), where 
the constant sound speed $\ciso=\Hiso\Omega_k$, $\Hiso=\hiso
r_0(r/r_0)^{3/2}$ is the isothermal scale-height, and $\hiso$ is the
characteristic aspect-ratio at $r_0$. 
%We set $\ciso=\Hiso
%\Omega_k$, where $\Hiso\propto r^{3/2}$ is the isothermal scale-height
%and we define $\hiso=\Hiso/r_0$. 
The dimensionless vertical 
co-ordinate is now $Z=z/\Hiso$. The isothermal atmosphere is
exponential, $g(Z)=\exp{(-Z^2/2)}$, so there is no surface. In
practice we choose a finite vertical domain, i.e. $Z=Z_s$ represents a
constant number of isothermal scale-heights above the midplane.

In the linear code we simply replace expressions for the entropy
and pressure length-scales by those corresponding to the isothermal
disk: the function $H\to\Hiso$ and $g(Z)$ becomes the Gaussian above. 
%% We set parameter values to mimic the large-$n$ polytrope in the 
%% previous section:  $Z_s=3$ and $\hiso=0.05$,
%% so that for both setups the pressure is reduced by roughly the 
%% same factor in going from the midplane to the upper boundary, and the
%% midplane temperature is roughly equal at $r_0$. 
{\bf
We choose $Z_s=3$ and $\hiso=0.05$, so the isothermal disk has roughly
the same temperature as that in the midplane of the large-$n$ 
polytrope considered above (at $r_0$).} In going from the
midplane to the upper boundary, the density is also reduced by
approximately the same factor for both cases.

We obtain $\omega/m\Omega_0=0.9860$, $\nu/\Omega_0=0.1008$ and
$\avg{\theta_m}=0.39$. The perturbations plotted in
Fig. \ref{isothermal_case} are similar to case 3a, so we expect these
are features of the RWI in nonhomentropic flow, rather than
associated with the chosen parameter values. The perturbed meridional
flow shown in Fig. \ref{polyn10} (bottom panel) is in qualitative
agreement with the large-$n$ polytrope. 
The result is, however, quite different to
isothermal linear perturbations, for which \cite{meheut12} found the 
vertical velocity appears to have a node at $r_0$ ({\bf see their Fig. 3d where the vertical velocity changes sign across co-rotation radius}
, i.e. the fluid column is hydrostatic there). 
Note that both $\gamma/\Gamma$
and the growth rate are slightly smaller than the nonhomentropic case 3a,
but here the vortical motion is more prominent. 
%contribution from higher m?

\begin{figure}[!t]
  \centering
  \includegraphics[scale=.425,clip=true,trim=0cm 0.cm 0cm
    0cm]{figures/iso_wq_3d_real_zslices.ps}
  \caption{Pressure (top, $W$), density (middle, $Q$) and entropy
    (bottom, $S$) for a globally isothermal background. 
    \label{isothermal_case}}
\end{figure}

  Fig. \ref{polyn10_vz} shows the vertical 
  velocity at the vortex core as a function of height. The strictly
  isothermal background {\bf (thick solid)} has a slightly larger $\dd v_z$ than the
  large-$n$ polytrope {\bf (thick dashed)}. This is consistent 
  with previous findings that vertical motions oppose the RWI
  \citep{lin12c}, as the former case has a smaller growth rate than
  the latter. The {\bf thick lines are} qualitatively similar to case
  3a in Fig. \ref{compare_dvz}, but these are not directly comparable 
  because the present case differs in both the background structure and adiabatic index  to 
  those in Fig. \ref{compare_dvz}.

  \begin{figure}[!t]
    \centering
    \includegraphics[scale=.425,clip=true,trim=0cm 0cm 0cm
      0.25cm]{figures/temp_dvz3}
    \caption{
      {\bf
        Vertical velocity as a function of $z$ at the vortex core
        $(r_0,\phi_0)$, for the $n=10$ 
        polytropic disk equilibrium (dashed) and a strictly isothermal
        equilibrium (solid) shown in 
        Fig. \ref{polyn10}, with free upper boundaries (thick lines). Corresponding thin 
        lines impose zero vertical velocity at $z=Z_s$ (growth rates increased by less than $0.5\%$ from the free boundary condition).
        Notice that changing upper disk boundary conditions only 
        affected the solution near $z=Z_s$ (cf. Fig. \ref{compare_dvz}). 
        This is consistent \cite{lin12c}, who found the
        influence of upper disk boundary condition to diminish with
        increasing polytropic index $n$. 
      }
      \label{polyn10_vz}}
\end{figure}

  Finally, we illustrate again a correlation 
  between meridional vortical flow and a tilted column of negative
  vertical vorticity perturbation in Fig. \ref{iso_vort3d_pz}. The figure is
  {\bf qualitatively} similar to that for polytropic backgrounds (case 3a in
  Fig. \ref{nonhomentropic_case_vort3d_pz}).{\bf  We find an average tilt
    of $1-~\avg{\cos{\theta}}_Z=0.0084
    \ll1$, so the vorticity column is nearly vertical.}

  %% but the tilt magnitude
  %%   is larger for the isothermal background. 
  %Note that we chose
  %$r=1.03r_0$ as the radial slice, so that $\delta\omega_z <0$ along
  %the vertical line $(\phi_0,z)$. 
  %%  A plot for $r=1.02r_0$ also display
  %% the tilt, but in that case $\delta\omega_z$ was found to be positive
  %% at

\begin{figure}[!t]
  \centering
  \includegraphics[scale=.425,clip=true,trim=0cm 0cm 0cm
    0.25cm]{figures/iso_vort3d_pz2.ps}
  \caption{
      Vertical vorticity perturbation, $\delta\omega_z$, in the
      $(\phi,z)$ plane at $r=1.03r_0$ for the strictly isothermal
      background. Regions of $\delta\omega\leq0$ are delineated by
      while lines. The center of meridional vortical motion identified
      in Fig. \ref{polyn10} occurs at height $z\sim
      H_\mathrm{iso}$. {\bf The azimuthal range
        $\phi-\phi_0\in[-0.5,0.5]\pi/m$ corresponds to anti-cyclonic
        motion about the vortex core.
        [ A plot for $r=1.02r_0$ also display
      tilted lines of constant $\delta\omega_z$, but in that case $\delta\omega_z > 0$
      at $(\phi_0, H_\mathrm{iso})$.]
      } 
    \label{iso_vort3d_pz}}
\end{figure}

\subsection{A nonlinear simulation}
%boundary conditions
%compare Q at different heights
%compare W in rz plane - vortical motion
We have also performed global 3D hydrodynamic simulations
using the \zeus  finite-difference code \citep{hays06}. As the focus
of this work is the linear problem, though, we defer a full discussion
of these nonlinear simulations to a follow-up paper. Our priority here
is to verify the  vortical motion in the meridional plane, which
appears characteristic in  the linear RWI solution for  nonhomentropic
flow.

\subsubsection{Setup}
We use spherical polar co-ordinates $(\rsph,\theta,\phi)$ to describe
the disk, taken to be initially strictly isothermal as described above.  The
computational domain is $\rsph\in[0.2,2.0]r_0$,
$\theta\in[\theta_\mathrm{min},\pi/2]$, $\phi\in[0,2\pi]$ and is
divided into $(512,48,512)$ zones, with
$\tan{(\pi/2-\theta_\mathrm{min})}=3\hiso$ and $r_0=10$. 
The grid is logarithmically spaced in radius and uniformly spaced in 
the angular co-ordinates. Boundary conditions 
are outflow in $\rsph$, reflection in $\theta$ and periodic in $\phi$.  
Additional damping to meridional velocities near radial boundaries are
employed to reduce reflections \citep{valborro07}.

%The disk is initially strictly isothermal as described in the previous section, and the vertical domain is such that 
%$\tan{(\pi/2-\theta_\mathrm{min})}=3\hiso$.

After some experimentation, we found it was most convenient to start with a smooth disk.
In this case, a surface density $\Sigma\propto r^{-3/2}$, and tapered toward 
the inner boundary \citep[as used in][]{lin12b}. We introduce 
the density bump at $r=r_0$ via source terms in the mass, momentum and thermal energy equations, over a time-scale of $10P_0$, 
where $P_0 \equiv 2\pi/\Omega_k(r_0) $. %% is the Keplerian orbital period at $r_0$
This reduces numerical transients associated with initialization with {\bf a localized bump which has} large radial gradients. 
%At $t=10P_0$ we add small-amplitude random
%radial velocity perturbations.

We choose the bump amplitude $A=1.25$ 
and isothermal aspect-ratio $\hiso=0.1$, as employed by \cite{meheut12} so that we can check our results against theirs.  
We measure perturbations with respect to azimuthally averaged hydrodynamic quantities at $t=10P_0$.

\subsubsection{Results and comparison to linear flow}
%complex freq. averaged over r_sph\in[0.8,1.2]r_0 is
%<growth rate>/omega0=      0.18870212
%<pattern speed>/m.omega0=      0.98988571
%
%complex freq. averaged over r_sph\in[0.9,1.1]r_0
%<growth rate>/omega0=      0.19579173
%<pattern speed>/m.omega0=      0.98709079
%theta (3dness) is 0.34
We focus on the earliest stage of the instability, when perturbation
amplitudes are small so comparison with linear calculations can be
made. Fig. \ref{hydro_polar_dens} shows the snapshot to be
examined, {\bf taken at $t=23P_0$}. A $m=4$ mode has developed {\bf
  from numerical noise.} Notice the double-peak in
density perturbation, which is also present in
Fig. \ref{isothermal_case}. Using the method described in Appendix
\ref{instant_growth}, we estimated the $m=4$ mode growth rate and
frequency to be $\nu/\Omega_0\simeq 0.189$ and
$\omega/m\Omega_0\simeq 0.990$, in agreement with \cite{meheut12}.
%consistent with linear calculations of %\cite{meheut12}.  
Although they assumed barotropic perturbations, 
whereas we simulate adiabatic evolution, our linear calculations
indicate growth rates are largely unaffected by entropy gradients (Table \ref{linsims}).

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.62,clip=true,trim=0cm 0.cm 0cm
    0cm]{figures/polar_dens015}
  \caption{Nonlinear hydrodynamic simulation of the RWI in a
    nonhomentropic 3D disk, initially isothermal but
    evolved adiabatically. The axes are in units of $r_0$. The relative density perturbation 
    near the midplane, scaled by 100, is shown. This quantity is
    proportional to the $Q$ used in linear calculations. The snapshot
    corresponds {\bf to the linear phase of the instability}. The drawn line   
    defines the vortex azimuth $\phi_0$ in Fig. \ref{lin_nonlin}---\ref{lin_nonlin_rz}.
    \label{hydro_polar_dens}}
\end{figure}

%use linear results for `big disk', theta_m is 0.39 
We have also computed this mode using the linear code as modified for
strictly isothermal equilibria, with a solid upper boundary. We obtain
growth rate and mode frequency $\nu/\Omega_0 = 0.1937$ and $\omega/m\Omega_0 = 
0.9896$, respectively. This is close to the nonlinear simulation. \cite{meheut12} suggested
smaller growth rate in the latter may be due to numerical viscosity. 
Fig. \ref{lin_nonlin} compares the density
perturbation $Q$ computed from the hydrodynamic simulation and linear code. They are broadly
consistent. The linear code also produces a bias toward the
over-density ahead of the vortex core at the midplane. Away from the
midplane, the center of the anti-cyclonic motion has shifted
downstream. 
This shows that, even within the linear regime, the vortex has
non-negligible vertical structure in the density perturbation (by comparing the two heights in Fig. \ref{lin_nonlin}). 
% for initially isothermal dis

\begin{figure}[!t]
  \centering
  \includegraphics[scale=.425,clip=true,trim=0cm 3.05cm 0cm
    1.3cm]{figures/streamxy_Q015_z0.0.ps}\includegraphics[scale=.425,clip=true,trim=2.2cm 
    3.05cm 0cm 1.3cm]{figures/iso_vel3d_rp_Q_HR_z0.0.ps}\\
  \includegraphics[scale=.425,clip=true,trim=0cm .0cm 0cm
    1.3cm]{figures/streamxy_Q015_z2.0.ps}\includegraphics[scale=.425,clip=true,trim=2.2cm 
    .0cm 0cm 1.3cm]{figures/iso_vel3d_rp_Q_HR_z2.0.ps}
  \caption{Normalized density perturbation, $Q$, associated with the
    RWI computed from a nonlinear hydrodynamic simulation (left) and
    the linear code (right), at the midplane (top) and at $2$
    scale-heights away from the midplane (bottom). The perturbed
    velocity field is also shown. The azimuthal wavenumber is $m=4$.  
  \label{lin_nonlin}}
\end{figure}

We compare meridional flows in Fig. \ref{lin_nonlin_rz}. The perturbed flow
is mostly horizontal in both cases.   
The nonlinear simulation also produce vortical motion in the same sense as
the linear calculation. For the \texttt{ZEUS} calculation, 
  we find the maximum magnitude of 
  vertical Mach number is $\sim 1\%$ with a density-weighted average
  value of $0.15\%$ in the shell $r_\mathrm{sph}\in[0.9,1.1]r_0$. 
The asymmetry of the pressure perturbation
about $r_0$ is captured by the linear code as well. Disagreement toward
the upper boundary is not unexpected, since the linear code assumes
the upper boundary is at a constant number of scale-heights above the
midplane, whereas the spherical grid imposes constant opening angle. 
However, both plots indicate $W$ increases away from the midplane in
the region exterior to $r_0$.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=.425,clip=true,trim=0cm 0.48cm 0cm
    0.4cm]{figures/hydro_vel3d_rz_W.ps}
  \includegraphics[scale=.425,clip=true,trim=0.0cm 
    .48cm 0.cm 0.4cm]{figures/iso_vel3d_rz_W_HR.ps}
  \caption{The perturbed velocity field projected onto the meridional
    plane at the vortex azimuth $\phi_0$, associated with the RWI
    calculated from a nonlinear hydrodynamic simulation (top) and the
    linear code (bottom). The average three-dimensionality, as
      measured by the ratio of vertical to meridional flow speeds,
      $\avg{\theta_m}$, is $0.39$ and $0.34$ in the linear and 
      nonlinear calculation, respectively. 
    A map of the normalized pressure
    perturbation is also shown. 
  \label{lin_nonlin_rz}}
\end{figure}

### 2013-10-05 paper/linear.tex
first commit

\section{Linear problem}\label{linear}
We consider axisymmetric Eulerian perturbations to the above
equilibrium in the form $\real[\dd\rho(z)\exp{\imgi(k_xx + \sigma
    t)}]$ and similarly for other fluid variables. Here, $k_x$ is a constant radial wavenumber and 
$\sigma=-(\omega +\imgi\gamma)$ is a complex 
frequency, where $-\omega$ is the real mode frequency and $\gamma$ is
the growth rate. Hereafter, we surpress the exponential factor, as
well as the real part notation.  We further write 
\begin{align}
  &z=\hat{z}H,\quad k_x =  k_x/H,\quad \sigma = \hat{\sigma}\Omega, \quad \delta\bm{v} = \csmid
  \delta\hat{\bm{v}}, \\ 
  &\delta\bm{B} = B\delta\hat{\bm{B}},\quad
  \delta\rho = \rho\hat{W}/\hat{c}_s^2,\quad \delta\Phi =
  \csmid^2\delta\hat{\Phi}, 
\end{align}
where we have introduced the non-dimensional enthalpy perturbation
$\hat{W}$ and the sound-speed $\hat{c}_s=\csmid^{-1}\sqrt{dP/d\rho}$ can be 
obtained from the equation of state. The background frequencies are
written $S=\hat{S}\Omega$ and $\Omega_z=\hat{\Omega}_z\Omega$ and the 
resistivity as $\eta = \hat{\eta}H^2\Omega$.

We will now drop the $\hat{\phantom{a}}$ notation. Henceforth it is
understood that all variables have been appropriately normalized. The
linearized continuity equation is 
\begin{align}
  \frac{\imgi\sigma}{fc_s^2}W + \imgi k_x \dd v_x
  +\left(\ln\rho\right)^\prime\dd v_z + \dd v_z^\prime = 0,\label{lin_cont}
\end{align}
where $^\prime$ denotes $d/dz$. The linearized equations of motion are
\begin{align}
  &\imgi\sigma \dvx - 2\dvy = f\left[v_A^2\dbx^\prime - \imgi k_x
    \left(\w + v_A^2\dbz
    \right)\right],\\
  &\imgi\sigma\dvy + \frac{\kappa^2}{2}\dvx = fv_A^2\dby^\prime,\\
  & \imgi\sigma\dvz = -f\w^\prime,
\end{align}
where the Alfven speed $v_A =
\left(\beta\rho\right)^{-1/2}$, the effective enthalpy perturbation
$\w = W + \dphi$ and the epicycle frequency $\kappa=\sqrt{2(2-S)}$. 
The linearized induction equation is
\begin{align}
&  \imgi\bar{\sigma}\dbx = f\dvx^\prime +
  \eta\dbx^{\prime\prime}+\eta^\prime\dbx^\prime - \imgi k_x
  \eta^\prime \dbz,\label{induct_x}\\
&\imgi\bar{\sigma}\dby = f\dvy^\prime - S\dbx +
  \eta\dby^{\prime\prime}+\eta^\prime\dby^\prime,\label{induct_y}\\
& \imgi\bar{\sigma}\dbz = -\imgi f k_x \dvx + \eta\dbz^{\prime\prime}, \label{induct_vert}
\end{align}
where $\imgi\bar{\sigma} = \imgi\sigma + \eta k_x^2$, and the
divergence-free condition is $\imgi k_x\dbx + \dbz^\prime=0$. Finally,
the linearized Poisson equation is
\begin{align}
  \dphi^{\prime\prime} - k_x^2\dphi = \frac{\rho}{c_s^2f^2Q}W.  \label{lin_poisson}
\end{align}

We now eliminate $\dd\bm{B}$ and $\dvz$ between the linearized
equations to obtain a system of ordinary differential equations for
$\bm{U}=\left(\dvx,\dvy,W,\dphi\right)$. We detail this technical
procedure in Appendix \ref{reduction}. The schematic problem is
\begin{align}
  L_{11}\dvx + L_{12}\dvy + L_{13}W + L_{14}\dphi &= 0, \label{lin1}\\
  L_{21}\dvx + L_{22}\dvy + L_{23}W + L_{24}\dphi &= 0, \label{lin2}\\
  L_{31}\dvx \quad\quad\quad\,\,\,\,\,\,\, + L_{33}W + L_{34}\dphi &=0,\label{lin3}\\
  \phantom{L_{31}\dvx + L_{22}\dvy +} L_{43}W + L_{44}\dphi
  &=0\label{lin4}, 
\end{align}
where the differential operators $L_{1j}$, $L_{2j}$ and $L_{3j}$ can
be read off Eq. \ref{final_vx}, Eq. \ref{final_vy} and 
Eq. \ref{final_w} respectively, and $L_{4j}$ from
Eq. \ref{lin_poisson}.

Note that $L_{31}\propto k_x$, so that if $k_x=0$ 
then Eq. \ref{lin1}---\ref{lin2} are decoupled from 
Eq. \ref{lin3}---\ref{lin4}.

\subsection{Boundary conditions}
We assume $\bm{U}$ is an even function of $z$, so that $d\bm{U}/dz=0$
at $z=0$. This implies that $\dbx=\dby=0$ at the midplane, consistent
with a highly resistive dead zone. At the upper disk boundary
$z=\zmax$ we assume 
$\dvz=\dbx=\dby=0$, and 
\begin{align}
  \dphi^\prime(\zmax) + k_x\dphi(\zmax) = 0
\end{align}
for the potential perturbation \citep[see][]{goldreich65a}.

\subsection{Numerical procedure}
We use a pseudo-spectral approach to solve the set of linearized
equations. Let
\begin{align}\label{cheby_expand}
  \bm{U}(z) 
  = \sum_{k=1}^{N_z} \bm{U}_k\psi_k(z/\zmax), 
\end{align}
where $  \psi_k  = T_{2(k-1)},  $ and  $T_l$ is a Chebyshev polynomial of the first
kind of order $l$ \citep{stegun65}. %The number of basis functions is
%$N_z$ and $l_\mathrm{max}=2(N_Z-1)$ is the highest polynomial order. 
Note that we only use even Chebyshev polynomials which automatically satisifies the symmetry 
condition at $z=0$. Henceforth we only consider $z\geq 0$. 
The pseudo-spectral coefficients $\bm{U}_n$  are obtained by demanding
the set of linear equations to be satisfied at $N_z$ collocation
points along the vertical direction. We choose these points as the
extrema of $T_{2(N_z-1)}$ plus end points.

This procedure discretizes the linear equations to a matrix equation,
\begin{align}\label{matrix_eqn}
\bm{M}\bm{w} = 0, 
\end{align}
where $\bm{M}$ is a $4N_z\times 4 N_z$ matrix representing the $L_{ij}$ 
and upper disk boundary conditions, 
and $\bm{w}$ is a vector storing the pseudo-spectral coefficients. 
Starting with an initial guess for $\sigma$, non-trivial solutions to
Eq. \ref{matrix_eqn} are obtained by varying $\sigma$ using Newton-Raphson iteration 
such that $\mathrm{det}\bm{M}=0$ \citep[details can be found in][]{lin12}.

\subsection{Diagnostics}

### 2013-10-05 paper/paper.tex
first commit

%\documentclass[12pt]{aastex}
\documentclass[iop]{emulateapj}
%\documentclass{article}
\usepackage{draftwatermark}
\usepackage{paralist}
\shorttitle{MRI in massive disks}
\shortauthors{M-K. Lin}

\usepackage{amsmath}
%\usepackage{url}
\usepackage{bm}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
\newcommand{\zhat}{\hat{z}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\zmax}{Z_s}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\w}{ \tilde{W}}
\newcommand{\dphi}{\dd \Phi}
\defcitealias{lin12}{L12}

\begin{document}

 \title{Resistive magneto-rotational instability in massive
  protoplanetary disks}

\author{Min-Kai Lin}
\affil{Canadian Institute for Theoretical Astrophysics,
  60 St. George Street, Toronto, ON, M5S 3H8, Canada}
\email{mklin924@cita.utoronto.ca}

\begin{abstract}

\end{abstract}

\section{Introduction}

\input{setup}
\input{linear}
%\input{results}
%\input{iso}
%\input{summary}

%% \acknowledgments
%% {\bf
%% I thank the referee, P. Barge, for suggesting the idea of a tilted vorticity column. }
%% I also thank S.-J. Paardekooper for comments on the first version of this paper.

\appendix
\input{appendix}

\bibliographystyle{apj}
\bibliography{ref}

\end{document}

### 2013-10-05 paper/setup.tex
first commit

\section{Local disk model}
We study the local stability of an inviscid, self-gravitating and
magnetized barotropic fluid disk orbiting a central mass with
potential $\Phi_*(r,z)$, where $(r,\varphi,z)$ are cylindrical
co-ordinates from the central mass. We use the shearing box approximation     
\citep{goldreich65b} and consider a small patch of the disk at
$r=r_0$. The local frame rotates at angular velocity 
$\Omega_0=\Omega(r_0,0)$ about the central mass, where $r\Omega^2 =
\p\Phi_*/\p r$. We also define $S=-r\p\Omega/\p r$ as the local shear
rate and $\Omega_z^2\equiv\p^2\Phi_*/\p z^2$ as the square of the
local vertical frequency.

A Cartesian co-ordinate system $(x,y,z)$ is set
up in this local frame, corresponding to the radial, azimuthal and vertical
directions of the global disk, respectively. The shearing box fluid
equations read 
\begin{align} 
  &\frac{\p\rho}{\p t}+\nabla\cdot(\rho\bm{v}) = 0,\\
  &\frac{\p \bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
  2\Omega_0\hat{\bm{z}}\times\bm{v} = - \frac{1}{\rho}\nabla\Pi +
  \frac{1}{\rho\mu_0}\bm{B}\cdot\nabla\bm{B}
  -\nabla\Phi,\\
  &\frac{\p\bm{B}}{\p t}= \nabla\times\left(\bm{v}\times\bm{B} -
  \eta\nabla\times\bm{B}\right), 
\end{align}
where $\rho$ is the density field; $\bm{v}$ is the total velocity in
the local frame; $\bm{B}$ is the magnetic field which satisfies
$\nabla\cdot\bm{B}=0$; $\Pi \equiv P +
|\bm{B}|^2/2\mu_0$ is the total pressure ($\mu_0$ is the vacuum
permeability) and $P=P(\rho)$ is the gas pressure given by a
barotropic equation of state chosen later.

The total potential is $\Phi = \Phi_\mathrm{ext} + \Phi_d$, where
\begin{align}
  \Phi_\mathrm{ext}(x,z) = -\Omega_0 S_0 x^2 +
  \frac{1}{2}\Omega_{z0}^2z^2 
\end{align}
is the effective external potential (central plus centrifugal) in the
shearing box approximation, where $S_0\equiv S(r_0,0)$ and
$\Omega_{z0}\equiv\Omega_z(r_0,0)$; 
%% \begin{align}
%%   S\equiv -r\frac{\p\Omega}{\p r}, \quad\quad \Omega_{z0}^2\equiv
%%   \frac{\p^2\Phi_*}{\p z^2} 
%% \end{align}
%is the local shear rate and vertical frequency respectively; 
and the gas potential $\Phi_d$ statisfies Poisson's equation
\begin{align}
  \nabla^2\Phi_d = 4\pi G \rho, 
\end{align}
where $G$ is the gravitational constant. For clarity, hereafter we
drop the subscript $0$ on the frequencies.

\subsection{Self-gravitating equilibria} 
The unperturbed disk is steady and described by
$\rho=\rho(z)$, $\bm{B} = B\hat{\bm{z}}$ where $B$ is a constant, and
velocity field $\bm{v} = -Sx\hat{\bm{y}}$. We have assumed a thin disk 
and neglected the radial component of the self-gravitational force in
the unperturbed disk.

The equilibrium density field is obtained by solving
\begin{align}
  &0=\frac{1}{\rho}\frac{d P}{dz} + \Omega_z^2z + \frac{d\Phi_d}{dz},\label{eqm_eqns1}\\
  &\frac{d^2\Phi_d}{dz^2} = 4\pi G \rho.\label{eqm_eqns2}
\end{align}
We consider \begin{inparaenum}[(i)]
\item isothermal disks with $P=c_{s0}^2\rho$\label{iso_eos}; 
\item polytropic disks with $P=K\rho^2$ where $K=c_{s0}^2/2\rho_0$ and $\rho_0\equiv\rho(0)$
\end{inparaenum}.
In both cases $c_{s0}$ is the midplane sound speed. For the polytropic
disk we define the disk thickness as $H$ such that $\rho(H)=0$. In the
isothermal case we define a characterisic disk thickness as
$H=c_{s0}/\Omega$, which is the pressure scale-height in the
non-self-gravitating limit.

We solve for $\hat{\rho}\equiv\rho/\rho_0$ subject to $\hat{\rho}=1$
and $d\hat{\rho}/dz=0$  at $z=0$. This is done numerically for
isothermal disks and analytically for the polytropic disk (see
Appendix \ref{appen1}). Examples of density profiles are shown in
Fig. \ref{eqm_den}.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_iso_density}
  \includegraphics[width=\linewidth]{figures/compare_poly_density}
  \caption{caption\label{eqm_den}}
\end{figure}

\subsection{Resistivity profile}
Following \cite{fleming03}, we adopt the resistivity profile
\begin{align}
  \eta(z) =
  \sqrt{2}\eta_0\left[\exp{\left(-g_+\right)}+\exp{\left(-g_-\right)}\right]^{-1/2},  
\end{align}
where
\begin{align}
  &g_\pm(z) =  \frac{\Sigma_\pm(z)-\Sigma_0}{\Sigma_*}, \\
  &\Sigma_\pm(z) = \int_{\pm z}^\infty\rho(z^\prime)dz^\prime, \label{sigma_pm}
\end{align}
and $\Sigma_0\equiv\Sigma_{\pm}(0)$, so that $g_\pm(0)=0$ and $\eta_0 
= \eta(0)$. The constant $\Sigma_*$ is chosen such that 
\begin{align}
  \cosh{\left(\frac{\Sigma_0}{\Sigma_*}\right)} =
  \left[\frac{\eta_0}{\eta(\infty)}\right]^2,
\end{align}
and we define $\eta_0/\eta(\infty)\equiv A_\eta$ as the conductivity
boost factor from the midplane to the disk surface. We remark that
once the equilibrium $\rho$ and $d\rho/dz$ are obtained from
Eq. \ref{eqm_eqns1} --- \ref{eqm_eqns2}, the integration for
Eq. \ref{sigma_pm} can be done implicitly by using the Poisson
equation.

\subsection{System parameters}
The strength of self-gravity is parametrized by
\begin{align}
  Q \equiv \frac{\Omega^2}{4\pi G\rho_0}
\end{align}
\citep{mamat10}. The plasma $\beta$ measures the inverse strength of
the magnetic field,
\begin{align}
  \beta \equiv \frac{\csmid^2}{v_{A0}^2} =
  \frac{\csmid^2\mu_0\rho_0}{B^2},  
\end{align}
where $v_{A0}$ is the midplane Alfven speed. The strength of
resistivity is measured by the midplane Elsasser number 
\begin{align}
  \Lambda \equiv \frac{v_{A0}^2}{\eta_0\Omega} = \frac{f^2 R_m}{\beta}, 
\end{align}
where the second equality defines the magnetic Reynolds number $R_m$
and $f\equiv \csmid/H\Omega$ is a numerical factor of order unity.

### 2013-10-05 paper/summary.tex
first commit

\section{Summary and discussion}\label{summary}
In this paper, we have examined the linear stability of radially
structured three-dimensional disks with non-uniform entropy
distribution. These calculations may be considered as an extension to 
the 2D Rossby wave instability \citep{li00} by adding the vertical
dimension, or to the barotropic RWI calculations of \citetalias{lin12} by
adding an energy equation with a simpler numerical method.

We adopted polytropic disk equilibria so that
the magnitude of entropy gradients can be conveniently parametrized
by $\Delta\gamma\equiv\gamma/\Gamma$, and we focused on the effect of
$\Delta\gamma\geq1$. When the background density and
velocity field is fixed through $\Gamma$, we found increasing $\Delta
\gamma$ has negligible effect on the instability growth rate. However,
{\bf the magnitude of} pressure and density perturbations increase with height, and the 
meridional flow associated with the vortex core is qualitatively
changed, with the introduction of {\bf meridional} vortical
motion.

{\bf Meridional vortical motion was found to correlate with a small 
  tilt of a fluid column with negative vertical vorticity
  perturbation.  In standard hydrodynamics, vorticity tilting
  can originate from a contribution of the
  $\bm{\omega}\cdot\nabla\bm{v}$ term in the evolution equation of the
  vorticity independently of the baroclinic source term $\nabla \rho
  \times\nabla p$. However, given the tilt is absent in our homentropic
  calculations, we associate the tilt with the baroclinic source term, which
  produces azimuthal vorticity. } 
We also found
that the vertical velocity at the vortex core is no longer linear in
$z$, as for homentropic flow.

In our second set of experiments, we fixed $\gamma$ and decreased
$\Gamma$. We found that by making the flow nonhomentropic, the
co-rotation region became \emph{more} three-dimensional, despite the
decrease in growth rate. This result
is opposite to \citetalias{lin12} where lowering $\Gamma$ made the
flow less three-dimensional. This implies that entropy gradients  
play an important role in the vertical structure of the perturbations. 
%This will likely have impact
%on the motion of dust particles within Rossby vortices, in particular 
%as the flow is dominated by bouyancy forces away from the midplane.

We also considered isothermal equilibria. A linear calculation with 
$\Gamma=1.1$ and one with a strictly isothermal setup ($\Gamma\equiv
1$) were consistent. Both produced prominent meridional 
vortical motion. In order to verify this feature, we ran a nonlinear
simulation of the RWI in an initially isothermal disk, but evolved
adiabatically. We indeed identified said  vortical motion.    
%although the agreement with linear calculations is
%qualitative. This might imply that nonlinear effects become important
%for the RWI vertical structure earlier than in the horizontal
%direction. 
Keeping in mind that the setup for linear and nonlinear simulations 
were not identical (e.g. numerical grid, boundary treatment), similarities
between them, such as mode frequency, growth rate and horizontal flow,
are satisfactory.

Vortical motion in the meridional plane thus appears characteristic of 
the \emph{linear} RWI in nonhomentropic disks. Whether or not this is 
significant for the vortex evolution can only be answered by detailed 
long term nonlinear simulations. If this vortical motion is present 
in the nonlinear regime then it may prevent dust particles from reaching the disk surface, 
which occurs for homentropic flow \citep{meheut12c}.

However, given this meridional vortical motion is absent in the homentropic 
linear solution, it may eventually vanish because of entropy
mixing, if no mechanism is present to maintain entropy gradients. 
For example, the background entropy increases with height but 
the linear entropy perturbation becomes more negative with height, {\bf 
and its magnitude grows exponentially in time}. 
{\bf Indeed,  recent 3D fully compressible simulations in
  nonhomentropic disks shows that well into the nonlinear regime,
  Rossby vortices are columnar
  \citep{richard13}}. On the other hand, \cite{meheut12} observed
strong meridional vortical motion in their homentropic hydrodynamic
simulations; we conclude they are of nonlinear origin.

In the linear solutions, we often observe perturbation magnitudes
increase away from the midplane in nonhomentropic
disks\footnote{This reminds us of the off-midplane vortices
  discovered by \cite{barranco05} in nonlinear local simulations, 
%  which suggest the non-zero vertical bouyancy frequency to be
%  stabilizing   against secondary instabilities \citep{lesur09}. 
  but the  
  setup considered in that study is very different from the
  present work. Nevertheless, in both cases the vertical entropy
  gradient is stabilizing away from the midplane. 
  %%  Nevertheless, a positive vertical bouyancy frequency
  %% may stabilize 3D vortices \citep{lesur09}.
} 
(e.g. Fig. \ref{lin_nonlin_rz}). Then the RWI may not be as robust
against vertical boundary conditions as it is to radial boundary
conditions. This could pose difficulty for the RWI to develop in dead
zones of real protoplanetary disks, which are expected to be confined
from above and below by magnetically turbulent layers
\citep{oishi09}. The vertical boundary condition set by these layers
may or may not be compatible with the linear RWI solution.

\subsection{Caveats and outlooks}
%need a good guess
One trade-off for the simplicity of our numerical method for 
linear simulations is that a trial eigenfrequency must be
guessed. This is not a significant obstacle for the problem at
hand, because previous RWI studies provide an important guide
\citep[][]{li00}. Otherwise, zeros of the complex function
$\mathcal{D}(\sigma)=\mathrm{det}\,\bm{U}$ need to be located with more 
rigorous methods \citep[e.g.][]{kojima86,valborro07}. We have also exploited
previous findings that the PPI and RWI are predominantly two-dimensional 
\citep{papaloizou85, goldreich86, kojima89, umurhan10, meheut12, lin12, lin12c}, which
enabled the use of a small number of basis functions. 
However, there could exist parameter regimes where the RWI
has significant vertical structure, rendering our solution method
inefficient.

%extension to baroclinic equilibria 
%simple polytrope background
Our conclusions are limited to polytropic backgrounds. While this was 
convenient for numerical experiments, it is an over-simplification
of protoplanetary disks, which are 
expected to have complicated vertical structure \citep{terquem08}. In
particular, we found that entropy gradients plays a role in the
vertical structure of the {\bf linear} RWI, and even a small entropy gradient can noticeably
modify the vertical flow (\S\ref{nz_nonzero}).
Thus, a realistic model for entropy evolution is needed. 
%This is important because even a small entropy gradient can noticeably 
%modify the vertical flow (\S\ref{nz_nonzero}).

It would also be of interest to generalize the calculations
to \emph{baroclinic} equilibria\footnote{In fact, baroclinic tori were
  briefly considered by \cite{frank88}.}, 
for which $\p_z\Omega\neq0$. This may well be the 
case when the equilibrium pressure depends on both the density and temperature. 
%This requires 
%derivation of the linearized equations including terms involving $\p_z\Omega$. 
Complications from baroclinic instabilities may arise, however \citep{knobloch86, umurhan12, nelson12}.
%% This may involve adding heating or cooling
%% sources to the energy equation. 
%source terms - friction, cooling
%new physics - self-gravity

We have neglected gas self-gravity in this study. Our models therefore
assume that the Toomre parameter is much larger than unity in both the 
unperturbed \emph{and} perturbed states. {\bf However, self-gravity
  may affect the RWI even when the Toomre parameter is not small
  \citep{lovelace12}}. 
Previous studies
have found  higher $m$ RWI modes are favored when disk self-gravity is
included \citep{lyra08,lin11a}. Recent 3D simulations of the RWI in a
locally isothermal disk show that vertical self-gravity can noticeably
enhance the density perturbation near the midplane,
even though the initial disk was considered low mass \citep{lin12b}. %%  The effects
%%  of self-gravity %% should be considered in conjunction with disk
%%  thermodynamics. %In a nonlinear simulation, the latter may not hold even if the former
%does.
%Previous studies of the RWI also find higher $m$ modes modes are
%favored when disk self-gravity is included
%\citep{lyra08,lin11a}.

In principle, one can express the Poisson integral as a matrix
operator and  incorporate it into our formalism. {\bf The linear} problem is
further complicated by the need of a numerical solution to the
equilibrium equations describing a radially structured,
self-gravitating 3D disk \citep{muto11}.  Such a linear calculation is
beyond the scope of this paper, but will be inevitable for
understanding  the RWI in 3D self-gravitating disks. Perhaps a simpler
starting point, to gain first insight, is  direct hydrodynamic 
simulations {\bf including disk gravity}. This is indeed the approach taken in our follow-up paper.    
%\clearpage
%since the ZEUS-MP code already includes a Poisson solver (as used in Lin 2012b).

%have shown that
%self-gravity reduces the RWI growth rate through the potential perturbation, 
%but may be destabilizing through modification of the background disk \citep{lin11a}.

### 2013-10-05 non-ideal/functions.f90

@@ -1,0 +1,598 @@
+SUBROUTINE F (NEQ, x, Y, YDOT, RPAR, IPAR)
+  use global
+  implicit none
+  integer :: NEQ, IPAR
+  real*8 :: x, Y(NEQ), YDOT(NEQ), RPAR
+  real*8 :: logrho, phidot
+  
+  logrho = Y(1)
+  phidot = Y(2)
+  
+  YDOT(1) = -omegaz2*x - phidot
+  YDOT(2) = exp(logrho)/bigQ
+  
+  return
+end subroutine F
+
+subroutine jac
+  use global
+  implicit none
+  return
+end subroutine jac
+
+subroutine setup_isosg
+  use global
+  implicit none
+  integer, parameter :: NEQ = 2, ITOL = 1 , MF = 10, LIW = 60, IOPT = 0
+  integer, parameter :: LRW = 2*(20 + 16*NEQ)
+  integer :: i, ITASK, ISTATE, IWORK(LIW), IPAR
+  real*8, parameter :: RTOL = 1d-15, ATOL=1d-15
+  real*8 :: Y(NEQ), x, xOUT, RWORK(LRW), RPAR
+  external :: F, JAC
+  
+  Y(1) = 0d0 !log rho_hat = 0 at z=0
+  Y(2) = 0d0 !dphi/dz =0 at z=0
+  
+  dnorm(1)   = 1d0
+  drho(1)    = 0d0
+  
+  ITASK = 1
+  ISTATE= 1
+  
+  do i = 1, nz-1
+     x    = zaxis(i)
+     xOUT = zaxis(i+1)
+     
+     call DVODE (F, NEQ, Y, x, xOUT, ITOL, RTOL, ATOL, ITASK, &
+          ISTATE, IOPT, RWORK, LRW, IWORK, LIW, JAC, MF, &
+          RPAR, IPAR)
+     
+     if(ISTATE .ne. 2) then
+        print*, 'setup unsuccessful'
+        stop
+     endif
+     
+     dnorm(i+1) = exp(Y(1))
+     drho(i+1)  = -omegaz2*zaxis(i+1) - Y(2)
+
+  enddo
+  d2rho= ( drho**2d0 - (omegaz2 + dnorm/bigQ)/fQ**2d0 )
+  valf = 1d0/(beta*dnorm)
+  csq  = 1d0 
+  return
+end subroutine setup_isosg
+
[…]

### 2013-10-05 non-ideal/sgmri.f90

@@ -1,0 +1,241 @@
+module global
+  logical :: refine, prin
+  character*3 :: eos, method
+  character*4 :: var 
+  integer, parameter :: nvar=4
+  integer :: nz, nk, big_nz, ntrials , kcount 
+  real*8, parameter :: omega=1.0, bigG = 1.0, mstar = 1.0, r0=1.0
+  real*8, parameter :: pi = 2d0*acos(0d0) 
+  real*8 :: bigQ, beta, shear, kmin, kmax, kx, dk, qmin, qmax, bmin, &
+       bmax, var_min, var_max 
+  real*8 :: zmax, dz    
+  real*8 :: kappa, omegak, omegaz2, fQ, Rm, amp, zstar, Rm_max, Rm_min
+  real*8 :: condition, min_growth, max_growth, sig_re, sig_im    
+  real*8, allocatable :: dnorm(:), drho(:), d2rho(:), valf(:), zaxis(:),&
+       kaxis(:), csq(:), eta(:), deta(:), d2eta(:) 
+  real*8, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)  
+  complex*16, parameter :: ii = (0d0, 1d0)      
+  complex*16 :: sig 
+  complex*16, allocatable :: vx(:), vy(:), bz(:)
+  complex*16, allocatable :: den(:), pot(:), wvector(:), sig_trials(:)
+end module global
+
+program sgmri
+  use global 
+  implicit none 
+  integer :: i, j, k  
+  integer :: haf_lmax, lmax  
+  character*2 :: str_mode 
+  real*8  :: zbar, T_l, dT_l, d2T_l, lmode, dummy
+  real*8, external :: density, dlogrho, alfven
+  namelist /params/ bigQ, beta, kx, var, shear, kmin, kmax, qmin,&
+       & qmax, bmin, bmax, eos 
+  namelist /resis/ Rm, amp, Rm_min, Rm_max
+  namelist /grid/ zmax, nz, nk  
+  namelist /eigen/ method, min_growth, max_growth, ntrials , refine, prin
+  
+  write(6,fmt='(A)') '--------------- welcome ---------------'
+  
+  !read input parameters.
+  open(7, file="input")
+  read(7, nml=params)
+  read(7, nml=resis)
+  read(7, nml=grid)
+  read(7, nml=eigen) 
+  close(7)
+
+  if( (eos.eq.'pol').and.(zmax .ge. 1d0)) then
[…]

### 2013-10-05 paper/makepdf.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+latex paper.tex
+dvips -o paper.ps paper.dvi
+ps2pdf -sPAPERSIZE=a4 paper.ps
+echo 'done'

### 2013-10-05 paper/makepdf.sh
clean up

@@ -1,5 +1,0 @@
-#!/bin/bash
-latex paper.tex
-dvips -o paper.ps paper.dvi
-ps2pdf -sPAPERSIZE=a4 paper.ps
-echo 'done'

### 2013-10-06 paper/linear.tex
minor paper update

@@
 \begin{align}
-  &z=\hat{z}H,\quad k_x =  k_x/H,\quad \sigma = \hat{\sigma}\Omega, \quad \delta\bm{v} = \csmid
+  &z=\hat{z}H,\quad k_x =  k_x/H,\quad \sigma = \hat{\sigma}\Omega,
+  \quad \delta\bm{v} = \csmid 
   \delta\hat{\bm{v}}, \\ 
@@
 equations to obtain a system of ordinary differential equations for
-$\bm{U}=\left(\dvx,\dvy,W,\dphi\right)$. We detail this technical
-procedure in Appendix \ref{reduction}. The schematic problem is
+$\bm{U}=\left(\dvx,\dvy,W,\dphi\right)$. We detail the steps 
+in Appendix \ref{reduction}. The problem is schematically 
 \begin{align}
@@
 
-Note that $L_{31}\propto k_x$, so that if $k_x=0$ 
-then Eq. \ref{lin1}---\ref{lin2} are decoupled from 
-Eq. \ref{lin3}---\ref{lin4}.
+Note that $L_{31}\propto k_x$, so for $k_x=0$ 
+Eq. \ref{lin1}---\ref{lin2} are decoupled from 
+Eq. \ref{lin3}---\ref{lin4}. In this case, a solution is $\dphi=W=0$,
+with $\dvx$ and $\dvy$ determined by Eq. \ref{lin1}---\ref{lin2},
+which is the incompressible MRI problem. This means that the density
+and potential perturbations can only influence the MRI for
+$k_x\neq0$. However, SG will still have an effect through the
+background equilibrium for $k_x=0$.
 
@@
 %$N_z$ and $l_\mathrm{max}=2(N_Z-1)$ is the highest polynomial order. 
-Note that we only use even Chebyshev polynomials which automatically satisifies the symmetry 
-condition at $z=0$. Henceforth we only consider $z\geq 0$. 
+Note that we only use even Chebyshev polynomials so the symmetry 
+condition at $z=0$ is satisfied. Henceforth we only consider $z\geq 0$. 
 The pseudo-spectral coefficients $\bm{U}_n$  are obtained by demanding
@@
 \subsection{Diagnostics}
+We visualize results in terms of energy densities. We define
+\begin{align}
+  &E_1 \equiv \frac{|\delta \bm{B}|^2}{2\beta},\\
+  &E_2 = \frac{\rho}{\hat{c}_s^2}|\real\left(W\dphi^*\right)|,\\
+  &E_3 = \frac{1}{2}\rho|\delta\bm{v}|^2,\\
+  &E_4 = \frac{\rho|W|^2}{2\hat{c}_s^2},
+\end{align}
+as the perturbed magnetic, gravitational, kinetic and thermal
+energies, respectively, which are functions of $z$. We
+then normalize each by $\mathrm{max}(E_1)$.

### 2013-10-06 paper/setup.tex
minor paper update

@@
 Appendix \ref{appen1}). Examples of density profiles are shown in
-Fig. \ref{eqm_den}.
+Fig. \ref{eqm_den}. The isothermal disk shows that self-gravity
+compresses the disk and reduces the true pressure scale-height. The
+polytropic disk shows little difference between different strengths of
+self-gravity because the vertical co-ordiante is already normalized by
+the disk thickness.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_poly_density}
-  \caption{caption\label{eqm_den}}
+  \caption{Equilibrium density field from solving Eq. \ref{eqm_eqns1}
+    --- \ref{eqm_eqns2} subject to an isothermal (top) and polytropic
+    (bottom) equation of state. Note that the vertical co-ordinate is
+    normalized differently for convenience. In the isothermal disk,
+    $H$ is the characteristic scale-height in the
+    non-self-gravitating limit; in the polytropic disk $H$ is the
+    disk thickness. \label{eqm_den}}
 \end{figure}
@@
 \end{align}
-and we define $\eta_0/\eta(\infty)\equiv A_\eta$ as the conductivity
+and we define $\eta_0/\eta(\infty)\equiv A$ as the conductivity
 boost factor from the midplane to the disk surface. We remark that

### 2013-10-07 paper/results.tex
minor paper update

\section{Results}

\subsection{Reference cases}
%ideal, non sg

%ideal, sg

%non ideal smooth, non sg

%non ideal struct. non sg

%non ideal smooth, sg

%non ideal struc, sg

%grav instab.

### 2013-10-23 non-ideal/sgmri.f90
code bug fix, poly elsass fig

@@ -182,22 +182,31 @@ program sgmri
   
   !solve the eigenvalue problem for each k (= bigQ, beta or kx)
   open(99, file='basic.dat')
+  open(100, file='params.dat')
+  
   if(nk .eq. 1) then
+
      call get_basic 
+     
+     write(100, fmt='(5(e22.15,x))') zmax, dble(nz), fQ, shear, Rm
+     
      do j=1, nz 
         write(99, fmt='(7(e22.15,x))') zaxis(j), dnorm(j), valf(j),&
              & csq(j), eta(j), drho(j), deta(j)
      enddo
+     
+     
   endif
-
+  
   open(20, file='eigenvalues.dat')  
   do i=1, nk 
      kcount = i      
      if(nk .gt. 1) then
+        
         if(var.eq.'grav') then
            bigQ = kaxis(i)
            write(6,fmt='(A,A,A,f5.2,A)') '---------------', var, '=',bigQ,'---------------'
-           call get_basic 
+           call get_basic
         endif
         if(var.eq.'beta') then
            beta = 10d0**kaxis(i)
@@ -214,10 +223,14 @@ program sgmri
            write(6,fmt='(A,A,a,f5.2,A)') '---------------', var, '=', Rm,'----------------'
            call get_basic
         endif 
+
+        write(100, fmt='(5(e22.15,x))') zmax, dble(nz), fQ, shear, Rm
         do j=1, nz 
            write(99, fmt='(7(e22.15,x))') zaxis(j), dnorm(j), valf(j)&
                 &, csq(j), eta(j), drho(j), deta(j)
         enddo
+
+        
      endif
     
      write(str_mode, fmt='(I2)') i 
@@ -238,4 +251,5 @@ program sgmri
   enddo
   close(20)!eigenvalue.dat
   close(99)!basic.dat 
+  close(100)!params.dat
 end program sgmri

### 2013-10-23 paper/setup.tex
minor paper update
minor tex update

@@
 
-\subsection{Resistivity profile}
+\subsection{Resistivity profile}\label{resis_profile}
 Following \cite{fleming03}, we adopt the resistivity profile
@@
 
+Example resistivity profiles are shown in Fig. \ref{eqm_resis} for the isothermal disk. 
+The midplane Elsasser number is $\Lambda(0)=0.1$ and the conductivity is boosted by a factor $A=10$ 
+from the midplane to the upper disk boundary. The figure compares a strongly self-gravitating case
+($Q=0.2$) and the non-self-gravitating limit ($Q=10$).
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/elsasser_iso}
+  \caption{Elsasser number as prescribed in \S\ref{resis_profile} for
+    isothermal disks with $Q=0.2$ (solid) 
+    and $Q=10$ (dotted). The vertical line
+    indicates $\Lambda=1$ in each case. \label{eqm_resis}}
+\end{figure}
+
 \subsection{System parameters}

### 2013-10-23 non-ideal/functions.f90
kx=0 special code

@@ -404,20 +404,20 @@ subroutine construct_matrix(big_matrix)
 
    big_matrix(1:nz, 1:nz)         = L11
    big_matrix(1:nz, nz+1 : 2*nz)  = L12
-   big_matrix(1:nz, 2*nz+1:3*nz)  = L13
-   big_matrix(1:nz, 3*nz+1:4*nz)  = L14 
+!   big_matrix(1:nz, 2*nz+1:3*nz)  = L13
+!   big_matrix(1:nz, 3*nz+1:4*nz)  = L14 
 
    big_matrix(nz+1:2*nz, 1:nz)         = L21
    big_matrix(nz+1:2*nz, nz+1 : 2*nz)  = L22
-   big_matrix(nz+1:2*nz, 2*nz+1:3*nz)  = L23
-   big_matrix(nz+1:2*nz, 3*nz+1:4*nz)  = L24
+!   big_matrix(nz+1:2*nz, 2*nz+1:3*nz)  = L23
+!   big_matrix(nz+1:2*nz, 3*nz+1:4*nz)  = L24
 
-   big_matrix(2*nz+1:3*nz, 1:nz)         = L31
-   big_matrix(2*nz+1:3*nz, 2*nz+1:3*nz)  = L33
-   big_matrix(2*nz+1:3*nz, 3*nz+1:4*nz)  = L34
+!   big_matrix(2*nz+1:3*nz, 1:nz)         = L31
+!   big_matrix(2*nz+1:3*nz, 2*nz+1:3*nz)  = L33
+!   big_matrix(2*nz+1:3*nz, 3*nz+1:4*nz)  = L34
 
-   big_matrix(3*nz+1:4*nz, 2*nz+1:3*nz)  = L43
-   big_matrix(3*nz+1:4*nz, 3*nz+1:4*nz)  = L44
+!   big_matrix(3*nz+1:4*nz, 2*nz+1:3*nz)  = L43
+!   big_matrix(3*nz+1:4*nz, 3*nz+1:4*nz)  = L44
   return
 end subroutine construct_matrix
 
@@ -587,8 +587,8 @@ subroutine maximize_growth_rate
   call eigenvalue_problem(output)
   vx(:)  = wvector(1:nz)
   vy(:)  = wvector(nz+1:2*nz)
-  den(:) = wvector(2*nz+1:3*nz)
-  pot(:) = wvector(3*nz+1:4*nz)
[…]

### 2013-10-23 non-ideal/sgmri.f90

@@ -2,7 +2,7 @@ module global
   logical :: refine, prin
   character*3 :: eos, method
   character*4 :: var 
-  integer, parameter :: nvar=4
+  integer, parameter :: nvar=2
   integer :: nz, nk, big_nz, ntrials , kcount 
   real*8, parameter :: omega=1.0, bigG = 1.0, mstar = 1.0, r0=1.0
   real*8, parameter :: pi = 2d0*acos(0d0) 

### 2013-10-23 non-ideal/sgmri.f90
minor code clean up

@@ -99,18 +99,6 @@ program sgmri
   omegak  = sqrt(bigG*mstar/r0**3d0)/omega 
   kappa   = sqrt(2.0*(2.0 - shear))
   omegaz2 = omegak**2d0 
-  if(eos .eq. 'pol') then
-     fQ      = sqrt(bigQ)*acos(omegaz2*bigQ/(1d0 + omegaz2*bigQ)) !H*Omega/cs0 
-     fQ      = 1d0/fQ 
-  else if((eos.eq.'iso').or.(eos.eq.'uns')) then
-     fQ = 1d0
-  endif
- 
-  !output fiducial parameters
-  open(10, file='params.dat')
-  write(10, fmt='(5(e22.15,x))') zmax, dble(nz), fQ, shear, Rm
-  close(10)
-
  
   big_nz = nvar*nz 
 

### 2013-11-12 paper/results.tex
paper update, begun results section

@@
 
-\subsection{Reference cases}
-%ideal, non sg
+\subsection{MRI in self-gravitating polytropic disks with uniform
+  resistivity}  
+We first calculate the MRI as a function of $Q$ in a polytropic disk
+with uniform resistivity ($A=1$). Here, we consider $Q\in[0.2,4]$,
+corresponding to stronly self-gravitating to non-self-gravitating
+disks, for mid-plane Elsasser numbers $\Lambda_0\in[0.3,10]$. We fix $k_x=0.1$ and 
+$\beta=100$.
 
-%ideal, sg
+It is useful to first see how self-gravity may still play a role even
+though density and potential perturbations are expected to be
+negligible because $k_x\ll1$, i.e. the linear response is
+non-self-gravitating. 
+%even though the linear response is expected to be incompressible
+%because $k_x\ll 1$, i.e. non-self-gravitating. 
+\cite{sano99} found that the MRI wavelength, normalized by $H$, can be estimated as
+\begin{align}
+  \lambda=\mathrm{max}\left(\lambda_\mathrm{ideal},\lambda_\mathrm{resis}\right), 
+\end{align}
+where
+\begin{align}
+  \lambda_\mathrm{ideal} = f v_A = \frac{f}{\sqrt{\beta\rho}} 
+\end{align}
+is the ideal MRI wavelength, and 
+\begin{align}
+  \lambda_\mathrm{resis} =\frac{\eta}{f v_A}=\frac{f}{\Lambda_0}\sqrt{\frac{\rho}{\beta}}
+\end{align}
+is the MRI wavelength in the limit of high resistivity.
 
-%non ideal smooth, non sg
+\subsection{Resistive MRI in massive polytropic disks with layered
+  resistivity}
 
-%non ideal struct. non sg
+\subsection{GI in magnetized isothermal disks with uniform
+  resistivity}
 
-%non ideal smooth, sg
-
-%non ideal struc, sg
-
-%grav instab.
+\subsection{GI in magnetized isothermal disks with layered
+  resistivity}

### 2013-11-12 paper/setup.tex
paper update, begun results section

@@
 Example resistivity profiles are shown in Fig. \ref{eqm_resis} for the isothermal disk. 
-The midplane Elsasser number is $\Lambda(0)=0.1$ and the conductivity is boosted by a factor $A=10$ 
-from the midplane to the upper disk boundary. The figure compares a strongly self-gravitating case
+The conductivity is boosted by a factor $A=10$ from the midplane to
+the upper disk boundary. The figure compares a strongly self-gravitating case
 ($Q=0.2$) and the non-self-gravitating limit ($Q=10$).
@@
 \begin{align}
-  \Lambda \equiv \frac{v_{A0}^2}{\eta_0\Omega} = \frac{f^2 R_m}{\beta}, 
+  \Lambda_0 \equiv \frac{v_{A0}^2}{\eta_0\Omega} = \frac{f^2 R_m}{\beta}, 
 \end{align}
@@
 and $f\equiv \csmid/H\Omega$ is a numerical factor of order unity.
+The Elsasser number as a function of height is $\Lambda(z)\equiv
+v_A^2/\eta\Omega$.

### 2013-11-13 paper/appendix.tex
text update

@@
 \end{align}
+Given a fixed mid-plane temperature, the function $f(Q)\equiv
+\csmid/\Omega H$ is an inverse measure of the disk thickness, and $f$ increases with decreasing $Q$,
+as shown in Fig. \ref{plot_fq}. This corresponds to a thinner disk
+with increasing strength of vertical self-gravity.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/plot_fq}
+  \caption{The function $f(Q)$ describing vertical hydrodstatic 
+    equilibrium in self-gravitating polytropic disks (solid line). The
+    horizontal dashed line is the asymptotic value of $1/\sqrt{2}$ for
+    large $Q$.  
+    \label{plot_fq}}
+\end{figure}
 
@@
 \begin{align}
-  \hat{D}_0 = \eta D_0, \quad \hat{D}_1 = \eta^\prime D_0 + \eta
-  D_1,\quad \hat{D}_2 = \eta^{\prime\prime} D_0 + 2\eta^\prime D_1 +
+  \overline{D}_0 = \eta D_0, \quad \overline{D}_1 = \eta^\prime D_0 + \eta
+  D_1,\quad \overline{D}_2 = \eta^{\prime\prime} D_0 + 2\eta^\prime D_1 +
   \eta D_2. 
@@
 \begin{align}\label{bz_eq}
-  -\sigma\dbz^{(n)} = f k_x \dvx^{(n)} + \frac{k_x\beta\rho}{f}\hat{D}_n\left(\imgi\sigma\dvx - 2\dvy +
+  -\sigma\dbz^{(n)} = f k_x \dvx^{(n)} + \frac{k_x\beta\rho}{f}\overline{D}_n\left(\imgi\sigma\dvx - 2\dvy +
   \imgi f k_x\w \right). 
@@
   \imgi\sigma \frac{f}{\beta\rho}\dbx =
-  \frac{f^2}{\beta\rho}\dvx^\prime + \hat{D}_1\left(\imgi\sigma\dvx -
+  \frac{f^2}{\beta\rho}\dvx^\prime + \overline{D}_1\left(\imgi\sigma\dvx -
   2\dvy + \imgi f k_x\w\right). \label{bx_expression}
@@
   0 = \frac{f^2}{\beta\rho}\dvx^{\prime\prime} -
-  \frac{f^2}{\beta\rho}k_x^2\dvx + \left(\hat{D}_2 - k_x^2 \hat{D}_0
+  \frac{f^2}{\beta\rho}k_x^2\dvx + \left(\overline{D}_2 - k_x^2 \overline{D}_0
     - \imgi\sigma D_0\right)\left( \imgi\sigma\dvx -
@@
   \imgi\bar{\sigma}\dby = f \dvy^\prime + \frac{\imgi S}{\sigma}\left[
-    f\dvx^\prime + \frac{\beta\rho}{f}\hat{D}_1\left(\imgi\sigma\dvx -
+    f\dvx^\prime + \frac{\beta\rho}{f}\overline{D}_1\left(\imgi\sigma\dvx -
   2\dvy + \imgi f k_x\w \right)\right] +
-  \frac{\beta\rho}{f}\hat{D}_1\left(\imgi\sigma\dvy +
+  \frac{\beta\rho}{f}\overline{D}_1\left(\imgi\sigma\dvy +
   \frac{\kappa^2}{2}\dvx\right). \label{by_expression}
@@
     \frac{\imgi\sbar\kappa^2}{2}D_0\dvx \notag\\
-    & + \left(\hat{D}_2 -
-    \frac{\sbar^\prime}{\sbar}\hat{D}_1\right)\left[\imgi\left(\sigma
+    & + \left(\overline{D}_2 -
+    \frac{\sbar^\prime}{\sbar}\overline{D}_1\right)\left[\imgi\left(\sigma
       - \frac{2S}{\sigma}\right)\dvy + \left(\frac{\kappa^2}{2} -

### 2013-11-13 paper/results.tex
text update

@@
 We first calculate the MRI as a function of $Q$ in a polytropic disk
-with uniform resistivity ($A=1$). Here, we consider $Q\in[0.2,4]$,
+with uniform resistivity ($A=1$). We consider $Q\in[0.2,4]$,
 corresponding to stronly self-gravitating to non-self-gravitating
@@
 %because $k_x\ll 1$, i.e. non-self-gravitating. 
-\cite{sano99} found that the MRI wavelength, normalized by $H$, can be estimated as
+\cite{sano99} found that the MRI wavelength, normalized by $2H$, can be estimated as
 \begin{align}
@@
 \begin{align}
-  \lambda_\mathrm{ideal} = f v_A = \frac{f}{\sqrt{\beta\rho}} 
+  \lambda_\mathrm{ideal} = \pi f v_A = \frac{\pi f}{\sqrt{\beta\rho}} 
 \end{align}
@@
 \begin{align}
-  \lambda_\mathrm{resis} =\frac{\eta}{f v_A}=\frac{f}{\Lambda_0}\sqrt{\frac{\rho}{\beta}}
+  \lambda_\mathrm{resis} =\frac{\pi\eta}{f v_A}=\frac{\pi f}{\Lambda_0}\sqrt{\frac{\rho}{\beta}}
 \end{align}
-is the MRI wavelength in the limit of high resistivity.
+%factor pi comes form converting vertical wavenumber to
+%wavelength. linear theory works with wavenumber, physical
+%interpretation in terms of wavelengths
+is the MRI wavelength in the limit of high resistivity. MRI requires
+$\lambda\lesssim 1$, so that the unstable wavelength fits within the
+disk column. As shown in Appendix \ref{appen1} $f$ increases with the
+strength of self-gravity. This is expected to increase $\lambda$ and
+weaken the MRI.
+
+We show in Fig. the growth rates as a function of $Q$ and
+$\Lambda_0$. In the limit of ideal MHD, $\Lambda_0>1$, there is
+negligible dependence on $Q$ because $f$ is of order unity for the
+range of $Q$ considered so $\lambda_\mathrm{ideal}$ remain less than
+  unity. However, in the resistive limit, the dependence of
+  $\lambda_\mathrm{resis}$ on $f$ (and hence $Q$) is amplified by
+  $\Lambda_0 < 1$.
 

### 2013-11-14 paper/results.tex
general backup

--- 2013-10-07 paper/results.tex
@@
 
-\subsection{Reference cases}
-%ideal, non sg
+\subsection{MRI in self-gravitating polytropic disks with uniform
+  resistivity}  
+We first calculate the MRI as a function of $Q$ in a polytropic disk
+with uniform resistivity ($A=1$). We consider $Q\in[0.2,4]$,
+corresponding to stronly self-gravitating to non-self-gravitating
+disks, for mid-plane Elsasser numbers $\Lambda_0\in[0.3,5]$. 
+We fix $k_x=0.1$ and $\beta=100$. The numerical resolution is
+$N_z=256$.
 
-%ideal, sg
+Fig. \ref{compare_growth_poly_uniresis} plots the MRI growth rates as
+a function of $Q$ and $\Lambda_0$. As expected growth rates generally
+decrease with $\Lambda_0$. In the limit of ideal MHD, $\Lambda_0>1$,
+there is negligible dependence on $Q$. In the resistive limit,
+$\Lambda_0<1$, growth rates decrease noticeably for $Q<1$. This shows
+show that self-gravity can still affect the MRI even  
+though density and potential perturbations are expected to be
+negligible because $k_x\ll1$ (i.e. the linear response is
+non-self-gravitating).
 
-%non ideal smooth, non sg
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_growth_poly_uniresis}
+  \caption{MRI growth rates as a function of $Q$ for mid-planet
+    Elsasser numbers $\Lambda_0=0.3$ (solid), $0.5$ (dotted), $1.0$
+    (dashed) and $5.0$ (dash-dot).  
+    \label{compare_growth_poly_uniresis}}
+\end{figure}
 
-%non ideal struct. non sg
+% (We find the weak variations in
+%$\gamma$ for small $Q$ is associated with a change in the number of
+%nodes in the magnetic field perturbation.)
+%because $f$ is of order unity for the
+%range of $Q$ considered so $\lambda_\mathrm{ideal}$ remain less than
+%  unity.
+%However, in the resistive limit, the dependence of
+%  $\lambda_\mathrm{resis}$ on $f$ (and hence $Q$) is amplified by
+%  $\Lambda_0 < 1$.
 
-%non ideal smooth, sg
+\cite{sano99} found that for instability to occur the linear mode
+dimensional wavelength $\lambda$ should fit inside the disk. That is,   
+\begin{align}\label{sano_crit}
+  \frac{\lambda}{2H} \equiv
+  \mathrm{max}\left(\lambda_\mathrm{ideal},\lambda_\mathrm{resis}\right)\lesssim
+  1, 
+\end{align}
+where the non-dimensional wavelengths are given by 
+\begin{align}\label{lambda_ideal}
+  \lambda_\mathrm{ideal} = \frac{4\pi}{\sqrt{15}} f v_A =
+  \frac{4\pi f}{\sqrt{15\beta\rho}}
+\end{align}
+in the limit of ideal MHD, and 
+\begin{align}\label{lambda_resis}
+  \lambda_\mathrm{resis} = \frac{2\pi}{\sqrt{3}}\frac{\eta}{v_A f} =
+  \frac{2\pi f}{\Lambda_0}\sqrt{\frac{\rho}{3\beta}} 
+\end{align}
+%factor pi comes form converting vertical wavenumber to
+%wavelength. linear theory works with wavenumber, physical
+%interpretation in terms of wavelengths
+in the limit of high resistivity. 
+%For $\Lambda_0\gtrsim 1$, we find
+%the condition $\lambda < 2H$ is well-satisifed throughout most of the
+%disk regardless of $Q$.     
+For polytropic disks the normalized density $\rho$ is weakly dependent
+on $Q$ (Fig. \ref{eqm_den}) so self-gravity only affects the
+MRI through the factor $f$, which increases with decreasing $Q$ (see
+Fig. \ref{plot_fq} in Appendix \ref{appen1}). This suggests that
+sufficiently strong self-gravity can stabilize the MRI by making
+$\lambda > 2H$.
 
-%non ideal struc, sg
+%Physically, the
+%disk becomes thinner with increasing self-gravity, and the MRI will be
+%surpressed when the disk is too thin to accommodate unstable modes. 
+%MRI requires
+%$\lambda\lesssim 1$, so that the unstable wavelength fits within the
+%disk column. As shown in Appendix \ref{appen1}, $f$ increases with the
+%strength of self-gravity. This is expected to increase $\lambda$ and
+%weaken the MRI.
 
-%grav instab.
+However, $f$ does not vary much for the range of $Q$ considered here. 
+In the ideal limit, we find $\lambda < 2H$ throughout most of the disk
+for all values of $Q$, so that self-gravity does not affect growth
+rates significantly. Weak variations in $\gamma$ were found to be
+associated with changes in the
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/lambda_poly_uniresis}
+  \caption{Wavelength of the most unstable resistive MRI modes as given by
+    Eq. \ref{sano_crit}---\ref{lambda_resis}, normalized by the 
+    disk thickness, as a function of height for a range of Toomre $Q$
+    values.  
+    \label{lambda_poly_resis}}
+\end{figure}
+
+\subsection{Resistive MRI in massive polytropic disks with layered
+  resistivity}
+
+\subsection{GI in magnetized isothermal disks with uniform
+  resistivity}
+
+\subsection{GI in magnetized isothermal disks with layered
+  resistivity}

### 2013-11-23 paper/setup.tex
updated tex

--- 2013-10-23 paper/setup.tex
@@
 We study the local stability of an inviscid, self-gravitating and
-magnetized barotropic fluid disk orbiting a central mass with
+magnetized fluid disk orbiting a central mass with
 potential $\Phi_*(r,z)$, where $(r,\varphi,z)$ are cylindrical
@@
 the local frame; $\bm{B}$ is the magnetic field which satisfies
-$\nabla\cdot\bm{B}=0$; $\Pi \equiv P +
+$\nabla\cdot~\bm{B}=0$; $\Pi \equiv P +
 |\bm{B}|^2/2\mu_0$ is the total pressure ($\mu_0$ is the vacuum
-permeability) and $P=P(\rho)$ is the gas pressure given by a
-barotropic equation of state chosen later.
+permeability). We choose a barotropic equation of state, specified
+below, so that the gas pressure is given by $P=P(\rho)$. 
+%is the gas pressure given by a
+%barotropic equation of state chosen later.
 
@@
 $\rho=\rho(z)$, $\bm{B} = B\hat{\bm{z}}$ where $B$ is a constant, and
-velocity field $\bm{v} = -Sx\hat{\bm{y}}$. We have assumed a thin disk 
-and neglected the radial component of the self-gravitational force in
+velocity field $\bm{v} = -Sx\hat{\bm{y}}$. We assume a thin disk 
+and neglect the radial component of the self-gravitational force in
 the unperturbed disk.
@@
 \item isothermal disks with $P=c_{s0}^2\rho$\label{iso_eos}; 
-\item polytropic disks with $P=K\rho^2$ where $K=c_{s0}^2/2\rho_0$ and $\rho_0\equiv\rho(0)$
-\end{inparaenum}.
-In both cases $c_{s0}$ is the midplane sound speed. For the polytropic
-disk we define the disk thickness as $H$ such that $\rho(H)=0$. In the
-isothermal case we define a characterisic disk thickness as
-$H=c_{s0}/\Omega$, which is the pressure scale-height in the
-non-self-gravitating limit.
+\item polytropic disks with $P=K\rho^2$ with $K=c_{s0}^2/2\rho(0)$;
+\end{inparaenum}
+where $c_{s0}$ is the global sound speed in the isothermal disk, and
+is the mid-plane sound speed in the polytropic disk. For the polytropic
+disk the disk thickness $H$ is such that $\rho(H)=0$. Since the
+isothermal disk has no surface, we define $H$ such that
+$\rho(H)=\epsilon\rho(0)$ with $\epsilon=10^{-2}$.
 
-We solve for $\hat{\rho}\equiv\rho/\rho_0$ subject to $\hat{\rho}=1$
-and $d\hat{\rho}/dz=0$  at $z=0$. This is done numerically for
-isothermal disks and analytically for the polytropic disk (see
-Appendix \ref{appen1}). Examples of density profiles are shown in
-Fig. \ref{eqm_den}. The isothermal disk shows that self-gravity
-compresses the disk and reduces the true pressure scale-height. The
-polytropic disk shows little difference between different strengths of
-self-gravity because the vertical co-ordiante is already normalized by
-the disk thickness.
+%and $\rho_0\equiv\rho(0)$
+
+We solve for $\hat{\rho}\equiv\rho/\rho_0$, where
+the mid-plane density $\rho_0\equiv\rho(0)$ is parameterized via the
+Toomre $Q$ parameter such that $\rho_0=\Omega^2/4\pi G Q$. Boundary
+conditions are $\hat{\rho}=1$ and $d\hat{\rho}/dz=0$  at $z=0$. This
+is done numerically for isothermal disks and analytically for the
+polytropic disk (see Appendix \ref{appen1}). Examples of density
+profiles are shown in Fig. \ref{eqm_den}. The density field is weakly
+dependent on the strength of self-gravity provided the vertical
+co-ordinate is appropriately normalized.
+
+%As the strength of self-gravity increases, the
+%disk becomes thinner ($H$ decreases)
+
+%The isothermal disk shows that self-gravity
+%compresses the disk and reduces the true pressure scale-height. The
+%polytropic disk shows little difference between different strengths of
+%self-gravity because the vertical co-ordiante is already normalized by
+%the disk thickness.
 
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_iso_density}
-  \includegraphics[width=\linewidth]{figures/compare_poly_density}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.5cm 0cm 0cm]{figures/compare_iso_density}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0.9cm]{figures/compare_poly_density}
   \caption{Equilibrium density field from solving Eq. \ref{eqm_eqns1}
     --- \ref{eqm_eqns2} subject to an isothermal (top) and polytropic
-    (bottom) equation of state. Note that the vertical co-ordinate is
-    normalized differently for convenience. In the isothermal disk,
-    $H$ is the characteristic scale-height in the
-    non-self-gravitating limit; in the polytropic disk $H$ is the
-    disk thickness. \label{eqm_den}}
+    (bottom) equation of state. Note that the normalization for the
+    horizontal axis also depends on the strength of self-gravity,
+    i.e. $H=H(Q)$ and is a decreasing function of $Q$. 
+%Note that the vertical co-ordinate is
+%    normalized differently for convenience. In the isothermal disk,
+%    $H$ is the characteristic scale-height in the
+%    non-self-gravitating limit; in the polytropic disk $H$ is the
+%    disk thickness. 
+\label{eqm_den}}
 \end{figure}
@@
 Eq. \ref{eqm_eqns1} --- \ref{eqm_eqns2}, the integration for
-Eq. \ref{sigma_pm} can be done implicitly by using the Poisson
+Eq. \ref{sigma_pm} can be performed implicitly by using Poisson's 
 equation.
 
-Example resistivity profiles are shown in Fig. \ref{eqm_resis} for the isothermal disk. 
-The midplane Elsasser number is $\Lambda(0)=0.1$ and the conductivity is boosted by a factor $A=10$ 
-from the midplane to the upper disk boundary. The figure compares a strongly self-gravitating case
-($Q=0.2$) and the non-self-gravitating limit ($Q=10$).
+We use the Elsasser number $\Lambda$ as a non-dimensional measure of
+conductivity,
+\begin{align} 
+  \Lambda \equiv \frac{v_A^2}{\eta\Omega},
+\end{align}
+where $v_A \equiv B/\sqrt{\mu_0\rho}$ is the Alfven speed. Example
+profiles of $\Lambda$ are shown in Fig. \ref{eqm_resis} for the
+isothermal disk. The disk may be considered ideal where $\Lambda
+\gtrsim 1$. Since $\rho(z/H)$ is weakly dependent on  self-gravity, the
+ideal region only expands slightly as $Q$ is decreased.
+
+%A similar plot is obtained for polytropic disks.
+
+%The conductivity is boosted by a factor $A=10$ from the midplane to
+%the upper disk boundary. The figure compares a strongly self-gravitating case
+%($Q=0.2$) and the non-self-gravitating limit ($Q=10$).
 
@@
 \subsection{System parameters}
-The strength of self-gravity is parametrized by
+We summarize here the main parameters describing the resistive,
+self-gravitating shearing box. The strength of self-gravity is
+parametrized by 
 \begin{align}
@@
 \begin{align}
-  \Lambda \equiv \frac{v_{A0}^2}{\eta_0\Omega} = \frac{f^2 R_m}{\beta}, 
+  \Lambda_0 \equiv\Lambda(0) =  \frac{v_{A0}^2}{\eta_0\Omega} = \frac{f^2 R_m}{\beta}, 
 \end{align}
-where the second equality defines the magnetic Reynolds number $R_m$
-and $f\equiv \csmid/H\Omega$ is a numerical factor of order unity.
+where the third equality defines the magnetic Reynolds number
+$R_m=H^2\Omega/\eta_0$ and $f\equiv \csmid/H\Omega$ is a numerical
+factor of order unity. 
+%The Elsasser number as a function of height is $\Lambda(z)\equiv
+%v_A^2/\eta\Omega$.

### 2013-12-05 paper/results.tex

@@
 \section{Results}
+We consider Keplerian disks so that $S = 1.5$ and $\kappa = \Omega =
+\Omega_z.$ In the linear simulations below we fix $\beta=100$ and the
+vertical extent of the disk is $Z_s=0.99$ and $Z_s=1$ in the
+polytropic and isothermal disk, respectively. The numerical resolution
+is $N_z=256$.
 
@@
 corresponding to stronly self-gravitating to non-self-gravitating
-disks, for mid-plane Elsasser numbers $\Lambda_0\in[0.3,5]$. 
-We fix $k_x=0.1$ and $\beta=100$. The numerical resolution is
-$N_z=256$.
+disks, for mid-plane Elsasser numbers $\Lambda_0\in[0.3,10]$. We fix
+$k_x=0.1$.  
+%We fix $k_x=0.1$ and $\beta=100$. The disk vertical extent is
+%$Z_s=0.99$ and adopt a numerical resolution of $N_z=256$.
 
@@
 %disk regardless of $Q$.     
-For polytropic disks the normalized density $\rho$ is weakly dependent
+The normalized polytropic disk density $\rho$ is weakly dependent
 on $Q$ (Fig. \ref{eqm_den}) so self-gravity only affects the
@@
 sufficiently strong self-gravity can stabilize the MRI by making
-$\lambda > 2H$.
+$ 2H<\lambda $.
 
@@
 
-However, $f$ does not vary much for the range of $Q$ considered here. 
+%However, $f$ does not vary much for the range of $Q$ considered here. 
 In the ideal limit, we find $\lambda < 2H$ throughout most of the disk
-for all values of $Q$, so that self-gravity does not affect growth
-rates significantly. Weak variations in $\gamma$ were found to be
-associated with changes in the
+for the values of $Q$ considered, so that self-gravity does not affect
+growth rates significantly. Fig. \ref{compare_result_lambda10} shows
+the magnetic energies for $\Lambda_0=10$ and a range of Toomre $Q$
+values. The number of vertical nodes decrease with $Q$, 
+%i.e. increasing the strength of vertical self-gravity makes the disk
+%unable to accommodate many wavelengths because it becomes thinner. 
+i.e. the disk accommodates fewer wavelengths because increasing
+vertical self-gravity makes it thinner. 
+%As the strength of 
+%self-gravity increase, the disk becomes thinner so it can only
+%accommodate one wavelength.
+Weak variations in growth rates in the ideal limit, seen in 
+Fig. \ref{compare_growth_poly_uniresis}, were found to be associated
+with changes in the mode character: $\gamma$ decreases
+slightly when the mode begins to acquire an additional vertical node as $Q$ is
+increased.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_result_lambda10}
+  \caption{Magnetic energies as a function of height, in the limit of
+    ideal MHD, for various strengths of self-gravity.   
+    \label{compare_result_lambda10}}
+\end{figure}
+
+Self-gravity appreciably decreases the MRI growth rates in the
+resistive limit. We plot in Fig. \ref{lambda_poly_resis}
+Eq. \ref{sano_crit} for $\Lambda_0=0.3$. Even in the
+non-self-gravitating disk ($Q=4$) the instability criterion is only
+marginally satisfied in the bulk of the disk. As $Q$ decreases, the
+Eq. \ref{sano_crit} is violated and the growth rate is dramatically
+reduced. This is seen for $Q=0.2$ where $\lambda \geq 2H$ throughout
+the disk. (The instability is not completely surpressed since
+Eq. \ref{sano_crit} is only an approximation.) Although the function
+$f$ does not change significantly for the range of $Q$ considered, the
+dependence of $\lambda_\mathrm{resis}$ on $f$ (and therefore $Q$) is
+amplified by the denominator $\Lambda_0<1$ in the resistive case. For resistively cases
+we find no nodes in the magnetic energy $E_1$ except at end points,
+i.e. only the longest wavelength survives Ohmic dissipation.
+
+%only one wavelength. no nodes except endpoints.
+
+%although $f$ does not vary significantly, the dependence of
+%$\lambda_\mathrm{resis}$ on $f$ is amplified by $\Lambda_0<1$ in the
+%denominator of Eq. \ref{lambda_resis}. Even in the
 
@@
   \includegraphics[width=\linewidth]{figures/lambda_poly_uniresis}
-  \caption{Wavelength of the most unstable resistive MRI modes as given by
+  \caption{Approximate wavelengths of the most unstable resistive MRI modes as given by
     Eq. \ref{sano_crit}---\ref{lambda_resis}, normalized by the 
@@
 
-\subsection{Resistive MRI in massive polytropic disks with layered
+\subsection{Resistive MRI in self-gravitating polytropic disks with layered
   resistivity} 
+Here we consider disks with mid-plane Elsasser number $\Lambda_0=0.1$
+and layered resistivity with $A=10^2$. We use $k_x=0.1$ as before.  
+Fig. \ref{poly_layer} compares the magnetic 
+energies for $Q=0.2,\,1$ and $4$. They have similar growth rates, $\gamma/\Omega
+= 0.53,\,0.64$ and $0.66$, respectively. In the non-self-gravitating
+limit ($Q=4$), the MRI is effectively surpressed for $z\lesssim0.5H$, whereas
+in the massive disk ($Q=0.2$) the mode occupies a wider vertical
+extent because its wavelength (in units of $H$) is larger. This
+suggests that in massive disks the MRI is not well localized to the
+active layer, as the instability takes on a more global character
+compared to non-self-gravitating disks.        
+%d is not as well localized
 
-\subsection{GI in magnetized isothermal disks with uniform
-  resistivity}
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_results_poly_layer_amp100}
+  \caption{Magnetic energies as a function of height, for layered polytropic disks with
+    such that the conductivity increases by a 
+    factor of $10^2$ in going from the mid-plane to the upper disk
+    boundary. The vertical lines indicate $\Lambda=1$ for each value
+    of $Q$.
+    \label{poly_layer}}
+\end{figure}
 
-\subsection{GI in magnetized isothermal disks with layered
-  resistivity}
+%\subsection{GI in magnetized isothermal disks with uniform
+%  resistivity}
+
+%\subsection{GI in magnetized isothermal disks with layered
+%  resistivity}
+
+\subsection{Influence of self-gravity on MRI at larger horizontal
+  wavenumbers} 
+As remarked above, for $k_xH\ll 1$ the most unstable linear mode is
+the incompressible MRI with vanishing density and potential
+perturbtions. For self-gravity to play a role in the linear
+pertubrations one must consider $k_xH \neq 0$, which we consider
+here.
+
+%to get compressibility effects, need kx non zero 
+%technically not fastest, but still pretty fast
+
+%\subsection{Influence of vertical magnetic field on gravitational
+%  instability}

### 2013-12-18 paper/appendix.tex
paper update

@@
     \label{plot_fq}}
+\end{figure}
+
+\section{Relation between $Q$ and the Toomre parameter}\label{q3d2d}
+The Toomre parameter defined for razor-thin disks is $Q_\mathrm{2D}\equiv \kappa c_s/\pi G
+\Sigma$,  where $\Sigma$ is the total column density. To relate our
+self-gravity parameter $Q$ and $Q_\mathrm{2D}$, we replace $c_s$ by
+$\overline{c_s}\equiv\int\rho c_s dz/\int\rho dz$, and replace
+$\kappa$ by $\Omega$, so that 
+\begin{align}
+  Q_\mathrm{2D} = 2Qf \frac{\int_0^1
+    \rho c_s
+    dz}{\left(\int_0^1 \rho dz\right)^2},   
+\end{align}
+where each term on the right-hand-side is non-dimensionalized. 
+We display this relation in Fig. \ref{plot_q3d2d} for the isothermal and
+polytropic disks.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/q2d_iso}
+  \caption{Relation between the self-gravity parameter $Q$ and the
+    classic Toomre parameter $Q_\mathrm{2D}$.   
+    \label{plot_q3d2d}}
 \end{figure}
@@
 \end{align}
+
+\section{Modifications for incompressible flow} 
+An incompressible flow satisfies 
+\begin{align}
+  \frac{\p \rho}{\p t} + \bm{v}\cdot\nabla\rho &= 0 \quad \mathrm{and} \quad
+  \nabla\cdot\bm{v} = 0
+\end{align}
+independently in the equilibrium and perturbed flows. Linearizing the
+above we obtain 
+\begin{align}
+  \frac{\dd\rho}{\rho} &= \frac{\ii }{\sigma}\dd v_z
+  \left(\ln\rho\right)^\prime,\\
+  0 &= \ii k_x \delta v_x + \dd v_z^\prime. 
+\end{align}
+The Eulerian density and pressure perturbation does not satisify the
+barotropic relation assumed for the unperturbed flow. Instead, the
+linearized pressure force is
+\begin{align}
+  -\frac{\nabla \dd P}{\rho} + \frac{\dd\rho}{\rho^2}\nabla P &=
+  -\nabla W + \nabla\ln\rho\left(\frac{c_s^2\dd\rho}{\rho} - W\right),
+  \notag\\
+  & = -\nabla W + \nabla\ln\rho\left(\frac{\ii c_s^2}{\sigma}\dd v _z \ln\rho^\prime  - W\right)
+\end{align}
+where $W=\delta P/\rho$ as before. The linearized vertical momentum
+equation acquires the second term, which gives the vertical velocity as 
+\begin{align}
+\dd v_z =
+\frac{\ii\sigma\left[\tilde{W}^\prime+\left(\ln\rho\right)^\prime W\right]}{\sigma^2 -
+  c_s^2\left(\ln{\rho}\right)^{\prime 2}}. 
+\end{align}
+Differentiating this and using $\delta v_z^\prime = -\ii k_x \delta
+v_x$, we obtain, after non-dimensionalizing, 
+\begin{align}
+  k_x \dd v_x\left[\sigma^2 -
+  f^2 c_s^2\left(\ln{\rho}\right)^{\prime 2}\right] + 
+  \sigma\left[\tilde{W}^{\prime\prime} +
+    \left(\ln\rho\right)^{\prime\prime}W +
+    \left(\ln\rho\right)^{\prime}W^\prime\right]f \notag\\ +
+  \frac{\sigma c_s^2 \left[\tilde{W}^\prime+\left(\ln{\rho}\right)^\prime
+    W\right]\left(\ln{\rho}\right)^\prime}{\sigma^2 -
+  f^2c_s^2\left(\ln{\rho}\right)^{\prime 2}}\left[\left(\ln{c_s^2}\right)^\prime\left(\ln{\rho}\right)^\prime
+      + 2\left(\ln{\rho}\right)^{\prime\prime}\right]f^3=0. 
+\end{align}
+The dimensionless linearized Poisson equation becomes
+\begin{align}
+  \dd\Phi^{\prime\prime} - k_x^2\dd\Phi +
+  \frac{\rho\left(\ln{\rho}\right)^\prime\left[\tilde{W}^\prime+\left(\ln\rho\right)^\prime
+      W\right]}{Q \left[\sigma^2 -
+  f^2c_s^2\left(\ln{\rho}\right)^{\prime 2}\right]} = 0.  
+\end{align}
+%boundary condition modifications
+The boundary condition for the gravitational potential is also
+modified to 
+\begin{align}
+  \dd\Phi^\prime + k_x\dd\Phi +
+  \frac{\rho\left[\tilde{W}^\prime+\left(\ln\rho\right)^\prime
+      W\right]}{Q\left[\sigma^2 - f^2c_s^2\left(\ln{\rho}\right)^{\prime 2}\right]}=0.
+\end{align}

### 2013-12-18 paper/linear.tex
general backup
updated tex
paper update

@@
 \begin{align}
-  &z=\hat{z}H,\quad k_x =  k_x/H,\quad \sigma = \hat{\sigma}\Omega,
+  &z=\hat{z}H,\quad k_x =  \hat{k}_x/H,\quad \sigma = \hat{\sigma}\Omega,
   \quad \delta\bm{v} = \csmid 
@@
 We will now drop the $\hat{\phantom{a}}$ notation. Henceforth it is
-understood that all variables have been appropriately normalized. The
+understood that all variables have been appropriately normalized
+(i.e. $\Omega\to 1$ and $H\to 1$), unless otherwise stated. The
 linearized continuity equation is 
@@
 
-Note that $L_{31}\propto k_x$, so for $k_x=0$ 
+Note that $L_{13},\,L_{14},\,L_{23},\,L_{24}$ and $L_{31}$ are
+proportional to $k_x$. Then for for $k_x=0$ 
 Eq. \ref{lin1}---\ref{lin2} are decoupled from 
 Eq. \ref{lin3}---\ref{lin4}. In this case, a solution is $\dphi=W=0$,
-with $\dvx$ and $\dvy$ determined by Eq. \ref{lin1}---\ref{lin2},
-which is the incompressible MRI problem. This means that the density
-and potential perturbations can only influence the MRI for
-$k_x\neq0$. However, SG will still have an effect through the
-background equilibrium for $k_x=0$.
+with $\dvx$ and $\dvy$ determined by $L_{11}\dvx + L_{12}\dvy = 0$ and 
+$L_{21}\dvx + L_{22}\dvy=0$, which is the incompressible MRI problem. 
+This means that the density and potential perturbations can only
+influence the MRI for $k_x\neq0$. Nevertheless, for $k_x=0$,
+self-gravity can still have an effect through the background
+equilibrium.
 
@@
 \begin{align}
-  &E_1 \equiv \frac{|\delta \bm{B}|^2}{2\beta},\\
-  &E_2 = \frac{\rho}{\hat{c}_s^2}|\real\left(W\dphi^*\right)|,\\
-  &E_3 = \frac{1}{2}\rho|\delta\bm{v}|^2,\\
-  &E_4 = \frac{\rho|W|^2}{2\hat{c}_s^2},
+  &E_m \equiv \frac{|\delta \bm{B}|^2}{2\beta},\\
+  &E_g = \frac{\rho}{c_s^2}|\real\left(W\dphi^*\right)|,\\
+  &E_k = \frac{1}{2}\rho|\delta\bm{v}|^2,\\
+  &E_t = \frac{\rho|W|^2}{2c_s^2},
 \end{align}
 as the perturbed magnetic, gravitational, kinetic and thermal
-energies, respectively, which are functions of $z$. We
-then normalize each by $\mathrm{max}(E_1)$.
+energies, respectively, which are functions of $z$. The total energy
+is $E=E_m+E_g+E_k+E_t$ and we then normalize $E_i$ by
+$\mathrm{max}(E)$.

### 2013-12-18 paper/results.tex
paper update

@@
 
-\subsection{Influence of self-gravity on MRI at larger horizontal
-  wavenumbers} 
+\subsection{Non-zero radial wavenumbers}  
 As remarked above, for $k_xH\ll 1$ the most unstable linear mode is
@@
 perturbtions. For self-gravity to play a role in the linear
-pertubrations one must consider $k_xH \neq 0$, which we consider
-here.
+pertubrations we consider $k_xH \neq 0$ and also an isothermal
+disk. From \cite{mamat10}, we expect GI to set in
+for $Q\lesssim 0.2$. 
+%We expect modes associated with GI have $k_x\sim 1/H$, for modes
+%with smaller $k_x$ are stabilized by rotation and
+
+In Fig. \ref{compare_growth_varQ} we plot the growth rate of 
+unstable modes in disks with uniform resistivity and varying $Q$.  
+The appearence of a bump at $k_xH\sim1$ for $Q< 0.2$ signifies GI. 
+%We checked that the perturbed energy is dominated
+%by gravitational energy.   
+%Because the most unstable MRI mode occurs at $k_xH=0$ while the most
+%unstable GI mode occurs at $k_xH\sim1$, 
+We expect a sufficiently massive magnetized disk to undergo both MRI
+at long radial lengthscales ($k_xH\ll 1$) and GI at intermediate
+radial lengthscales ($k_xH\sim 1$). However, unless the system
+parameters are finely tuned, the MRI and GI modes do not grow
+at the same rate. For example, at $Q=0.17$ GI is expected but
+increasing it slightly to $Q=0.2$, MRI dominates. This is largely due
+to promoting GI with decreasing $Q$, although lowering $Q$ also
+stabilizes the MRI.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_growth_varQ}
+  \caption{Linear growth rates as a function of horizontal wavenumber
+    $k_x$ in a resistive disk with $\Lambda_0=0.1$ and varying
+    strengths of self-gravity. 
+    \label{compare_growth_varQ}}
+\end{figure}
+
+%Strictly speaking, given a set of disk parameters, one is only
+%concerned with the most unstable linear mode.
+
+%The plot indicates the peak GI growth rate is
+%$\gamma\sim 0.3\Omega$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_growth_varLam}
+  \caption{Linear mode growth rates as a function of horizontal wavenumber
+    $k_x$ in a self-gravitating disk with $Q=0.18$ and uniform resitivity 
+     as measured by the mid-plane Elsasser number $\Lambda_0$.
+    \label{compare_growth_varLam}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/result_Q0d18_lamda0d2}
+  \caption{Perturbation energies for a linear mode in a
+    self-gravitating disk ($Q=0.18$) with uniform resistivity (such
+    that $\Lambda_0=0.2$). 
+    The mangetic
+    ($E_m$, solid), gravitational ($E_k$, dotted ), kinetic ($E_k$,
+    dashed) and thermal ($E_t$, dash-dot) energies are normalized by
+    the maximum of the total energy.  
+    \label{result_Q0.18_lambda0.2}}
+\end{figure}
 

### 2014-02-10 paper/appendix.tex
major update to tex

--- 2013-11-13 paper/appendix.tex
@@
 
+\section{Relation between $Q$ and the Toomre parameter}\label{q3d2d}
+The Toomre parameter defined for razor-thin disks is 
+\begin{align}
+  Q_\mathrm{2D}\equiv \frac{\kappa c_s}{\pi G\Sigma},  
+\end{align}
+where $\Sigma$ is the total column density. To relate our
+self-gravity parameter $Q$ and $Q_\mathrm{2D}$, we replace $c_s$ by
+$\overline{c_s}\equiv\int\rho c_s dz/\int\rho dz$, and replace
+$\kappa$ by $\Omega$, so that 
+\begin{align}
+  Q_\mathrm{2D} = 2Qf \frac{\int_0^1
+    \hat{\rho} \hat{c}_s
+    d\hat{z}}{\left(\int_0^1 \hat{\rho} d\hat{z}\right)^2},   
+\end{align}
+where each term on the right-hand-side is non-dimensionalized (see
+\S\ref{non-dim}). We display this relation in Fig. \ref{plot_q3d2d}
+for the isothermal and polytropic disks.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/q2d_iso}
+  \caption{Relation between the self-gravity parameter $Q$ and the
+    classic Toomre parameter $Q_\mathrm{2D}$.   
+    \label{plot_q3d2d}}
+\end{figure}
+
 \section{Reduction to linear hydrodynamics}\label{reduction}
+
 Our task here is to remove the magnetic field and vertical velocity
-perturbations from the non-dimensional linearized equations. 
-Let us first define operators
+perturbations from the linearized equations. Let us first define operators
 \begin{align}
@@
 \end{align}
-Denoting the $n^\mathrm{th}$ vertical derivative as $^{(n)}$, the
+And we define the variables
+\begin{align}
+  &U \equiv \imgi\sigma\dvx - 2\Omega\dvy + \imgi k_x \w,\\
+  &V \equiv \imgi\sigma\dvy + \frac{\kappa^2}{2\Omega}\dvx.
+\end{align}
+
+We first express the continuity equation in terms of horizontal
+velocity, density and potential perturbations. 
+The vertical velocity perturbation is 
+\begin{align}
+  \dvz = \frac{\imgi}{\sigma}\left(\w^\prime + \epsilon V\right),  
+\end{align}
+where the linearized $y$ momentum equation was used (i.e. eliminating
+$\dby^\prime$ between Eq. \ref{lin_vy} and Eq. \ref{lin_vz}). 
+Inserting this into the linearized continuity equation
+(Eq. \ref{lin_cont}), we obtain
+\begin{align}
+0 = W^{\prime\prime} + \left(\ln{\rho}\right)^\prime W^\prime +
+\frac{\sigma^2}{c_s^2}W +  \dphi^{\prime\prime} + \left(\ln{\rho}\right)^\prime \dphi^\prime
+ + \sigma k_x \dvx + \epsilon D_1 V.
+%  0=  W^{\prime\prime} + \left(\ln{\rho}\right)^\prime W^\prime +
+%  \frac{1}{c_s^2f^2}\left(\frac{\rho}{Q} + \sigma^2\right)W +
+% \left(\ln{\rho}\right)^\prime\dphi^\prime + k_x^2\dphi + \frac{\sigma
+%    k_x}{f}\dvx.\label{final_w}
+\end{align}
+
+Next, we examine separately the cases of a vertical field with
+variable resistivity and that of a tilted field with uniform
+resistivity. We remark that a similar procedure can be performed in
+the general case of a tilted field with variable resistivity.
+
+\subsection{Vertical field with variable resistivity}
+First consider $\epsilon = 0$ and $\eta=\eta(z)$ in the linearized
+equations. Denoting the $n^\mathrm{th}$ vertical derivative as $^{(n)}$, the
 equations of motion give
 \begin{align}
-  &\dbx^{(n)} = \frac{\beta\rho}{f}D_{n-1}\left(\imgi\sigma\dvx - 2\dvy +
-  \imgi f k_x\w\right) + \imgi k_x \dbz^{(n-1)},\label{bx_eq}\\
-  &\dby^{(n)} = \frac{\beta\rho}{f}D_{n-1}\left(\imgi\sigma\dvy +
-  \frac{\kappa^2}{2}\dvx\right),\label{by_eq}
+  &\dbx^{(n)} = \frac{\mu_0\rho}{B_z}D_{n-1}U + \imgi k_x \dbz^{(n-1)},\label{bx_eq}\\
+  &\dby^{(n)} = \frac{\mu_0\rho}{B_z}D_{n-1}V,\label{by_eq}
 \end{align}
@@
 We insert the expression for $\dbx^\prime$ from Eq. \ref{bx_eq} and the
-expressoin for $\dbz^{\prime\prime}$ from the $z$ component of the
+expression for $\dbz^{\prime\prime}$ from the $z$ component of the
 linearized induction equation (Eq. \ref{induct_vert}) to obtain
 \begin{align}\label{bz_eq}
-  -\sigma\dbz^{(n)} = f k_x \dvx^{(n)} + \frac{k_x\beta\rho}{f}\overline{D}_n\left(\imgi\sigma\dvx - 2\dvy +
-  \imgi f k_x\w \right). 
+  -\sigma\dbz^{(n)} = k_x B_z \dvx^{(n)} + k_x\frac{\mu_0\rho}{B_z}\overline{D}_nU. 
 \end{align}
 
-\subsection{Eliminating $\dbx$}
+%\subsection{Eliminating $\dbx$}
 Inserting the above expressions for $\dbx^{\prime\prime}$,
@@
 \begin{align}
-  \imgi\sigma \frac{f}{\beta\rho}\dbx =
-  \frac{f^2}{\beta\rho}\dvx^\prime + \overline{D}_1\left(\imgi\sigma\dvx -
-  2\dvy + \imgi f k_x\w\right). \label{bx_expression}
+  \imgi\sigma \dbx =
+  B_z\dvx^\prime + \frac{\mu_0\rho}{B_z}\overline{D}_1U. \label{bx_expression}
 \end{align}
-($\bar{\sigma}\neq0$ has been assumed to obtain this.) 
-We multiply this expression by $\rho$, differentiate with respect to
-$z$ and eliminate the resulting $\dbx^\prime$ using
-Eq. \ref{bx_eq}, to obtain
+($\bar{\sigma}\neq0$ has been assumed to obtain this.) We
+differentiate this expression with respect to $z$ and eliminate the
+resulting $\dbx^\prime$ using Eq. \ref{bx_eq}, to obtain
 \begin{align}
-  0 = \frac{f^2}{\beta\rho}\dvx^{\prime\prime} -
-  \frac{f^2}{\beta\rho}k_x^2\dvx + \left(\overline{D}_2 - k_x^2 \overline{D}_0
-    - \imgi\sigma D_0\right)\left( \imgi\sigma\dvx -
-  2\dvy + \imgi f k_x\w\right). \label{final_vx} 
+%  0 = B_z\dvx^{\prime\prime} -
+%  k_x^2B_z\dvx + \frac{\mu_0\rho}{B_z}\left(\overline{D}_2 - k_x^2 \overline{D}_0
+%    - \imgi\sigma D_0\right)U. \label{final_vx} 
+  0 = v_A^2\left(\dvx^{\prime\prime} -
+  k_x^2\dvx\right) + \left(\overline{D}_2 - k_x^2 \overline{D}_0
+    - \imgi\sigma D_0\right)U. \label{final_vx} 
 \end{align}
 
-\subsection{Eliminating $\dby$}
-We follow a similar procedure as above to remove $\dby$. We first use
-the divergence-free condition to eliminate $\dbx$ from the
+%\subsection{Eliminating $\dby$}
+We follow a similar procedure as above to remove $\dby$. 
+We use Eq. \ref{bx_expression} and Eq. \ref{by_eq} to 
+eliminate $\dbx,\dby^\prime$ and $\dby^{\prime\prime}$ from the
 right-hand-side of the $y$-induction equation (Eq. \ref{induct_y}), 
+%\begin{align}
+%  &\imgi\bar{\sigma}\dby = f\dvy^\prime - \imgi
+%  S\frac{\dbz^\prime}{k_x} +
+%  \eta\dby^{\prime\prime}+\eta^\prime\dby^\prime.  
+%\end{align}
+%We can now insert expressions for the magnetic field derivatives using
+%Eq. \ref{by_eq} and \ref{bz_eq} to obtain an expression for $\dby$, 
 \begin{align}
-  &\imgi\bar{\sigma}\dby = f\dvy^\prime - \imgi
-  S\frac{\dbz^\prime}{k_x} +
-  \eta\dby^{\prime\prime}+\eta^\prime\dby^\prime.  
-\end{align}
-We can now insert expressions for the magnetic field derivatives using
-Eq. \ref{by_eq} and \ref{bz_eq} to obtain an expression for $\dby$, 
-\begin{align}
-  \imgi\bar{\sigma}\dby = f \dvy^\prime + \frac{\imgi S}{\sigma}\left[
-    f\dvx^\prime + \frac{\beta\rho}{f}\overline{D}_1\left(\imgi\sigma\dvx -
-  2\dvy + \imgi f k_x\w \right)\right] +
-  \frac{\beta\rho}{f}\overline{D}_1\left(\imgi\sigma\dvy +
-  \frac{\kappa^2}{2}\dvx\right). \label{by_expression}
+  \imgi\bar{\sigma}\dby = B_z\dvy^\prime + \frac{\imgi S}{\sigma}\left(
+    B_z\dvx^\prime + \frac{\mu_0\rho}{B_z}\overline{D}_1U\right) +
+  \frac{\mu_0\rho}{B_z}\overline{D}_1V. \label{by_expression}
 \end{align}
@@
 \begin{align}
-  0 =& \frac{f^2}{\beta\rho}\left(\dvy^{\prime\prime} -
-  \frac{\bar{\sigma}^\prime}{\bar{\sigma}}\dvy^\prime\right) 
-  + \sbar\sigma D_0\dvy  + \frac{\imgi 
-    S}{\sigma}\frac{f^2}{\beta\rho}\left(\dvx^{\prime\prime} -
-    \frac{\sbar^\prime}{\sbar}\dvx^\prime\right) -
-    \frac{\imgi\sbar\kappa^2}{2}D_0\dvx \notag\\
-    & + \left(\overline{D}_2 -
-    \frac{\sbar^\prime}{\sbar}\overline{D}_1\right)\left[\imgi\left(\sigma
-      - \frac{2S}{\sigma}\right)\dvy + \left(\frac{\kappa^2}{2} -
-      S\right)\dvx - \frac{Sfk_x}{\sigma}\w\right]. \label{final_vy}
+0 = v_A^2\left(\dvy^{\prime\prime} -
+\frac{\sbar^\prime}{\sbar}\dvy^\prime\right)  
++ \frac{\imgi S v_A^2}{\sigma}\left(\dvx^{\prime\prime} -
+\frac{\sbar^\prime}{\sbar}\dvx^\prime\right) + \frac{\imgi
+  S}{\sigma}\left(\overline{D}_2 -
+    \frac{\sbar^\prime}{\sbar}\overline{D}_1 \right)U 
++\left(\overline{D}_2 -
+    \frac{\sbar^\prime}{\sbar}\overline{D}_1 - \imgi\sbar D_0 \right)V. \label{final_vy}
+%0 = B_z\left(\dvy^{\prime\prime} -
+%\frac{\sbar^\prime}{\sbar}\dvy^\prime\right)  
+%+ \frac{\imgi S B_z}{\sigma}\left(\dvx^{\prime\prime} -
+%\frac{\sbar^\prime}{\sbar}\dvx^\prime\right) + \frac{\imgi
+%  S}{\sigma}\frac{\mu_0\rho}{B_z} \left(\overline{D}_2 -
+%    \frac{\sbar^\prime}{\sbar}\overline{D}_1 \right)U 
+%+\frac{\mu_0\rho}{B_z} \left(\overline{D}_2 -
+%    \frac{\sbar^\prime}{\sbar}\overline{D}_1 - \imgi\sbar D_0 \right)V. \label{final_vy}
+%
+%  0 =& \frac{f^2}{\beta\rho}\left(\dvy^{\prime\prime} -
+%  \frac{\bar{\sigma}^\prime}{\bar{\sigma}}\dvy^\prime\right) 
+%  + \sbar\sigma D_0\dvy  + \frac{\imgi 
+%    S}{\sigma}\frac{f^2}{\beta\rho}\left(\dvx^{\prime\prime} -
+%    \frac{\sbar^\prime}{\sbar}\dvx^\prime\right) -
+%    \frac{\imgi\sbar\kappa^2}{2}D_0\dvx \notag\\
+%    & + \left(\overline{D}_2 -
+%    \frac{\sbar^\prime}{\sbar}\overline{D}_1\right)\left[\imgi\left(\sigma
+%      - \frac{2S}{\sigma}\right)\dvy + \left(\frac{\kappa^2}{2} -
+%      S\right)\dvx - \frac{Sfk_x}{\sigma}\w\right]. \label{final_vy}
 \end{align}
 
-\subsection{Elminating $\dvz$}
-The vertical velocity perturbation is
+Eq. \ref{final_vx} and \ref{final_vy} constitutes the first two
+linearized equations to be solved.
+
+\subsection{Tilted field with uniform resistivity}
+Here we allow $\epsilon\neq 0$ but take $\eta$ to be constant. We
+first obtain expressions for $\dbx$ and $\dby$. Differentiating
+the $x$ momentum equation and replacing the resuting $\dbz^\prime$
+using the divergence-free condition and $\dby^\prime$ using the $y$
+momentum equation, we obtain an expression for $\dbx^{\prime\prime}$
+which can be inserted into the $x$ induction equation. This gives
 \begin{align}
-  \dvz = \frac{\imgi f}{\sigma}\w^\prime. 
+  \imgi\sigma\dbx = B_z\dvx^\prime +
+  \frac{\eta\mu_0\rho}{B_z}\left( D_1 U + \imgi\epsilon k_x D_0
+  V\right).\label{bx_tilted}  
 \end{align}
-Inserting this into the linearized continuity equation
-(Eq. \ref{lin_cont}) and using the Poisson equation, we obtain
+We can insert this into the $y$ induction equation to obtain
 \begin{align}
-0=  W^{\prime\prime} + \left(\ln{\rho}\right)^\prime W^\prime +
-  \frac{1}{c_s^2f^2}\left(\frac{\rho}{Q} + \sigma^2\right)W +
-  \left(\ln{\rho}\right)^\prime\dphi^\prime + k_x^2\dphi + \frac{\sigma
-  k_x}{f}\dvx.\label{final_w}
+  \imgi\sbar\dby = -B_y\Delta + B_z\dvy^\prime + \frac{\imgi
+    S}{\sigma}\left[\frac{\eta\mu_0\rho}{B_z}\left( D_1 U + \imgi\epsilon k_x D_0
+  V\right)\right] + \frac{\eta\mu_0\rho}{B_z}D_1V,\label{by_tilted}
 \end{align} 
+where we have also used the derivative of the $y$ momentum equation to
+eliminate  $\dby^{\prime\prime}$. Recall $\Delta \equiv i k_x \dvx +
+\dvz^\prime$, so that
+\begin{align}
+  \Delta = \imgi k_x \dvx + \frac{i}{\sigma}\left(\w^{\prime\prime} +
+  \epsilon V^\prime\right) = -\left[\frac{\imgi\sigma W}{c_s^2} +
+    \frac{\imgi\left(\ln{\rho}\right)^\prime}{\sigma}\left(\w^\prime +
+    \epsilon V\right)\right], 
+\end{align}
+where the second equality results from the continuity equation.
+
+Now consider
+\begin{align}
+  \dbx^\prime - \imgi k_x \dbz = \frac{\mu_0\rho}{B_z}D_0U +
+  \imgi\epsilon k_x \dby = \frac{\sigma}{\sbar}\dbx^\prime +
+  \frac{\imgi k_x^2B_z}{\sbar}\dvx,
+\end{align}
+where the first equality corresponds to the $x$ momentum equation and
+the second equality results from replacing $\dbz$ using the $z$
+induction equation. We can now use the above expressions for $\dbx$
+and $\dby$ (Eq. \ref{bx_tilted}--\ref{by_tilted}) to obtain
+\begin{align}
+0 = &v_A^2\left[k_x^2\left(1+\epsilon^2\right)\dvx - \frac{\epsilon k_x
+  S}{\sigma}\dvx^\prime - \dvx^{\prime\prime}\right] + \imgi\epsilon
+k_x v_A^2 \dvy^\prime + \frac{\epsilon^2 k_x
+  v_A^2}{\sigma}\left(\w^{\prime\prime} + \epsilon V^\prime\right) 
+\notag\\ &-\left[\eta\left(D_2 + \frac{\epsilon k_x S}{\sigma}D_1\right) -
+  \imgi\sbar D_0 \right]U %\notag\\
+- \frac{\imgi\epsilon^2k_x^2S}{\sigma}\eta D_0V.
+\end{align}
+%We can now use the above expression for $\dbx,\dby$ to eliminate
+%$\dbx^\prime$ and $\$ in the $x$ momentum equation  
+%\begin{align}
+%  \dbx^\prime - \imgi k_x\dbz = \frac{}{}
+%\end{align}
+
+Similarly, we differentiate Eq. \ref{by_tilted} and use the $y$
+momentum equation to eliminate $\dby^\prime$ to obtain
+\begin{align}
+0 = &v_A^2\dvy^{\prime\prime} + \frac{\imgi S}{\sigma}v_A^2
+\dvx^{\prime\prime} +\frac{\imgi S}{\sigma}\eta D_2U +
+\left\{\eta\left(D_2 - \frac{\epsilon k_x S}{\sigma} D_1\right) +
+  \imgi\left[\frac{\epsilon^2v_A^2\left(\ln{\rho}\right)^{\prime\prime}}{\sigma}
+    - \sbar\right]D_0\right\}V \notag\\ &+ \imgi\epsilon
+  v_A^2\left\{\frac{\sigma}{c_s^2}\left[W^\prime -
+    \left(\ln{c_s^2}\right)^\prime W\right] +
+    \frac{1}{\sigma}\left[\left(\ln{\rho}\right)^\prime\w^{\prime\prime}
+      +\left(\ln{\rho}\right)^{\prime\prime}\w^{\prime} \right]\right\}.    
+\end{align}
+
+%\subsection{Non-dimensionalization}
+%In practice we solve the linearized equations in non-dimensional form,
+%by defining
+%We further write 
+%\begin{align}
+%  &z=\hat{z}H,\quad k_x =  \hat{k}_x/H,\quad \sigma = \hat{\sigma}\Omega,
+%  \quad \delta\bm{v} = \csmid 
+%  \delta\hat{\bm{v}}, \\ 
+%  &\delta\bm{B} = B\delta\hat{\bm{B}},\quad
+%  \delta\rho = \rho\hat{W}/\hat{c}_s^2,\quad \delta\Phi =
+%  \csmid^2\delta\hat{\Phi}, 
+%\end{align}
+%where we have introduced the non-dimensional enthalpy perturbation
+%$\hat{W}$ and the sound-speed $\hat{c}_s=\csmid^{-1}\sqrt{dP/d\rho}$ can be 
+%obtained from the equation of state. The background frequencies are
+%written $S=\hat{S}\Omega$ and $\Omega_z=\hat{\Omega}_z\Omega$ and the 
+%resistivity as $\eta = \hat{\eta}H^2\Omega$.
+
+%We will now drop the $\hat{\phantom{a}}$ notation. Henceforth it is
+%understood that all variables have been appropriately normalized
+%(i.e. $\Omega\to 1$ and $H\to 1$), unless otherwise stated.

### 2014-02-10 paper/linear.tex
major update to tex

--- 2013-10-05 paper/linear.tex
@@
 equilibrium in the form $\real[\dd\rho(z)\exp{\imgi(k_xx + \sigma
-    t)}]$ and similarly for other fluid variables. Here, $k_x$ is a constant radial wavenumber and 
-$\sigma=-(\omega +\imgi\gamma)$ is a complex 
-frequency, where $-\omega$ is the real mode frequency and $\gamma$ is
-the growth rate. Hereafter, we surpress the exponential factor, as
-well as the real part notation.  We further write 
+    t)}]$ and similarly for other fluid variables. Here, $k_x$ is a
+constant radial wavenumber and $\sigma=-(\omega +\imgi\gamma)$ is a
+complex  frequency, where $-\omega$ is the real mode frequency and $\gamma$ is
+the growth rate. We take $k_x>0$ without loss of generality. 
+Hereafter, we surpress the exponential factor, as
+well as the real part notation.
+
+The linearized continuity equation is 
 \begin{align}
-  &z=\hat{z}H,\quad k_x =  k_x/H,\quad \sigma = \hat{\sigma}\Omega, \quad \delta\bm{v} = \csmid
-  \delta\hat{\bm{v}}, \\ 
-  &\delta\bm{B} = B\delta\hat{\bm{B}},\quad
-  \delta\rho = \rho\hat{W}/\hat{c}_s^2,\quad \delta\Phi =
-  \csmid^2\delta\hat{\Phi}, 
-\end{align}
-where we have introduced the non-dimensional enthalpy perturbation
-$\hat{W}$ and the sound-speed $\hat{c}_s=\csmid^{-1}\sqrt{dP/d\rho}$ can be 
-obtained from the equation of state. The background frequencies are
-written $S=\hat{S}\Omega$ and $\Omega_z=\hat{\Omega}_z\Omega$ and the 
-resistivity as $\eta = \hat{\eta}H^2\Omega$.
-
-We will now drop the $\hat{\phantom{a}}$ notation. Henceforth it is
-understood that all variables have been appropriately normalized. The
-linearized continuity equation is 
-\begin{align}
-  \frac{\imgi\sigma}{fc_s^2}W + \imgi k_x \dd v_x
+  \frac{\imgi\sigma}{c_s^2}W + \imgi k_x \dd v_x
   +\left(\ln\rho\right)^\prime\dd v_z + \dd v_z^\prime = 0,\label{lin_cont}
 \end{align}
-where $^\prime$ denotes $d/dz$. The linearized equations of motion are
+where $^\prime$ denotes $d/dz$ and $W=\delta P/\rho =
+c_s^2\delta\rho/\rho$ is the enthalpy perturbation.  
+The linearized equations of motion are
 \begin{align}
-  &\imgi\sigma \dvx - 2\dvy = f\left[v_A^2\dbx^\prime - \imgi k_x
-    \left(\w + v_A^2\dbz
-    \right)\right],\\
-  &\imgi\sigma\dvy + \frac{\kappa^2}{2}\dvx = fv_A^2\dby^\prime,\\
-  & \imgi\sigma\dvz = -f\w^\prime,
+  &\imgi\sigma \dvx - 2\Omega\dvy  = - \imgi
+  k_x\w + \frac{B_z}{\mu_0\rho} \left[\dbx^\prime - \imgi 
+    k_x\left(\dbz + \epsilon\dby\right)\right],\\
+  &\imgi\sigma\dvy + 
+    \frac{\kappa^2}{2\Omega}\dvx = \frac{B_z}{\mu_0\rho}\dby^\prime\label{lin_vy},\\ 
+  & \imgi\sigma\dvz = -\w^\prime - \frac{B_y}{\mu_0\rho}\dby^\prime\label{lin_vz},  
 \end{align}
-where the Alfven speed $v_A =
-\left(\beta\rho\right)^{-1/2}$, the effective enthalpy perturbation
-$\w = W + \dphi$ and the epicycle frequency $\kappa=\sqrt{2(2-S)}$. 
-The linearized induction equation is
+where the effective enthalpy perturbation
+$\w = W + \dphi$. The linearized induction equation is
 \begin{align}
-&  \imgi\bar{\sigma}\dbx = f\dvx^\prime +
+&  \imgi\bar{\sigma}\dbx = B_z\dvx^\prime +
   \eta\dbx^{\prime\prime}+\eta^\prime\dbx^\prime - \imgi k_x
   \eta^\prime \dbz,\label{induct_x}\\
-&\imgi\bar{\sigma}\dby = f\dvy^\prime - S\dbx +
+&\imgi\bar{\sigma}\dby = B_z\dvy^\prime -B_y\Delta - S\dbx +
   \eta\dby^{\prime\prime}+\eta^\prime\dby^\prime,\label{induct_y}\\
-& \imgi\bar{\sigma}\dbz = -\imgi f k_x \dvx + \eta\dbz^{\prime\prime}, \label{induct_vert}
+& \imgi\bar{\sigma}\dbz = -\imgi  k_xB_z \dvx +
+\eta\dbz^{\prime\prime}, \label{induct_vert}
+%& \imgi\bar{\sigma}\dbz = -\imgi  k_x \dvx  -\imgi k_x\eta \dbx^{\prime}, \label{induct_vert}
 \end{align} 
-where $\imgi\bar{\sigma} = \imgi\sigma + \eta k_x^2$, and the
-divergence-free condition is $\imgi k_x\dbx + \dbz^\prime=0$. Finally,
-the linearized Poisson equation is
+where $\imgi\bar{\sigma} = \imgi\sigma + \eta k_x^2$,
+$\Delta\equiv \nabla\cdot\bm{\delta v}  = \imgi k_x\dvx + \dvz^\prime$,   
+and the divergence-free condition is $\imgi k_x\dbx +
+\dbz^\prime=0$. Finally, the linearized Poisson equation is 
 \begin{align}
-  \dphi^{\prime\prime} - k_x^2\dphi = \frac{\rho}{c_s^2f^2Q}W.  \label{lin_poisson}
+  \dphi^{\prime\prime} - k_x^2\dphi = \frac{\Omega^2\rho}{c_s^2Q\rho_0}W.  \label{lin_poisson}
 \end{align}
 
-We now eliminate $\dd\bm{B}$ and $\dvz$ between the linearized
+We eliminate $\dd\bm{B}$ and $\dvz$ between the linearized
 equations to obtain a system of ordinary differential equations for
-$\bm{U}=\left(\dvx,\dvy,W,\dphi\right)$. We detail this technical
-procedure in Appendix \ref{reduction}. The schematic problem is
+$\bm{U}=\left(\dvx,\dvy,W,\dphi\right)$. We detail the steps 
+in Appendix \ref{reduction} for two cases considered in this paper:
+%We consider two cases:
+\begin{enumerate}
+\item Purely vertical field with constant or variable
+  resistivity (see below), so that $\epsilon = 0$ and $\eta=\eta(z)$. 
+\item Tilted field with uniform resistivity
+  so that $\epsilon \neq 0$ and $\eta=\rm{constant}$.   
+\end{enumerate}
+Schematically, the numerical problem is to solve 
 \begin{align}
@@
   L_{21}\dvx + L_{22}\dvy + L_{23}W + L_{24}\dphi &= 0, \label{lin2}\\
-  L_{31}\dvx \quad\quad\quad\,\,\,\,\,\,\, + L_{33}W + L_{34}\dphi &=0,\label{lin3}\\
+  L_{31}\dvx +L_{32}\dvy + L_{33}W + L_{34}\dphi &=0,\label{lin3}\\
   \phantom{L_{31}\dvx + L_{22}\dvy +} L_{43}W + L_{44}\dphi
@@
 \end{align}
-where the differential operators $L_{1j}$, $L_{2j}$ and $L_{3j}$ can
-be read off Eq. \ref{final_vx}, Eq. \ref{final_vy} and 
-Eq. \ref{final_w} respectively, and $L_{4j}$ from
-Eq. \ref{lin_poisson}.
+where differential operators $L_{1j},\,L_{2j}$ and $L_{3j}$
+($j=1,2,3,4$) can be read off Appendix \ref{reduction} and $L_{4j}$
+($j=3,4$) from the linearized Poisson equation above. We remark that
+the more general problem of tilted field and variable resistivity
+%($\epsilon\neq0$ and $\eta=\eta(z)$
+can also be reduced to the above form.
 
-Note that $L_{31}\propto k_x$, so that if $k_x=0$ 
-then Eq. \ref{lin1}---\ref{lin2} are decoupled from 
-Eq. \ref{lin3}---\ref{lin4}.
+%where the differential operators $L_{1j}$, $L_{2j}$ and $L_{3j}$ can
+%be read off Eq. \ref{final_vx}, Eq. \ref{final_vy} and 
+%Eq. \ref{final_w} respectively, and $L_{4j}$ from
+%Eq. \ref{lin_poisson}.
+
+%Note that $L_{13},\,L_{14},\,L_{23},\,L_{24}$ and $L_{31}$ are
+%proportional to $k_x$. Then for for $k_x=0$ 
+%Eq. \ref{lin1}---\ref{lin2} are decoupled from 
+%Eq. \ref{lin3}---\ref{lin4}. In this case, a solution is $\dphi=W=0$,
+%with $\dvx$ and $\dvy$ determined by $L_{11}\dvx + L_{12}\dvy = 0$ and 
+%$L_{21}\dvx + L_{22}\dvy=0$, which is the incompressible MRI problem. 
+%This means that the density and potential perturbations can only
+%influence the MRI for $k_x\neq0$. Nevertheless, for $k_x=0$,
+%self-gravity can still have an effect through the background
+%equilibrium.
 
 \subsection{Boundary conditions}
-We assume $\bm{U}$ is an even function of $z$, so that $d\bm{U}/dz=0$
-at $z=0$. This implies that $\dbx=\dby=0$ at the midplane, consistent
-with a highly resistive dead zone. At the upper disk boundary
-$z=\zmax$ we assume 
+At the vertical disk boundaries $z=\pm\zmax$ we assume  
 $\dvz=\dbx=\dby=0$, and 
 \begin{align}
-  \dphi^\prime(\zmax) + k_x\dphi(\zmax) = 0
+  \dphi^\prime(\zmax) \pm k_x\dphi(\zmax) = 0
 \end{align}
-for the potential perturbation \citep[see][]{goldreich65a}.
+for the potential perturbation \citep[see][]{goldreich65a}, where the
+upper (lower) sign applies to the upper (lower) disk surface. 
+For Case 1 (purely vertical field) we additionally assume $\bm{U}$ is
+an even function of $z$, so that $d\bm{U}/dz=0$ 
+at $z=0$.
+
+%This implies that $\dbx=\dby=0$ at the midplane, consistent
+%with a highly resistive dead zone.
 
 \subsection{Numerical procedure}
-We use a pseudo-spectral approach to solve the set of linearized
+We use a pseudo-spectral method to solve the set of linearized
 equations. Let
@@
 \end{align}
-where $  \psi_k  = T_{2(k-1)},  $ and  $T_l$ is a Chebyshev polynomial of the first
-kind of order $l$ \citep{stegun65}. %The number of basis functions is
+where 
+\begin{align}
+  \psi_k  = 
+  \begin{cases}
+    T_{2(k-1)} & \text{Vertical field (case 1)},\\
+    T_{k-1}   & \text{Tilted field (case 2)},
+  \end{cases}
+\end{align}
+and  $T_l$ is a Chebyshev polynomial of the first kind of order $l$
+\citep{stegun65}. %Note that in case 1
+%The number of basis functions is
 %$N_z$ and $l_\mathrm{max}=2(N_Z-1)$ is the highest polynomial order. 
-Note that we only use even Chebyshev polynomials which automatically satisifies the symmetry 
-condition at $z=0$. Henceforth we only consider $z\geq 0$. 
+For a vertical field we only use even Chebyshev polynomials 
+to satisfy the symmetry condition at $z=0$, so in case 1 we only
+consider $z\geq 0$.
+
 The pseudo-spectral coefficients $\bm{U}_n$  are obtained by demanding
@@
 points along the vertical direction. We choose these points as the
-extrema of $T_{2(N_z-1)}$ plus end points.
+extrema of $T_{l_\mathrm{max}}$ plus end points, where
+$l_\mathrm{max}$ is the highest polynomial order.
 
-This procedure discretizes the linear equations to a matrix equation,
+The above procedure discretizes the linear equations to a matrix equation,
 \begin{align}\label{matrix_eqn}
@@
 
+\subsubsection{Non-dimensionalization}\label{non-dim}
+In practice we solve the linearized equations in non-dimensional form,
+by defining
+%We further write 
+\begin{align}
+  &z=\hat{z}H,\quad k_x =  \hat{k}_x/H,\quad \sigma = \hat{\sigma}\Omega,
+  \quad \delta\bm{v} = \csmid 
+  \delta\hat{\bm{v}}, \\ 
+  &\delta\bm{B} = B_z\delta\hat{\bm{B}},\quad
+  \delta\rho = \rho\hat{W}/\hat{c}_s^2,\quad \delta\Phi =
+  \csmid^2\delta\hat{\Phi},  
+\end{align} 
+where $\hat{c}_s=c_s/\csmid$. We also non-dimensionalize other
+background quantities, i.e. $\hat{v}_A=v_A/\csmid$,
+$\hat{S}=S/\Omega$, $\hat{\kappa}=\kappa/\Omega$,
+$\hat{\Omega}_z=\Omega_z/\Omega$ and $\hat{\eta} = \eta/(H^2\Omega)$.
+
 \subsection{Diagnostics}
+We visualize results in terms of dimensionless energy densities. We
+define 
+\begin{align}
+  &E_m \equiv \frac{|\delta \hat{\bm{B}}|^2}{2\beta},\\
+  &E_g =
+  \frac{\hat{\rho}}{\hat{c}_s^2}\left|\real\left(\hat{W}\dd\hat{\Phi}^*\right)\right|,\\  
+  &E_k = \frac{1}{2}\hat{\rho}|\delta\hat{\bm{v}}|^2,\\
+  &E_t = \frac{\hat{\rho}|\hat{W}|^2}{2\hat{c}_s^2},
+\end{align}
+as the perturbed magnetic, gravitational, kinetic and thermal
+energies, respectively, which are functions of $z$. The total energy
+is $E=E_m+E_g+E_k+E_t$ and we then normalize $E_i$ by
+$\mathrm{max}(E)$.

### 2014-02-10 paper/results.tex
major update to tex

--- 2013-11-14 paper/results.tex
@@
-\section{Results}
+\section{Results: MRI in self-gravitating disks}
 
-\subsection{MRI in self-gravitating polytropic disks with uniform
+\subsection{Influence of self-gravity on the MRI through the
+  background equilibrium}
+
+In this section we focus on the MRI. We consider polytropic disks,
+which have a well-defined thickness, and a vertical field. 
+We set the upper disk boundary at $Z_s=0.99H$ and fix $\beta=100$ and
+$k_xH=0.1$. The numerical resolution is $N_z=256$.
+
+%kx -> 0 so incompressible
+
+\subsubsection{Uniform
   resistivity}  
-We first calculate the MRI as a function of $Q$ in a polytropic disk
-with uniform resistivity ($A=1$). We consider $Q\in[0.2,4]$,
-corresponding to stronly self-gravitating to non-self-gravitating
-disks, for mid-plane Elsasser numbers $\Lambda_0\in[0.3,5]$. 
-We fix $k_x=0.1$ and $\beta=100$. The numerical resolution is
-$N_z=256$.
+We first calculate the MRI as a function of $Q$ with uniform
+resistivity ($A=1$) for mid-plane Elsasser numbers $\Lambda_0\in[0.3,10]$.
 
@@
 a function of $Q$ and $\Lambda_0$. As expected growth rates generally
-decrease with $\Lambda_0$. In the limit of ideal MHD, $\Lambda_0>1$,
-there is negligible dependence on $Q$. In the resistive limit,
-$\Lambda_0<1$, growth rates decrease noticeably for $Q<1$. This shows
-show that self-gravity can still affect the MRI even  
-though density and potential perturbations are expected to be
-negligible because $k_x\ll1$ (i.e. the linear response is
+decrease with increasing $\Lambda_0$. In the limit of ideal MHD ($\Lambda_0>1$), 
+there is negligible dependence on $Q$. In the resistive limit
+($\Lambda_0<1$), however, growth rates decrease noticeably for $Q<1$. This shows
+show that disk self-gravity affects the MRI through the background
+equilibrium, even though density and potential perturbations are  
+negligible (i.e. the linear response is
 non-self-gravitating).
@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_growth_poly_uniresis}
+  \includegraphics[width=\linewidth]{figures/compare_growth_poly_uniresis2}
   \caption{MRI growth rates as a function of $Q$ for mid-planet
@@
 
-% (We find the weak variations in
-%$\gamma$ for small $Q$ is associated with a change in the number of
-%nodes in the magnetic field perturbation.)
-%because $f$ is of order unity for the
-%range of $Q$ considered so $\lambda_\mathrm{ideal}$ remain less than
-%  unity.
-%However, in the resistive limit, the dependence of
-%  $\lambda_\mathrm{resis}$ on $f$ (and hence $Q$) is amplified by
-%  $\Lambda_0 < 1$.
-
 \cite{sano99} found that for instability to occur the linear mode
-dimensional wavelength $\lambda$ should fit inside the disk. That is,   
+wavelength $\lambda$ should fit inside the disk. That is,   
 \begin{align}\label{sano_crit}
-  \frac{\lambda}{2H} \equiv
+  \lambda \equiv
   \mathrm{max}\left(\lambda_\mathrm{ideal},\lambda_\mathrm{resis}\right)\lesssim
-  1, 
+  2H, 
 \end{align}
-where the non-dimensional wavelengths are given by 
+where the MRI wavelengths are given by 
 \begin{align}\label{lambda_ideal}
-  \lambda_\mathrm{ideal} = \frac{4\pi}{\sqrt{15}} f v_A =
-  \frac{4\pi f}{\sqrt{15\beta\rho}}
+  \frac{\lambda_\mathrm{ideal}}{2H} = \frac{4\pi}{\sqrt{15}} f \hat{v}_A =
+  \frac{4\pi f}{\sqrt{15\beta\hat{\rho}}}
 \end{align}
@@
 \begin{align}\label{lambda_resis}
-  \lambda_\mathrm{resis} = \frac{2\pi}{\sqrt{3}}\frac{\eta}{v_A f} =
-  \frac{2\pi f}{\Lambda_0}\sqrt{\frac{\rho}{3\beta}} 
+  \frac{\lambda_\mathrm{resis}}{2H} = \frac{2\pi}{\sqrt{3}}\frac{\hat{\eta}}{\hat{v}_A f} =
+  \frac{2\pi f}{\Lambda_0}\sqrt{\frac{\hat{\rho}}{3\beta}} 
 \end{align}
-%factor pi comes form converting vertical wavenumber to
-%wavelength. linear theory works with wavenumber, physical
-%interpretation in terms of wavelengths
 in the limit of high resistivity.
-%For $\Lambda_0\gtrsim 1$, we find
-%the condition $\lambda < 2H$ is well-satisifed throughout most of the
-%disk regardless of $Q$.     
-For polytropic disks the normalized density $\rho$ is weakly dependent
-on $Q$ (Fig. \ref{eqm_den}) so self-gravity only affects the
+
+Since $\hat{\rho}$ is weakly dependent
+on $Q$ (Fig. \ref{eqm_den}), self-gravity only affects the
 MRI through the factor $f$, which increases with decreasing $Q$ (see
@@
 sufficiently strong self-gravity can stabilize the MRI by making
-$\lambda > 2H$.
+$ 2H<\lambda $.
 
-%Physically, the
-%disk becomes thinner with increasing self-gravity, and the MRI will be
-%surpressed when the disk is too thin to accommodate unstable modes. 
-%MRI requires
-%$\lambda\lesssim 1$, so that the unstable wavelength fits within the
-%disk column. As shown in Appendix \ref{appen1}, $f$ increases with the
-%strength of self-gravity. This is expected to increase $\lambda$ and
-%weaken the MRI.
+In the ideal limit, we find $\lambda < 2H$ throughout most of the disk
+for the values of $Q$ considered, so that self-gravity does not affect
+growth rates significantly. However, the wavelength of the instability
+increases. This is shown in Fig. \ref{compare_result_lambda10} which
+plots the magnetic energies for $\Lambda_0=10$ and a range of Toomre $Q$
+values. The number of vertical nodes decrease with $Q$, 
+i.e. the disk accommodates fewer wavelengths because increasing
+vertical self-gravity makes it thinner.
 
-However, $f$ does not vary much for the range of $Q$ considered here. 
-In the ideal limit, we find $\lambda < 2H$ throughout most of the disk
-for all values of $Q$, so that self-gravity does not affect growth
-rates significantly. Weak variations in $\gamma$ were found to be
-associated with changes in the
+%Weak variations in growth rates in the ideal limit, seen in 
+%Fig. \ref{compare_growth_poly_uniresis}, were found to be associated
+%with changes in the mode character: $\gamma$ decreases
+%slightly when the mode begins to acquire an additional vertical node as $Q$ is
+%increased.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_result_lambda10}
+  \caption{Magnetic energies as a function of height, in the limit of
+    ideal MHD, for various strengths of self-gravity.   
+    \label{compare_result_lambda10}}
+\end{figure}
+
+Self-gravity appreciably decreases the MRI growth rates in the
+resistive limit. Fig. \ref{lambda_poly_resis} plots 
+Eq. \ref{sano_crit} for $\Lambda_0=0.3$. In the non-self-gravitating
+disk ($Q=4$) the instability criterion is marginally satisfied and the
+MRI operates. As $Q$ decreases, 
+Eq. \ref{sano_crit} is violated and the MRI growth rate is
+significantly reduced. This is seen for $Q=0.2$ where $\lambda \geq 2H$ throughout
+the disk. (The instability is not completely surpressed since
+Eq. \ref{sano_crit} is only an approximation.) Although the function
+$f$ does not change significantly for the range of $Q$ considered, the
+dependence of $\lambda_\mathrm{resis}$ on $f$ (and therefore $Q$) is
+amplified by the denominator $\Lambda_0<1$ in the resistive case. For
+resistive cases we find no nodes in the magnetic energy $E_1$ except
+at end points, i.e. only the longest wavelength survives.
 
@@
 
-\subsection{Resistive MRI in massive polytropic disks with layered
+\subsubsection{Layered
   resistivity} 
+Here we consider disks with mid-plane Elsasser number $\Lambda_0=0.1$
+and a variable resistivity profile with
+$A=10^2$. Fig. \ref{poly_layer} compares the magnetic  
+energies for $Q=0.2,\,1$ and $4$. They have similar growth rates, $\gamma/\Omega
+= 0.53,\,0.64$ and $0.66$, respectively. In the non-self-gravitating
+limit ($Q=4$), the MRI is effectively surpressed for
+$z\lesssim0.5H$. This is consistent with layered accretion. However,
+in the massive disk ($Q=0.2$) the mode occupies a wider vertical
+extent because its wavelength (in units of $H$) is larger. This
+suggests that in massive disks the MRI is not well localized to the
+active layer, as the instability takes on a more global character
+compared to non-self-gravitating disks.        
+%d is not as well localized
 
-\subsection{GI in magnetized isothermal disks with uniform
-  resistivity}
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_results_poly_layer_amp100}
+  \caption{Magnetic energies as a function of height, for layered polytropic disks with
+    such that the conductivity increases by a 
+    factor of $10^2$ in going from the mid-plane to the upper disk
+    boundary. The vertical lines indicate $\Lambda=1$ for each value
+    of $Q$.
+    \label{poly_layer}}
+\end{figure}
 
-\subsection{GI in magnetized isothermal disks with layered
-  resistivity}
+\subsection{Influence of self-gravity on the MRI through the linear
+  response}
+
+Our goal here is to examine whether or not self-gravity can amplify
+the density perturbations caused by the MRI. We choose a massive 
+isothermal disk with $Q=0.2$, which is still expected to be marginally
+stable to gravitational instability \citep[][who find
+a critical value of  $Q\simeq 0.2$]{mamat10}.  
+The upper disk boundary is set to $Z_s=H$. The field is vertical.
+
+%resolution Nz=256
+
+\subsubsection{Ideal case}
+We first consider the limit of ideal MHD by adopting a uniform
+resistivity with $\Lambda_0=100$. %Non-zero density perturbations
+%require $k_x\neq0$ so we consider linear modes as a function of $k_x$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_energy_ideal}
+  \caption{Maximum gravitational potential energy perturbation
+    associated with the MRI in a self-gravitating disk with $Q=0.2$ in
+    the limit of ideal MHD ($\Lambda_0=10^2$). 
+    \label{gravity_energy}}
+\end{figure}
+
+\subsubsection{Qualitative interpretation} 
+In order to make sense of the above results, we consider ideal MHD and
+a vertical field. For this discussion we will ignore stratification
+and set $d/dz\to \imgi k_z$ where $k_z$ is a vertical 
+wavenumber. Equivalently, we are considering regions close to the disk
+midplane. The governing equations are then
+
+\begin{align}
+  &  0= v_A^2k^2\dvx + \imgi\sigma\left(\imgi\sigma \dvx - 2\Omega\dvy + \imgi k_x \w\right),\label{simp1}\\
+  &  0= v_A^2k_z^2\left(\dvy + \frac{\imgi S
+  }{\sigma}\dvx\right) + \imgi\sigma\left(\imgi\sigma\dvy +
+  \frac{\kappa^2}{2\Omega}\dvx\right),\\
+  & 0 = -k_z^2\w + \frac{\sigma^2}{c_s^2}W + \sigma k_x \dvx, \label{simp3}\\
+  & 0 = k^2\dphi + \frac{\Omega^2}{c_s^2Q}W,\label{simp4}
+\end{align}
+where $k^2 = k_z^2 + k_x^2$.
+
+We imagine an iterative procedure to solve the above equations,
+starting from the \emph{Cowling approximation}, defined to be the
+solution to Eq. \ref{simp1}---\ref{simp3} with potential perturbation
+set to zero. This is the standard MRI and we denote the solution as
+$\dvx^{(0)}$, $\dvy^{(0)}$ and $W^{(0)}$.
+
+We now include self-gravity. In general, the MRI will have some
+density perturbation so $W^{(0)}\neq0$. The Poisson equation implies
+an associated potential perturbation,    
+\begin{align} 
+  \dphi = -\frac{\Omega^2}{c_s^2 Q k^2} W^{(0)}.
+\end{align}
+Physically, we expect $k^2\geq0$, so that a positive (negative) density
+perturbation causes a negative (positive) potential perturbation. We
+then regard $\dphi$ to alter the Cowling solution, so
+that $\dvx^{0} \to \dvx^{0} + \dvx^{(1)}$ and similarly for $\dvy$ and
+$W$. Then 
+\begin{align}
+  &   k_x\sigma \dphi = v_A^2k^2\dvx^{(1)} + \imgi\sigma\left[\imgi\sigma
+  \dvx^{(1)} - 2\Omega\dvy^{(1)} + \imgi k_x W^{(1)}\right], \label{simp_pert1}\\ 
+  &  0= v_A^2k_z^2\left[\dvy^{(1)} + \frac{\imgi S
+    }{\sigma}\dvx^{(1)}\right] + \imgi\sigma\left[\imgi\sigma\dvy^{(1)} +
+  \frac{\kappa^2}{2\Omega}\dvx^{(1)}\right],\label{simp_pert2}\\
+  & k_z^2\dphi  = \left(\frac{\sigma^2}{c_s^2}-k_z^2\right)W^{(1)} +
+  \sigma k_x \dvx^{(1)}.\label{simp_pert3} 
+\end{align}
+Now, if the perturbations to the magnetic field remain
+unchanged, i.e. the mode remains close to the standard MRI, then
+$\dvx^{(1)} \sim 0$ and $\dvy^{(1)}\sim0$, so Eq. \ref{simp_pert2} is
+satisfied. Eq. \ref{simp_pert1} then require $\dphi +
+W^{(1)}\sim0$. This is compatible with Eq. \ref{simp_pert3} if 
+\begin{align}
+  \left|k_z^2\right| \gg \left|\frac{\sigma^2}{c_s^2}\right|. \label{cond}
+\end{align}
+These assumptions imply
+\begin{align}
+  W^{(1)} \sim \frac{\Omega^2}{c_s^2 Q k^2} W^{(0)},\label{feedback}
+\end{align} 
+which indicates a non-zero density perturbation due to the
+MRI can be amplified by self-gravity. 
+%However, the magnetic field
+%perturbations remain that in the
+
+By introducing $\hat{k}_z=k_zH$, Eq. \ref{cond} can be
+non-dimensionalized to 
+\begin{align}
+  \left|\frac{\hat{\sigma}^2}{f^2\hat{c}_s^2\hat{k}_z^2}\right|\ll 1.  
+\end{align}
+
+%\begin{align}
+%  
+%\end{align}
+%However, in the full problem $\dphi$ appears in the $\mathcal{G}$
+
+%We write the horizontal velocities and enthalpy perturbations as
+%\begin{align}
+%  \dvx &= u + \Delta u,\\
+%  \dvy &= v + \Delta v,\\
+%  W    &= w + \Delta w,
+%\end{align}
+%where $u$, $v$ and $w$ is the solution to
+%Eq. \ref{simp1}---\ref{simp4} in the Cowling approximation, in which
+%the Poisson equation is ignored and the potential perturbation set to
+%zero.

### 2014-02-10 paper/setup.tex
paper update
major update to tex

@@
 permeability). We choose a barotropic equation of state, specified
-below, so that the gas pressure is given by $P=P(\rho)$. 
+below, so that the gas pressure is given by $P=P(\rho)$. The
+resistivity $\eta$ is either uniform or a prescribed function of
+height. 
+%The %resistivity is allowed to be non-uniform, $\eta=eta(z)$.
+
 %is the gas pressure given by a
 %barotropic equation of state chosen later.  
+%no curvature effects - magnetic tension neglected (limits strength of
+%toroidal field that can be considered
 
@@
 
-\subsection{Self-gravitating equilibria} 
+%We consider two cases
+%\begin{enumerate}
+%\item Purely vertical field with constant or variable
+%  resistivity, so that $\epsilon = 0$ and $\eta=\eta(z)$. 
+%\item Tilted field with uniform resistivity
+%  so that $\epsilon \neq 0$ and $\eta=\rm{constant}$.   
+%\end{enumerate}
+
+\subsection{Equilibrium disk} 
 The unperturbed disk is steady and described by
-$\rho=\rho(z)$, $\bm{B} = B\hat{\bm{z}}$ where $B$ is a constant, and
-velocity field $\bm{v} = -Sx\hat{\bm{y}}$. We assume a thin disk 
-and neglect the radial component of the self-gravitational force in
+$\rho=\rho(z)$, $\bm{B} = B_z\hat{\bm{z}} + B_y\hat{\bm{y}}$ where
+$B_{y,z}$ are constants and the toroidal field strength is
+$B_y=\epsilon B_z$. The equilibrium 
+velocity field is $\bm{v} = -Sx\hat{\bm{y}}$. We consider Keplerian
+disks so that $S = 3\Omega/2$ and the epicyclic frequency 
+$\kappa\equiv\sqrt{2\Omega(2\Omega-S)}=\Omega=\Omega_z$. 
+%$\Omega = \Omega_z.$ 
+We assume a thin disk and neglect the radial component of the
+self-gravitational force in 
 the unperturbed disk.
@@
 \item isothermal disks with $P=c_{s0}^2\rho$\label{iso_eos}; 
-\item polytropic disks with $P=K\rho^2$ with $K=c_{s0}^2/2\rho(0)$;
+\item polytropic disks with $P=K\rho^2$ with $K=c_{s0}^2/2\rho_0$;
 \end{inparaenum}
-where $c_{s0}$ is the global sound speed in the isothermal disk, and
+where $\rho_0\equiv\rho(0)$. The sound speed $c_s\equiv\sqrt{dP/d\rho}$
+so that $c_{s0}$ is the global sound speed in the isothermal disk, and
 is the mid-plane sound speed in the polytropic disk. For the polytropic
@@
 
-We solve for $\hat{\rho}\equiv\rho/\rho_0$, where
-the mid-plane density $\rho_0\equiv\rho(0)$ is parameterized via the
-Toomre $Q$ parameter such that $\rho_0=\Omega^2/4\pi G Q$. Boundary
-conditions are $\hat{\rho}=1$ and $d\hat{\rho}/dz=0$  at $z=0$. This
+We solve for $\hat{\rho}\equiv\rho/\rho_0$ with 
+%$\rho_0\equiv\rho(0)$ is the mid-plane density.
+%is parameterized via the
+%Toomre $Q$ parameter such that $\rho_0=\Omega^2/4\pi G Q$. 
+boundary conditions $\hat{\rho}=1$ and $d\hat{\rho}/dz=0$  at $z=0$. This
 is done numerically for isothermal disks and analytically for the
 polytropic disk (see Appendix \ref{appen1}). Examples of density
-profiles are shown in Fig. \ref{eqm_den}. The density field is weakly
-dependent on the strength of self-gravity provided the vertical
-co-ordinate is appropriately normalized.
+profiles are shown in Fig. \ref{eqm_den}. The normalized density field
+is weakly dependent on the strength of self-gravity provided the
+vertical co-ordinate is appropriately scaled as well.
 
@@
 
+%\subsection{}
+
 \subsection{Resistivity profile}\label{resis_profile}
-Following \cite{fleming03}, we adopt the resistivity profile
+In this work we consider either constant resistivity or a
+resistivity prescription such that $\eta(z)$ increases away from the
+midplane. In the latter case, we follow \cite{fleming03} and adopt the
+resistivity profile 
 \begin{align}
@@
 \end{align}
-where $v_A \equiv B/\sqrt{\mu_0\rho}$ is the Alfven speed. Example
+where $v_A \equiv B_z/\sqrt{\mu_0\rho}$ is the vertical Alfven speed. Example
 profiles of $\Lambda$ are shown in Fig. \ref{eqm_resis} for the
 isothermal disk. The disk may be considered ideal where $\Lambda
-\gtrsim 1$. Since $\rho(z/H)$ is weakly dependent on  self-gravity, the
+\gtrsim 1$. Since $\hat{\rho}(z/H)$ is weakly dependent on  self-gravity, the
 ideal region only expands slightly as $Q$ is decreased.
@@
 \end{align}
-\citep{mamat10}. The plasma $\beta$ measures the inverse strength of
-the magnetic field,
+\citep{mamat10}, which is used to set the mid-plane density $\rho_0$. 
+The relation between $Q$ and the Toomre parameter for gravitational
+instability of razor-thin disks is described in Appendix \ref{q3d2d}.
+
+The plasma $\beta$ measures the inverse strength of the  
+magnetic field, 
 \begin{align}
   \beta \equiv \frac{\csmid^2}{v_{A0}^2} =
-  \frac{\csmid^2\mu_0\rho_0}{B^2},  
+  \frac{\csmid^2\mu_0\rho_0}{B_z^2},  
 \end{align}
@@
 factor of order unity.
+
 %The Elsasser number as a function of height is $\Lambda(z)\equiv

### 2014-02-12 paper/results.tex
backup

@@
 \section{Results: MRI in self-gravitating disks}
+In this section we focus on the MRI. We use the vertical field setup
+of case 1. The numerical resolution is $N_z=256$.
 
@@
 
-In this section we focus on the MRI. We consider polytropic disks,
-which have a well-defined thickness, and a vertical field. 
-We set the upper disk boundary at $Z_s=0.99H$ and fix $\beta=100$ and
-$k_xH=0.1$. The numerical resolution is $N_z=256$.
+Here we consider polytropic disks, which have a well-defined
+thickness. Numerically, the upper disk boundary is set to
+$Z_s=0.99H$. We fix $\beta=100$ and $k_xH=0.1$.
 
@@
   response}  
-
 Our goal here is to examine whether or not self-gravity can amplify
@@
 a critical value of  $Q\simeq 0.2$]{mamat10}.   
-The upper disk boundary is set to $Z_s=H$. The field is vertical.
+The upper disk boundary is set to $Z_s=H$.
 
@@
 %require $k_x\neq0$ so we consider linear modes as a function of $k_x$.   
+Fig. \ref{gravity_energy} shows the maximum gravitational potential
+energy perturbation, as a function of $k_x$ for several values of
+$\beta$. The potential perturbation is negligible for all cases when 
+$k_xH\lesssim 0.5$, since the MRI becomes incompressible as $k_x\to
+0$.
 
@@
 
+For $\beta\gg 1$, i.e. a weak field, density perturbations are
+negligible and the standard MRI operates. However, as $\beta$ is
+lowered we find non-negligible potential perturbation for
+$k_xH=O(1)$. This suggets that in a strongly magnetized disk that
+still permits the MRI, the associated density perturbation can be
+important when the disk is self-gravitating.
+
+Fig. \ref{mri_massive} shows an example of a MRI mode with
+non-negligible density perturbation. The growth rate $\gamma\simeq
+0.4\Omega$ is comparable to that of the most unstable mode
+($\gamma\simeq 0.6\Omega$ for $k_x\simeq0$). 
+We therefore expect significant density perturbations to grow
+dynamically, even though the disk is expected to be gravitationally
+stable.
+
+%dynamical
+%timescales 
+%($\gamma\sim0.4\Omega$)
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/result_ideal_sg}
+  \caption{Example of a linear MRI mode with non-zero density and
+    potential perturbation, in a strongly magnetized massive disk
+    ($\beta=10$ and $Q=0.2$) in the limit of ideal MHD
+    ($\Lambda_0=10^2$).   
+    \label{mri_massive}}
+\end{figure}
+
+\subsubsection{Resistive case}
+
 \subsubsection{Qualitative interpretation} 
-In order to make sense of the above results, we consider ideal MHD and
-a vertical field. For this discussion we will ignore stratification
-and set $d/dz\to \imgi k_z$ where $k_z$ is a vertical 
-wavenumber. Equivalently, we are considering regions close to the disk
-midplane. The governing equations are then
+In order to make sense of the above results, we return to ideal MHD. 
+We consider regions close to the disk midplane ($z\sim 0$), where
+self-gravity is expected to be most important. For this discussion we
+will ignore stratification and set $d/dz\to \imgi k_z$ where $k_z$ is
+a vertical wavenumber. The governing equations are then
 
@@
   }{\sigma}\dvx\right) + \imgi\sigma\left(\imgi\sigma\dvy +
-  \frac{\kappa^2}{2\Omega}\dvx\right),\\
+  \frac{\kappa^2}{2\Omega}\dvx\right),\label{simp2}\\
   & 0 = -k_z^2\w + \frac{\sigma^2}{c_s^2}W + \sigma k_x \dvx, \label{simp3}\\
@@
 \end{align}
-These assumptions imply
+For the ideal MRI,  we assume $k_z\sim \Omega/v_A$. Then
+$|\sigma^2/c_s^2k_z^2|\sim |\sigma^2/\Omega^2\beta|\ll1$ because
+$|\sigma|\lesssim \Omega$ and we are considering $\beta\gtrsim
+10$. Thus Eq. \ref{cond} is generally satisfied.
+
+The above assumptions imply
 \begin{align}
@@
 which indicates a non-zero density perturbation due to the
-MRI can be amplified by self-gravity. 
-%However, the magnetic field
-%perturbations remain that in the
+MRI can be amplified by self-gravity. Now, for $k_xH\sim 1$ we have
+$|k_z^2/k_x^2|\sim \beta/f\gg1$ because $f=O(1)$ and $\beta\gtrsim10$
+for the cases considered above. Then
+\begin{align}
+  \left|\frac{W^{(1)}}{W^{(0)}}\right| \sim \frac{1}{Q\beta}, 
+\end{align}
+ suggesting stronger amplification of the density field by 
+ self-gravity with increasing field strength (decreasing
+ $\beta$).
 
-By introducing $\hat{k}_z=k_zH$, Eq. \ref{cond} can be
-non-dimensionalized to 
+ For weak fields, the MRI has small vertical wavelength,
+ implying rapid density variations with height. The magnitude of the
+ corresponding gravitational potential is then expected to be small
+ because of internal cancellations, since the potential is essentially
+ the integral of the density field.
+
+% not expect significant potential perturbations
+
+%Conversely, for strong fields, only a single MRI wavelength occupies
+%the fluid column.
+
+The above arguments also apply to the resistive
+limit. Eq. \ref{simp3}---\ref{simp4} are unchanged, while resistive
+terms appearing in Eq. \ref{simp1}---\ref{simp2} only involve the
+potential perturbation through $\w$. For the resistive MRI we assume
+$k_z\sim v_A/\eta$ and $|\sigma|\sim v_A^2/\eta = \Lambda\Omega$
+\citep{sano99}. Then $|\sigma^2/c_s^2k_z^2|\sim 1/\beta \ll 1$ so
+Eq. \ref{cond} is satisfied. Noting that $k_z^2\sim
+\Lambda^2\Omega^2\beta/c_s^2$, the feedback equation becomes
 \begin{align}
-  \left|\frac{\hat{\sigma}^2}{f^2\hat{c}_s^2\hat{k}_z^2}\right|\ll 1.  
+  \left|\frac{W^{(1)}}{W^{(0)}}\right| \sim
+  \frac{1}{Q\left(f^2\hat{k}_x^2 + \beta\Lambda^2\right)},
 \end{align}
-
-%\begin{align}
-%  
-%\end{align}
-%However, in the full problem $\dphi$ appears in the $\mathcal{G}$
-
-%We write the horizontal velocities and enthalpy perturbations as
-%\begin{align}
-%  \dvx &= u + \Delta u,\\
-%  \dvy &= v + \Delta v,\\
-%  W    &= w + \Delta w,
-%\end{align}
-%where $u$, $v$ and $w$ is the solution to
-%Eq. \ref{simp1}---\ref{simp4} in the Cowling approximation, in which
-%the Poisson equation is ignored and the potential perturbation set to
-%zero.
+so increasing the resistivity (decreasing $\Lambda$) should enhance
+density perturbations.

### 2014-02-12 paper/results.tex
tex update

@@
 the density perturbations caused by the MRI. We choose a massive 
-isothermal disk with $Q=0.2$, which is still expected to be marginally
+isothermal disk with $Q=0.2$ (corresponding to $Q_\mathrm{2D}=0.72$),
+which is still expected to be marginally 
 stable to gravitational instability \citep[][who find
@@
 \subsubsection{Resistive case}
+We repeat the above calculation for resistive disks. We fix
+$\beta=100$. Fig. \ref{gravity_energy_resis}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_energy_resis}
+  \caption{Maximum gravitational potential energy perturbation
+    associated with the MRI in a self-gravitating disk with $Q=0.2$,
+    for a range of midplane Elsasser numbers.   
+    \label{gravity_energy_resis}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/result_resis_sg}
+  \caption{Example of a linear resistive MRI mode with significant
+    gravitational potential perturbation.  
+    \label{mri_massive_resis}}
+\end{figure}
 

### 2014-02-21 paper/results.tex
marginal notes

@@
     associated with the MRI in a self-gravitating disk with $Q=0.2$ in
-    the limit of ideal MHD ($\Lambda_0=10^2$). 
+    the limit of ideal MHD ($\Lambda_0=10^2$), for a range of field
+    strengths $\beta$.  
     \label{gravity_energy}}
@@
   \includegraphics[width=\linewidth]{figures/result_ideal_sg}
-  \caption{Example of a linear MRI mode with non-zero density and
+  \caption{Example of a linear MRI mode with non-negligible density and
     potential perturbation, in a strongly magnetized massive disk
@@
 \subsubsection{Resistive case}
-We repeat the above calculation for resistive disks. We fix
-$\beta=100$. Fig. \ref{gravity_energy_resis}
+We repeat the above calculation for resistive disks, but fix
+$\beta=100$. Fig. \ref{gravity_energy_resis} shows the maximum
+perturbed gravitational energy increases with 
+resistivity. Interestingly, in the highly resistive case
+the MRI can be energeically dominated by self-gravity. An example is shown in
+Fig. \ref{mri_massive_resis}, which  plots the energies assocaited with the 
+linear MRI mode with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
+disk. The gravitational potential energy significantly exceeds the
+magnetic energy near the midplane ($z\lesssim0.3H$).
 
@@
     associated with the MRI in a self-gravitating disk with $Q=0.2$,
-    for a range of midplane Elsasser numbers.   
+    for a range of midplane Elsasser numbers $\Lambda_0$ at fixed
+    field strength $\beta=100$.    
     \label{gravity_energy_resis}}
 \end{figure}
+
+We also calculated the mode in Fig. \ref{mri_massive_resis} with the \emph{Cowling 
+  approximation}, where the Poisson equation is ignored in the 
+linearized equations and the potential perturbation set to zero  
+(formally, we set $Q\to\infty$ in Eq. \ref{lin_poisson}).  
+We compare the magnetic energy in the Cowling approximation to that of
+the full problem in Fig. \ref{mri_massive_cowling}. The growth rate
+increases when self-gravity is included in the linear response, as
+expected, but only slightly. $\gamma$ and $E_m$ are similar in both cases,
+indicating that the instability in the self-gravitating calculation is
+still due to the MRI.
 
@@
   \caption{Example of a linear resistive MRI mode with significant
-    gravitational potential perturbation.  
+    gravitational potential energy perturbation.  The vertical line
+    inidicates $\Lambda=1$.  
     \label{mri_massive_resis}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_result_cowling}
+  \caption{Magnetic energy associated with the linear mode shown in
+    Fig. \ref{mri_massive_resis} (solid) compared with that computed
+    under the Cowling approximation (dotted). 
+    \label{mri_massive_cowling}}
 \end{figure}
@@
 We imagine an iterative procedure to solve the above equations,
-starting from the \emph{Cowling approximation}, defined to be the
-solution to Eq. \ref{simp1}---\ref{simp3} with potential perturbation
-set to zero. This is the standard MRI and we denote the solution as
+starting from the Cowling approximation. %defined to be the
+%solution to Eq. \ref{simp1}---\ref{simp3} with potential perturbation
+%set to zero. 
+This is the standard MRI and we denote the solution as
 $\dvx^{(0)}$, $\dvy^{(0)}$ and $W^{(0)}$.
 
-We now include self-gravity. In general, the MRI will have some
-density perturbation so $W^{(0)}\neq0$. The Poisson equation implies
+We now include self-gravity. In general, the compressible MRI has
+$W^{(0)}\neq0$. The Poisson equation implies
 an associated potential perturbation,    

### 2014-03-17 paper/linear.tex
marginal notes
tex updates

@@
 
-\subsection{Boundary conditions}
-At the vertical disk boundaries $z=\pm\zmax$ we assume  
-$\dvz=\dbx=\dby=0$, and 
+\subsection{Domain and boundary conditions}
+
+\subsubsection{Case 1: vertical field}
+We impose $\bm{U}$ to be an even function of $z$, so that
+$d\bm{U}/dz=0$ at $z=0$. This reduces the computational domain to
+$z\in[0,Z_s]$ and permits higher numerical resolution. At the disk
+surface we set %$\dbx=\dby=\dvz=0$ 
 \begin{align}
-  \dphi^\prime(\zmax) \pm k_x\dphi(\zmax) = 0
+  \dbx(Z_s) = \dby(Z_s) = \dvz (Z_s) = 0, 
 \end{align}
-for the potential perturbation \citep[see][]{goldreich65a}, where the
-upper (lower) sign applies to the upper (lower) disk surface. 
-For Case 1 (purely vertical field) we additionally assume $\bm{U}$ is
-an even function of $z$, so that $d\bm{U}/dz=0$ 
-at $z=0$.
+so the field remains vertical. The potential boundary
+condition is 
+\begin{align}
+  \dphi^\prime(\zmax) + k_x\dphi(\zmax) = 0 \label{pot_bc1}
+\end{align}
+\citep{goldreich65a}.
+
+\subsubsection{Case 2: tilted field} 
+In this more general setup the computational domain is
+$z\in[-Z_s, Z_s]$. At the disk surfaces we adopt the `halo' model of
+\cite{gammie94}, so that
+\begin{align}
+  \Delta(\pm Z_s) & = 0, \\
+  \dby(\pm Z_s) & = 0,\\
+  \dbz(\pm Z_s) \mp \imgi \dbx(\pm Z_s) &= 0,
+\end{align}
+and the potential boundary condition is 
+\begin{align}
+  \dphi^\prime(\pm \zmax) \pm k_x\dphi(\pm\zmax) =  - \left. \frac{\Omega^2
+    \rho \xi_z}{\rho_0Q}\right|_{\pm\zmax}, \label{pot_bc2}
+\end{align}
+where $\xi_z=\dvz/\imgi\sigma$ is the vertical Lagrangian
+displacement. In practice,  Eq. \ref{pot_bc1} and \ref{pot_bc2} are
+very similar because $\rho(\pm\zmax)\ll\rho_0$.
+
+%We adopt two kinds of vertical boundary conditions depending on the
+%field geometry. 
+%In both case 1 and case 2 the potential boundary
+%condition is 
+%\begin{align}
+%  \dphi^\prime(\zmax) \pm k_x\dphi(\zmax) = 0
+%\end{align}
+
+%At the vertical disk boundaries $z=\pm\zmax$ we assume  
+%$\dvz=\dbx=\dby=0$, and
+
+%for the potential perturbation \citep[see][]{goldreich65a}, where the
+%upper (lower) sign applies to the upper (lower) disk surface. 
+%For Case 1 (purely vertical field) we additionally assume $\bm{U}$ is
+%an even function of $z$, so that $d\bm{U}/dz=0$ 
+%at $z=0$.
 
@@
   \begin{cases}
-    T_{2(k-1)} & \text{Vertical field (case 1)},\\
-    T_{k-1}   & \text{Tilted field (case 2)},
+    T_{2(k-1)} & B_y \equiv 0 \text{ (case 1)},\\
+    T_{k-1}   & B_y\neq 0 \text{ (case 2)},
   \end{cases}
@@
 and  $T_l$ is a Chebyshev polynomial of the first kind of order $l$
-\citep{stegun65}. %Note that in case 1
+\citep{stegun65}. Note that for case 1 the midplane symmetry condition
+is taken care of by the choice of basis functions. 
+%Note that in case 1
 %The number of basis functions is
 %$N_z$ and $l_\mathrm{max}=2(N_Z-1)$ is the highest polynomial order. 
-For a vertical field we only use even Chebyshev polynomials 
-to satisfy the symmetry condition at $z=0$, so in case 1 we only
-consider $z\geq 0$.
+%For a vertical field we only use even Chebyshev polynomials 
+%to satisfy the symmetry condition at $z=0$, so in case 1 we only
+%consider $z\geq 0$.
 
@@
 extrema of $T_{l_\mathrm{max}}$ plus end points, where
-$l_\mathrm{max}$ is the highest polynomial order.
+$l_\mathrm{max}$ is the highest polynomial order. Our standard
+resolution is $N_z=256$.
 
@@
 \subsubsection{Non-dimensionalization}\label{non-dim}
-In practice we solve the linearized equations in non-dimensional form,
+We solve the linearized equations in non-dimensional form,
 by defining
@@
 energies, respectively, which are functions of $z$. The total energy
-is $E=E_m+E_g+E_k+E_t$ and we then normalize $E_i$ by
-$\mathrm{max}(E)$.
+is $E=E_m+E_g+E_k+E_t$. We use $\avg{\cdot}$ to denote an average over
+$z$. 
+%and we then normalize $E_i$ by
+%$\mathrm{max}(E)$.

### 2014-03-17 paper/paper.tex
minor paper update
paper update
major update to tex
marginal notes
tex updates

@@
 %\documentclass{article}
-\usepackage{draftwatermark}
+%\usepackage{draftwatermark}
 \usepackage{paralist}
-\shorttitle{MRI in massive disks}
+\shorttitle{Magnetized massive disks}
 \shortauthors{M-K. Lin}
@@
 \newcommand{\dbz}{\dd B_z}
-\newcommand{\w}{ \tilde{W}}
+\newcommand{\w}{ \widetilde{W}}
+%\newcommand{\sbar}{\bar{\sigma}}
 \newcommand{\dphi}{\dd \Phi}
@@
 
- \title{Resistive magneto-rotational instability in massive
-  protoplanetary disks}
+\title{Linear stability of magnetized massive protoplanetary disks}
 
-\author{Min-Kai Lin}
+\author{Min-Kai Lin (+Menou, Fromang, Youdin)}
 \affil{Canadian Institute for Theoretical Astrophysics,
@@
 \section{Introduction}
+Understanding the evolution of protoplanetatry disks (PPDs) is key to planet
+formation theory. A central aspect of PPD evolution is the mechanism
+of angular momentum transport.
 
@@
 \input{linear}
-%\input{results}
-%\input{iso}
-%\input{summary}
-
-%% \acknowledgments
-%% {\bf
-%% I thank the referee, P. Barge, for suggesting the idea of a tilted vorticity column. }
-%% I also thank S.-J. Paardekooper for comments on the first version of this paper.
+\input{results}
+\input{results_gi}
+\input{results_azi}
 

### 2014-03-17 paper/results.tex
tex updates

@@
-\section{Results: MRI in self-gravitating disks}
+\section{MRI in self-gravitating disks}
 In this section we focus on the MRI. We use the vertical field setup
-of case 1. The numerical resolution is $N_z=256$.
+of case 1. We first consider MRI modes with negligble densit/potential
+perturbations to see the effect of self-gravity on the MRI through the
+background stratification, then go on to consider MRI modes with
+significant density/potential perturbations in massive disks.
 
@@
 
-Here we consider polytropic disks, which have a well-defined
-thickness. Numerically, the upper disk boundary is set to
-$Z_s=0.99H$. We fix $\beta=100$ and $k_xH=0.1$.
+Here we use polytropic disks, which have a well-defined
+thickness. The upper disk boundary is set to 
+$Z_s=0.99H$. We fix $\beta=100$ and $k_xH=0.1$ unless otherwise
+stated.
 
@@
 suggests that in massive disks the MRI is not well localized to the
-active layer, as the instability takes on a more global character
-compared to non-self-gravitating disks.        
+active layer.
+
+%MRI penetrates 
 %d is not as well localized
@@
 \end{figure}
+
+\subsubsection{Dependence on $k_x$}
+The above experiments show that with increasing 
+disk self-gravity, the MRI becomes more global in the vertical
+direction. We find a similar result in the horizontal direction. 
+In Fig. \ref{compare_growth_poly_kx} we plot MRI growth rates as a
+function of $k_x$ for a range of $Q$ values. Increasing self-gravity
+decreases the cut-off horizontal wavenumber for the MRI, implying that
+small-scale MRI does not operate in massive disks.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_growth_poly_varQ_kx}
+  \caption{MRI growth rates in self-gravitating polytropic disks, as a
+    function of the horizontal wavenumber $k_x$. The disk is ideal
+    ($\Lambda_0=10^2,\, A=1$) with $\beta = 40$. These modes have negligible
+    density/potential perturbations.  
+    \label{compare_growth_poly_kx}}
+\end{figure}
+
+%these modes are still non-sg
+%use beta = 40 so only a small range of kx need to be simulated to
+%demonstrate this effect
+%{\bf cut-off $k_xH$ decreases with increasing SG -> small-scale MRI
+%  stabilized by self-gravity}
 
@@
 %require $k_x\neq0$ so we consider linear modes as a function of $k_x$.   
-Fig. \ref{gravity_energy} shows the maximum gravitational potential
+Fig. \ref{gravity_energy} shows the mean gravitational potential
 energy perturbation, as a function of $k_x$ for several values of
@@
   \includegraphics[width=\linewidth]{figures/compare_energy_ideal}
-  \caption{Maximum gravitational potential energy perturbation
-    associated with the MRI in a self-gravitating disk with $Q=0.2$ in
+  \caption{Average gravitational potential energy perturbation
+    associated with the MRI, compared to the average magnetic energy
+    perturbation, in a self-gravitating disk with $Q=0.2$ in 
     the limit of ideal MHD ($\Lambda_0=10^2$), for a range of field
@@
 We repeat the above calculation for resistive disks, but fix
-$\beta=100$. Fig. \ref{gravity_energy_resis} shows the maximum
+$\beta=100$. Fig. \ref{gravity_energy_resis} shows the 
 perturbed gravitational energy increases with   
 resistivity. Interestingly, in the highly resistive case
-the MRI can be energeically dominated by self-gravity. An example is shown in
-Fig. \ref{mri_massive_resis}, which  plots the energies assocaited with the 
-linear MRI mode with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
-disk. The gravitational potential energy significantly exceeds the
-magnetic energy near the midplane ($z\lesssim0.3H$).
+the MRI can be energeically dominated by self-gravity, i.e. the case
+with $\Lambda_0=0.1$ and $k_xH\simeq1.3$. We also calculated this mode
+in the \emph{Cowling  
+  approximation}, where the Poisson equation is ignored in the 
+linearized equations and the potential perturbation set to zero  
+(formally letting $Q\to\infty$ in Eq. \ref{lin_poisson}).
+Fig. \ref{mri_massive_cowling} compares the magnetic energy in the
+Cowling approximation to that of 
+the full problem in . The growth rate increases when self-gravity is
+included in the linear response, as expected, but only
+slightly. $\gamma$ and $E_m$ are similar in both cases, indicating
+that the instability in the self-gravitating calculation is 
+still the MRI.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_energy_resis}
-  \caption{Maximum gravitational potential energy perturbation
-    associated with the MRI in a self-gravitating disk with $Q=0.2$,
+  \caption{Average gravitational potential energy perturbation
+    associated with the MRI, compared to the average magnetic energy
+    perturbation, in a self-gravitating disk with $Q=0.2$,
     for a range of midplane Elsasser numbers $\Lambda_0$ at fixed
@@
 
-We also calculated the mode in Fig. \ref{mri_massive_resis} with the \emph{Cowling 
-  approximation}, where the Poisson equation is ignored in the 
-linearized equations and the potential perturbation set to zero  
-(formally, we set $Q\to\infty$ in Eq. \ref{lin_poisson}).  
-We compare the magnetic energy in the Cowling approximation to that of
-the full problem in Fig. \ref{mri_massive_cowling}. The growth rate
-increases when self-gravity is included in the linear response, as
-expected, but only slightly. $\gamma$ and $E_m$ are similar in both cases,
-indicating that the instability in the self-gravitating calculation is
-still due to the MRI.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/result_resis_sg}
-  \caption{Example of a linear resistive MRI mode with significant
-    gravitational potential energy perturbation.  The vertical line
-    inidicates $\Lambda=1$.  
-    \label{mri_massive_resis}}
-\end{figure}
-
 \begin{figure}
   \includegraphics[width=\linewidth]{figures/compare_result_cowling}
-  \caption{Magnetic energy associated with the linear mode shown in
-    Fig. \ref{mri_massive_resis} (solid) compared with that computed
+  \caption{Magnetic energy associated with the linear mode with
+    largest gravitational-to-magnetic energy ratio in
+    Fig. \ref{gravity_energy_resis} (solid) compared with that computed
     under the Cowling approximation (dotted). 
@@
 \end{figure}
+
+%An example is shown in
+%Fig. \ref{mri_massive_resis}, which  plots the energies assocaited with the 
+%linear MRI mode with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
+%disk. The gravitational potential energy significantly exceeds the
+%magnetic energy near the midplane ($z\lesssim0.3H$).
+
+%\begin{figure}
+%  \includegraphics[width=\linewidth]{figures/result_resis_sg}
+%  \caption{Example of a linear resistive MRI mode with significant
+%    gravitational potential energy perturbation.  The vertical line
+%    inidicates $\Lambda=1$.  
+%    \label{mri_massive_resis}}
+%\end{figure}
 
@@
 then regard $\dphi$ to alter the Cowling solution, so
-that $\dvx^{0} \to \dvx^{0} + \dvx^{(1)}$ and similarly for $\dvy$ and
+that $\dvx^{(0)} \to \dvx^{(0)} + \dvx^{(1)}$ and similarly for $\dvy$ and
 $W$. Then 

### 2014-03-17 paper/results_azi.tex
tex updates

\section{Effect of an azimuthal field}
We now extend the previous investigation to disks with an azimuthal
field, so that $B_y\neq 0$.

%cannot consider superthermal fields
%compressibility important f

%\subsection{}

### 2014-03-17 paper/results_gi.tex
tex updates

\section{Gravitationally unstable disks} 
Gravitational instability (GI) becomes possible in a sufficiently
massive and/or cold disk. We explore how the emergence of GI affects
the MRI, and vice versa. To do so, we compute unstbale modes for  
isothermal disks with $Q < 0.2 $ ($Q_\mathrm{2D}\lesssim 0.67$) which
permits gravitational instability as well as the MRI.

\subsection{Ideal disks}
We first consider ideal MHD by setting $\Lambda_0=100$ and $A=1$. 
Fig. \ref{compare_growth3} show growth rates for 
modes with $k_xH=1$ as a function of $\beta$ in disks with $Q=0.18$,
$Q=0.14$ and $Q=0.12$. The colorbar quantifies the importance of
self-gravity with 
\begin{align}
 \tau \equiv \frac{\avg{E_g}}{\avg{E_g} + \avg{E_m}}. 
\end{align} 
Thus, modes with $\tau  = 1$ are
energetically dominated by self-gravity (GI) and $\tau\ll1$ represents a mode
dominated by magnetic perturbations (MRI).

All three cases display distinct GI modes (red/brown branch). The GI
growth rates are $\gamma\simeq 0.25\Omega,\,0.6\Omega,\,0.8\Omega$ for
$Q=0.18,\,0.14,\,0.12$, respecitvely. GI is stabilized by magnetic
pressure for sufficiently small $\beta$, and the critical field
strength increases with the self-gravity \citep{nakamura83}. For
$Q=0.18$, GI is stabilized for $\beta \lesssim 15$. Nevertheless, 
the MRI branch for $\beta < 15$ becomes self-gravitating, so that
density perturbations are still expected even though  GI does not
formally operate.

%critical beta increases with $Q$

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 2.1cm 0cm
    0cm]{figures/compare_growth3_kx1_Q0d18.ps} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 2.1cm 0cm
    0.5cm]{figures/compare_growth3_kx1_Q0d14.ps} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
    0.5cm]{figures/compare_growth3_kx1_Q0d12.ps} 
  \caption{Growth rates for linear modes with $k_xH=1$ in an isothermal disk with
    $Q=0.18$ (top), $Q=0.14$ (middle) and $Q=0.14$ (bottom). The
    colorbar measures the importance of 
    self-gravity.  
    \label{compare_growth3}}
\end{figure}

The GI and MRI branches only interact when their growth rates
are similar. This is seen in Fig. \ref{compare_growth3} for $Q=0.18$
where the GI branch approaches an MRI branch at $\beta\simeq 25,\,
\gamma\simeq 0.2\Omega$. 
%Although the MRI is still most unstable, it is
%self-gravitating for $\beta\lesssim 50$ (uppermost curve). 
Some modes may have been missed in our
numerical search of eigenfrequencies. Thus, `breaks' along the
GI branch for $Q=0.18$ and $Q=0.12$ does not prove that the GI and MRI branches do not
intersect. However, the continous variation of $\tau$ with respect to
$\beta$ suggests that unstable modes can transition smoothly from
magnetically-dominated to gravitationally-dominated at low $\beta$.

For $Q=0.12$ the GI growth rate exceeds all MRI modes. Notice the most
unstable MRI branch becomes self-gravitating for $20 \lesssim
\beta\lesssim 40$, but magnetic perturbations remain negligible
for GI. 
%we generally find MRI modes aquire significant potential
%perturbations but GI modes always have negligible magnetic
%perturbation. 
Fig. \ref{compare_growth3_Q01d2} show growth rates
as a function of $k_xH$ for the $Q=0.12,\,\beta=20$ disk. All
perturbations with $k_xH \lesssim 3.5$ are unstable. MRI operates on the largest
length scale, followed by GI at $0.7 \lesssim k_xH 1.7$. The 
transition between MRI to GI is smoother with increasing $k_xH$ as the
MRI becomes compressible.

%Modes with $k_xH
% \gtrsim 1.7$ are neither pure MRI or GI. 
%GI wave lengths fit inside channel solutions

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_growth3_Q0d12_B20.ps}  
  \caption{Growth rates of unstable modes in the massive isothermal
    disk with $Q=0.12$ and $\beta=20$, as a function of the horizontal
    wavenumber $k_x$. The colorbar measures the importance of
    self-gravity. The solid line corresponds to MRI modes in the
    Cowling approximation. 
    \label{compare_growth3_Q01d2}}
\end{figure}

\subsection{Layered disks}
Can a disk with a dense, resistive midplane and a highly conducting
atmosphere  be subject to both MRI and GI but at different heights?

%GI in the midplne and MRI in the surface?

### 2014-03-17 paper/setup.tex
tex updates

@@
 \end{align}
-where $v_A \equiv B_z/\sqrt{\mu_0\rho}$ is the vertical Alfven speed. Example
-profiles of $\Lambda$ are shown in Fig. \ref{eqm_resis} for the
-isothermal disk. The disk may be considered ideal where $\Lambda
+where $v_A \equiv B_z/\sqrt{\mu_0\rho}$ is the vertical Alfven speed. 
+%Example
+%profiles of $\Lambda$ are shown in Fig. \ref{eqm_resis} for the
+%isothermal disk. 
+The disk may be considered ideal where $\Lambda
 \gtrsim 1$. Since $\hat{\rho}(z/H)$ is weakly dependent on  self-gravity, the
-ideal region only expands slightly as $Q$ is decreased.
+ideal region only changes slightly as $Q$ is decreased.
 
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/elsasser_iso}
-  \caption{Elsasser number as prescribed in \S\ref{resis_profile} for
-    isothermal disks with $Q=0.2$ (solid) 
-    and $Q=10$ (dotted). The vertical line
-    indicates $\Lambda=1$ in each case. \label{eqm_resis}}
-\end{figure}
+%\begin{figure}
+%  \includegraphics[width=\linewidth]{figures/elsasser_iso}
+%  \caption{Elsasser number as prescribed in \S\ref{resis_profile} for
+%    isothermal disks with $Q=0.2$ (solid) 
+%    and $Q=10$ (dotted). The vertical line
+%    indicates $\Lambda=1$ in each case. \label{eqm_resis}}
+%\end{figure}
 

### 2014-03-20 paper/results.tex
tex update

@@
 
-Fig. \ref{mri_massive} shows an example of a MRI mode with
-non-negligible density perturbation. The growth rate $\gamma\simeq
-0.4\Omega$ is comparable to that of the most unstable mode
-($\gamma\simeq 0.6\Omega$ for $k_x\simeq0$). 
-We therefore expect significant density perturbations to grow
-dynamically, even though the disk is expected to be gravitationally
-stable.
+%Fig. \ref{mri_massive} shows an example of a MRI mode with
+%non-negligible density perturbation. The growth rate $\gamma\simeq
+%0.4\Omega$ is comparable to that of the most unstable mode
+%($\gamma\simeq 0.6\Omega$ for $k_x\simeq0$). 
+%We therefore expect significant density perturbations to grow
+%dynamically, even though the disk is expected to be gravitationally
+%stable.
 
-%dynamical
-%timescales 
-%($\gamma\sim0.4\Omega$)
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/result_ideal_sg}
-  \caption{Example of a linear MRI mode with non-negligible density and
-    potential perturbation, in a strongly magnetized massive disk
-    ($\beta=10$ and $Q=0.2$) in the limit of ideal MHD 
-    ($\Lambda_0=10^2$).   
-    \label{mri_massive}}
-\end{figure}
+%\begin{figure}
+%  \includegraphics[width=\linewidth]{figures/result_ideal_sg}
+%  \caption{Example of a linear MRI mode with non-negligible density and
+%    potential perturbation, in a strongly magnetized massive disk
+%    ($\beta=10$ and $Q=0.2$) in the limit of ideal MHD 
+%    ($\Lambda_0=10^2$).   
+%    \label{mri_massive}}
+%\end{figure}
 
@@
 
-%An example is shown in
-%Fig. \ref{mri_massive_resis}, which  plots the energies assocaited with the 
-%linear MRI mode with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
-%disk. The gravitational potential energy significantly exceeds the
-%magnetic energy near the midplane ($z\lesssim0.3H$).
+An example is shown in
+Fig. \ref{mri_massive_resis}, which  plots the energies assocaited with the 
+linear MRI mode with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
+disk. The gravitational potential energy exceeds the
+magnetic energy near the midplane ($z\lesssim0.2H$). The growth rate $\gamma=0.25\Omega$
+is not much smaller than that of the most unstable mode
+($\gamma=0.36\Omega$) so significant density perturbations are
+expected to grow on dynamical timescales.
 
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figures/result_resis_sg}
-%  \caption{Example of a linear resistive MRI mode with significant
-%    gravitational potential energy perturbation.  The vertical line
-%    inidicates $\Lambda=1$.  
-%    \label{mri_massive_resis}}
-%\end{figure}
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/result_resis_sg}
+  \caption{Example of a linear resistive MRI mode with significant
+    gravitational potential energy perturbation.  The vertical line
+    inidicates $\Lambda=1$.  
+    \label{mri_massive_resis}}
+\end{figure}
 

### 2014-03-20 paper/results_azi.tex
tex update

@@
 \section{Effect of an azimuthal field}
-We now extend the previous investigation to disks with an azimuthal
-field, so that $B_y\neq 0$.
+In this section we add an azimuthal field so that $B_y\neq 0$,  
+parameterized by $\epsilon \equiv B_y/B_z$. Note that we continute to
+use $B_z$ for normalizations and $\beta$ is associated with the
+vertical Alfven speed. We also extend the previous calculations to the
+full disk $z\in[-Z_s,Z_s]$, which allows us to compare the effect of
+self-gravity on MRI modes with different symmetries across the
+midplane.
 
+\subsection{Fiducial case}
+Here we consider an isothermal disk with $Q=0.2$
+($Q_\mathrm{2D}=0.72$) and $\beta=10$ in the limit of ideal MHD
+($\Lambda_0=100$). We do not expect gravitational instability, since
+Fig. \ref{compare_growth3} shows that even for $Q=0.18$, GI is
+surpressed for $\beta \lesssim 15$. We consider azimuthal field
+strengths with $\epsilon=[0,3]$.
+
+Fig. shows the MRI growth rates for the above system as a function of
+$k_x$. The top panel are modes with even midplane symmetry
+$W^\prime(0)\sim 0$ for $k_x\to0$, and the bottom panel are modes with odd midplane 
+symmetry $W(0)\sim 0$ for $k_x\to0$. The latter set of modes were
+excluded in the previous sections by midplane boundary
+conditions. The plot also include growth rates computed in the Cowling
+approximation. As expected, none of the modes are energetically
+dominated by self-gravity.
+
+%general stabilization of MRI
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 2cm 0cm 0cm]{figures/compare_growth3_tilted_even.ps} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0.52cm]{figures/compare_growth3_tilted_odd.ps}
+  \caption{Growth rates in massive isothermal disks with $Q=~0.2$ and
+    $\beta=10$ for a range of azimuthal field strengths $B_y/B_z$. The
+    dots are solutions computed from the full problem, with the
+    colorbar measuring the gravitational potential perturbation, while
+    the solid curves are computed from the Cowling approximation. The
+    topic panel correspond to even modes and the
+    bottom panel correspond to odd modes, which have
+    $W^\prime(0)\sim0$ and $W(0)\sim0$, respectively, for $k_x\to0$.
+    \label{compare_growth3_tilted}}
+\end{figure}
+
+%We consider $\epsilon\equiv B_y/B_z \in[0,3]$. The
+
+Consider first the even modes in the top panel of
+Fig. \ref{compare_growth3_tilted}. Self-gravity increasingly stabilizes
+modes with small $k_x$ as $B_y$ is increased, but the effect is
+negligible. SG destabilizes modes with  $k_xH\gtrsim
+O(1)$. Consequently the cut-off wavenumber is larger when SG is
+included. The destabilization effect weakens with $B_y$ because the
+total magnetic pressure increases. %which generally opposes GI 
+The Cowling approximation becomes 
+increasingly good for even modes as $B_y$ is increased.
+
+%cowling is good with increasing azi field for even modes. 
 %cannot consider superthermal fields
@@
 
-%\subsection{}
+The odd modes in the bottom panel of Fig. \ref{compare_growth3_tilted}
+display opposite behavior. Self-gravity has no effect when
+the field is purely vertical ($B_y=0$). SG stabilizes all wavelengths,
+and is most effective with increasing $B_y$ and $k_xH =  O(1)$. 
+%We therefore expect odd modes to become less important as the azimuthal
+%field strength is increased.
+
+%compare eigenfunctions
+
+With $B_y\neq 0$, the MRI is compressible even in the Cowling
+approximation \citep{pessah05}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.5cm 0cm 0cm]{figures/result_tilted_kx0d1.ps} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0.cm]{figures/result_tilted_kx1d2.ps}
+  \caption{caption
+    \label{result_tilted}}
+\end{figure}
+
+\subsection{Resistive disks}
+
+\subsection{Disks with GI}

### 2014-03-27 paper/paper.tex
tex update
tex updates

@@
 \newcommand{\dphi}{\dd \Phi}
+\newcommand{\actaa}{Acta Astron.}
 \defcitealias{lin12}{L12}
@@
 
-\author{Min-Kai Lin (+Menou, Fromang, Youdin)}
+\author{Min-Kai Lin }%(+Menou, Fromang, Youdin)}
 \affil{Canadian Institute for Theoretical Astrophysics,
@@
 \begin{abstract}
+  Magneto-rotational instability (MRI) and gravitational instability
+  (GI) are the two principle routes to turbulent angular momentum
+  transport in accretion disks. Protoplanetary disks may develop
+  both. The goal of this paper is to reinvigorate interest in the study of
+  magnetized  massive protoplanetary disks, starting from the basic
+  issue of stability. The linear stability of a local patch of a
+  self-gravitating, magnetized three-dimensional (3D) disk to
+  axisymmetric perturbations is examined. %Addition effects such as
+  %resistivity and azimuthal fields are considered.
 
+%We examine the local linear stability of
+%  magnetized, self-gravitating and differentially rotating disks to
+%  axisymmetric perturbations. 
+%
+%not many studies model both
+%revisits 
 \end{abstract}
@@
 \section{Introduction}
-Understanding the evolution of protoplanetatry disks (PPDs) is key to planet
-formation theory. A central aspect of PPD evolution is the mechanism
-of angular momentum transport.
+Astrophysical disks host a wide range of fluid instabilities. Among
+them, the magneto-rotational instability \citep[MRI,][]{chandrasekhar61,
+  balbus91,balbus98} %balbus %chandrasekhar ref
+and gravitational instability \citep[GI, ][]{toomre64,goldreich65a,goldreich65b} 
+provide robust pathways to turbulent angular momentum transport that
+enables mass accretion \citep[][ 
+  and references therein]{balbus99,armitage11,turner14}.   
+%gammie, lin/pringle, paczynski 
+%relevance of MRI and GI to planet formation 
+They are also relevant to planet formation theory. For example,
+the strength of MRI turbulence directly affect planetesimal dynamics
+in protoplanetary disks \citep{yang12,gressel12}; while GI can
+potentially form giant planets directly through disk fragmentation
+\citep{boss97,boss98,gammie01,voro13,helled14}. 
+%gammie,
+%boss %some review articles (PP6)  
+%shakura73, pacz78, lin87,
+
+Accretion disks such as those surrounding black holes can develop both
+MRI and GI \citep{menou01,goodman03}.  
+Protoplanetary disks (PPDs) are also expected to be massive and  
+magnetized in its earliest evolutionary phase \citep{inutsuka10}.  
+The interplay between MRI and GI has been invoked to explain outbursts
+in circumstellar disks \citep{armitage01, zhu10a, zhu10b,
+  martin12b}, and predicts similar phenomenon in circumplanetary disks
+\citep{lubow12}. This results from the development of `dead zones' 
+--- magnetically inactive, laminar regions near the disk midplane --- with
+magnetized layers above and below 
+\citep{gammie96,martin12,landry13}. Mass accumulation in the dead zone can lead to
+GI and trigger MRI through heating. In these models, the condition required
+for MRI is realized through GI, but the MRI is unaffected by disk
+self-gravity.
+
+PPDs subject to both MRI and GI are often modeled through separate
+turbulent viscosities in a hydrodynamical framework 
+\citep{terquem08}. This implicitly assumes that the development of MRI 
+and GI can be assessed independently. 
+%That is, MRI is unaffected by
+%disk self-gravity; and GI is unaffected by weak magnetization. 
+Circumstellar disk models that explicitly combine the equations of magneto-hydrodynamics
+and self-gravity have been limited to a few early simulations 
+\citep{fromang04,fromang04b,fromang05}. It will be neccessary to %no
+%major difference % kim/ostriker/stone
+revisit and extend these pioneering calculations to fully explore the
+impact of MRI and GI on the structure and evolution of PPDs. In
+preparation of this, it is important to have a thorough understanding
+of the stability properties of such systems.  
+%non-ideal effects important
+
+%why is GI and MRI independent 
+%compressibility is not fundamental to MRI 
+%horizontal wavenumber (gammie/balbus) 
+%The effect of self-gravity on the linear MRI has not been considered, 
+%since fluid compressibility is not fundamental to the MRI.
+
+Since compressibility is not fundamental for the MRI, much of the
+early stability calculations assume incompressible perturbations
+\citep{goodman94,jin96}. However, recent works indicate
+compressibility may be important under certain conditions,
+such as strong fields \citep{kim00, pessah05,bonanno07}.   
+Previous MRI studies have also focused on modes with vanishing
+radial wavenumber, because they are the most unstable
+\citep{sano99,reyes01}. Self-gravity has minimal effect on such
+perturbations in a rotating disk. However, modes
+with radial lengthscales on the order of the disk scale height may be
+subject to self-gravity. It is of interest to generalize the MRI with
+non-zero radial waveumbers to massive disks.
+
+%Simiarly, studies of GI in accretion disks have mostly neglected
+%magnetization \citep{mamat10,kim12}. 
+%particularly in the context of
+
+%Studies of GI in accretion disks do not often include magnetization
+%\citep{mamat10,kim12}.
+
+The effect of a magnetic field on the GI of rotating disks has been
+considered by some authors.  
+% \cite{lynden-bell66} found that a toroidal
+%field can destabilize non-axisymmetric disturbances
+Recently, \cite{lizano10} generalized the classic Toomre stability
+criterion for razor-thin disks to include a vertical field. For
+circumstellar disks, the authors concluded that the field is
+stabilizing. This is consistent with previous analysis by
+\cite{nakamura83} for three-dimensional (3D) uniformly rotating
+disks. However, the GI of 3D differentially rotating disks have
+mostly neglected magnetic fields \citep{mamat10,kim12}, but such disks
+are subject to the MRI if magnetized. 
+%how is the dispersion relation for GI affected by the presence of MRI?
+%It is interesting to ask whether
+%or not the presence of MRI affects GI. 
+%
+%It is desirable to clarify
+%whether or not the presence of MRI modes affect the dispersion
+%relation for GI.
+
+%although MRI and GI are physically distinct 
+%explicitly show
+
+This paper marks the beginning of our study of magnetized,
+self-gravitating PPDs. We start from linear calculations. 
+Linear stability analyses have the advantage that a wide range of
+parameters can be studied with negligible computational cost. This
+allows one to indentify conditions, if any, under which MRI and GI
+cannot be consdered independent. It is also important to have such 
+calculations at hand to help benchmark and guide non-linear
+simulations.
+
+%This paper marks the beginning of our study of magnetized,
+%self-gravitating PPDs and is organized as follows.
+
+This paper is organized as follows. \S\ref{setup} lists the governing
+equations and describes the disk equilibra under consideration. 
+The linear problem is formulated in \S\ref{linear}. The impact
+of self-gravity on the MRI is discussed in
+\S\ref{result1}; gravitationally unstable disks are considered in
+\S\ref{result2} and equilibra with an azimuthal field is explored in
+\S\ref{result3}. We summarize results in \S\ref{summary}, including a
+discussion of caveats and future directions.
 
@@
 \input{results_azi} 
+\input{summary}
 

### 2014-03-27 paper/results_azi.tex
tex updates

--- 2014-03-17 paper/results_azi.tex
@@
-\section{Effect of an azimuthal field}
-We now extend the previous investigation to disks with an azimuthal
-field, so that $B_y\neq 0$.
+\section{Effect of an azimuthal field}\label{result3}
+In this section we add an azimuthal field so that $B_y\neq 0$,  
+parameterized by $\epsilon \equiv B_y/B_z$. However, we continute to
+use $B_z$ for normalizations and $\beta$ is associated with the
+vertical Alfven speed. We also extend the previous calculations to the
+full disk $z\in[-Z_s,Z_s]$, which allows us to compare the effect of
+self-gravity on MRI modes with different symmetries across the
+midplane.
 
-%cannot consider superthermal fields
-%compressibility important f
+\subsection{Ideal disks without GI} 
+Here we consider an isothermal disk with $Q=0.2$
+($Q_\mathrm{2D}=0.72$) and $\beta=10$ in the limit of ideal MHD
+($\Lambda_0=100$). Gravitational instability is not expected because 
+Fig. \ref{compare_growth3} shows that even for $Q=0.18$, GI is 
+surpressed for $\beta \lesssim 15$. We consider azimuthal field
+strengths with $\epsilon=[0,3]$.
 
-%\subsection{}
+Fig. shows the MRI growth rates for the above system as a function of
+$k_x$. The top panel are modes with even midplane symmetry: 
+$W^\prime(0)\sim 0$ for $k_x\to0$, and the bottom panel are modes with odd midplane 
+symmetry: $W(0)\sim 0$ for $k_x\to0$. The latter set of modes were 
+excluded in the previous sections by midplane boundary 
+conditions. The plot also include growth rates computed in the Cowling
+approximation. As expected, none of the modes are energetically
+dominated by self-gravity, $\avg{E_g}<\avg{E_m}$.
+
+%general stabilization of MRI
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 2cm 0cm 0cm]{figures/compare_growth3_tilted_even.ps} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0.52cm]{figures/compare_growth3_tilted_odd.ps}
+  \caption{Growth rates in massive isothermal disks with $Q=~0.2$ and
+    $\beta=10$ for a range of azimuthal field strengths $B_y/B_z$. The
+    dots are solutions computed from the full problem, with the
+    colorbar measuring the gravitational potential perturbation, while
+    the solid curves are computed from the Cowling approximation. The
+    top panel correspond to even modes and the
+    bottom panel correspond to odd modes, which have
+    $W^\prime(0)\sim0$ and $W(0)\sim0$, respectively, for $k_x\to0$.
+    \label{compare_growth3_tilted}}
+\end{figure}
+
+%We consider $\epsilon\equiv B_y/B_z \in[0,3]$. The
+
+Consider first the even modes in the top panel of
+Fig. \ref{compare_growth3_tilted}. Self-gravity slightly stabilizes
+modes with small $k_x$ as $B_y$ is increased. SG destabilizes modes
+with  $k_xH\gtrsim O(1)$, and is most effective for purely
+vertical fields: for $\epsilon=0,\, k_xH\simeq 1.4$, SG increases the
+growth rate by $\sim 30\%$. Consequently the cut-off wavenumber is
+larger when SG is included. This destabilization effect weakens with
+$B_y$ because the total magnetic pressure increases, which generally
+opposes GI, although the modes are fundamentally associated with MRI.
+
+The odd modes in the bottom panel of Fig. \ref{compare_growth3_tilted} 
+display opposite behavior. Self-gravity has negligible effect when 
+the field is purely vertical ($B_y=0$). SG stabilizes all wavelengths,
+and is most effective with increasing $B_y$ and $k_xH =  O(1)$.
+
+%With $B_y\neq 0$, the MRI is compressible even in the Cowling
+%approximation \citep{pessah05}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.5cm 0cm
+    0cm]{figures/result_tilted_cowling.ps}  
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+    0.cm]{figures/result_tilted_fullsg.ps} 
+  \caption{caption
+    \label{result_tilted}}
+\end{figure}
+
+\subsection{Resistive disks with GI}
+
+%\subsection{GI}

### 2014-03-27 paper/results_gi.tex
tex update
tex updates

@@
-\section{Gravitationally unstable disks} 
+\section{Gravitationally unstable disks} \label{result2}
 Gravitational instability (GI) becomes possible in a sufficiently
@@
 
-\subsection{Ideal disks}
+\subsection{Co-existence of MRI and GI}
 We first consider ideal MHD by setting $\Lambda_0=100$ and $A=1$. 
@@
 where the GI branch approaches an MRI branch at $\beta\simeq 25,\,
-\gamma\simeq 0.2\Omega$. 
-%Although the MRI is still most unstable, it is
-%self-gravitating for $\beta\lesssim 50$ (uppermost curve). 
-Some modes may have been missed in our
-numerical search of eigenfrequencies. Thus, `breaks' along the
-GI branch for $Q=0.18$ and $Q=0.12$ does not prove that the GI and MRI branches do not
-intersect. However, the continous variation of $\tau$ with respect to
-$\beta$ suggests that unstable modes can transition smoothly from
-magnetically-dominated to gravitationally-dominated at low $\beta$.
+\gamma\simeq 0.2\Omega$. In fact following the red curve to smaller
+$\beta$ indicates GI transitions to MRI. The
+`gaps' in the GI and MRI branches for $Q=0.18$ and $Q=0.12$ may be due
+to avoided crossing of modes, as seen in stars
+\citep[e.g.][]{aizenman77} and accretion tori 
+\citep[e.g.][]{christo93}. However, we cannot exclude the possibility 
+that some modes may have been missed in the numerical search of
+eigenfrequencies.
 
-For $Q=0.12$ the GI growth rate exceeds all MRI modes. Notice the most
-unstable MRI branch becomes self-gravitating for $20 \lesssim
-\beta\lesssim 40$, but magnetic perturbations remain negligible
-for GI. 
-%we generally find MRI modes aquire significant potential
-%perturbations but GI modes always have negligible magnetic
-%perturbation. 
-Fig. \ref{compare_growth3_Q01d2} show growth rates
-as a function of $k_xH$ for the $Q=0.12,\,\beta=20$ disk. All
-perturbations with $k_xH \lesssim 3.5$ are unstable. MRI operates on the largest
-length scale, followed by GI at $0.7 \lesssim k_xH 1.7$. The 
-transition between MRI to GI is smoother with increasing $k_xH$ as the
-MRI becomes compressible.
+Thus, our results do not rigorously prove that the GI and MRI branches
+do not intersect.  Nevertheless, the continous variation of $\tau$ with
+respect to $\beta$ strongly suggest that unstable modes can transition
+smoothly from magnetically-dominated to gravitationally-dominated at
+low $\beta$.
 
-%Modes with $k_xH
-% \gtrsim 1.7$ are neither pure MRI or GI. 
-%GI wave lengths fit inside channel solutions
+\subsubsection{Case study} 
+In reality, perturbations with a range of $k_x$ will be present for a
+given set of disk parameters. Fig. \ref{compare_growth3_Q01d2} show
+growth rates as a function of $k_x$ for an isothermal disk with
+$Q=0.12,\,\beta=20$. MRI and GI are active with comparable growth
+rates. All perturbations with $k_xH \lesssim 3.5$ grow dynamically
+($\gamma\gtrsim 0.1\Omega$, or $\lesssim$ 1.6 orbits).
+
+We also plot in Fig. \ref{compare_growth3_Q01d2} growth rates obtained 
+from the Cowling approximation, which isolates MRI; and that from a
+high-resistivity run, which isolates GI by allowing the field lines to
+slip through the fluid. We refer to these as  
+pure MRI and pure GI, respectively. 
+For $k_xH\lesssim 0.2$, growth rates are equal to
+those on the pure MRI and pure GI branches. That is, MRI and GI
+operate independently until their growth rates become equal as a
+function of $k_x$.
+
+The dispersion relation $\gamma(k_x)$ increasingly deviates from the
+pure GI/MRI curves with $k_x$, implying increasing interaction between
+magnetic and density perturbations. 
+Comparing pure GI (dashed line)  and the gravitationally-dominated
+portions of $\gamma(k_x)$ indicate inclusion of magnetic field
+stabilizes high-$k_x$ pure GI. (Note also the slight decrease in the most unstable
+$k_x$.) Since axisymmetric GI is unaffected by shear, this
+stabilization is due to magnetic pressure and is consistent with
+\citep{lizano10}.
+
+%3D geometry doesn't affect 
+%axisymmetric GI doesn't care about shear 
+%
+%two legs of GI are not symmetric 
+%MRI become self-gravitating
+%GI with increasing magnetization
+
+Comparing pure MRI (solid line) and the magnetically-dominated
+portions of $\gamma(k_x)$ show that self-gravity can enhance or reduce
+MRI growth rates depending on $k_x$, but the effect is small.
+
+%magnetic pressure, small scale more effective 
+%, reflected in the full range of $\tau$ displayed by the
+%modes. 
+%mode classification becomes difficult. 
+%the point is: dispersion relation cannot be estimated by computing
+%the MRI and GI separately
 
@@
     self-gravity. The solid line corresponds to MRI modes in the
-    Cowling approximation. 
+    Cowling approximation. The dashed line corresponds to pure GI
+    modes, obtained by including a high resisitivity in the full problem. 
     \label{compare_growth3_Q01d2}}
@@
 
-\subsection{Layered disks}
-Can a disk with a dense, resistive midplane and a highly conducting
-atmosphere  be subject to both MRI and GI but at different heights?
+%\subsection{Layered disks}
+%Can a disk with a dense, resistive midplane and a highly conducting
+%atmosphere  be subject to both MRI and GI but at different heights?
 

### 2014-03-27 paper/setup.tex
tex updates

@@
-\section{Local disk model}
+\section{Local disk model}\label{setup}
 We study the local stability of an inviscid, self-gravitating and
@@
 $\nabla\cdot~\bm{B}=0$; $\Pi \equiv P +
-|\bm{B}|^2/2\mu_0$ is the total pressure ($\mu_0$ is the vacuum
-permeability). We choose a barotropic equation of state, specified
+|\bm{B}|^2/2\mu_0$ is the total pressure, and $\mu_0$ is the vacuum
+permeability. We choose a barotropic equation of state, specified
 below, so that the gas pressure is given by $P=P(\rho)$. The
@@
 \end{inparaenum}
-where $\rho_0\equiv\rho(0)$. The sound speed $c_s\equiv\sqrt{dP/d\rho}$
+where $\rho_0\equiv\rho(0)$ is the mid-plane density.
+The sound speed $c_s\equiv\sqrt{dP/d\rho}$ 
 so that $c_{s0}$ is the global sound speed in the isothermal disk, and
@@
 isothermal disk has no surface, we define $H$ such that
-$\rho(H)=\epsilon\rho(0)$ with $\epsilon=10^{-2}$.
+$\rho(H)=\epsilon\rho(0)$ with $\epsilon=10^{-2}$. A non-dimensional
+measure of the disk thickness is given by 
+\begin{align}
+  f^{-1} = \frac{H\Omega}{\csmid}, 
+\end{align}
+and $f$ will appear in subsequent discussions.
 
@@
 is weakly dependent on the strength of self-gravity provided the
-vertical co-ordinate is appropriately scaled as well.
+$z$-axis is appropriately scaled.
 
@@
     horizontal axis also depends on the strength of self-gravity,
-    i.e. $H=H(Q)$ and is a decreasing function of $Q$. 
+    i.e. $H=H(Q)$ and is an increasing function of $Q$. 
 %Note that the vertical co-ordinate is
@@
 \subsection{Resistivity profile}\label{resis_profile}
-In this work we consider either constant resistivity or a
+We consider constant resistivity or 
 resistivity prescription such that $\eta(z)$ increases away from the
@@
 The disk may be considered ideal where $\Lambda
-\gtrsim 1$. Since $\hat{\rho}(z/H)$ is weakly dependent on  self-gravity, the
-ideal region only changes slightly as $Q$ is decreased.
+\gtrsim 1$. 
+%Since $\hat{\rho}(z/H)$ is weakly dependent on  self-gravity, the
+%ideal region only changes slightly as $Q$ is decreased.
 
@@
 \begin{align}
-  \Lambda_0 \equiv\Lambda(0) =  \frac{v_{A0}^2}{\eta_0\Omega} = \frac{f^2 R_m}{\beta}, 
+  \Lambda_0 \equiv\Lambda(0) =  \frac{v_{A0}^2}{\eta_0\Omega}. %= \frac{f^2 R_m}{\beta}, 
 \end{align}
-where the third equality defines the magnetic Reynolds number
-$R_m=H^2\Omega/\eta_0$ and $f\equiv \csmid/H\Omega$ is a numerical
-factor of order unity.
+%where the third equality defines the magnetic Reynolds number
+%$R_m=H^2\Omega/\eta_0$ and
+%where $f\equiv \csmid/H\Omega$ is a numerical
+%factor of order unity. %measures thickness
 

### 2014-03-31 paper/linear.tex
tex update
tex updates

@@
 the growth rate. We take $k_x>0$ without loss of generality. 
-Hereafter, we surpress the exponential factor, as
-well as the real part notation.
+Hereafter, we surpress the exponential factor and the real part notation.
 
@@
 where the effective enthalpy perturbation
-$\w = W + \dphi$. The linearized induction equation is
+$\w = W + \dphi$. The components of the linearized induction equation
+are 
 \begin{align}
@@
 \item Purely vertical field with constant or variable
-  resistivity (see below), so that $\epsilon = 0$ and $\eta=\eta(z)$. 
+  resistivity, so that $\epsilon = 0$ and $\eta=\eta(z)$. 
 \item Tilted field with uniform resistivity
@@
 \end{align}
-where differential operators $L_{1j},\,L_{2j}$ and $L_{3j}$
+where the differential operators $L_{1j},\,L_{2j}$ and $L_{3j}$
 ($j=1,2,3,4$) can be read off Appendix \ref{reduction} and $L_{4j}$
 ($j=3,4$) from the linearized Poisson equation above. We remark that
-the more general problem of tilted field and variable resistivity
+the case of a tilted field and variable resistivity
 %($\epsilon\neq0$ and $\eta=\eta(z)$
@@
 
-\subsubsection{Case 1: vertical field}
+\subsubsection{Case 1: vertical field}%because SG can only destabilize
+                                %symmetric density perts
 We impose $\bm{U}$ to be an even function of $z$, so that
@@
 \begin{align}
-  \dphi^\prime(\zmax) + k_x\dphi(\zmax) = 0 \label{pot_bc1}
+  \dphi^\prime(\zmax) + k_x\dphi(\zmax) = 0 \label{pot_bc1},
 \end{align}
-\citep{goldreich65a}.
+and is adapted from \cite{goldreich65a}. The general form of this
+potential boundary condition is used for case 2 below.
 
@@
 and  $T_l$ is a Chebyshev polynomial of the first kind of order $l$
-\citep{stegun65}. Note that for case 1 the midplane symmetry condition
+\citep{stegun65}. Note that for case 1 the mid-plane symmetry condition
 is taken care of by the choice of basis functions. 
@@
 $l_\mathrm{max}$ is the highest polynomial order. Our standard
-resolution is $N_z=256$.
+resolution is $N_z=256$ ($N_z=257$) for case 1 (case 2).
 
@@
 \begin{align}\label{matrix_eqn}
-\bm{M}\bm{w} = 0, 
+\bm{M}\bm{w} = \mathbf{0}, 
 \end{align}
 where $\bm{M}$ is a $4N_z\times 4 N_z$ matrix representing the $L_{ij}$ 
-and upper disk boundary conditions, 
+plus boundary conditions, 
 and $\bm{w}$ is a vector storing the pseudo-spectral coefficients. 
@@
   &E_g = 
-  \frac{\hat{\rho}}{\hat{c}_s^2}\left|\real\left(\hat{W}\dd\hat{\Phi}^*\right)\right|,\\  
+  \frac{\hat{\rho}}{2\hat{c}_s^2}\left|\real\left(\hat{W}\dd\hat{\Phi}^*\right)\right|,\\  
   &E_k = \frac{1}{2}\hat{\rho}|\delta\hat{\bm{v}}|^2,\\
@@
 as the perturbed magnetic, gravitational, kinetic and thermal
-energies, respectively, which are functions of $z$. The total energy
+energies, respectively, which are functions of $z$. Although we do not
+solve an energy equation, we nevertheless define $E_t$ as a measure of
+density perturbations \citep{kojima89}. The total energy
 is $E=E_m+E_g+E_k+E_t$. We use $\avg{\cdot}$ to denote an average over
 $z$. 
+%The colorbar quantifies the importance of
+%self-gravity with
+
+Since we will primarily be concerned with massive disks, we define
+\begin{align}
+ \tau \equiv \frac{\avg{E_g}}{\avg{E_g} + \avg{E_m}} 
+\end{align} 
+as a measure of the importance of self-gravity. Thus, modes with $\tau
+= 1$ are energetically dominated by self-gravity (GI) and modes with
+$\tau\ll1$ are dominated by magnetic perturbations (MRI).
+
 %and we then normalize $E_i$ by

### 2014-03-31 paper/paper.tex
tex updates

@@
 \newcommand{\avg}[1]{\langle{#1}\rangle}
+%\newcommand{\max}[1]{\mathrm{max}({#1})}
+%\newcommand{\min}[1]{\mathrm{min}({#1})}
 \newcommand{\zhat}{\hat{z}}
@@
   transport in accretion disks. Protoplanetary disks may develop 
-  both. The goal of this paper is to reinvigorate interest in the study of
+  both. This paper aims to reinvigorate interest in the study of
   magnetized  massive protoplanetary disks, starting from the basic
-  issue of stability. The linear stability of a local patch of a
-  self-gravitating, magnetized three-dimensional (3D) disk to
-  axisymmetric perturbations is examined. %Addition effects such as
+  issue of stability. The local linear stability of a 
+  self-gravitating, uniformly magnetized, differentially rotating, 
+  three-dimensional disk subject to axisymmetric perturbations is
+  calculated numerically. The formulation includes 
+  resistivity. It is found that the reduction in the disk scale height
+  by self-gravity (SG) can decrease MRI growth rates. SG causes the MRI to
+  become global in the vertical direction, and MRI modes
+  with small radial lengthscales are stabilized. In massive
+  disks with layered resistivity, the MRI is not well-localized
+  to regions where the Elsasser number exceeds unity. For MRI modes
+  with radial lengthscales on the order of the disk scale-height, SG
+  can enhance density perturbations, an effect that becomes 
+  significant in the presence of a strong toroidal field. In
+  gravitationally unstable disks where GI and MRI growth rates are
+  comparable, the character of unstable modes can transition smoothly
+  between MRI and GI. Implications for non-linear simulations are
+  discussed briefly.
+
+%layered dis   
+  %Addition effects such as
   %resistivity and azimuthal fields are considered. 
-
 %We examine the local linear stability of

### 2014-03-31 paper/results.tex
tex updates

@@
-\section{MRI in self-gravitating disks}
-In this section we focus on the MRI. We use the vertical field setup
-of case 1. We first consider MRI modes with negligble densit/potential
+\section{MRI in self-gravitating disks}\label{result1}
+In this section we focus on the MRI and use the vertical field setup
+of case 1. We first consider MRI modes with negligble density/potential
 perturbations to see the effect of self-gravity on the MRI through the
-background stratification, then go on to consider MRI modes with
-significant density/potential perturbations in massive disks.
+background stratification, then go on to examine MRI modes with
+density/potential perturbations in massive disks.
 
@@
   background equilibrium}
-
-Here we use polytropic disks, which have a well-defined
-thickness. The upper disk boundary is set to 
+Here we use polytropic disks, so that the disk thickness is
+well-defined. The upper disk boundary is set to 
 $Z_s=0.99H$. We fix $\beta=100$ and $k_xH=0.1$ unless otherwise
@@
   resistivity}   
-We first calculate the MRI as a function of $Q$ with uniform
-resistivity ($A=1$) for mid-plane Elsasser numbers $\Lambda_0\in[0.3,10]$.
-
-Fig. \ref{compare_growth_poly_uniresis} plots the MRI growth rates as
-a function of $Q$ and $\Lambda_0$. As expected growth rates generally
-decrease with increasing $\Lambda_0$. In the limit of ideal MHD ($\Lambda_0>1$), 
-there is negligible dependence on $Q$. In the resistive limit
-($\Lambda_0<1$), however, growth rates decrease noticeably for $Q<1$. This shows
+%We calculate the MRI as a function of $Q$ with uniform
+%resistivity ($A=1$) for mid-plane Elsasser numbers $\Lambda_0\in[0.3,10]$.    
+Fig. \ref{compare_growth_poly_uniresis} plots MRI growth rates as a
+function of $Q$ and $\Lambda_0$. The resistivity is uniform ($A=1$). 
+%As expected growth rates generally decrease with decreasing 
+%$\Lambda_0$ \citep{jin96}. 
+In the limit of ideal MHD ($\Lambda_0>1$), there is negligible
+dependence on $Q$. In the resistive limit 
+($\Lambda_0<1$), however, growth rates decrease noticeably for $Q<0.5$
+($Q_\mathrm{2D}\lesssim 1.5$). This shows 
 show that disk self-gravity affects the MRI through the background
 equilibrium, even though density and potential perturbations are  
-negligible (i.e. the linear response is
-non-self-gravitating).
+unimportant (i.e. the linear response is non-self-gravitating).
 
@@
   \includegraphics[width=\linewidth]{figures/compare_growth_poly_uniresis2}
-  \caption{MRI growth rates as a function of $Q$ for mid-planet
+  \caption{MRI growth rates as a function of $Q$ for mid-planee 
     Elsasser numbers $\Lambda_0=0.3$ (solid), $0.5$ (dotted), $1.0$
@@
 
-\cite{sano99} found that for instability to occur the linear mode
+\cite{sano99} found that for MRI to operate, its 
 wavelength $\lambda$ should fit inside the disk. That is,   
@@
 \end{align}
-in the limit of ideal MHD, and 
+for ideal MHD, and 
 \begin{align}\label{lambda_resis}
@@
 MRI through the factor $f$, which increases with decreasing $Q$ (see
-Fig. \ref{plot_fq} in Appendix \ref{appen1}). This suggests that
-sufficiently strong self-gravity can stabilize the MRI by making
-$ 2H<\lambda $.
+Fig. \ref{plot_fq} in Appendix \ref{appen1}).   
+This implies that sufficiently strong self-gravity can stabilize the
+MRI by making $ 2H<\lambda $.  %these perts are not gravitationally
+                               %unstable
 
 In the ideal limit, we find $\lambda < 2H$ throughout most of the disk
-for the values of $Q$ considered, so that self-gravity does not affect
-growth rates significantly. However, the wavelength of the instability
-increases. This is shown in Fig. \ref{compare_result_lambda10} which
-plots the magnetic energies for $\Lambda_0=10$ and a range of Toomre $Q$
-values. The number of vertical nodes decrease with $Q$, 
-i.e. the disk accommodates fewer wavelengths because increasing
-vertical self-gravity makes it thinner.
+for the values of $Q$ considered, so self-gravity does not affect
+growth rates significantly. However, the wavelength of the
+instability, in units of $H$, increases. This is shown in
+Fig. \ref{compare_result_lambda10} which plots the magnetic energies
+for $\Lambda_0=10$ and a range of $Q$ values. The number of vertical
+nodes decrease with $Q$, i.e. the disk accommodates fewer wavelengths
+because increasing vertical self-gravity makes it thinner.
 
@@
 %increased.
+
+%{\bf DISCUSS STRESS} Although for ideal MHD self-gravity makes no
+%difference to the MRI growth rate, we find the associated stress does
+%change with $Q$.
 
@@
 Eq. \ref{sano_crit} is only an approximation.) Although the function
-$f$ does not change significantly for the range of $Q$ considered, the
-dependence of $\lambda_\mathrm{resis}$ on $f$ (and therefore $Q$) is
-amplified by the denominator $\Lambda_0<1$ in the resistive case. For
-resistive cases we find no nodes in the magnetic energy $E_1$ except
-at end points, i.e. only the longest wavelength survives.
+$f(Q)$ does not change significantly for the range of $Q$ considered, the
+dependence of $\lambda_\mathrm{resis}$ on $f(Q)$ is amplified by the
+denominator $\Lambda_0<1$ in the resistive case. For resistive cases
+we find no nodes in the magnetic energy $E_m$ except at end points,
+i.e. only the longest wavelength survives.
 
@@
   \includegraphics[width=\linewidth]{figures/lambda_poly_uniresis}
-  \caption{Approximate wavelengths of the most unstable resistive MRI modes as given by
+  \caption{Approximate wavelengths of the most unstable MRI modes as given by
     Eq. \ref{sano_crit}---\ref{lambda_resis}, normalized by the 
-    disk thickness, as a function of height for a range of Toomre $Q$
-    values.  
+    disk thickness, as a function of height. MRI is expected to
+    operate if $\lambda/2H\lesssim 1$. 
     \label{lambda_poly_resis}}
@@
 function of $k_x$ for a range of $Q$ values. Increasing self-gravity
-decreases the cut-off horizontal wavenumber for the MRI, implying that
-small-scale MRI does not operate in massive disks.
+decreases the cut-off horizontal wavenumber for the MRI. We checked
+that these modes have negligible density perturbations. Then we can 
+understand this result by invoking the instability criteria for  
+incompressible MRI in an unstratified Keplerian disk,
+\begin{align}
+  v_A^2(k_z^2 + k_x^2) < 3\Omega^2,
+\end{align}
+where $k_z$ is a vertical wavenumber \citep{kim00}. Setting $k_z^2\sim
+\Omega^2/v_A^2$ and non-dimensionalizing, we find
+\begin{align} 
+  k_xH \lesssim \frac{\sqrt{\beta}}{f},
+\end{align}
+where order-unity factors have been dropped. Despite a simplistic
+approach, this demonstrates that with increasing self-gravity
+(increasing $f$), we expect that MRI modes with small radial
+lengthscales to be suppressed.
+
+%, implying that
+%small-scale MRI does not operate in massive disks.
 
@@
 Our goal here is to examine whether or not self-gravity can amplify
-the density perturbations caused by the MRI. We choose a massive 
-isothermal disk with $Q=0.2$ (corresponding to $Q_\mathrm{2D}=0.72$),
-which is still expected to be marginally 
+the density perturbations caused by the MRI. We compute unstable modes
+in a massive isothermal disk with $Q=0.2$ (corresponding to
+$Q_\mathrm{2D}=0.72$), which is still expected to be marginally  
 stable to gravitational instability \citep[][who find
@@
 
-\subsubsection{Ideal case}
-We first consider the limit of ideal MHD by adopting a uniform
+\subsubsection{Ideal disks}
+We first consider ideal MHD by adopting a uniform 
 resistivity with $\Lambda_0=100$. %Non-zero density perturbations
 %require $k_x\neq0$ so we consider linear modes as a function of $k_x$.   
-Fig. \ref{gravity_energy} shows the mean gravitational potential
-energy perturbation, as a function of $k_x$ for several values of
-$\beta$. The potential perturbation is negligible for all cases when 
-$k_xH\lesssim 0.5$, since the MRI becomes incompressible as $k_x\to
-0$.
+Fig. \ref{gravity_energy} plots MRI growth rates as a function of
+$k_x$ for several values of $\beta$. The curves are color-coded
+according to $\tau$. The potential perturbation is
+negligible for all cases when $k_xH\lesssim 0.5$, since the MRI
+becomes incompressible as $k_x\to 0$.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_energy_ideal}
-  \caption{Average gravitational potential energy perturbation
-    associated with the MRI, compared to the average magnetic energy
-    perturbation, in a self-gravitating disk with $Q=0.2$ in 
-    the limit of ideal MHD ($\Lambda_0=10^2$), for a range of field
-    strengths $\beta$.  
+  \caption{Growth rates of MRI modes in an isothermal self-gravitating
+    disk with $Q=0.2$ ($Q_\mathrm{2D}=0.72$) in the limit of ideal MHD
+    ($\Lambda_0=10^2$), for a range of field strengths $\beta$. The
+    colorbar measures the importance of self-gravity by $\tau$.  
+    %{\bf ADD BETA=8}
     \label{gravity_energy}}
@@
 negligible and the standard MRI operates. However, as $\beta$ is
-lowered we find non-negligible potential perturbation for
-$k_xH=O(1)$. This suggets that in a strongly magnetized disk that
-still permits the MRI, the associated density perturbation can be
-important when the disk is self-gravitating.
+lowered and the MRI growth rate reduced, we find non-negligible
+potential perturbation for $k_xH=O(1)$. This suggets that in a
+strongly magnetized disk that still permits the MRI, the associated
+density perturbation can be important when the disk is
+self-gravitating. %but formall stable
+
+%critical beta, below which MRI is surpressed, increases with the
+%strength of self-gravity.
 
@@
 
-\subsubsection{Resistive case}
+\subsubsection{Resistive disks}
 We repeat the above calculation for resistive disks, but fix
-$\beta=100$. Fig. \ref{gravity_energy_resis} shows the 
-perturbed gravitational energy increases with   
-resistivity. Interestingly, in the highly resistive case
-the MRI can be energeically dominated by self-gravity, i.e. the case
-with $\Lambda_0=0.1$ and $k_xH\simeq1.3$. We also calculated this mode
-in the \emph{Cowling  
+$\beta=100$ and vary the mid-plane Elasser number $\Lambda_0$. Growth
+rates are shown in Fig. \ref{gravity_energy_resis}. 
+Interestingly, in the highly resistive case
+the magnetic and gravitational energies can be comparable. For 
+$\Lambda_0=0.1$ and $k_xH\simeq1.3$, $\tau\sim 0.3$ which corresponds
+to $\avg{E_g}\sim 0.5\avg{E_m}$.  
+Fig. \ref{mri_massive_cowling}, we compare the magnetic
+energy of this mode to that computed in in the \emph{Cowling  
   approximation}, where the Poisson equation is ignored in the 
@@
 (formally letting $Q\to\infty$ in Eq. \ref{lin_poisson}).  
-Fig. \ref{mri_massive_cowling} compares the magnetic energy in the
-Cowling approximation to that of 
-the full problem in . The growth rate increases when self-gravity is
-included in the linear response, as expected, but only
-slightly. $\gamma$ and $E_m$ are similar in both cases, indicating
-that the instability in the self-gravitating calculation is 
-still the MRI.
+The growth rate increases slightly when self-gravity is included in the linear
+response, since self-gravity is usually destabilizing. However, the
+similarity between $\gamma$ and $E_m$ indicates that the instability
+in the self-gravitating calculation is fundamentally the MRI.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_energy_resis}
-  \caption{Average gravitational potential energy perturbation
-    associated with the MRI, compared to the average magnetic energy
-    perturbation, in a self-gravitating disk with $Q=0.2$,
-    for a range of midplane Elsasser numbers $\Lambda_0$ at fixed
-    field strength $\beta=100$.    
+  \caption{
+    Growth rates of MRI modes in an isothermal self-gravitating
+    disk with $Q=0.2$ ($Q_\mathrm{2D}=0.72$) at fixed $\beta=100$, 
+    for a range of mid-plane Elasser numbers. The resistivity is
+    uniform. The colorbar measures the importance of self-gravity by $\tau$. 
     \label{gravity_energy_resis}}
@@
     Fig. \ref{gravity_energy_resis} (solid) compared with that computed
-    under the Cowling approximation (dotted). 
+    under the Cowling approximation (dotted). %{\bf REDO FIGURE}
     \label{mri_massive_cowling}}
@@
 
-An example is shown in
-Fig. \ref{mri_massive_resis}, which  plots the energies assocaited with the 
-linear MRI mode with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
-disk. The gravitational potential energy exceeds the
+Fig. \ref{mri_massive_resis} plots the energies assocaited with the 
+MRI mode discussed above. 
+%with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
+%disk. 
+The gravitational energy exceeds the 
 magnetic energy near the midplane ($z\lesssim0.2H$). The growth rate $\gamma=0.25\Omega$
 is not much smaller than that of the most unstable mode
-($\gamma=0.36\Omega$) so significant density perturbations are
-expected to grow on dynamical timescales.
+($\gamma=0.36\Omega$ for $k_xH=0.1$), so significant density
+perturbations are expected to grow on dynamical timescales for this
+system, even though GI should not develop.
 
@@
   \includegraphics[width=\linewidth]{figures/result_resis_sg}
-  \caption{Example of a linear resistive MRI mode with significant
-    gravitational potential energy perturbation.  The vertical line
+  \caption{Example of a resistive MRI mode with significant
+    gravitational potential perturbation.  The vertical line
     inidicates $\Lambda=1$.  
@@
 \subsubsection{Qualitative interpretation} 
-In order to make sense of the above results, we return to ideal MHD. 
-We consider regions close to the disk midplane ($z\sim 0$), where
+To make sense of the above results, we first return to ideal MHD and 
+consider regions close to the disk midplane ($z\sim 0$), where
 self-gravity is expected to be most important. For this discussion we
-will ignore stratification and set $d/dz\to \imgi k_z$ where $k_z$ is
-a vertical wavenumber. The governing equations are then
+will ignore stratification and set $d^2/dz^2\to -k_z^2$. The governing
+equations are then
 
@@
 \end{align}
-where $k^2 = k_z^2 + k_x^2$.
-
-We imagine an iterative procedure to solve the above equations,
-starting from the Cowling approximation. %defined to be the
+where $k^2 = k_z^2 + k_x^2$. We imagine an iterative procedure to
+solve the above equations, starting from the Cowling approximation. 
+%defined to be the
 %solution to Eq. \ref{simp1}---\ref{simp3} with potential perturbation
@@
 \end{align}
-Physically, we expect $k^2\geq0$, so that a positive (negative) density
-perturbation causes a negative (positive) potential perturbation. We
-then regard $\dphi$ to alter the Cowling solution, so
-that $\dvx^{(0)} \to \dvx^{(0)} + \dvx^{(1)}$ and similarly for $\dvy$ and
-$W$. Then 
+Physically, we expect $k^2\geq0$, so that a local positive (negative) density
+perturbation causes a negative (positive) local potential
+perturbation. We then insert $\dphi$ back into the momentum and
+continuity equations, and ask how does this potential perturbation
+modify the Cowling solution? Writing $\dvx^{(0)} \to \dvx^{(0)} +
+\dvx^{(1)}$ and similarly for $\dvy $ and $ W$, we find 
+%regard $\dphi$ to alter the Cowling solution, so
+%that 
+%Then 
 \begin{align}
@@
 $|\sigma^2/c_s^2k_z^2|\sim |\sigma^2/\Omega^2\beta|\ll1$ because
-$|\sigma|\lesssim \Omega$ and we are considering $\beta\gtrsim
-10$. Thus Eq. \ref{cond} is generally satisfied.
+$|\sigma|\lesssim \Omega$ and we are considering $\beta\gtrsim 10$. 
+Thus Eq. \ref{cond} is generally satisfied.
 
@@
 
- For weak fields, the MRI has small vertical wavelength,
- implying rapid density variations with height. The magnitude of the
- corresponding gravitational potential is then expected to be small
- because of internal cancellations, since the potential is essentially
- the integral of the density field.
-
 % not expect significant potential perturbations 
-
 %Conversely, for strong fields, only a single MRI wavelength occupies
@@
 
-The above arguments also apply to the resistive
+ The above arguments can be adapted to the resistive
  limit. Eq. \ref{simp3}---\ref{simp4} are unchanged, while resistive
@@
  density perturbations.
+
+ For weak fields, the MRI has small vertical wavelength,
+ implying rapid density variations with height. The magnitude of the
+ corresponding gravitational potential is then expected to be small
+ because of internal cancellations, since the potential is essentially
+ the integral of the density field.

### 2014-03-31 paper/results_azi.tex
tex updates

@@
 
-\subsection{Ideal disks without GI} 
+\subsection{Ideal disks with MRI} 
 Here we consider an isothermal disk with $Q=0.2$
@@
 surpressed for $\beta \lesssim 15$. We consider azimuthal field
-strengths with $\epsilon=[0,3]$.
+strengths with $\epsilon\in[0,3]$.
 
@@
 $k_x$. The top panel are modes with even midplane symmetry: 
-$W^\prime(0)\sim 0$ for $k_x\to0$, and the bottom panel are modes with odd midplane 
-symmetry: $W(0)\sim 0$ for $k_x\to0$. The latter set of modes were 
+$W^\prime(0)= 0$ for $\epsilon\to0$, and the bottom panel are modes with odd midplane 
+symmetry: $W(0) = 0$ for $\epsilon\to0$. The latter set of modes were 
 excluded in the previous sections by midplane boundary 
 conditions. The plot also include growth rates computed in the Cowling
-approximation. As expected, none of the modes are energetically
-dominated by self-gravity, $\avg{E_g}<\avg{E_m}$.
+approximation. As expected,  $\avg{E_g}<\avg{E_m}$, so none of the
+modes are energetically dominated by self-gravity.
+
+%no GI for odd modes
 
@@
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 2cm 0cm 0cm]{figures/compare_growth3_tilted_even.ps} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0.52cm]{figures/compare_growth3_tilted_odd.ps}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 2cm 0cm
+    0cm]{figures/compare_growth3_tilted_even.ps}  
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+    0.52cm]{figures/compare_growth3_tilted_odd.ps} 
   \caption{Growth rates in massive isothermal disks with $Q=~0.2$ and
@@
     bottom panel correspond to odd modes, which have
-    $W^\prime(0)\sim0$ and $W(0)\sim0$, respectively, for $k_x\to0$.
+    $W^\prime(0)=0$ and $W(0)=0$, respectively, for $\epsilon\to0$.
     \label{compare_growth3_tilted}}
@@
 Consider first the even modes in the top panel of 
-Fig. \ref{compare_growth3_tilted}. Self-gravity slightly stabilizes
-modes with small $k_x$ as $B_y$ is increased. SG destabilizes modes
-with  $k_xH\gtrsim O(1)$, and is most effective for purely
-vertical fields: for $\epsilon=0,\, k_xH\simeq 1.4$, SG increases the
-growth rate by $\sim 30\%$. Consequently the cut-off wavenumber is
-larger when SG is included. This destabilization effect weakens with
-$B_y$ because the total magnetic pressure increases, which generally
-opposes GI, although the modes are fundamentally associated with MRI.
+Fig. \ref{compare_growth3_tilted}. %Self-gravity slightly stabilizes
+%modes with small $k_x$ as $B_y$ is increased. 
+Self-gravity destabilizes modes with  $k_xH\gtrsim O(1)$. Consequently, the
+cut-off wavenumber is larger when SG is included. %get smaller scale
+                                %MRI 
+Destablization is most effective for purely vertical fields: with
+$\epsilon=0,\, 
+k_xH\simeq 1.4$, SG increases the growth rate by $\sim 30\%$. Although
+these modes are fundamentally magnetic, this is consistent with
+\cite{goldreich65a}, who find SG can only destabilize 
+symmetric density perturbations, which is the case for $B_y=0$ here.  
+With increasing $\epsilon$, we find the density perturbation $W$ deviates from an even 
+function, $|W(0)|$ decreases and $\max{(|W|)}$ moves off of the mid-plane.  
+%density pert at the midplane goes down
+%We expect this to reduce the destabilization
+%effect of SG,  
+Together with the increased total magnetic pressure with 
+$\epsilon$, destabilization by SG weakens. 
+Thus, the Cowling approximation becomes increasingly good with stronger
+$B_y$ for the even modes.
 
-The odd modes in the bottom panel of Fig. \ref{compare_growth3_tilted} 
-display opposite behavior. Self-gravity has negligible effect when 
-the field is purely vertical ($B_y=0$). SG stabilizes all wavelengths,
-and is most effective with increasing $B_y$ and $k_xH =  O(1)$.
+%This destabilization effect weakens with
+%$B_y$ because the total magnetic pressure increases, which generally
+%opposes GI, although the modes are fundamentally associated with MRI.
 
-%With $B_y\neq 0$, the MRI is compressible even in the Cowling
-%approximation \citep{pessah05}.
+The modes in the bottom panel of Fig. \ref{compare_growth3_tilted} 
+display opposite behavior. For $B_y=0$, $W$ is odd and 
+self-gravity has no effect. When $B_y>0$, we find $W$ deviates from an odd
+function and the mid-plane density perturbation $|W(0)|$ increases. 
+SG is stabilizing for these modes at all wavelengths, and is most
+effective at $k_xH =  O(1)$.  
+%destabilizing effect is absent  
+%for $\epsilon=0,\, k_xH\simeq 1.2$, SG decreases the growth rate by $\sim
+%13\%$.
+
+Fig. \ref{result_tilted} show eigenfunctions for $\epsilon=3$ and
+$k_xH=1.2$ with and without the Cowling approximation. These differ
+from the even modes considered above in that the horizontal magnetic
+perturbations maximize at $z=0$ (cf. Fig. \ref{mri_massive_resis}).  
+SG significantly enhances the mid-plane density perturbation, making the
+gravitational potential energy comparable to the magnetic energy, 
+which becomes more confined near the mid-plane.
 
@@
     0.cm]{figures/result_tilted_fullsg.ps} 
-  \caption{caption
+  \caption{Energy densities for a MRI mode in an isothermal ideal disk
+    with an azimuthal field $B_y = 3B_z$, computed in the Cowling
+    approximation (top) and with full self-gravity (bottom). These
+    modes correspond to those in the bottom panel of
+    Fig. \ref{compare_growth3_tilted}.    
     \label{result_tilted}}
@@
 
+%strong field, compressibility 
+To interpret the above result for odd modes, note 
+that compressibility affects the MRI in the presence of an 
+azimuthal field, even in a non-self-gravitating disk. If the perturbed 
+disk remains in vertical hydrostatic equilibrium, then 
+\begin{align} 
+  |W|\sim \frac{B_y}{\mu_0\rho}|\dby|,
+\end{align}
+to order of magnitude in a non-SG disk, so a strong toroidal field can
+cause a large density perturbation
+\citep{pessah05}. Fig. \ref{result_tilted} shows that this is
+amplified by self-gravity.
+
+We checked that for the modes in Fig. \ref{result_tilted} the vertical
+velocity is small, $|\dvz|/\left(|\dvx|^2+|\dvy|^2\right)^{1/2}
+\lesssim 0.2$, so the above argument applies. Compressibilty is 
+enhanced by both the azimuthal field and self-gravity, which tends
+to stabilize the MRI \citep{kim00}. Its effect is important here  
+because the azimuthal Alfven speed is sonic. The destabilization
+effect of SG is absent (or unimportant) because the density
+perturbation is not symmetric.
+
+%The results for even and odd modes are do not contradict each
+%other. Self-gravity tends to be destabilizing, but this requires the
+%density perturbation to be symmetric, as is the case for the even
+%modes. 
+%For the MRI however, compressibility has a stabilizing effect
+
+%scale separation for odd modes? because modes with kx=1 will be
+%quenched
+
+% for $\beta=10$ and
+%$\epsilon=3$. 
+%because magnetic energy is concentrated near midplane    
+%magnetic pressure, need thermal pressure 
+%This
+%can subsequently be amplitfied by its self-gravity. 
+%That is, compressibility is enhanced by both
+%For $\epsilon=3$ and $\beta=10$ the azimuthal Alven speed is
+%sonic. Compressibility becomes important for strong toroidal
+%fields. If the system remains in approximately vertical hydrod
+
 \subsection{Resistive disks with GI}
+Here we examine an isothermal, resistive disk with $Q=0.18$ and
+$\Lambda_0=0.1$. GI is permitted, and has a similar growth rate as the
+MRI.
 

### 2014-03-31 paper/results_gi.tex
tex updates

@@
 modes with $k_xH=1$ as a function of $\beta$ in disks with $Q=0.18$,
-$Q=0.14$ and $Q=0.12$. The colorbar quantifies the importance of
-self-gravity with 
-\begin{align}
- \tau \equiv \frac{\avg{E_g}}{\avg{E_g} + \avg{E_m}}. 
-\end{align} 
-Thus, modes with $\tau  = 1$ are
-energetically dominated by self-gravity (GI) and $\tau\ll1$ represents a mode
-dominated by magnetic perturbations (MRI).
+$Q=0.14$ and $Q=0.12$.
 
@@
 Comparing pure MRI (solid line) and the magnetically-dominated
-portions of $\gamma(k_x)$ show that self-gravity can enhance or reduce
-MRI growth rates depending on $k_x$, but the effect is small.
+portions of $\gamma(k_x)$ show that self-gravity increases MRI growth
+rates at large $k_x$, but the effect is small. Note that this
+destabilization is through the linear response, rather than through
+the background stratification.
 
@@
     Cowling approximation. The dashed line corresponds to pure GI
-    modes, obtained by including a high resisitivity in the full problem. 
+    modes, obtained by including a high resisitivity in the full
+    problem. 
+%    {\bf REDO FIGURE}
     \label{compare_growth3_Q01d2}}

### 2014-03-31 paper/setup.tex
tex updates

@@
 We study the local stability of an inviscid, self-gravitating and
-magnetized fluid disk orbiting a central mass with
+magnetized fluid disk orbiting a central star with
 potential $\Phi_*(r,z)$, where $(r,\varphi,z)$ are cylindrical
-co-ordinates from the central mass. We use the shearing box approximation     
-\citep{goldreich65b} and consider a small patch of the disk at
-$r=r_0$. The local frame rotates at angular velocity 
-$\Omega_0=\Omega(r_0,0)$ about the central mass, where $r\Omega^2 =
+co-ordinates from the star. We use the shearing box approximation     
+\citep{goldreich65b} to consider a small patch of the disk at
+a fiducial radius $r=r_0$. The local frame rotates at angular velocity 
+$\Omega_0=\Omega(r_0,0)$ about the star, where $r\Omega^2 =
 \p\Phi_*/\p r$. We also define $S=-r\p\Omega/\p r$ as the local shear
@@
 The unperturbed disk is steady and described by
-$\rho=\rho(z)$, $\bm{B} = B_z\hat{\bm{z}} + B_y\hat{\bm{y}}$ where
+$\rho=~\rho(z)$, $\bm{B} = B_z\hat{\bm{z}} + B_y\hat{\bm{y}}$ where
 $B_{y,z}$ are constants and the toroidal field strength is
@@
 isothermal disk has no surface, we define $H$ such that
-$\rho(H)=\epsilon\rho(0)$ with $\epsilon=10^{-2}$. A non-dimensional
+$\rho(H)=10^{-2}\rho_0$. A non-dimensional
 measure of the disk thickness is given by 
@@
 \subsection{Resistivity profile}\label{resis_profile}
-We consider constant resistivity or 
+We consider constant resistivity or a 
 resistivity prescription such that $\eta(z)$ increases away from the
-midplane. In the latter case, we follow \cite{fleming03} and adopt the
+mid-plane. In the latter case, we follow \cite{fleming03} and adopt the
 resistivity profile 
@@
 and we define $\eta_0/\eta(\infty)\equiv A$ as the conductivity 
-boost factor from the midplane to the disk surface. We remark that
-once the equilibrium $\rho$ and $d\rho/dz$ are obtained from
+boost factor from the mid-plane to the disk surface. We remark that
+once $\rho$ and $d\rho/dz$ are obtained from
 Eq. \ref{eqm_eqns1} --- \ref{eqm_eqns2}, the integration for
@@
 %isothermal disk. 
-The disk may be considered ideal where $\Lambda
-\gtrsim 1$. 
+Because of the density stratification, the Elsasser number 
+increases with height even for constant resistivity. The disk may be
+considered ideal where $\Lambda \gtrsim 1$. 
 %Since $\hat{\rho}(z/H)$ is weakly dependent on  self-gravity, the
@@
 \subsection{System parameters}
-We summarize here the main parameters describing the resistive,
-self-gravitating shearing box. The strength of self-gravity is
+The main parameters describing the resistive, self-gravitating
+shearing box is as follows. The strength of self-gravity is
 parametrized by 
@@
 The relation between $Q$ and the Toomre parameter for gravitational
-instability of razor-thin disks is described in Appendix \ref{q3d2d}.
+instability of razor-thin disks, $Q_\mathrm{2D}$, is described in
+Appendix \ref{q3d2d}.
 
@@
 \end{align}
-where $v_{A0}$ is the midplane Alfven speed. The strength of
-resistivity is measured by the midplane Elsasser number 
+where $v_{A0}$ is the midplane Alfven speed. Note that we use 
+the vertical field for this definition throughout the paper. The
+strength of conductivity is measured by the midplane Elsasser number  
 \begin{align}

### 2014-03-31 paper/summary.tex
tex updates

@@
 \section{Summary and discussion}\label{summary}
-In this paper, we have examined the linear stability of radially
-structured three-dimensional disks with non-uniform entropy
-distribution. These calculations may be considered as an extension to 
-the 2D Rossby wave instability \citep{li00} by adding the vertical
-dimension, or to the barotropic RWI calculations of \citetalias{lin12} by
-adding an energy equation with a simpler numerical method.
+In this paper, we have performed explicit stability calculations of
+magnetized, self-gravitating disks in the local approximation.
 
-We adopted polytropic disk equilibria so that
-the magnitude of entropy gradients can be conveniently parametrized
-by $\Delta\gamma\equiv\gamma/\Gamma$, and we focused on the effect of
-$\Delta\gamma\geq1$. When the background density and
-velocity field is fixed through $\Gamma$, we found increasing $\Delta
-\gamma$ has negligible effect on the instability growth rate. However,
-{\bf the magnitude of} pressure and density perturbations increase with height, and the 
-meridional flow associated with the vortex core is qualitatively
-changed, with the introduction of {\bf meridional} vortical
-motion.
+%layered accretion incompatable in massive disks
 
-{\bf Meridional vortical motion was found to correlate with a small 
-  tilt of a fluid column with negative vertical vorticity
-  perturbation.  In standard hydrodynamics, vorticity tilting
-  can originate from a contribution of the
-  $\bm{\omega}\cdot\nabla\bm{v}$ term in the evolution equation of the
-  vorticity independently of the baroclinic source term $\nabla \rho
-  \times\nabla p$. However, given the tilt is absent in our homentropic
-  calculations, we associate the tilt with the baroclinic source term, which
-  produces azimuthal vorticity. } 
-We also found
-that the vertical velocity at the vortex core is no longer linear in
-$z$, as for homentropic flow.
-
-In our second set of experiments, we fixed $\gamma$ and decreased
-$\Gamma$. We found that by making the flow nonhomentropic, the
-co-rotation region became \emph{more} three-dimensional, despite the
-decrease in growth rate. This result
-is opposite to \citetalias{lin12} where lowering $\Gamma$ made the
-flow less three-dimensional. This implies that entropy gradients  
-play an important role in the vertical structure of the perturbations. 
-%This will likely have impact
-%on the motion of dust particles within Rossby vortices, in particular 
-%as the flow is dominated by bouyancy forces away from the midplane.
-
-We also considered isothermal equilibria. A linear calculation with 
-$\Gamma=1.1$ and one with a strictly isothermal setup ($\Gamma\equiv
-1$) were consistent. Both produced prominent meridional 
-vortical motion. In order to verify this feature, we ran a nonlinear
-simulation of the RWI in an initially isothermal disk, but evolved
-adiabatically. We indeed identified said  vortical motion.    
-%although the agreement with linear calculations is
-%qualitative. This might imply that nonlinear effects become important
-%for the RWI vertical structure earlier than in the horizontal
-%direction. 
-Keeping in mind that the setup for linear and nonlinear simulations 
-were not identical (e.g. numerical grid, boundary treatment), similarities
-between them, such as mode frequency, growth rate and horizontal flow,
-are satisfactory.
-
-Vortical motion in the meridional plane thus appears characteristic of 
-the \emph{linear} RWI in nonhomentropic disks. Whether or not this is 
-significant for the vortex evolution can only be answered by detailed 
-long term nonlinear simulations. If this vortical motion is present 
-in the nonlinear regime then it may prevent dust particles from reaching the disk surface, 
-which occurs for homentropic flow \citep{meheut12c}.
-
-However, given this meridional vortical motion is absent in the homentropic 
-linear solution, it may eventually vanish because of entropy
-mixing, if no mechanism is present to maintain entropy gradients. 
-For example, the background entropy increases with height but 
-the linear entropy perturbation becomes more negative with height, {\bf 
-and its magnitude grows exponentially in time}. 
-{\bf Indeed,  recent 3D fully compressible simulations in
-  nonhomentropic disks shows that well into the nonlinear regime,
-  Rossby vortices are columnar
-  \citep{richard13}}. On the other hand, \cite{meheut12} observed
-strong meridional vortical motion in their homentropic hydrodynamic
-simulations; we conclude they are of nonlinear origin.
-
-In the linear solutions, we often observe perturbation magnitudes
-increase away from the midplane in nonhomentropic
-disks\footnote{This reminds us of the off-midplane vortices
-  discovered by \cite{barranco05} in nonlinear local simulations, 
-%  which suggest the non-zero vertical bouyancy frequency to be
-%  stabilizing   against secondary instabilities \citep{lesur09}. 
-  but the  
-  setup considered in that study is very different from the
-  present work. Nevertheless, in both cases the vertical entropy
-  gradient is stabilizing away from the midplane. 
-  %%  Nevertheless, a positive vertical bouyancy frequency
-  %% may stabilize 3D vortices \citep{lesur09}.
-} 
-(e.g. Fig. \ref{lin_nonlin_rz}). Then the RWI may not be as robust
-against vertical boundary conditions as it is to radial boundary
-conditions. This could pose difficulty for the RWI to develop in dead
-zones of real protoplanetary disks, which are expected to be confined
-from above and below by magnetically turbulent layers
-\citep{oishi09}. The vertical boundary condition set by these layers
-may or may not be compatible with the linear RWI solution.
+%why did previous MRI simulations find no difference between SG and
+%nonSG? prob because the mri mode has too small vertical wavelength
 
 \subsection{Caveats and outlooks}
-%need a good guess
-One trade-off for the simplicity of our numerical method for 
-linear simulations is that a trial eigenfrequency must be
-guessed. This is not a significant obstacle for the problem at
-hand, because previous RWI studies provide an important guide
-\citep[][]{li00}. Otherwise, zeros of the complex function
-$\mathcal{D}(\sigma)=\mathrm{det}\,\bm{U}$ need to be located with more 
-rigorous methods \citep[e.g.][]{kojima86,valborro07}. We have also exploited
-previous findings that the PPI and RWI are predominantly two-dimensional 
-\citep{papaloizou85, goldreich86, kojima89, umurhan10, meheut12, lin12, lin12c}, which
-enabled the use of a small number of basis functions. 
-However, there could exist parameter regimes where the RWI
-has significant vertical structure, rendering our solution method
-inefficient.
+We discuss below two major extensions to our linear model that
+should be undertaken, before performing non-linear simulations of
+magnetized, self-gravitating disks, which is our eventual goal.
 
-%extension to baroclinic equilibria 
-%simple polytrope background
-Our conclusions are limited to polytropic backgrounds. While this was 
-convenient for numerical experiments, it is an over-simplification
-of protoplanetary disks, which are 
-expected to have complicated vertical structure \citep{terquem08}. In
-particular, we found that entropy gradients plays a role in the
-vertical structure of the {\bf linear} RWI, and even a small entropy gradient can noticeably
-modify the vertical flow (\S\ref{nz_nonzero}).
-Thus, a realistic model for entropy evolution is needed. 
-%This is important because even a small entropy gradient can noticeably 
-%modify the vertical flow (\S\ref{nz_nonzero}).
+\emph{Beyond the shearing box.} The shearing box ignores the curvature
+of toroidal field lines present in the global disk
+geometry. \cite{pessah05} have shown that new effects on the MRI arise
+when the curvature of a super-thermal toroidal field is accounted for;
+although \citeauthor{pessah05} focused on modes with large (small)
+vertical (radial) wavenumbers, for which we do not expect self-gravity
+to be important. Since compressibilty becomes important for strong
+toroidal fields, the effect of self-gravity may become even more
+significant when super-thermal toroidal fields are
+considered. However, global disk models will be neccessary
+to self-consistently probe this regime.
 
-It would also be of interest to generalize the calculations
-to \emph{baroclinic} equilibria\footnote{In fact, baroclinic tori were
-  briefly considered by \cite{frank88}.}, 
-for which $\p_z\Omega\neq0$. This may well be the 
-case when the equilibrium pressure depends on both the density and temperature. 
-%This requires 
-%derivation of the linearized equations including terms involving $\p_z\Omega$. 
-Complications from baroclinic instabilities may arise, however \citep{knobloch86, umurhan12, nelson12}.
-%% This may involve adding heating or cooling
-%% sources to the energy equation. 
-%source terms - friction, cooling
-%new physics - self-gravity
+%
+%This
+%prevents consideration of super-thermal toroidal fields
+%\citep{pessah05}.
 
-We have neglected gas self-gravity in this study. Our models therefore
-assume that the Toomre parameter is much larger than unity in both the 
-unperturbed \emph{and} perturbed states. {\bf However, self-gravity
-  may affect the RWI even when the Toomre parameter is not small
-  \citep{lovelace12}}. 
-Previous studies
-have found  higher $m$ RWI modes are favored when disk self-gravity is
-included \citep{lyra08,lin11a}. Recent 3D simulations of the RWI in a
-locally isothermal disk show that vertical self-gravity can noticeably
-enhance the density perturbation near the midplane,
-even though the initial disk was considered low mass \citep{lin12b}. %%  The effects
-%%  of self-gravity %% should be considered in conjunction with disk
-%%  thermodynamics. %In a nonlinear simulation, the latter may not hold even if the former
-%does.
-%Previous studies of the RWI also find higher $m$ modes modes are
-%favored when disk self-gravity is included
-%\citep{lyra08,lin11a}.
+%tension stabilizes GI  
+%This prevents us from considering super-thermal
+%toroidal fields.  
+%go to cylindrical disk geometry (fu & lai) 
+%tension stabilizes GI
 
-In principle, one can express the Poisson integral as a matrix
-operator and  incorporate it into our formalism. {\bf The linear} problem is
-further complicated by the need of a numerical solution to the
-equilibrium equations describing a radially structured,
-self-gravitating 3D disk \citep{muto11}.  Such a linear calculation is
-beyond the scope of this paper, but will be inevitable for
-understanding  the RWI in 3D self-gravitating disks. Perhaps a simpler
-starting point, to gain first insight, is  direct hydrodynamic 
-simulations {\bf including disk gravity}. This is indeed the approach taken in our follow-up paper.    
-%\clearpage
-%since the ZEUS-MP code already includes a Poisson solver (as used in Lin 2012b).
-
-%have shown that
-%self-gravity reduces the RWI growth rate through the potential perturbation, 
-%but may be destabilizing through modification of the background disk \citep{lin11a}.
+%spiral arms, GI angular mometum distribution 
+\emph{Beyond axisymmetry.} It is well known that self-gravitating
+disks can be unstable to non-axisymmetric perturbations but stable
+against axisymmetric perturbations, although the threshold Toomre
+parameter for both is still of order unity. %papaloizou/lin,
+                                %sajvoini??

### 2014-04-02 paper/results.tex
tex updates

@@
 
-Fig. \ref{mri_massive_resis} plots the energies assocaited with the 
+Fig. \ref{mri_massive_resis} plots the energies associated with the 
 MRI mode discussed above. 
@@
 MRI can be amplified by self-gravity. Now, for $k_xH\sim 1$ we have
-$|k_z^2/k_x^2|\sim \beta/f\gg1$ because $f=O(1)$ and $\beta\gtrsim10$
+$|k_z^2/k_x^2|\sim \beta/f^2\gg1$ because $f=O(1)$ and $\beta\gtrsim10$
 for the cases considered above. Then
@@
  The above arguments can be adapted to the resistive
- limit. Eq. \ref{simp3}---\ref{simp4} are unchanged, while resistive
+ disk. Eq. \ref{simp3}---\ref{simp4} are unchanged, while resistive
  terms appearing in Eq. \ref{simp1}---\ref{simp2} only involve the
@@
 
- For weak fields, the MRI has small vertical wavelength,
- implying rapid density variations with height. The magnitude of the
- corresponding gravitational potential is then expected to be small
- because of internal cancellations, since the potential is essentially
- the integral of the density field.
+ %mass contained in small wavelength is small
+
+ For weak fields in an ideal disk, the MRI has a small vertical
+ wavelength, $\lambda \ll H$. The perturbed mass contained within
+ $\sim \lambda$ is small and its potential
+ perturbation is unimportant. Furthermore, considering the stratified
+ disk, $\lambda\ll H$ imply rapid variations in the density
+ perturbation across the disk height, averaging to zero, so the
+ magnitude of the associated potential perturbation is small. 
+% because the gravitational potential is the integral of the density
+% field, i
+ Self-gravity plays does not affect the MRI in this regime.
+
+ %, i.e. internal cancellations would occur when such an integral is
+ % performed.
+
+ However, a strong field and/or large resistivity increases the MRI
+ vertical wavelength. When the vertical scale of the MRI becomes
+ comparable to the disk thickness, i.e. $\lambda\sim H$, the
+ perturbed mass across the disk height can contribute to a net potential
+ perturbation. We therefore expect self-gravity to be important for
+ the MRI when it is weak.
+
+%The perturbed mass
+% within the unstable mode corresponds to that in the within 
+%This can give rise to a net
+% potential perturbation  
+ %Self-gravity can then be effective, since it
+ % operates globally in the vertical direction. 
+%GI of 3D disks can be
+ %analysed in vertically-integrated
+ %systems 
+%coupling to SG 
+%That is when the MRI is weak

### 2014-04-02 paper/results_azi.tex
tex updates

@@
 
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 2cm 0cm
+    0cm]{figures/compare_growth3_tilted_resis_ep0.ps}  
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 2cm 0cm
+    0.52cm]{figures/compare_growth3_tilted_resis_ep1.ps}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+    0.52cm]{figures/compare_growth3_tilted_resis_ep2.ps}
+  \caption{Growth rates in isothermal resistive disks with $Q=0.18$
+    ($Q_\mathrm{2D}=0.67$) and $\Lambda_0=0.1$.  
+    \label{compare_growth3_tilted_resis}}
+\end{figure}
+
 %\subsection{GI}

### 2014-04-02 paper/results_gi.tex
tex updates

@@
 \section{Gravitationally unstable disks} \label{result2}
-Gravitational instability (GI) becomes possible in a sufficiently
-massive and/or cold disk. We explore how the emergence of GI affects
-the MRI, and vice versa. To do so, we compute unstbale modes for  
-isothermal disks with $Q < 0.2 $ ($Q_\mathrm{2D}\lesssim 0.67$) which
-permits gravitational instability as well as the MRI.
+Gravitational instability becomes possible in a sufficiently
+massive and/or cold disk. Here, we explore whether or not GI and MRI
+can interact by computing unstable modes for isothermal disks with $Q < 0.2 $
+($Q_\mathrm{2D}\lesssim ~0.67$) which permits GI, as shown below.
+We consider ideal disks with $\Lambda_0=100$ and $A=1$, unless
+otherwise stated.
 
 \subsection{Co-existence of MRI and GI} 
-We first consider ideal MHD by setting $\Lambda_0=100$ and $A=1$. 
-Fig. \ref{compare_growth3} show growth rates for 
-modes with $k_xH=1$ as a function of $\beta$ in disks with $Q=0.18$,
-$Q=0.14$ and $Q=0.12$.
-
+Fig. \ref{compare_growth3} show growth rates for modes with $k_xH=1$
+as a function of $\beta$ in disks with $Q=0.18$, $Q=0.14$ and
+$Q=0.12$.   
 All three cases display distinct GI modes (red/brown branch). The GI 
@@
 $Q=0.18,\,0.14,\,0.12$, respecitvely. GI is stabilized by magnetic
-pressure for sufficiently small $\beta$, and the critical field
-strength increases with the self-gravity \citep{nakamura83}. For
-$Q=0.18$, GI is stabilized for $\beta \lesssim 15$. Nevertheless, 
+pressure for sufficiently small $\beta$, and the critical
+field strength increases with increasing self-gravity \citep{nakamura83}. 
+For $Q=0.18$, GI is stabilized for $\beta \lesssim 15$. Nevertheless, 
 the MRI branch for $\beta < 15$ becomes self-gravitating, so that
-density perturbations are still expected even though  GI does not
+density perturbations will still develop, even though GI does not
 formally operate.
@@
     0.5cm]{figures/compare_growth3_kx1_Q0d12.ps} 
-  \caption{Growth rates for linear modes with $k_xH=1$ in an isothermal disk with
+  \caption{Growth rates for modes with $k_xH=1$ in an isothermal disk with
     $Q=0.18$ (top), $Q=0.14$ (middle) and $Q=0.14$ (bottom). The
     colorbar measures the importance of  
-    self-gravity.  
+    self-gravity with $\tau$.   
     \label{compare_growth3}}
@@
 where the GI branch approaches an MRI branch at $\beta\simeq 25,\,
-\gamma\simeq 0.2\Omega$. In fact following the red curve to smaller
+\gamma\simeq 0.2\Omega$. In fact, following the red curve to smaller
 $\beta$ indicates GI transitions to MRI. The
 `gaps' in the GI and MRI branches for $Q=0.18$ and $Q=0.12$ may be due
-to avoided crossing of modes, as seen in stars
+to the phenomenon of avoided crossing, as seen in stars
 \citep[e.g.][]{aizenman77} and accretion tori
-\citep[e.g.][]{christo93}. However, we cannot exclude the possibility 
-that some modes may have been missed in the numerical search of
-eigenfrequencies.
+\citep[e.g.][]{christo93}, where physically distinct modes approach
+one another in frequency and exchange character. However, we cannot
+exclude the possibility  that some modes may have been missed in a 
+numerical search of eigenfrequencies.
 
 Thus, our results do not rigorously prove that the GI and MRI branches
-do not intersect.  Nevertheless, the continous variation of $\tau$ with
-respect to $\beta$ strongly suggest that unstable modes can transition
-smoothly from magnetically-dominated to gravitationally-dominated at
-low $\beta$.
+do not intersect.  Nevertheless, the continous variation of $\tau$
+strongly suggest that unstable modes can transition smoothly from 
+MRI to GI and vice versa, especially at low $\beta$.
 
@@
 given set of disk parameters. Fig. \ref{compare_growth3_Q01d2} show
-growth rates as a function of $k_x$ for an isothermal disk with
-$Q=0.12,\,\beta=20$. MRI and GI are active with comparable growth
+growth rates as a function of $k_x$ in a disk with
+$Q=0.12,\,\beta=20$, where MRI and GI have comparable growth
 rates. All perturbations with $k_xH \lesssim 3.5$ grow dynamically
@@
 
-The dispersion relation $\gamma(k_x)$ increasingly deviates from the
-pure GI/MRI curves with $k_x$, implying increasing interaction between
-magnetic and density perturbations. 
-Comparing pure GI (dashed line)  and the gravitationally-dominated
-portions of $\gamma(k_x)$ indicate inclusion of magnetic field
-stabilizes high-$k_x$ pure GI. (Note also the slight decrease in the most unstable
-$k_x$.) Since axisymmetric GI is unaffected by shear, this
-stabilization is due to magnetic pressure and is consistent with
-\citep{lizano10}.
+The dispersion relation $\gamma(k_x)$ deviates from the
+pure GI/MRI curves with increasing $k_x$, implying stronger interaction between
+magnetic and density perturbations. Comparing pure GI (dashed line)
+and the gravitationally-dominated portions of $\gamma(k_x)$ shows that 
+inclusion of magnetic field stabilizes high-$k_x$ pure GI. (Note also
+the slight decrease in the most unstable $k_x$.) %Since axisymmetric GI
+%is unaffected by shear, 
+This stabilization is due to magnetic pressure \citep{lizano10}, which
+is consistent with stabilizing the small wavelength GI only.
 
@@
 portions of $\gamma(k_x)$ show that self-gravity increases MRI growth
-rates at large $k_x$, but the effect is small. Note that this
+rates at large $k_x$, but the effect is small. %can be used for code
+                                %test 
+Note that this 
 destabilization is through the linear response, rather than through
-the background stratification.
+the background stratification (which is stabilizing). 
+%For the $k_x$ in the GI
+%interval, MRI is weakly stabilized by self-gravity.
 
@@
 
+%implications for simulations -> need to resolve small-scales
+
+%Fig. \ref{compare_growth3_Q01d2} again shows smooth
+%transitions between MRI and GI. This means the MRI and GI do indeed
+%interact, but
+
+%In particular, modes with $1.5 \lessim
+%k_xH\lesssim2$ is neither pure
+
+%demonstrates that mode classification
+%becomes ambigious when both GI and MRI are present.
+
 %\subsection{Layered disks}

### 2014-04-02 paper/summary.tex
tex updates

@@
 \section{Summary and discussion}\label{summary}
-In this paper, we have performed explicit stability calculations of
-magnetized, self-gravitating disks in the local approximation.
+In this paper, we have performed axisymmetric linear stability
+calculations of magnetized, self-gravitating disks in the local
+approximation. Our models include resistivity and azimuthal fields.  
+We have identified regimes under which the magneto-rotational
+instability (MRI) is affected by disk self-gravity (SG).  
+%what have we extended?
 
-%layered accretion incompatable in massive disks
+For a vertical field, the requirement for the MRI to operate is that its
+vertical wavelength $\lambda \lesssim H$. While $\lambda$ is independent of
+the strength of self-gravity $Q$, the disk thickness $H=H(Q)$ %lambda
+%depends on omega, which might be
+                                %lowered in a massive disk (sub-keplerian)
+decreases with increasing SG, which reduces the MRI growth rates.   
+%This
+%effect becomes noticeable when the disk is marginally stable to GI in
+%3D ($Q\sim 0.2$ or $Q_\mathrm{2D}\sim 0.7$).  
+Thus, a sufficiently massive disk can surpress the MRI. %but GI may
+                                %develop  instead
+The MRI is also restricted to larger radial
+scales as $Q$ is lowered. This means that the MRI becomes more global
+in self-gravitating disks. Finally, the condition $\lambda < H$ 
+requries $f(Q)/\sqrt{\beta}$ to be sufficiently small, so weaker
+fields are required to permit the MRI with increasing self-gravity.
 
-%why did previous MRI simulations find no difference between SG and
-%nonSG? prob because the mri mode has too small vertical wavelength
+%need larger boxes 
+%This suggests
+%that small-scale MRI turbulence may not develop in self-gravitating
+%disks.
+
+%correlation
+
+Interestingly, for layered resistivity we do not find layered
+perturbations when the disk is massive. This is 
+consistent with MRI becoming vertically global with increasing self-gravity.    
+For non-self-gravitating disks $\lambda\ll H$, so the MRI can be
+restricted to regions of size $L<H$, i.e. an active layer. This is not
+possible if $\lambda \sim H$, as found for massive disks. Hence we
+find magnetic perturbations penetrate into the high-resistivity dead
+zone (e.g. $Q=0.2$ in Fig. \ref{poly_layer}), and there is no distinct
+boundary between active and dead layers. This suggests that the
+picture of layered accretion  \citep[e.g.][]{fleming03} may not be applicable 
+to self-gravitating disks.  %strong magnetization
+
+%SG through 
+We find MRI perturbations with radial scale of $\sim H$ can acquire  
+density perturbations in a massive but Toomre-stable disk. 
+This occurs when the MRI is weak, for example with a strong field or
+high resisitivty. We argue in that case $\lambda\simH$, so the density
+perturbation caused by the MRI can be enhanced by self-gravity.
+
+%further enhancing density perturbations. 
+At this point it is worth mentioning previous non-linear calculations 
+performed by \cite{kim03} and \cite{fromang04a,fromang04} in the
+context of galactic and circumstellar disks, respectively. These
+authors simulated the MRI in self-gravitating disks. They find
+self-gravity has little impact on the MRI.
+
+%cite{kim03} and \cite{fromang04}
+
+%Numerical experiments
+%varying the field strength and resisitivity indicate this is again a
+%result
+
+%GI/MRI: avoided crossings
+
+%azimuthal fields
+
+%\subsection{Implications for non-linear simulations}
+%what scaes? resolution
+%why don't formang  and kim/ostriker see any difference?
 

### 2014-04-03 paper/paper.tex
tex updates

@@
   calculated numerically. The formulation includes 
-  resistivity. It is found that the reduction in the disk scale height
-  by self-gravity (SG) can decrease MRI growth rates. SG causes the MRI to
-  become global in the vertical direction, and MRI modes
+  resistivity. It is found that the reduction in the disk thickness
+  by self-gravity (SG) can decrease MRI growth rates; the MRI
+  becomes global in the vertical direction, and MRI modes 
   with small radial lengthscales are stabilized. In massive
@@
   to regions where the Elsasser number exceeds unity. For MRI modes
-  with radial lengthscales on the order of the disk scale-height, SG
+  with radial lengthscales on the order of the disk thickness, SG
   can enhance density perturbations, an effect that becomes 

### 2014-04-03 paper/results_azi.tex
tex updates

@@
 \section{Effect of an azimuthal field}\label{result3}
-In this section we add an azimuthal field so that $B_y\neq 0$,  
+In this section we use the setup of case 2 described in
+\S\ref{linear}, and examine the effect of an azimuthal field so that
+$B_y\neq 0$,   
 parameterized by $\epsilon \equiv B_y/B_z$. However, we continute to
@@
 \subsection{Ideal disks with MRI} 
-Here we consider an isothermal disk with $Q=0.2$
-($Q_\mathrm{2D}=0.72$) and $\beta=10$ in the limit of ideal MHD
-($\Lambda_0=100$). Gravitational instability is not expected because 
-Fig. \ref{compare_growth3} shows that even for $Q=0.18$, GI is 
-surpressed for $\beta \lesssim 15$. We consider azimuthal field
-strengths with $\epsilon\in[0,3]$.
+We consider an isothermal disk with $Q=0.2$ ($Q_\mathrm{2D}=~0.72$) and
+$\beta=10$ in the limit of ideal MHD ($\Lambda_0=100$). Gravitational
+instability is not expected because Fig. \ref{compare_growth3} shows
+that even for $Q=0.18$, GI is surpressed for $\beta \lesssim 15$. %We
+%consider azimuthal field strengths with $\epsilon\in[0,3]$.
 
-Fig. shows the MRI growth rates for the above system as a function of
-$k_x$. The top panel are modes with even midplane symmetry: 
-$W^\prime(0)= 0$ for $\epsilon\to0$, and the bottom panel are modes with odd midplane 
-symmetry: $W(0) = 0$ for $\epsilon\to0$. The latter set of modes were 
-excluded in the previous sections by midplane boundary 
-conditions. The plot also include growth rates computed in the Cowling
+Fig. \ref{compare_growth3_tilted} show MRI growth rates for
+$B_y/B_z=0,\,1,\,2$ and $3$. We divide the modes into two catagories
+depending on the extremum of magnetic energy at the midplane. 
+The top panel are modes where $E_m$ has a local minimum at $z=0$ and 
+the bottom panel are modes where $E_m$ has a local maximum at $z=0$. 
+The latter set of modes were excluded in the previous sections by
+midplane boundary conditions. 
+%The top panel are modes with even midplane symmetry: 
+%$W^\prime(0)= 0$ for $\epsilon\to0$, and the bottom panel are modes with odd midplane 
+%symmetry: $W(0) = 0$ for $\epsilon\to0$. 
+We also plot growth rates computed in the Cowling
 approximation. As expected,  $\avg{E_g}<\avg{E_m}$, so none of the
@@
 %no GI for odd modes 
-
 %general stabilization of MRI
@@
 
-Consider first the even modes in the top panel of 
-Fig. \ref{compare_growth3_tilted}. %Self-gravity slightly stabilizes
-%modes with small $k_x$ as $B_y$ is increased. 
-Self-gravity destabilizes modes with  $k_xH\gtrsim O(1)$. Consequently, the
+Consider first modes in the top panel of 
+Fig. \ref{compare_growth3_tilted}. As with previous results,  
+self-gravity destabilizes modes with  $k_xH\gtrsim O(1)$. Consequently, the
 cut-off wavenumber is larger when SG is included. %get smaller scale
@@
 Destablization is most effective for purely vertical fields: with
-$\epsilon=0,\, 
-k_xH\simeq 1.4$, SG increases the growth rate by $\sim 30\%$. Although
-these modes are fundamentally magnetic, this is consistent with
-\cite{goldreich65a}, who find SG can only destabilize 
-symmetric density perturbations, which is the case for $B_y=0$ here.  
-With increasing $\epsilon$, we find the density perturbation $W$ deviates from an even 
-function, $|W(0)|$ decreases and $\max{(|W|)}$ moves off of the mid-plane.  
+$\epsilon=0,\, k_xH\simeq 1.4$, SG increases the growth rate by $\sim
+30\%$. For $B_y=0$ we find the density perturbation $W(z)$ is an even
+function. Although these modes are fundamentally magnetic, this is consistent with
+\cite{goldreich65a}, who showed that SG can only destabilize 
+symmetric density perturbations.  
+With increasing $\epsilon$, we find $W$ deviates from an even
+function. %,% $|W(0)|$ decreases and $\max{(|W|)}$ 
+%moves off of the mid-plane.    
 %density pert at the midplane goes down
@@
 Thus, the Cowling approximation becomes increasingly good with stronger
-$B_y$ for the even modes.
+$B_y$ for these modes.
 
@@
 The modes in the bottom panel of Fig. \ref{compare_growth3_tilted} 
-display opposite behavior. For $B_y=0$, $W$ is odd and 
-self-gravity has no effect. When $B_y>0$, we find $W$ deviates from an odd
-function and the mid-plane density perturbation $|W(0)|$ increases. 
-SG is stabilizing for these modes at all wavelengths, and is most
-effective at $k_xH =  O(1)$.  
+display opposite behavior. For $B_y=0$ we find $W$ is odd, and
+self-gravity has no effect. When $B_y>0$, $W$ deviates from an
+odd function and the mid-plane density perturbation $|W(0)|$
+increases. SG is stabilizing for these modes at all wavelengths, and
+is most effective at $k_xH =  O(1)$.  
 %destabilizing effect is absent  
@@
 Fig. \ref{result_tilted} show eigenfunctions for $\epsilon=3$ and
-$k_xH=1.2$ with and without the Cowling approximation. These differ
-from the even modes considered above in that the horizontal magnetic
-perturbations maximize at $z=0$ (cf. Fig. \ref{mri_massive_resis}).  
-SG significantly enhances the mid-plane density perturbation, making the
+$k_xH=1.1$ with and without the Cowling approximation. 
+%These differ  
+%from the even modes considered above in that the magnetic energy
+%maximize at $z=0$ (cf. Fig. \ref{mri_massive_resis}).  
+SG significantly enhances the midplane density perturbation, making the
 gravitational potential energy comparable to the magnetic energy, 
@@
 %strong field, compressibility 
-To interpret the above result for odd modes, note 
-that compressibility affects the MRI in the presence of an 
-azimuthal field, even in a non-self-gravitating disk. If the perturbed 
-disk remains in vertical hydrostatic equilibrium, then 
+To interpret the above result for modes with magnetic energy
+concentrated at the midplane, we note that compressibility affects the
+MRI in the presence of an azimuthal field even in a
+non-self-gravitating disk. If the perturbed disk remains in vertical
+hydrostatic equilibrium, then  
 \begin{align} 
@@
 \end{align}
-to order of magnitude in a non-SG disk, so a strong toroidal field can
-cause a large density perturbation
-\citep{pessah05}. Fig. \ref{result_tilted} shows that this is
-amplified by self-gravity.
+to order of magnitude in a non-SG disk. Thus a strong azimuthal field can 
+cause a large density perturbation \citep{pessah05}. We checked that
+for the modes in Fig. \ref{result_tilted}, vertical velocities are 
+small, $|\dvz|/\left(|\dvx|^2+|\dvy|^2\right)^{1/2} \lesssim 0.2$.
 
-We checked that for the modes in Fig. \ref{result_tilted} the vertical
-velocity is small, $|\dvz|/\left(|\dvx|^2+|\dvy|^2\right)^{1/2}
-\lesssim 0.2$, so the above argument applies. Compressibilty is 
-enhanced by both the azimuthal field and self-gravity, which tends
-to stabilize the MRI \citep{kim00}. Its effect is important here  
-because the azimuthal Alfven speed is sonic. The destabilization
-effect of SG is absent (or unimportant) because the density
-perturbation is not symmetric.
+Compressibilty is enhanced by an azimuthal field, which has a
+stabilizing effect on the MRI \citep{kim00}. %because...?
+This effect is significant for $\epsilon=3$ because the azimuthal
+Alfven speed is sonic.    
+%maybe blaes 
+Fig. \ref{result_tilted} indicates that self-gravity further enhances
+compressibility, and therefore stabilization. We suspect this is
+overwhelmed by the destabilization effect of SG because the density
+perturbation has an anit-symmetric component.
+
+%compressibility -> pressure restoring force
 
@@
 
+%\subsection{Ideal disks with GI (TBC)}
+
 \subsection{Resistive disks with GI}
-Here we examine an isothermal, resistive disk with $Q=0.18$ and
-$\Lambda_0=0.1$. GI is permitted, and has a similar growth rate as the
-MRI.
+Here we consider an isothermal, resistive disk with $Q=0.18$,
+$\Lambda_0=0.1$ and $\beta=100$. This system permits GI and MRI. 
+%GI is permitted, and has a similar
+%growth rate as the  
+%MRI. 
+Fig. \ref{compare_growth3_tilted_resis} show growth rates for
+$\epsilon=0,\,1$ and $2$. For $B_y=0$, MRI and GI are decoupled except 
+for a narrow range of $k_x$ in which the lower MRI modes transitions to
+GI. Notice that the upper MRI modes intersect the GI branch. There is no
+interaction because the upper MRI modes have anti-symmetric $W(z)$ whereas
+the GI modes have symmetric $W(z)$.
+
+%GI, and there are MRI modes available on that scale
 
@@
   \caption{Growth rates in isothermal resistive disks with $Q=0.18$
-    ($Q_\mathrm{2D}=0.67$) and $\Lambda_0=0.1$.  
+    ($Q_\mathrm{2D}=0.67$), $\beta=100$ and $\Lambda_0=0.1$. For
+    $B_y=0$, the upper MRI modes have anti-symmetric density
+    perturbations, $W(0) = 0$; while the lower MRI modes have symmetric
+    density perturbations, $W^\prime(0) = 0$. For $B_y/B_z=2$ the
+    overstable modes have non-zero real frequencies. 
     \label{compare_growth3_tilted_resis}}
@@
 
+Introducing $B_y = B_z$ leads to an exchange in the mode
+characters. For $k_xH\lesssim 0.9$ the modes on the two MRI branches
+are similar to the vertical field case. However, for $k_xH\gtrsim0.9$ 
+the upper MRI mode transitions to GI, for which $E_m(0)$ is a
+minimum; and the lower MRI mode has $E_m(0)$ being a maximum. 
+%interchange possible because of GI branch
+
+%In both cases where the gravitational energy is significant, the
+%density perturbation is symmetric.
+
+Increasing the azimuthal field further to $B_y=2B_z$ we find
+overstable MRI modes \citep{gammie96} with non-negligible real
+frequencies. An example is shown in
+Fig. \ref{result_tilted_overstable}. Notice the density/potential
+perturbation is off-set from the midplane. This is not possible for
+pure GI \citep{goldreich65a}. Thus, these overstable MRI modes indeed
+become self-gravitating, before being stabilized. 
+%these may lead to significant vertical motions
+
+Notice in Fig. \ref{compare_growth3_tilted_resis} the disappearence of
+magnetic modes as $B_y$ is increased. For $B_y=2B_z$, MRI and GI
+operate at distinct scales. This implies that regions unstable to GI
+cannot develop MRI.
+
+%This may lead to 
+%the lower MRI mode transitions to the top MRI mode
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+    0cm]{figures/result_tilted_resis.ps}
+  \caption{Overstable MRI mode in an isothermal resistive disk with
+    $Q=0.18 (Q_\mathrm{2D} = 0.67)$, $\Lambda_0=0.1$ and $\beta=100$. 
+    The mode has a real frequency $\omega = 0.059\Omega$, or
+    $\omega/\gamma \simeq 0.2 $.   
+    \label{result_tilted_overstable}}
+\end{figure}
+
 %\subsection{GI}

### 2014-04-03 paper/results_gi.tex
tex updates

@@
 $Q=0.18,\,0.14,\,0.12$, respecitvely. GI is stabilized by magnetic
-pressure for sufficiently small $\beta$, and the critical
+pressure for sufficiently small $\beta$. The critical
 field strength increases with increasing self-gravity \citep{nakamura83}. 

### 2014-04-03 paper/summary.tex
tex updates

@@
 This occurs when the MRI is weak, for example with a strong field or
-high resisitivty. We argue in that case $\lambda\simH$, so the density
+high resisitivty. We argue in that case $\lambda\sim H$, so the density
 perturbation caused by the MRI can be enhanced by self-gravity.
@@
 %further enhancing density perturbations. 
-At this point it is worth mentioning previous non-linear calculations 
-performed by \cite{kim03} and \cite{fromang04a,fromang04} in the
-context of galactic and circumstellar disks, respectively. These
-authors simulated the MRI in self-gravitating disks. They find
-self-gravity has little impact on the MRI.
+At this point it is worth mentioning previous non-linear simulations
+of magnetized self-gravitating galactic and circumstellar disks
+\citep{kim03,fromang04a,fromang04}. These authors find self-gravity
+does not enhance MRI density flucations significantly. However, they
+employed ideal MHD simulations with gas-to-magnetic pressure ratios of
+order $10^2$ to $10^3$. According  
+to our results, self-gravity is not expected to affect the MRI in this
+regime of $\beta$. On the other hand, \cite{fromang04} found
+self-gravity makes MRI turbulence more coherent. This may be related
+to our finding that small radial scale MRI is surpressed when
+self-gravity is included in the background equilibria.
 
+Physically, we expect MRI to only interact with self-gravity when
+their spatial scales are similar. Because self-gravity acts globally in the
+vertical direction, for it to affect the MRI, future non-linear
+simluations should consider parameter regimes in which the MRI is
+vertically global. Indeed, in the setup of \cite{kim03}, the disk
+scale height exceeds the MRI vertical wavelength and self-gravity has
+little effect.
+
+%We suggest that in
+%order for self-gravity to be significant,
+%We suggest that in
+%order for self-gravity to influence MRI,      
 %cite{kim03} and \cite{fromang04}
-
 %Numerical experiments
@@
 
-%GI/MRI: avoided crossings
+%GI/MRI: avoided crossings, relevance to simulations
+Curiously, when GI and MRI are simultaneously supported, we find
+unstable modes transition between MRI and GI. Such modes
+have comparable potential and magnetic energy perturbations. These
+transitions occur smoother with decreasing $\beta$
+(Fig. \ref{compare_growth3}) or increasing $k_x$
+(Fig. \ref{compare_growth3_Q01d2}). The latter implies that, in order
+to capture magnetic-gravitational interactions represented by these
+intermediate modes, non-linear simulations must resolve radial scales
+smaller than the most unstable GI mode. For example, 
+Fig. \ref{compare_growth3_Q01d2} suggest radial scales down to $\sim H/2$
+should be resolved.
+
+%This is
+%consistent with the MRI only affected by self-gravity for strong
+%fields 
+%mode classificication ambgious 
+%This is consistent /expected
+%numerical simulations should measure a continous range of modes
+%This is
+%consistent self-gravity only affecting the MRI when it is weak.  
+%mag field only affect GI through pressure
+%mri influenced by GI at large kx - > small scale -> need resolution
 
@@
 
-%\subsection{Implications for non-linear simulations}
-%what scaes? resolution
-%why don't formang  and kim/ostriker see any difference?
+We examined the effect of an additional azimuthal field, while
+keeping the vertical field at fixed strength. In this case, we also
+relaxed the equitorial symmetry condition applied previously and
+considered the full disk column. Self-gravity affects the MRI
+differently depending on the midplane symmetry of the MRI density
+perturbation. Self-gravity destabilizes MRI modes where the magnetic
+energy reaches a minimum at $z=0$, which have a symmetric
+density perturbation in the limit $B_y\to0$. However, 
+self-gravity stabilizes MRI modes where the magnetic energy maximizes
+at $z=0$. These modes have an anti-symmetric density
+perturbation in the limit $B_y\to0$. This stabilization effect is
+stronger for increasing $B_y$. Previous linear calculations indicate
+increased compressibility associated with a toroidal field stabilizes
+the MRI \citep{kim03}. We conjecture that self-gravity can
+further enhance this effect.   
+%can't do WKB (can't tell symmetries)
+
+%lesure 14 (strong azimuthal field)
+%implications non-linear MRI simulations with a strong toroidal field
+%would over-estimate growth rates (and perhaps turbulence strengths)
+%if SG is not accounted for. not applicable to self-gravitating disks     
+%previous linear calcs show compressibility stabilizes MRI 
+%additional pressure stabilizing MRI
 
@@
 We discuss below two major extensions to our linear model that
-should be undertaken, before performing non-linear simulations of
+should be undertaken, before embarking on non-linear simulations of
 magnetized, self-gravitating disks, which is our eventual goal.
@@
 disks can be unstable to non-axisymmetric perturbations but stable
-against axisymmetric perturbations, although the threshold Toomre
-parameter for both is still of order unity. %papaloizou/lin,
-                                %sajvoini??
+against axisymmetric perturbations \citep{papaloizou89,papaloizou91}.

### 2014-04-03 paper/results_azi.tex
first rough draft done

@@
 minimum; and the lower MRI mode has $E_m(0)$ being a maximum. 
+%We
+%believe this interchange is enabled by the presence of GI. 
 %interchange possible because of GI branch
@@
 Notice in Fig. \ref{compare_growth3_tilted_resis} the disappearence of
-magnetic modes as $B_y$ is increased. For $B_y=2B_z$, MRI and GI
-operate at distinct scales. This implies that regions unstable to GI
-cannot develop MRI.
+magnetic modes between $0.8\lesssim k_xH\lesssim 1.5$ as $B_y$ is
+increased. For $B_y=2B_z$, MRI and GI operate at distinct radial
+scales. This implies that perturbations unstable to GI cannot develop
+MRI.
 
@@
   \caption{Overstable MRI mode in an isothermal resistive disk with
-    $Q=0.18 (Q_\mathrm{2D} = 0.67)$, $\Lambda_0=0.1$ and $\beta=100$. 
+    $Q=0.18$ ($Q_\mathrm{2D} = 0.67$), $\Lambda_0=0.1$ and $\beta=100$. 
     The mode has a real frequency $\omega = 0.059\Omega$, or

### 2014-04-03 paper/summary.tex
first rough draft done

@@
 
-Physically, we expect MRI to only interact with self-gravity when
+Physically, we expect MRI to interact with self-gravity when
 their spatial scales are similar. Because self-gravity acts globally in the
@@
 scale height exceeds the MRI vertical wavelength and self-gravity has
-little effect.
+little impact.
 
@@
 
+%not looked at how MRI affects GI -> it doesn't because GI isn't
+%concerned with shear. only magnetic pressure. maybe a plot of GI with
+%magnetic perturbations?
+
 %spiral arms, GI angular mometum distribution 
-\emph{Beyond axisymmetry.} It is well known that self-gravitating
-disks can be unstable to non-axisymmetric perturbations but stable
-against axisymmetric perturbations \citep{papaloizou89,papaloizou91}.
+%global modes 
+%compressible term in y-induction equation 
+\emph{Beyond axisymmetry.} Axisymmetric perturbations, as we have
+assumed, preclude gravitational torques \citep{lynden-bell72}. 
+%In 
+%order to investigate how angular momentum transport 
+The local non-axisymmetric stability of magnetized self-gravitating
+thin disks has been considered by several authors
+\citep{elmegreen87,gammie96,fan97,kim01}. However, two-dimensional models
+exclude the MRI. It will be neccessary to generalize these studies to
+3D in order to investigate the impact of the MRI on angular momentum
+transport by gravitational instability. Furthermore, self-gravitating
+disks can develop global spiral instabilities while stable against local 
+axisymmetric perturbations \citep{papaloizou89,papaloizou91}. Global
+non-axisymmetric linear models will be desirable to support non-linar
+simulations of this kind \citep{fromang04c,fromang05}.
+
+%Our linear problem assumes a non-zero
+%vertical field is always present. It is of interest to 
+%consider the opposite limit of purely toroidal fields. Of course, one
+%must consider the non-axisymmetric MRI \cite{}
+%The local non-axisymmetric stability of magnetized, self-gravitating
+%disks has been treated in the shearing sheet framework \citep{} 
+%framework by decomposing
+%perturbations into shearing waves %ref.  
+%
+%global angular momentum exchange
+%It is necessary to generalize the current models to non-axisymmetric
+%perturbations. 
+%It would be interesting to examine global spiral instabilities in a 
+%magnetized disk. 
+%In particular, to examine global spiral instabilities
+%in a magnetized disk. 
+%fu lai

### 2014-04-03 paper/linear.tex
tex updates

@@
 
-\subsection{Domain and boundary conditions}
+\subsection{Domain and boundary conditions}\label{domain}
 
-\subsubsection{Case 1: vertical field}%because SG can only destabilize
-                                %symmetric density perts
+\subsubsection{Case 1: vertical field}
 We impose $\bm{U}$ to be an even function of $z$, so that
-$d\bm{U}/dz=0$ at $z=0$. This reduces the computational domain to
-$z\in[0,Z_s]$ and permits higher numerical resolution. At the disk
-surface we set %$\dbx=\dby=\dvz=0$ 
+$d\bm{U}/dz=0$ at $z=0$. This permits higher numerical resolution by
+reducing the computational domain to $z\in~[0,Z_s]$.  
+At the upper disk boundary $z=Z_s$ we set 
 \begin{align}
@@
 
-%We adopt two kinds of vertical boundary conditions depending on the
-%field geometry. 
-%In both case 1 and case 2 the potential boundary
-%condition is 
-%\begin{align}
-%  \dphi^\prime(\zmax) \pm k_x\dphi(\zmax) = 0
-%\end{align}
-
-%At the vertical disk boundaries $z=\pm\zmax$ we assume  
-%$\dvz=\dbx=\dby=0$, and
-
-%for the potential perturbation \citep[see][]{goldreich65a}, where the
-%upper (lower) sign applies to the upper (lower) disk surface. 
-%For Case 1 (purely vertical field) we additionally assume $\bm{U}$ is
-%an even function of $z$, so that $d\bm{U}/dz=0$ 
-%at $z=0$.
-
-%This implies that $\dbx=\dby=0$ at the midplane, consistent
-%with a highly resistive dead zone.
-
 \subsection{Numerical procedure}
@@
 and  $T_l$ is a Chebyshev polynomial of the first kind of order $l$
-\citep{stegun65}. Note that for case 1 the mid-plane symmetry condition
+\citep{stegun65}. Note that for case 1 the midplane symmetry condition
 is taken care of by the choice of basis functions.
-%Note that in case 1
-%The number of basis functions is
-%$N_z$ and $l_\mathrm{max}=2(N_Z-1)$ is the highest polynomial order. 
-%For a vertical field we only use even Chebyshev polynomials 
-%to satisfy the symmetry condition at $z=0$, so in case 1 we only
-%consider $z\geq 0$.
 
@@
 the set of linear equations to be satisfied at $N_z$ collocation
-points along the vertical direction. We choose these points as the
+points along the vertical direction, here chosen to be the 
 extrema of $T_{l_\mathrm{max}}$ plus end points, where 
@@
 \end{align} 
-where $\hat{c}_s=c_s/\csmid$. We also non-dimensionalize other
+where $\hat{c}_s=c_s/\csmid$. We also non-dimensionalize 
 background quantities, i.e. $\hat{v}_A=v_A/\csmid$,
@@
 $z$.
-%The colorbar quantifies the importance of
-%self-gravity with
 
@@
 $\tau\ll1$ are dominated by magnetic perturbations (MRI).
-
-%and we then normalize $E_i$ by
-%$\mathrm{max}(E)$.

### 2014-04-03 paper/paper.tex
tex updates

@@
 them, the magneto-rotational instability \citep[MRI,][]{chandrasekhar61,
-  balbus91,balbus98} %balbus %chandrasekhar ref
+  balbus91,balbus98} 
 and gravitational instability \citep[GI, ][]{toomre64,goldreich65a,goldreich65b} 
@@
   and references therein]{balbus99,armitage11,turner14}.   
-%gammie, lin/pringle, paczynski 
-%relevance of MRI and GI to planet formation 
 They are also relevant to planet formation theory. For example,
@@
 \citep{boss97,boss98,gammie01,voro13,helled14}.
-%gammie,
-%boss %some review articles (PP6)  
-%shakura73, pacz78, lin87,
 
@@
 and GI can be assessed independently. 
-%That is, MRI is unaffected by
-%disk self-gravity; and GI is unaffected by weak magnetization. 
 Circumstellar disk models that explicitly combine the equations of magneto-hydrodynamics
 and self-gravity have been limited to a few early simulations 
-\citep{fromang04,fromang04b,fromang05}. It will be neccessary to %no
-%major difference % kim/ostriker/stone
+\citep{fromang04,fromang04b,fromang05}. It will be neccessary to 
 revisit and extend these pioneering calculations to fully explore the
@@
 of the stability properties of such systems.
-%non-ideal effects important
-
-%why is GI and MRI independent 
-%compressibility is not fundamental to MRI 
-%horizontal wavenumber (gammie/balbus) 
-%The effect of self-gravity on the linear MRI has not been considered, 
-%since fluid compressibility is not fundamental to the MRI.
 
@@
 with radial lengthscales on the order of the disk scale height may be
-subject to self-gravity. It is of interest to generalize the MRI with
+subject to self-gravity. It is therefore of interest to generalize the MRI with
 non-zero radial waveumbers to massive disks.
 
-%Simiarly, studies of GI in accretion disks have mostly neglected
-%magnetization \citep{mamat10,kim12}. 
-%particularly in the context of
-
-%Studies of GI in accretion disks do not often include magnetization
-%\citep{mamat10,kim12}.
-
 The effect of a magnetic field on the GI of rotating disks has been
-considered by some authors.  
-% \cite{lynden-bell66} found that a toroidal
-%field can destabilize non-axisymmetric disturbances
-Recently, \cite{lizano10} generalized the classic Toomre stability
+considered recently by \cite{lizano10}, who generalized the Toomre stability
 criterion for razor-thin disks to include a vertical field. For
@@
 are subject to the MRI if magnetized.
-%how is the dispersion relation for GI affected by the presence of MRI?
-%It is interesting to ask whether
-%or not the presence of MRI affects GI. 
-%
-%It is desirable to clarify
-%whether or not the presence of MRI modes affect the dispersion
-%relation for GI.
 
-%although MRI and GI are physically distinct 
-%explicitly show
-
-This paper marks the beginning of our study of magnetized,
-self-gravitating PPDs. We start from linear calculations. 
-Linear stability analyses have the advantage that a wide range of
-parameters can be studied with negligible computational cost. This
-allows one to indentify conditions, if any, under which MRI and GI
-cannot be consdered independent. It is also important to have such 
-calculations at hand to help benchmark and guide non-linear
+This work marks the beginning of our study of magnetized,
+self-gravitating PPDs. We start from linear calculations, which  
+have the advantage that a wide range of parameters can be studied at
+negligible computational cost. This  allows us to indentify
+conditions, if any, under which MRI and GI 
+cannot be considered independent. It is also important to have such 
+calculations to benchmark and guide future non-linear
 simulations.
-
-%This paper marks the beginning of our study of magnetized,
-%self-gravitating PPDs and is organized as follows.
 
@@
 The linear problem is formulated in \S\ref{linear}. The impact
-of self-gravity on the MRI is discussed in
-\S\ref{result1}; gravitationally unstable disks are considered in
-\S\ref{result2} and equilibra with an azimuthal field is explored in
-\S\ref{result3}. We summarize results in \S\ref{summary}, including a
-discussion of caveats and future directions.
+of self-gravity on the MRI with a vertical field is discussed in
+\S\ref{result1}, gravitationally unstable disks are considered in
+\S\ref{result2}, and equilibra including an azimuthal field is explored in
+\S\ref{result3}. We summarize results in \S\ref{summary} with a 
+discussion of important extensions to our current models.
 

### 2014-04-03 paper/results.tex
tex updates

@@
   background equilibrium}
-Here we use polytropic disks, so that the disk thickness is
-well-defined. The upper disk boundary is set to 
-$Z_s=0.99H$. We fix $\beta=100$ and $k_xH=0.1$ unless otherwise
-stated.
-
-%kx -> 0 so incompressible
+Here we use polytropic disks, which have a well-defined disk
+thickness. The upper disk boundary is set to $Z_s=0.99H$. We fix
+$\beta=100$ and $k_xH=0.1$ unless otherwise stated.
 
@@
   resistivity}   
-%We calculate the MRI as a function of $Q$ with uniform
-%resistivity ($A=1$) for mid-plane Elsasser numbers $\Lambda_0\in[0.3,10]$.    
 Fig. \ref{compare_growth_poly_uniresis} plots MRI growth rates as a
 function of $Q$ and $\Lambda_0$. The resistivity is uniform ($A=1$). 
-%As expected growth rates generally decrease with decreasing 
-%$\Lambda_0$ \citep{jin96}. 
-In the limit of ideal MHD ($\Lambda_0>1$), there is negligible
-dependence on $Q$. In the resistive limit 
+For ideal MHD ($\Lambda_0>1$), there is negligible dependence on
+$Q$. In the resistive limit   
 ($\Lambda_0<1$), however, growth rates decrease noticeably for $Q<0.5$
-($Q_\mathrm{2D}\lesssim 1.5$). This shows 
-show that disk self-gravity affects the MRI through the background
-equilibrium, even though density and potential perturbations are  
-unimportant (i.e. the linear response is non-self-gravitating).
+($Q_\mathrm{2D}\lesssim 1.5$).  Since we find density and potential
+perturbations to be negligible (i.e. the linear response is
+non-self-gravitating), this shows that disk self-gravity can affect the
+MRI through the background equilibrium.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_growth_poly_uniresis2}
-  \caption{MRI growth rates as a function of $Q$ for mid-planee 
-    Elsasser numbers $\Lambda_0=0.3$ (solid), $0.5$ (dotted), $1.0$
-    (dashed) and $5.0$ (dash-dot).  
+  \caption{MRI growth rates as a function of $Q$ and midplane 
+    Elsasser numbers, in polytropic disks with $\beta=100$ and uniform
+    resistivity.  
     \label{compare_growth_poly_uniresis}}
@@
 
-Since $\hat{\rho}$ is weakly dependent
+Because $\hat{\rho}$ is weakly dependent
 on $Q$ (Fig. \ref{eqm_den}), self-gravity only affects the
@@
 This implies that sufficiently strong self-gravity can stabilize the
-MRI by making $ 2H<\lambda $.  %these perts are not gravitationally
-                               %unstable
+MRI by making $ 2H<\lambda $.
 
@@
 for the values of $Q$ considered, so self-gravity does not affect
-growth rates significantly. However, the wavelength of the
-instability, in units of $H$, increases. This is shown in
+growth rates significantly. However, the ratio $\lambda/2H$ does
+increase with stronger self-gravity. Consequently, the wavelength of
+the instability, in units of $H$, increases. This is shown in
 Fig. \ref{compare_result_lambda10} which plots the magnetic energies
@@
 
-%Weak variations in growth rates in the ideal limit, seen in 
-%Fig. \ref{compare_growth_poly_uniresis}, were found to be associated
-%with changes in the mode character: $\gamma$ decreases
-%slightly when the mode begins to acquire an additional vertical node as $Q$ is
-%increased.
-
-%{\bf DISCUSS STRESS} Although for ideal MHD self-gravity makes no
-%difference to the MRI growth rate, we find the associated stress does
-%change with $Q$.
-
 \begin{figure}
   \includegraphics[width=\linewidth]{figures/compare_result_lambda10}
-  \caption{Magnetic energies as a function of height, in the limit of
-    ideal MHD, for various strengths of self-gravity.   
+  \caption{MRI magnetic energies in ideal polytropic disks
+    for different strengths of self-gravity.    
     \label{compare_result_lambda10}}
@@
 significantly reduced. This is seen for $Q=0.2$ where $\lambda \geq 2H$ throughout
-the disk. (The instability is not completely surpressed since
-Eq. \ref{sano_crit} is only an approximation.) Although the function
-$f(Q)$ does not change significantly for the range of $Q$ considered, the
-dependence of $\lambda_\mathrm{resis}$ on $f(Q)$ is amplified by the
-denominator $\Lambda_0<1$ in the resistive case. For resistive cases
-we find no nodes in the magnetic energy $E_m$ except at end points,
-i.e. only the longest wavelength survives.
+the disk. (The instability is not surpressed since
+Eq. \ref{lambda_ideal}---\ref{lambda_resis} is only exact for
+unstratified disks.) Although the function 
+$f(Q)$ does not change significantly for the range of $Q$ considered,
+the dependence of $\lambda_\mathrm{resis}$ on $f(Q)$ is amplified by the
+denominator $\Lambda_0<1$ in the resistive case. Modes in
+Fig. \ref{lambda_poly_resis} have no nodes in the magnetic energy
+$E_m$ except at $z\simeq0,\,H$, i.e. only the longest wavelength survives
+with large resistivity. 
+%self-gravity tips it over
 
@@
   resistivity} 
-Here we consider disks with mid-plane Elsasser number $\Lambda_0=0.1$
+Here we consider disks with midplane Elsasser number $\Lambda_0=0.1$
 and a variable resistivity profile with
@@
 limit ($Q=4$), the MRI is effectively surpressed for
-$z\lesssim0.5H$. This is consistent with layered accretion. However,
-in the massive disk ($Q=0.2$) the mode occupies a wider vertical
+$z\lesssim0.5H$. This is consistent with the picture of layered
+accretion proposed for non-self-gravitating disks \citep{gammie96,fleming03}. 
+However,in the massive disk ($Q=0.2$) the mode occupies a wider vertical
 extent because its wavelength (in units of $H$) is larger. This
-suggests that in massive disks the MRI is not well localized to the
-active layer.
-
-%MRI penetrates 
-%d is not as well localized
+suggests that in massive disks, the MRI is not well localized to a
+sub-layer within the height, even when the resisitvity has a layered
+structure.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_results_poly_layer_amp100}
-  \caption{Magnetic energies as a function of height, for layered polytropic disks with
-    such that the conductivity increases by a 
-    factor of $10^2$ in going from the mid-plane to the upper disk
+  \caption{Magnetic energies as a function of height, for polytropic disks
+    in which the conductivity increases by a
+    factor $A=10^2$ in going from the midplane to the upper disk
     boundary. The vertical lines indicate $\Lambda=1$ for each value 
@@
 direction. We find a similar result in the horizontal direction. 
-In Fig. \ref{compare_growth_poly_kx} we plot MRI growth rates as a
+Fig. \ref{compare_growth_poly_kx} show MRI growth rates as a
 function of $k_x$ for a range of $Q$ values. Increasing self-gravity
-decreases the cut-off horizontal wavenumber for the MRI. We checked
+decreases the cut-off radial wavenumber for the MRI. We checked
 that these modes have negligible density perturbations. Then we can 
@@
 approach, this demonstrates that with increasing self-gravity
-(increasing $f$), we expect that MRI modes with small radial
+(increasing $f$), we expect MRI modes with small radial
 lengthscales to be suppressed.
-
-%, implying that
-%small-scale MRI does not operate in massive disks.
 
@@
 
-%these modes are still non-sg
-%use beta = 40 so only a small range of kx need to be simulated to
-%demonstrate this effect
-%{\bf cut-off $k_xH$ decreases with increasing SG -> small-scale MRI
-%  stabilized by self-gravity}
-
 \subsection{Influence of self-gravity on the MRI through the linear
@@
 Our goal here is to examine whether or not self-gravity can amplify
-the density perturbations caused by the MRI. We compute unstable modes
+the density perturbations associated with the MRI. We compute unstable modes
 in a massive isothermal disk with $Q=0.2$ (corresponding to
@@
 We repeat the above calculation for resistive disks, but fix
-$\beta=100$ and vary the mid-plane Elasser number $\Lambda_0$. Growth
+$\beta=100$ and vary the midplane Elasser number $\Lambda_0$. Growth
 rates are shown in Fig. \ref{gravity_energy_resis}. 
@@
     disk with $Q=0.2$ ($Q_\mathrm{2D}=0.72$) at fixed $\beta=100$, 
-    for a range of mid-plane Elasser numbers. The resistivity is
+    for a range of midplane Elasser numbers. The resistivity is
     uniform. The colorbar measures the importance of self-gravity by $\tau$. 
@@
 % field, i
- Self-gravity plays does not affect the MRI in this regime.
+ Self-gravity does not affect the MRI in this regime.
 

### 2014-04-03 paper/setup.tex
tex updates

@@
 $\Omega_0=\Omega(r_0,0)$ about the star, where $r\Omega^2 =
-\p\Phi_*/\p r$. We also define $S=-r\p\Omega/\p r$ as the local shear
+\p\Phi_*/\p r$. We also define $S\equiv-r\p\Omega/\p r$ as the local shear
 rate and $\Omega_z^2\equiv\p^2\Phi_*/\p z^2$ as the square of the
@@
 height.
-%The %resistivity is allowed to be non-uniform, $\eta=eta(z)$.
-
-%is the gas pressure given by a
-%barotropic equation of state chosen later.  
-%no curvature effects - magnetic tension neglected (limits strength of
-%toroidal field that can be considered
 
@@
 is the effective external potential (central plus centrifugal) in the
-shearing box approximation, where $S_0\equiv S(r_0,0)$ and
+shearing box approximation, where $S_0\equiv ~S(r_0,0)$ and
 $\Omega_{z0}\equiv\Omega_z(r_0,0)$; 
-%% \begin{align}
-%%   S\equiv -r\frac{\p\Omega}{\p r}, \quad\quad \Omega_{z0}^2\equiv
-%%   \frac{\p^2\Phi_*}{\p z^2} 
-%% \end{align}
-%is the local shear rate and vertical frequency respectively; 
 and the gas potential $\Phi_d$ statisfies Poisson's equation
@@
 drop the subscript $0$ on the frequencies.
-
-%We consider two cases
-%\begin{enumerate}
-%\item Purely vertical field with constant or variable
-%  resistivity, so that $\epsilon = 0$ and $\eta=\eta(z)$. 
-%\item Tilted field with uniform resistivity
-%  so that $\epsilon \neq 0$ and $\eta=\rm{constant}$.   
-%\end{enumerate}
 
@@
 $\kappa\equiv\sqrt{2\Omega(2\Omega-S)}=\Omega=\Omega_z$.  
-%$\Omega = \Omega_z.$ 
 We assume a thin disk and neglect the radial component of the
-self-gravitational force in 
-the unperturbed disk.
+self-gravitational force in the unperturbed disk.
 
@@
 \end{inparaenum}
-where $\rho_0\equiv\rho(0)$ is the mid-plane density.
+where $\rho_0\equiv\rho(0)$ is the midplane density.
 The sound speed $c_s\equiv\sqrt{dP/d\rho}$ 
 so that $c_{s0}$ is the global sound speed in the isothermal disk, and
-is the mid-plane sound speed in the polytropic disk. For the polytropic
+is the midplane sound speed in the polytropic disk. For the polytropic
 disk the disk thickness $H$ is such that $\rho(H)=0$. Since the
@@
 
-%and $\rho_0\equiv\rho(0)$
-
 We solve for $\hat{\rho}\equiv\rho/\rho_0$ with 
-%$\rho_0\equiv\rho(0)$ is the mid-plane density.
-%is parameterized via the
-%Toomre $Q$ parameter such that $\rho_0=\Omega^2/4\pi G Q$. 
 boundary conditions $\hat{\rho}=1$ and $d\hat{\rho}/dz=0$  at $z=0$. This
@@
 
-%As the strength of self-gravity increases, the
-%disk becomes thinner ($H$ decreases)
-
-%The isothermal disk shows that self-gravity
-%compresses the disk and reduces the true pressure scale-height. The
-%polytropic disk shows little difference between different strengths of
-%self-gravity because the vertical co-ordiante is already normalized by
-%the disk thickness.
-
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.5cm 0cm 0cm]{figures/compare_iso_density}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0.9cm]{figures/compare_poly_density}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.5cm 0cm
+    0cm]{figures/compare_iso_density} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+    0.9cm]{figures/compare_poly_density} 
   \caption{Equilibrium density field from solving Eq. \ref{eqm_eqns1}
@@
     i.e. $H=H(Q)$ and is an increasing function of $Q$. 
-%Note that the vertical co-ordinate is
-%    normalized differently for convenience. In the isothermal disk,
-%    $H$ is the characteristic scale-height in the
-%    non-self-gravitating limit; in the polytropic disk $H$ is the
-%    disk thickness. 
     \label{eqm_den}}
@@
 
-%\subsection{}
-
 \subsection{Resistivity profile}\label{resis_profile}
-We consider constant resistivity or a 
+We adopt constant resistivity or a 
 resistivity prescription such that $\eta(z)$ increases away from the
-mid-plane. In the latter case, we follow \cite{fleming03} and adopt the
+midplane. In the latter case, we follow \cite{fleming03} and use 
 resistivity profile 
@@
 and we define $\eta_0/\eta(\infty)\equiv A$ as the conductivity 
-boost factor from the mid-plane to the disk surface. We remark that
+boost factor from the midplane to the disk surface. We remark that
 once $\rho$ and $d\rho/dz$ are obtained from
@@
 
-We use the Elsasser number $\Lambda$ as a non-dimensional measure of
+We use the Elsasser number $\Lambda$ a non-dimensional measure of
 conductivity,
@@
 where $v_A \equiv B_z/\sqrt{\mu_0\rho}$ is the vertical Alfven speed. 
-%Example
-%profiles of $\Lambda$ are shown in Fig. \ref{eqm_resis} for the
-%isothermal disk. 
 Because of the density stratification, the Elsasser number 
@@
 considered ideal where $\Lambda \gtrsim 1$.
-%Since $\hat{\rho}(z/H)$ is weakly dependent on  self-gravity, the
-%ideal region only changes slightly as $Q$ is decreased.
 
-%A similar plot is obtained for polytropic disks.
-
-%The conductivity is boosted by a factor $A=10$ from the midplane to
-%the upper disk boundary. The figure compares a strongly self-gravitating case
-%($Q=0.2$) and the non-self-gravitating limit ($Q=10$).
-
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figures/elsasser_iso}
-%  \caption{Elsasser number as prescribed in \S\ref{resis_profile} for
-%    isothermal disks with $Q=0.2$ (solid) 
-%    and $Q=10$ (dotted). The vertical line
-%    indicates $\Lambda=1$ in each case. \label{eqm_resis}}
-%\end{figure}
-
-\subsection{System parameters}
-The main parameters describing the resistive, self-gravitating
-shearing box is as follows. The strength of self-gravity is
+\subsection{Disk parameters}
+%The main parameters describing the resistive, self-gravitating
+%shearing box is as follows. 
+The strength of self-gravity is
 parametrized by 
@@
 \end{align}
-\citep{mamat10}, which is used to set the mid-plane density $\rho_0$. 
-The relation between $Q$ and the Toomre parameter for gravitational
+\citep{mamat10}, which is used to set the midplane density $\rho_0$. 
+A relation between $Q$ and the Toomre parameter for gravitational
 instability of razor-thin disks, $Q_\mathrm{2D}$, is described in
@@
 where $v_{A0}$ is the midplane Alfven speed. Note that we use 
-the vertical field for this definition throughout the paper. The
-strength of conductivity is measured by the midplane Elsasser number  
+the vertical field for this definition throughout this paper.
+
+The strength of conductivity is measured by the midplane Elsasser
+number   
 \begin{align}
-  \Lambda_0 \equiv\Lambda(0) =  \frac{v_{A0}^2}{\eta_0\Omega}. %= \frac{f^2 R_m}{\beta}, 
+  \Lambda_0 \equiv\Lambda(0) =  \frac{v_{A0}^2}{\eta_0\Omega}. 
 \end{align}
-%where the third equality defines the magnetic Reynolds number
-%$R_m=H^2\Omega/\eta_0$ and
-%where $f\equiv \csmid/H\Omega$ is a numerical
-%factor of order unity. %measures thickness
-
-%The Elsasser number as a function of height is $\Lambda(z)\equiv
-%v_A^2/\eta\Omega$.
+For non-uniform resistivity we also specify $A > 1$.

### 2014-04-05 paper/results.tex
tex updates

@@
 We first consider ideal MHD by adopting a uniform 
-resistivity with $\Lambda_0=100$. %Non-zero density perturbations
-%require $k_x\neq0$ so we consider linear modes as a function of $k_x$.   
+resistivity with $\Lambda_0=100$.  
 Fig. \ref{gravity_energy} plots MRI growth rates as a function of
 $k_x$ for several values of $\beta$. The curves are color-coded 
-according to $\tau$. The potential perturbation is
-negligible for all cases when $k_xH\lesssim 0.5$, since the MRI
-becomes incompressible as $k_x\to 0$.
+according $\tau$. (Recall $\tau\to1$ implies self-gravity dominates
+over magnetic perturbations, and $\tau\to0$ is the opposite limit.)  
+The potential perturbation is negligible for all cases when
+$k_xH\lesssim 0.5$, since the MRI becomes incompressible as $k_x\to
+0$.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_energy_ideal}
-  \caption{Growth rates of MRI modes in an isothermal self-gravitating
-    disk with $Q=0.2$ ($Q_\mathrm{2D}=0.72$) in the limit of ideal MHD
-    ($\Lambda_0=10^2$), for a range of field strengths $\beta$. The
+  \caption{Growth rates of MRI modes in isothermal self-gravitating
+    disks with $Q=0.2$ ($Q_\mathrm{2D}=0.72$) in the limit of ideal MHD
+    ($\Lambda_0=10^2$, $A=1$), for a range of field strengths $\beta$. The
     colorbar measures the importance of self-gravity by $\tau$.  
@@
 For $\beta\gg 1$, i.e. a weak field, density perturbations are
-negligible and the standard MRI operates. However, as $\beta$ is
+negligible and the incompressible MRI operates. However, as $\beta$ is
 lowered and the MRI growth rate reduced, we find non-negligible
-potential perturbation for $k_xH=O(1)$. This suggets that in a
+potential perturbation for $k_xH=O(1)$. This suggests that in a
 strongly magnetized disk that still permits the MRI, the associated
@@
 self-gravitating. %but formall stable
-
-%critical beta, below which MRI is surpressed, increases with the
-%strength of self-gravity.
-
-%Fig. \ref{mri_massive} shows an example of a MRI mode with
-%non-negligible density perturbation. The growth rate $\gamma\simeq
-%0.4\Omega$ is comparable to that of the most unstable mode
-%($\gamma\simeq 0.6\Omega$ for $k_x\simeq0$). 
-%We therefore expect significant density perturbations to grow
-%dynamically, even though the disk is expected to be gravitationally
-%stable.
-
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figures/result_ideal_sg}
-%  \caption{Example of a linear MRI mode with non-negligible density and
-%    potential perturbation, in a strongly magnetized massive disk
-%    ($\beta=10$ and $Q=0.2$) in the limit of ideal MHD 
-%    ($\Lambda_0=10^2$).   
-%    \label{mri_massive}}
-%\end{figure}
 
@@
 rates are shown in Fig. \ref{gravity_energy_resis}. 
-Interestingly, in the highly resistive case
-the magnetic and gravitational energies can be comparable. For 
-$\Lambda_0=0.1$ and $k_xH\simeq1.3$, $\tau\sim 0.3$ which corresponds
-to $\avg{E_g}\sim 0.5\avg{E_m}$.  
-Fig. \ref{mri_massive_cowling}, we compare the magnetic
-energy of this mode to that computed in in the \emph{Cowling  
+Interestingly, the highly resistive case $\Lambda_0=0.1$ has
+comparable magnetic and gravitational energies: at $k_xH\simeq1.3$ we
+find $\tau\sim 0.3$, which corresponds to $\avg{E_g}\sim 0.5\avg{E_m}$.  
+Fig. \ref{mri_massive_cowling} compares the magnetic
+energy of this mode to that computed in the \emph{Cowling  
   approximation}, where the Poisson equation is ignored in the 
@@
 (formally letting $Q\to\infty$ in Eq. \ref{lin_poisson}).  
-The growth rate increases slightly when self-gravity is included in the linear
-response, since self-gravity is usually destabilizing. However, the
-similarity between $\gamma$ and $E_m$ indicates that the instability
-in the self-gravitating calculation is fundamentally the MRI.
+The growth rate increases when self-gravity is included in the linear
+response, since self-gravity is usually destabilizing. However, 
+$\gamma$ and $E_m$ are very similar, indicating that the instability
+in the self-gravitating calculation is fundamentally still the MRI.
 
@@
 MRI mode discussed above. 
-%with $k_xH\sim1.3$ in the $\Lambda_0=0.1$  
-%disk. 
 The gravitational energy exceeds the 
-magnetic energy near the midplane ($z\lesssim0.2H$). The growth rate $\gamma=0.25\Omega$
-is not much smaller than that of the most unstable mode
-($\gamma=0.36\Omega$ for $k_xH=0.1$), so significant density
-perturbations are expected to grow on dynamical timescales for this
-system, even though GI should not develop.
+magnetic energy near the midplane ($z\lesssim0.2H$). The growth rate
+$\gamma=0.25\Omega$ is not much smaller than that of the most unstable
+mode ($\gamma=0.36\Omega$ for $k_xH=0.1$), so significant density
+perturbations will grow on dynamical timescales for this
+system, even though GI is not expected.
 
@@
   \caption{Example of a resistive MRI mode with significant
-    gravitational potential perturbation.  The vertical line
+    gravitational potential perturbation. The disk is isothermal. 
+    The vertical line
     inidicates $\Lambda=1$.  
@@
 where $k^2 = k_z^2 + k_x^2$. We imagine an iterative procedure to
-solve the above equations, starting from the Cowling approximation. 
-%defined to be the
-%solution to Eq. \ref{simp1}---\ref{simp3} with potential perturbation
-%set to zero. 
-This is the standard MRI and we denote the solution as
-$\dvx^{(0)}$, $\dvy^{(0)}$ and $W^{(0)}$.
+solve the above equations, starting from the Cowling approximation
+where $\delta\Phi=0$. This is the standard MRI and we denote the
+solution as $\dvx^{(0)}$, $\dvy^{(0)}$ and $W^{(0)}$. Eq. \ref{simp3} implies
+\begin{align}
+  W^{(0)} = \frac{c_s^2\sigma k_x \dvx^{(0)}}{c_s^2k_z^2 - \sigma^2}. 
+\end{align}
+We argue below that $c_s^2k_z^2\gg \sigma^2$ by taking $k_z\sim
+\Omega/v_A$. Then, recalling $W=c_s^2\delta\rho/\rho$, we can write
+\begin{align}
+  \frac{\delta\rho^{(0)}}{\rho} \sim
+  \frac{\sigma}{\Omega}\frac{1}{\beta}\left[\frac{k_x\dvx^{(0)}}{\Omega}\right].   
+\end{align}
+The MRI has, in general, a non-zero density perturbation. However, it
+is negligible for $k_x\to 0$ and/or a weak field ($\beta \gg 1$).
 
-We now include self-gravity. In general, the compressible MRI has
-$W^{(0)}\neq0$. The Poisson equation implies
-an associated potential perturbation,    
+% In general, the
+%compressible MRI has $W^{(0)}\neq0$, since
+
+We now include self-gravity.  The Poisson equation implies $W^{(0)}$
+has an associated potential perturbation,     
 \begin{align} 
@@
 \end{align}
-Physically, we expect $k^2\geq0$, so that a local positive (negative) density
+Physically, we expect $k^2\geq0$, so that a positive (negative) local density
 perturbation causes a negative (positive) local potential
@@
 \dvx^{(1)}$ and similarly for $\dvy $ and $ W$, we find
-%regard $\dphi$ to alter the Cowling solution, so
-%that 
-%Then 
+
 \begin{align}
@@
 Now, if the perturbations to the magnetic field remain
-unchanged, i.e. the mode remains close to the standard MRI, then
-$\dvx^{(1)} \sim 0$ and $\dvy^{(1)}\sim0$, so Eq. \ref{simp_pert2} is
+unchanged, i.e. the mode remains close to the standard MRI as observed
+in Fig. \ref{mri_massive_cowling}, then $\dvx^{(1)} \sim 0$ and
+$\dvy^{(1)}\sim0$, so Eq. \ref{simp_pert2} is 
 satisfied. Eq. \ref{simp_pert1} then require $\dphi +
@@
 \end{align}
-For the ideal MRI,  we assume $k_z\sim \Omega/v_A$. Then
+For the ideal MRI,  we take $k_z\sim \Omega/v_A$. Then
 $|\sigma^2/c_s^2k_z^2|\sim |\sigma^2/\Omega^2\beta|\ll1$ because
@@
 
-% not expect significant potential perturbations 
-%Conversely, for strong fields, only a single MRI wavelength occupies
-%the fluid column.
-
  The above arguments can be adapted to the resistive
@@
  terms appearing in Eq. \ref{simp1}---\ref{simp2} only involve the
- potential perturbation through $\w$. For the resistive MRI we assume
+ potential perturbation through $\w$. For the resistive MRI we take
  $k_z\sim v_A/\eta$ and $|\sigma|\sim v_A^2/\eta = \Lambda\Omega$
@@
 
- %mass contained in small wavelength is small
+%these arguments assume, of course, there is a non-zero density
+%perturbation in the first place for self-gravity to ehance.
 
- For weak fields in an ideal disk, the MRI has a small vertical
- wavelength, $\lambda \ll H$. The perturbed mass contained within
- $\sim \lambda$ is small and its potential
- perturbation is unimportant. Furthermore, considering the stratified
+ For weak fields in an ideal disk, the MRI has a vertical
+ wavelength $\lambda \ll H$. It will be almost incompressible so the 
+ `seed' density perturbation $W^{(0)}$ is small.  
+ The perturbed mass contained within 
+ $\sim \lambda$ is small and its potential is
+ unimportant. Furthermore, considering the stratified  
  disk, $\lambda\ll H$ imply rapid variations in the density
@@
  magnitude of the associated potential perturbation is small. 
-% because the gravitational potential is the integral of the density
-% field, i
  Self-gravity does not affect the MRI in this regime.
 
- %, i.e. internal cancellations would occur when such an integral is
- % performed.
+%the `seed' density pert is also small
+%weak amplification
 
@@
  perturbed mass across the disk height can contribute to a net potential
- perturbation. We therefore expect self-gravity to be important for
- the MRI when it is weak.
+ perturbation. We therefore expect a neccessary condition for
+ self-gravity to affect the MRI is for the latter to be weak. 
+%we are not considering GI yet
 
-%The perturbed mass
-% within the unstable mode corresponds to that in the within 
-%This can give rise to a net
-% potential perturbation  
- %Self-gravity can then be effective, since it
- % operates globally in the vertical direction. 
-%GI of 3D disks can be
- %analysed in vertically-integrated
- %systems 
-%coupling to SG 
-%That is when the MRI is weak
+%Our
+% interpretation above also assumes that the Cowling solution can
+% provide some `seed' density perturbation. 
+%Thus we also require the
+% MRI to be compressible. 
+%there has to be some  `seed' density perturbation provided by the
+%non-SG MRI

### 2014-04-05 paper/results_azi.tex
tex updates

@@
 self-gravity on MRI modes with different symmetries across the
-midplane.
+midplane. We use an isothermal disk throughout.
 
 \subsection{Ideal disks with MRI} 
-We consider an isothermal disk with $Q=0.2$ ($Q_\mathrm{2D}=~0.72$) and
+We consider disk with $Q=0.2$ ($Q_\mathrm{2D}=~0.72$) and
 $\beta=10$ in the limit of ideal MHD ($\Lambda_0=100$). Gravitational
 instability is not expected because Fig. \ref{compare_growth3} shows
-that even for $Q=0.18$, GI is surpressed for $\beta \lesssim 15$. %We
-%consider azimuthal field strengths with $\epsilon\in[0,3]$.
+that even for $Q=0.18$, GI is surpressed for $\beta \lesssim 15$.
 
@@
 midplane boundary conditions. 
-%The top panel are modes with even midplane symmetry: 
-%$W^\prime(0)= 0$ for $\epsilon\to0$, and the bottom panel are modes with odd midplane 
-%symmetry: $W(0) = 0$ for $\epsilon\to0$. 
 We also plot growth rates computed in the Cowling
@@
 modes are energetically dominated by self-gravity.
-
-%no GI for odd modes 
-%general stabilization of MRI
 
@@
     colorbar measuring the gravitational potential perturbation, while
-    the solid curves are computed from the Cowling approximation. The
-    top panel correspond to even modes and the
-    bottom panel correspond to odd modes, which have
-    $W^\prime(0)=0$ and $W(0)=0$, respectively, for $\epsilon\to0$.
+    the solid curves are computed from the Cowling approximation. 
+    For $B_y=0$, modes in top and bottom panels 
+    $W^\prime(0)=0$ and $W(0)=0$, respectively.
     \label{compare_growth3_tilted}}
 \end{figure}
-
-%We consider $\epsilon\equiv B_y/B_z \in[0,3]$. The
 
@@
 self-gravity destabilizes modes with  $k_xH\gtrsim O(1)$. Consequently, the
-cut-off wavenumber is larger when SG is included. %get smaller scale
-                                %MRI 
+cut-off wavenumber is larger when SG is included. 
 Destablization is most effective for purely vertical fields: with
@@
 symmetric density perturbations.  
-With increasing $\epsilon$, we find $W$ deviates from an even
-function. %,% $|W(0)|$ decreases and $\max{(|W|)}$ 
-%moves off of the mid-plane.    
-%density pert at the midplane goes down
-%We expect this to reduce the destabilization
-%effect of SG,  
+With increasing $B_y$, we find $W$ deviates from an even
+function. 
 Together with the increased total magnetic pressure with 
-$\epsilon$, destabilization by SG weakens. 
+$B_y$ (since $B_z$ is fixed), destabilization by SG weakens. 
 Thus, the Cowling approximation becomes increasingly good with stronger
 $B_y$ for these modes.
-
-%This destabilization effect weakens with
-%$B_y$ because the total magnetic pressure increases, which generally
-%opposes GI, although the modes are fundamentally associated with MRI.
 
@@
 self-gravity has no effect. When $B_y>0$, $W$ deviates from an
-odd function and the mid-plane density perturbation $|W(0)|$
+odd function and the midplane density perturbation $|W(0)|$
 increases. SG is stabilizing for these modes at all wavelengths, and
 is most effective at $k_xH =  O(1)$.
-%destabilizing effect is absent  
-%for $\epsilon=0,\, k_xH\simeq 1.2$, SG decreases the growth rate by $\sim
-%13\%$.
 
@@
 $k_xH=1.1$ with and without the Cowling approximation.
-%These differ  
-%from the even modes considered above in that the magnetic energy
-%maximize at $z=0$ (cf. Fig. \ref{mri_massive_resis}).  
 SG significantly enhances the midplane density perturbation, making the
 gravitational potential energy comparable to the magnetic energy, 
-which becomes more confined near the mid-plane.
+which becomes more confined near the midplane.
 
@@
 
-%strong field, compressibility 
 To interpret the above result for modes with magnetic energy
@@
 
-Compressibilty is enhanced by an azimuthal field, which has a
-stabilizing effect on the MRI \citep{kim00}. %because...?
+Compressibilty is enhanced by an azimuthal field, which is 
+stabilizing for the MRI \citep{kim00}. %because...?
 This effect is significant for $\epsilon=3$ because the azimuthal
@@
 compressibility, and therefore stabilization. We suspect this is
-overwhelmed by the destabilization effect of SG because the density
-perturbation has an anit-symmetric component.
-
-%compressibility -> pressure restoring force
-
-%The results for even and odd modes are do not contradict each
-%other. Self-gravity tends to be destabilizing, but this requires the
-%density perturbation to be symmetric, as is the case for the even
-%modes. 
-%For the MRI however, compressibility has a stabilizing effect
-
-%scale separation for odd modes? because modes with kx=1 will be
-%quenched
-
-% for $\beta=10$ and
-%$\epsilon=3$. 
-%because magnetic energy is concentrated near midplane    
-%magnetic pressure, need thermal pressure 
-%This
-%can subsequently be amplitfied by its self-gravity. 
-%That is, compressibility is enhanced by both
-%For $\epsilon=3$ and $\beta=10$ the azimuthal Alven speed is
-%sonic. Compressibility becomes important for strong toroidal
-%fields. If the system remains in approximately vertical hydrod
-
-%\subsection{Ideal disks with GI (TBC)}
+overwhelmed by the destabilization effect of SG, because the density
+perturbation has an anti-symmetric component.
 
 \subsection{Resistive disks with GI}
-Here we consider an isothermal, resistive disk with $Q=0.18$,
-$\Lambda_0=0.1$ and $\beta=100$. This system permits GI and MRI. 
-%GI is permitted, and has a similar
-%growth rate as the  
-%MRI. 
+Here we examine a resistive disk which permits MRI and GI by setting $Q=0.18$,
+$\Lambda_0=0.1$ and $\beta=100$. 
 Fig. \ref{compare_growth3_tilted_resis} show growth rates for
@@
 the GI modes have symmetric $W(z)$.
-
-%GI, and there are MRI modes available on that scale
 
@@
     ($Q_\mathrm{2D}=0.67$), $\beta=100$ and $\Lambda_0=0.1$. For
-    $B_y=0$, the upper MRI modes have anti-symmetric density
-    perturbations, $W(0) = 0$; while the lower MRI modes have symmetric
-    density perturbations, $W^\prime(0) = 0$. For $B_y/B_z=2$ the
+    $B_y=0$, the upper and lower MRI modes have anti-symmetric and
+    symmetric density perturbations, corresponding to $W(0)=0$ and
+    $W^\prime(0)=0$, respectively.  
+    %anti-symmetric density
+    %    perturbation $W(0) = 0$ and symmetric density perturbations,
+    %    $W^\prime(0) = 0$.  
+    For $B_y/B_z=2$ the
     overstable modes have non-zero real frequencies. 
@@
 the upper MRI mode transitions to GI, for which $E_m(0)$ is a
-minimum; and the lower MRI mode has $E_m(0)$ being a maximum. 
-%We
-%believe this interchange is enabled by the presence of GI. 
-%interchange possible because of GI branch
-
-%In both cases where the gravitational energy is significant, the
-%density perturbation is symmetric.
+minimum; and the lower MRI mode has $E_m(0)$ being a maximum. We find
+all perturbations with  $k_xH\gtrsim0.9$ have symmetric $W(z)$. 
+%small scale perts are all even
 
 Increasing the azimuthal field further to $B_y=2B_z$ we find
-overstable MRI modes \citep{gammie96} with non-negligible real
-frequencies. An example is shown in
+overstable MRI modes  with non-negligible real
+frequencies \citep{gammie96}. An example is shown in
 Fig. \ref{result_tilted_overstable}. Notice the density/potential
@@
 
-Notice in Fig. \ref{compare_growth3_tilted_resis} the disappearence of
+Notice also in Fig. \ref{compare_growth3_tilted_resis} the disappearence of
 magnetic modes between $0.8\lesssim k_xH\lesssim 1.5$ as $B_y$ is
-increased. For $B_y=2B_z$, MRI and GI operate at distinct radial
-scales. This implies that perturbations unstable to GI cannot develop
-MRI.
-
-%This may lead to 
-%the lower MRI mode transitions to the top MRI mode
+increased. For $B_y=2B_z$, MRI and GI are again independent, but
+because they operate at distinct radial scales. This implies that
+perturbations unstable to GI cannot develop MRI.
 
@@
 \end{figure}
-
-%\subsection{GI}

### 2014-04-05 paper/results_gi.tex
tex updates

@@
 pressure for sufficiently small $\beta$. The critical
-field strength increases with increasing self-gravity \citep{nakamura83}. 
+field strength for stabilizating GI increases with increasing
+self-gravity, consistent with \cite{nakamura83}.   
 For $Q=0.18$, GI is stabilized for $\beta \lesssim 15$. Nevertheless, 
 the MRI branch for $\beta < 15$ becomes self-gravitating, so that
-density perturbations will still develop, even though GI does not
+density perturbations still grow, even though GI does not
 formally operate.
@@
     0.5cm]{figures/compare_growth3_kx1_Q0d12.ps} 
-  \caption{Growth rates for modes with $k_xH=1$ in an isothermal disk with
-    $Q=0.18$ (top), $Q=0.14$ (middle) and $Q=0.14$ (bottom). The
+  \caption{Growth rates for modes with $k_xH=1$ in isothermal ideal
+    disks with $Q=0.18$ (top), $Q=0.14$ (middle) and $Q=0.14$ (bottom). The
     colorbar measures the importance of  
-    self-gravity with $\tau$.   
+    self-gravity by $\tau$.   
     \label{compare_growth3}}
@@
 The GI and MRI branches only interact when their growth rates
-are similar. This is seen in Fig. \ref{compare_growth3} for $Q=0.18$
-where the GI branch approaches an MRI branch at $\beta\simeq 25,\,
+are similar. This is seen in Fig. \ref{compare_growth3} for $Q=~0.18$
+where the GI branch approaches a MRI branch at $\beta\simeq 25,\,
 \gamma\simeq 0.2\Omega$. In fact, following the red curve to smaller
@@
 numerical search of eigenfrequencies.  
+%maybe ogilvie?
 
@@
 MRI to GI and vice versa, especially at low $\beta$.     
+%no unique transition
 
@@
 pure MRI and pure GI, respectively.  
-For $k_xH\lesssim 0.2$, growth rates are equal to
+For $k_xH\lesssim 0.7$, growth rates are equal to
 those on the pure MRI and pure GI branches. That is, MRI and GI
@@
 inclusion of magnetic field stabilizes high-$k_x$ pure GI. (Note also
-the slight decrease in the most unstable $k_x$.) %Since axisymmetric GI
-%is unaffected by shear, 
+the slight decrease in the most unstable $k_x$.) 
 This stabilization is due to magnetic pressure \citep{lizano10}, which
@@
 
-%3D geometry doesn't affect 
-%axisymmetric GI doesn't care about shear 
-%
-%two legs of GI are not symmetric 
-%MRI become self-gravitating
-%GI with increasing magnetization
-
 Comparing pure MRI (solid line) and the magnetically-dominated
 portions of $\gamma(k_x)$ show that self-gravity increases MRI growth
-rates at large $k_x$, but the effect is small. %can be used for code
-                                %test 
-Note that this 
-destabilization is through the linear response, rather than through
-the background stratification (which is stabilizing). 
-%For the $k_x$ in the GI
-%interval, MRI is weakly stabilized by self-gravity.
+rates at large $k_x$. This effect is small but noticeable, which can
+be used as a code test for non-linear simulations. Note that this
+destabilization is through the linear response, rather 
+than through the background stratification (which is stabilizing).
 
-%magnetic pressure, small scale more effective 
-%, reflected in the full range of $\tau$ displayed by the
-%modes. 
-%mode classification becomes difficult. 
-%the point is: dispersion relation cannot be estimated by computing
-%the MRI and GI separately
+%smooth transitions, mode exchange
 
@@
     wavenumber $k_x$. The colorbar measures the importance of
-    self-gravity. The solid line corresponds to MRI modes in the
+    self-gravity by $\tau$. The solid line corresponds to MRI modes in the
     Cowling approximation. The dashed line corresponds to pure GI
@@
     problem. 
-%    {\bf REDO FIGURE}
     \label{compare_growth3_Q01d2}}
 \end{figure}
-
-%implications for simulations -> need to resolve small-scales
-
-%Fig. \ref{compare_growth3_Q01d2} again shows smooth
-%transitions between MRI and GI. This means the MRI and GI do indeed
-%interact, but
-
-%In particular, modes with $1.5 \lessim
-%k_xH\lesssim2$ is neither pure
-
-%demonstrates that mode classification
-%becomes ambigious when both GI and MRI are present.
-
-%\subsection{Layered disks}
-%Can a disk with a dense, resistive midplane and a highly conducting
-%atmosphere  be subject to both MRI and GI but at different heights?
-
-%GI in the midplne and MRI in the surface?

### 2014-04-05 paper/summary.tex
tex updates

@@
 In this paper, we have performed axisymmetric linear stability
-calculations of magnetized, self-gravitating disks in the local
-approximation. Our models include resistivity and azimuthal fields.  
-We have identified regimes under which the magneto-rotational
-instability (MRI) is affected by disk self-gravity (SG).  
-%what have we extended?
+calculations of magnetized, self-gravitating, vertically stratified
+disks in the local approximation. Our models include resistivity and
+azimuthal fields. We have identified regimes under which the
+magneto-rotational instability (MRI) is affected by disk self-gravity
+(SG).
 
-For a vertical field, the requirement for the MRI to operate is that its
-vertical wavelength $\lambda \lesssim H$. While $\lambda$ is independent of
-the strength of self-gravity $Q$, the disk thickness $H=H(Q)$ %lambda
-%depends on omega, which might be
-                                %lowered in a massive disk (sub-keplerian)
-decreases with increasing SG, which reduces the MRI growth rates.   
-%This
-%effect becomes noticeable when the disk is marginally stable to GI in
-%3D ($Q\sim 0.2$ or $Q_\mathrm{2D}\sim 0.7$).  
-Thus, a sufficiently massive disk can surpress the MRI. %but GI may
-                                %develop  instead
+For a vertical field, the requirement for the MRI is that   
+its vertical wavelength $\lambda \lesssim H$. While $\lambda$ is independent of
+the strength of self-gravity $Q$, the disk thickness $H=H(Q)$ 
+decreases with increasing SG, which can reducie MRI growth rates. 
+Thus, a sufficiently massive disk can potentially surpress the MRI. 
 The MRI is also restricted to larger radial
@@
 fields are required to permit the MRI with increasing self-gravity.
-
-%need larger boxes 
-%This suggests
-%that small-scale MRI turbulence may not develop in self-gravitating
-%disks.
-
-%correlation
 
@@
 picture of layered accretion  \citep[e.g.][]{fleming03} may not be applicable 
-to self-gravitating disks.  %strong magnetization
+to self-gravitating disks.
 
 %SG through 
-We find MRI perturbations with radial scale of $\sim H$ can acquire  
-density perturbations in a massive but Toomre-stable disk. 
+We find MRI modes with radial scale of $\sim H$ can acquire  
+density perturbations in massive but Toomre-stable disks. 
 This occurs when the MRI is weak, for example with a strong field or
@@
 perturbation caused by the MRI can be enhanced by self-gravity.
+%Our 
+%interpretation involves the MRI to provide a seed density
+%perturbation %in the absence of self-gravity.
+
+%what is required? a seed pert, plus amplification
 
@@
 employed ideal MHD simulations with gas-to-magnetic pressure ratios of
-order $10^2$ to $10^3$. According  
-to our results, self-gravity is not expected to affect the MRI in this
+order $10^2$ to $10^3$. %MRI is basically incompressible 
+This is qualitatively consistent with our results, as 
+self-gravity is not expected to influence the MRI in this 
 regime of $\beta$. On the other hand, \cite{fromang04} found
@@
 
-%We suggest that in
-%order for self-gravity to be significant,
-%We suggest that in
-%order for self-gravity to influence MRI,      
-%cite{kim03} and \cite{fromang04}
-%Numerical experiments
-%varying the field strength and resisitivity indicate this is again a
-%result
-
 %GI/MRI: avoided crossings, relevance to simulations
@@
 should be resolved.
-
-%This is
-%consistent with the MRI only affected by self-gravity for strong
-%fields 
-%mode classificication ambgious 
-%This is consistent /expected
-%numerical simulations should measure a continous range of modes
-%This is
-%consistent self-gravity only affecting the MRI when it is weak.  
-%mag field only affect GI through pressure
-%mri influenced by GI at large kx - > small scale -> need resolution
 
@@
 
-%lesure 14 (strong azimuthal field)
-%implications non-linear MRI simulations with a strong toroidal field
-%would over-estimate growth rates (and perhaps turbulence strengths)
-%if SG is not accounted for. not applicable to self-gravitating disks     
-%previous linear calcs show compressibility stabilizes MRI 
-%additional pressure stabilizing MRI
-
 \subsection{Caveats and outlooks}
@@
 
-%
-%This
-%prevents consideration of super-thermal toroidal fields
-%\citep{pessah05}.
-
-%tension stabilizes GI  
-%This prevents us from considering super-thermal
-%toroidal fields.  
-%go to cylindrical disk geometry (fu & lai) 
-%tension stabilizes GI
-
-%not looked at how MRI affects GI -> it doesn't because GI isn't
-%concerned with shear. only magnetic pressure. maybe a plot of GI with
-%magnetic perturbations?
-
-%spiral arms, GI angular mometum distribution 
-%global modes 
-%compressible term in y-induction equation 
 \emph{Beyond axisymmetry.} Axisymmetric perturbations, as we have
 assumed, preclude gravitational torques \citep{lynden-bell72}. 
-%In 
-%order to investigate how angular momentum transport 
 The local non-axisymmetric stability of magnetized self-gravitating
 thin disks has been considered by several authors
-\citep{elmegreen87,gammie96,fan97,kim01}. However, two-dimensional models
+\citep{elmegreen87,gammie96b,fan97,kim01}. However, two-dimensional models
 exclude the MRI. It will be neccessary to generalize these studies to
@@
 simulations of this kind \citep{fromang04c,fromang05}.
-
-%Our linear problem assumes a non-zero
-%vertical field is always present. It is of interest to 
-%consider the opposite limit of purely toroidal fields. Of course, one
-%must consider the non-axisymmetric MRI \cite{}
-%The local non-axisymmetric stability of magnetized, self-gravitating
-%disks has been treated in the shearing sheet framework \citep{} 
-%framework by decomposing
-%perturbations into shearing waves %ref.  
-%
-%global angular momentum exchange
-%It is necessary to generalize the current models to non-axisymmetric
-%perturbations. 
-%It would be interesting to examine global spiral instabilities in a 
-%magnetized disk. 
-%In particular, to examine global spiral instabilities
-%in a magnetized disk. 
-%fu lai

### 2014-04-05 paper/appendix.tex
tex updates
first draft pre spell check

@@
 self-gravity parameter $Q$ and $Q_\mathrm{2D}$, we replace $c_s$ by
-$\overline{c_s}\equiv\int\rho c_s dz/\int\rho dz$, and replace
-$\kappa$ by $\Omega$, so that 
+$\overline{c_s}\equiv\int\rho c_s dz/\int\rho dz$, and 
+$\kappa$ by $\Omega$, giving
 \begin{align}
@@
 where each term on the right-hand-side is non-dimensionalized (see
-\S\ref{non-dim}). We display this relation in Fig. \ref{plot_q3d2d}
-for the isothermal and polytropic disks.
+\S\ref{non-dim}). Fig. \ref{plot_q3d2d} plots this relation 
+for isothermal and polytropic disks.
 
@@
   \includegraphics[width=\linewidth]{figures/q2d_iso}
-  \caption{Relation between the self-gravity parameter $Q$ and the
-    classic Toomre parameter $Q_\mathrm{2D}$.   
+  \caption{Relation between the self-gravity parameter $Q$ used in
+    this paper and the Toomre parameter $Q_\mathrm{2D}$ for razor-thin
+    disks. 
     \label{plot_q3d2d}}
@@
 \section{Reduction to linear hydrodynamics}\label{reduction} 
-
 Our task here is to remove the magnetic field and vertical velocity
@@
 variable resistivity and that of a tilted field with uniform
-resistivity. We remark that a similar procedure can be performed in
-the general case of a tilted field with variable resistivity.
+resistivity. (A similar procedure can be performed in
+the general case of a tilted field with variable resistivity)
 
@@
   \imgi\sbar\dby = -B_y\Delta + B_z\dvy^\prime + \frac{\imgi
-    S}{\sigma}\left[\frac{\eta\mu_0\rho}{B_z}\left( D_1 U + \imgi\epsilon k_x D_0
+    S}{\sigma}\left[ B_z \dvx^\prime+ \frac{\eta\mu_0\rho}{B_z}\left( D_1 U + \imgi\epsilon k_x D_0
   V\right)\right] + \frac{\eta\mu_0\rho}{B_z}D_1V,\label{by_tilted}
@@
     \left(\ln{c_s^2}\right)^\prime W\right] +
-    \frac{1}{\sigma}\left[\left(\ln{\rho}\right)^\prime\w^{\prime\prime}
+    \frac{1}{\sigma}\left[\left(\ln{\rho}\right)^\prime(\w^{\prime\prime}
+      + \epsilon V^\prime)
       +\left(\ln{\rho}\right)^{\prime\prime}\w^{\prime} \right]\right\}.    
@@
 %(i.e. $\Omega\to 1$ and $H\to 1$), unless otherwise stated.
+
+%tilted field
+% calculate b field vertical derivatives carefully
+% try no velocity pert at bc
+% try anti-aligned field

### 2014-04-05 paper/summary.tex
first draft pre spell check

@@
 fields are required to permit the MRI with increasing self-gravity.   
+%since MRI is 3D, strong SG makes it 2D
 
 Interestingly, for layered resistivity we do not find layered
-perturbations when the disk is massive. This is 
-consistent with MRI becoming vertically global with increasing self-gravity.    
+magnetic perturbations when the disk is massive. This is 
+consistent with the MRI becoming vertically global with increasing self-gravity.    
 For non-self-gravitating disks $\lambda\ll H$, so the MRI can be
-restricted to regions of size $L<H$, i.e. an active layer. This is not
-possible if $\lambda \sim H$, as found for massive disks. Hence we
+restricted to regions of size $L<H$, i.e. an active layer. This not
+compatible with $\lambda \sim H$, as found for massive disks. Hence we
 find magnetic perturbations penetrate into the high-resistivity dead
@@
 %SG through 
-We find MRI modes with radial scale of $\sim H$ can acquire  
+We find MRI modes with radial scales of $\sim H$ can acquire  
 density perturbations in massive but Toomre-stable disks. 
 This occurs when the MRI is weak, for example with a strong field or
-high resisitivty. We argue in that case $\lambda\sim H$, so the density
-perturbation caused by the MRI can be enhanced by self-gravity.
+high resisitivty. We argue in that case $\lambda\sim H$, so the MRI is
+compressible and the associated density
+perturbation can be enhanced by self-gravity.
 %Our 
@@
 \citep{kim03,fromang04a,fromang04}. These authors find self-gravity
-does not enhance MRI density flucations significantly. However, they
+did not enhance MRI density flucations significantly. However, they
 employed ideal MHD simulations with gas-to-magnetic pressure ratios of
@@
 self-gravity is not expected to influence the MRI in this  
-regime of $\beta$. On the other hand, \cite{fromang04} found
-self-gravity makes MRI turbulence more coherent. This may be related
+regime of $\beta$, except through the background state. 
+%both `seed' and amplfication are small 
+For example, \cite{fromang04} found
+MRI turbulence is more coherent in self-gravitating disks. This may be related
 to our finding that small radial scale MRI is surpressed when
@@
 simluations should consider parameter regimes in which the MRI is
-vertically global. Indeed, in the setup of \cite{kim03}, the disk
+vertically global. %this makes MRI more compressible
+Indeed, in the setup of \cite{kim03}, the disk
 scale height exceeds the MRI vertical wavelength and self-gravity has
@@
 Curiously, when GI and MRI are simultaneously supported, we find
-unstable modes transition between MRI and GI. Such modes
-have comparable potential and magnetic energy perturbations. These
+unstable modes transition between MRI and GI. There exists modes 
+with comparable potential and magnetic energy perturbations, which 
+demonstrates MRI and GI can interact. These 
 transitions occur smoother with decreasing $\beta$
@@
 (Fig. \ref{compare_growth3_Q01d2}). The latter implies that, in order
-to capture magnetic-gravitational interactions represented by these
+to capture the magneto-gravitational interactions represented by these
 intermediate modes, non-linear simulations must resolve radial scales
@@
 Fig. \ref{compare_growth3_Q01d2} suggest radial scales down to $\sim H/2$
-should be resolved.
+should be well-resolved.
 
@@
 considered the full disk column. Self-gravity affects the MRI
-differently depending on the midplane symmetry of the MRI density
-perturbation. Self-gravity destabilizes MRI modes where the magnetic
-energy reaches a minimum at $z=0$, which have a symmetric
+differently depending on its character. 
+%midplane symmetry of the MRI 
+%density
+%perturbation. 
+Self-gravity destabilizes MRI modes where the magnetic
+energy has a minimum at $z=0$, these modes have a symmetric
 density perturbation in the limit $B_y\to0$. However, 
-self-gravity stabilizes MRI modes where the magnetic energy maximizes
-at $z=0$. These modes have an anti-symmetric density
+self-gravity stabilizes MRI modes where the magnetic energy has a
+maximum at $z=0$, these modes have an anti-symmetric density
 perturbation in the limit $B_y\to0$. This stabilization effect is
-stronger for increasing $B_y$. Previous linear calculations indicate
+stronger for increasing $B_y$. Previous linear calculations show that 
 increased compressibility associated with a toroidal field stabilizes
-the MRI \citep{kim03}. We conjecture that self-gravity can
-further enhance this effect.   
+the MRI \citep{kim03}. We conjecture that self-gravity 
+further enhances this effect. Non-linear MRI simulations with strong
+toroidal fields that neglect self-gravity may over-estimate the
+strength of MRI turbulence.   
 %can't do WKB (can't tell symmetries)
@@
 of toroidal field lines present in the global disk 
-geometry. \cite{pessah05} have shown that new effects on the MRI arise
+geometry. \cite{pessah05} demonstrated new effects on the MRI 
 when the curvature of a super-thermal toroidal field is accounted for;
 although \citeauthor{pessah05} focused on modes with large (small)
-vertical (radial) wavenumbers, for which we do not expect self-gravity
-to be important. Since compressibilty becomes important for strong
-toroidal fields, the effect of self-gravity may become even more
-significant when super-thermal toroidal fields are
+vertical (radial) wavenumbers, for which we expect self-gravity
+can be ignored. 
+Since compressibilty becomes important for strong
+toroidal fields, the effect of self-gravity on modes with $k_xH\sim1$
+may become significant when super-thermal toroidal fields are
 considered. However, global disk models will be neccessary

### 2014-04-05 paper/appendix.tex
first draft, spell-checked

@@
   \includegraphics[width=\linewidth]{figures/plot_fq}
-  \caption{The function $f(Q)$ describing vertical hydrodstatic 
+  \caption{The function $f(Q)$ describing vertical hydrostatic 
     equilibrium in self-gravitating polytropic disks (solid line). The
@@
 first obtain expressions for $\dbx$ and $\dby$. Differentiating
-the $x$ momentum equation and replacing the resuting $\dbz^\prime$
+the $x$ momentum equation and replacing the resulting $\dbz^\prime$
 using the divergence-free condition and $\dby^\prime$ using the $y$

### 2014-04-05 paper/paper.tex
first draft, spell-checked

@@
   becomes global in the vertical direction, and MRI modes 
-  with small radial lengthscales are stabilized. In massive
+  with small radial length scales are stabilized. In massive
   disks with layered resistivity, the MRI is not well-localized
   to regions where the Elsasser number exceeds unity. For MRI modes
-  with radial lengthscales on the order of the disk thickness, SG
+  with radial length scales on the order of the disk thickness, SG
   can enhance density perturbations, an effect that becomes 
@@
 PPDs subject to both MRI and GI are often modeled through separate
-turbulent viscosities in a hydrodynamical framework 
+turbulent viscosity coefficients in a hydrodynamical framework 
 \citep{terquem08}. This implicitly assumes that the development of MRI 
@@
 and self-gravity have been limited to a few early simulations 
-\citep{fromang04,fromang04b,fromang05}. It will be neccessary to 
+\citep{fromang04,fromang04b,fromang05}. It will be necessary to 
 revisit and extend these pioneering calculations to fully explore the
@@
 perturbations in a rotating disk. However, modes
-with radial lengthscales on the order of the disk scale height may be
+with radial length scales on the order of the disk scale height may be
 subject to self-gravity. It is therefore of interest to generalize the MRI with
-non-zero radial waveumbers to massive disks.
+non-zero radial wavenumbers to massive disks.
 
@@
 have the advantage that a wide range of parameters can be studied at
-negligible computational cost. This  allows us to indentify
+negligible computational cost. This  allows us to identify
 conditions, if any, under which MRI and GI 
@@
 This paper is organized as follows. \S\ref{setup} lists the governing
-equations and describes the disk equilibra under consideration. 
+equations and describes the disk equilibria under consideration. 
 The linear problem is formulated in \S\ref{linear}. The impact
@@
 \S\ref{result1}, gravitationally unstable disks are considered in
-\S\ref{result2}, and equilibra including an azimuthal field is explored in
+\S\ref{result2}, and equilibria including an azimuthal field is explored in
 \S\ref{result3}. We summarize results in \S\ref{summary} with a 

### 2014-04-05 paper/results.tex
first draft, spell-checked

@@
 In this section we focus on the MRI and use the vertical field setup
-of case 1. We first consider MRI modes with negligble density/potential
+of case 1. We first consider MRI modes with negligible density/potential
 perturbations to see the effect of self-gravity on the MRI through the
@@
 significantly reduced. This is seen for $Q=0.2$ where $\lambda \geq 2H$ throughout
-the disk. (The instability is not surpressed since
+the disk. (The instability is not suppressed since
 Eq. \ref{lambda_ideal}---\ref{lambda_resis} is only exact for
@@
 = 0.53,\,0.64$ and $0.66$, respectively. In the non-self-gravitating
-limit ($Q=4$), the MRI is effectively surpressed for
+limit ($Q=4$), the MRI is effectively suppressed for
 $z\lesssim0.5H$. This is consistent with the picture of layered
@@
 suggests that in massive disks, the MRI is not well localized to a
-sub-layer within the height, even when the resisitvity has a layered
+sub-layer within the height, even when the resistivity has a layered
 structure.
@@
 (increasing $f$), we expect MRI modes with small radial
-lengthscales to be suppressed.
+length scales to be suppressed.
 
@@
 We repeat the above calculation for resistive disks, but fix
-$\beta=100$ and vary the midplane Elasser number $\Lambda_0$. Growth
+$\beta=100$ and vary the midplane Elsasser number $\Lambda_0$. Growth
 rates are shown in Fig. \ref{gravity_energy_resis}. 
@@
     disk with $Q=0.2$ ($Q_\mathrm{2D}=0.72$) at fixed $\beta=100$, 
-    for a range of midplane Elasser numbers. The resistivity is
+    for a range of midplane Elsasser numbers. The resistivity is
     uniform. The colorbar measures the importance of self-gravity by $\tau$. 
@@
     The vertical line
-    inidicates $\Lambda=1$.  
+    indicates $\Lambda=1$.  
     \label{mri_massive_resis}}
@@
  perturbed mass across the disk height can contribute to a net potential
- perturbation. We therefore expect a neccessary condition for
+ perturbation. We therefore expect a necessary condition for
  self-gravity to affect the MRI is for the latter to be weak. 

### 2014-04-05 paper/results_azi.tex
first draft, spell-checked

@@
 $B_y\neq 0$,   
-parameterized by $\epsilon \equiv B_y/B_z$. However, we continute to
+parametrized by $\epsilon \equiv B_y/B_z$. However, we continue to
 use $B_z$ for normalizations and $\beta$ is associated with the
@@
 instability is not expected because Fig. \ref{compare_growth3} shows
-that even for $Q=0.18$, GI is surpressed for $\beta \lesssim 15$.
+that even for $Q=0.18$, GI is suppressed for $\beta \lesssim 15$.
 
 Fig. \ref{compare_growth3_tilted} show MRI growth rates for
-$B_y/B_z=0,\,1,\,2$ and $3$. We divide the modes into two catagories
+$B_y/B_z=0,\,1,\,2$ and $3$. We divide the modes into two categories
 depending on the extremum of magnetic energy at the midplane. 
@@
 cut-off wavenumber is larger when SG is included. 
-Destablization is most effective for purely vertical fields: with
+Destabilization is most effective for purely vertical fields: with
 $\epsilon=0,\, k_xH\simeq 1.4$, SG increases the growth rate by $\sim
@@
 
-Compressibilty is enhanced by an azimuthal field, which is 
+Compressibility is enhanced by an azimuthal field, which is 
 stabilizing for the MRI \citep{kim00}. %because...?
@@
 
-Notice also in Fig. \ref{compare_growth3_tilted_resis} the disappearence of
+Notice also in Fig. \ref{compare_growth3_tilted_resis} the disappearance of
 magnetic modes between $0.8\lesssim k_xH\lesssim 1.5$ as $B_y$ is

### 2014-04-05 paper/results_gi.tex
first draft, spell-checked

@@
 growth rates are $\gamma\simeq 0.25\Omega,\,0.6\Omega,\,0.8\Omega$ for
-$Q=0.18,\,0.14,\,0.12$, respecitvely. GI is stabilized by magnetic
+$Q=0.18,\,0.14,\,0.12$, respectively. GI is stabilized by magnetic
 pressure for sufficiently small $\beta$. The critical
-field strength for stabilizating GI increases with increasing
+field strength for stabilizing GI increases with increasing
 self-gravity, consistent with \cite{nakamura83}.   
@@
 Thus, our results do not rigorously prove that the GI and MRI branches
-do not intersect.  Nevertheless, the continous variation of $\tau$
+do not intersect.  Nevertheless, the continuous variation of $\tau$
 strongly suggest that unstable modes can transition smoothly from 
@@
     Cowling approximation. The dashed line corresponds to pure GI
-    modes, obtained by including a high resisitivity in the full
+    modes, obtained by including a high resistivity in the full
     problem. 

### 2014-04-05 paper/setup.tex
first draft, spell-checked

@@
 $\Omega_{z0}\equiv\Omega_z(r_0,0)$; 
-and the gas potential $\Phi_d$ statisfies Poisson's equation
+and the gas potential $\Phi_d$ satisfies Poisson's equation
 \begin{align}
@@
 velocity field is $\bm{v} = -Sx\hat{\bm{y}}$. We consider Keplerian
-disks so that $S = 3\Omega/2$ and the epicyclic frequency 
+disks so that $S = 3\Omega/2$ and the epicycle frequency 
 $\kappa\equiv\sqrt{2\Omega(2\Omega-S)}=\Omega=\Omega_z$.  

### 2014-04-07 non-ideal/functions.f90
updated code source, idl source

@@ -8,8 +8,8 @@ SUBROUTINE F (NEQ, x, Y, YDOT, RPAR, IPAR)
   logrho = Y(1)
   phidot = Y(2)
   
-  YDOT(1) = -omegaz2*x - phidot
-  YDOT(2) = exp(logrho)/bigQ
+  YDOT(1) = -omegaz2*x/fQ**2d0 - phidot
+  YDOT(2) = exp(logrho)/(bigQ*fQ**2d0)
   
   return
 end subroutine F
@@ -19,6 +19,63 @@ subroutine jac
   implicit none
   return
 end subroutine jac
+
+SUBROUTINE FCN_FQ_ISO(N,X,FVEC,IFLAG)
+  use global
+  implicit none
+  integer :: N, IFLAG
+  real*8 :: X(N),FVEC(N)
+  integer, parameter :: NEQ = 2, ITOL = 1 , MF = 10, LIW = 60, IOPT = 0
+  integer, parameter :: LRW = 2*(20 + 16*NEQ)
+  integer :: i, ITASK, ISTATE, IWORK(LIW), IPAR
+  real*8, parameter :: RTOL = 1d-15, ATOL=1d-15, eps=1d-2
+  real*8 :: Y(NEQ), xSTART, xEND, RWORK(LRW), RPAR
+  external :: F, JAC
+
+  Y(1) = 0d0 !log rho_hat = 0 at z=0
+  Y(2) = 0d0 !dphi/dz =0 at z=0
+
+  ITASK = 1
+  ISTATE= 1
+
+  xSTART  = 0d0 
+  xEND    = x(1)
+
+  call DVODE (F, NEQ, Y, xSTART, xEND, ITOL, RTOL, ATOL, ITASK, &
+          ISTATE, IOPT, RWORK, LRW, IWORK, LIW, JAC, MF, &
+          RPAR, IPAR)
+
+  FVEC(1) = exp(Y(1)) - eps 
+
+  return 
+END SUBROUTINE FCN_FQ_ISO
+
+
+subroutine get_fQ_iso
+   use global
+   implicit none
+   integer, parameter :: N = 1, LWA = N*(3*N+13)
+   integer :: INFO
+   real*8, parameter :: TOL = 1d-15
+   real*8 :: x(N) , fvec(N), wa(LWA)
+   external :: FCN_FQ_ISO 
+
+   fQ   = 1d0 
+
+   x(1) = 2d0 !trial value of zmax in units of non-sg scale heights 
[…]

### 2014-04-07 non-ideal/sgmri.f90

@@ -1,9 +1,9 @@
 module global
-  logical :: refine, prin
+  logical :: refine, prin, use_old
   character*3 :: eos, method
-  character*4 :: var 
-  integer, parameter :: nvar=2
-  integer :: nz, nk, big_nz, ntrials , kcount 
+  character*4 :: var, vbc 
+  integer, parameter :: nvar=4
+  integer :: nz, nk, big_nz, ntrials , kcount, nuse_old 
   real*8, parameter :: omega=1.0, bigG = 1.0, mstar = 1.0, r0=1.0
   real*8, parameter :: pi = 2d0*acos(0d0) 
   real*8 :: bigQ, beta, shear, kmin, kmax, kx, dk, qmin, qmax, bmin, &
@@ -15,7 +15,7 @@ module global
        kaxis(:), csq(:), eta(:), deta(:), d2eta(:) 
   real*8, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)  
   complex*16, parameter :: ii = (0d0, 1d0)      
-  complex*16 :: sig 
+  complex*16 :: sig, sig_old
   complex*16, allocatable :: vx(:), vy(:), bz(:)
   complex*16, allocatable :: den(:), pot(:), wvector(:), sig_trials(:)
 end module global
@@ -25,13 +25,13 @@ program sgmri
   implicit none 
   integer :: i, j, k  
   integer :: haf_lmax, lmax  
-  character*2 :: str_mode 
+  character*3 :: str_mode 
   real*8  :: zbar, T_l, dT_l, d2T_l, lmode, dummy
   real*8, external :: density, dlogrho, alfven
   namelist /params/ bigQ, beta, kx, var, shear, kmin, kmax, qmin,&
        & qmax, bmin, bmax, eos 
   namelist /resis/ Rm, amp, Rm_min, Rm_max
-  namelist /grid/ zmax, nz, nk  
+  namelist /grid/ zmax, nz, nk, vbc, nuse_old   
   namelist /eigen/ method, min_growth, max_growth, ntrials , refine, prin
   
   write(6,fmt='(A)') '--------------- welcome ---------------'
@@ -84,7 +84,7 @@ program sgmri
         var_max = kmax
      endif
      if(var.eq.'resi') then
-        write(6,fmt='(A)') ' varying horizontal wave number'
+        write(6,fmt='(A)') ' varying resistivity wave number'
[…]

### 2014-04-07 non-ideal/cowling/functions.f90
added non-ideal/cowling

@@ -1,0 +1,713 @@
+SUBROUTINE F (NEQ, x, Y, YDOT, RPAR, IPAR)
+  use global
+  implicit none
+  integer :: NEQ, IPAR
+  real*8 :: x, Y(NEQ), YDOT(NEQ), RPAR
+  real*8 :: logrho, phidot
+  
+  logrho = Y(1)
+  phidot = Y(2)
+  
+  YDOT(1) = -omegaz2*x/fQ**2d0 - phidot
+  YDOT(2) = exp(logrho)/(bigQ*fQ**2d0)
+  
+  return
+end subroutine F
+
+subroutine jac
+  use global
+  implicit none
+  return
+end subroutine jac
+
+SUBROUTINE FCN_FQ_ISO(N,X,FVEC,IFLAG)
+  use global
+  implicit none
+  integer :: N, IFLAG
+  real*8 :: X(N),FVEC(N)
+  integer, parameter :: NEQ = 2, ITOL = 1 , MF = 10, LIW = 60, IOPT = 0
+  integer, parameter :: LRW = 2*(20 + 16*NEQ)
+  integer :: i, ITASK, ISTATE, IWORK(LIW), IPAR
+  real*8, parameter :: RTOL = 1d-15, ATOL=1d-15, eps=1d-2
+  real*8 :: Y(NEQ), xSTART, xEND, RWORK(LRW), RPAR
+  external :: F, JAC
+
+  Y(1) = 0d0 !log rho_hat = 0 at z=0
+  Y(2) = 0d0 !dphi/dz =0 at z=0
+
+  ITASK = 1
+  ISTATE= 1
+
+  xSTART  = 0d0 
+  xEND    = x(1)
+
+  call DVODE (F, NEQ, Y, xSTART, xEND, ITOL, RTOL, ATOL, ITASK, &
+          ISTATE, IOPT, RWORK, LRW, IWORK, LIW, JAC, MF, &
+          RPAR, IPAR)
+
+  FVEC(1) = exp(Y(1)) - eps 
+
+  return 
+END SUBROUTINE FCN_FQ_ISO
+
+
+subroutine get_fQ_iso
+   use global
+   implicit none
+   integer, parameter :: N = 1, LWA = N*(3*N+13)
+   integer :: INFO
+   real*8, parameter :: TOL = 1d-15
+   real*8 :: x(N) , fvec(N), wa(LWA)
+   external :: FCN_FQ_ISO 
+
+   fQ   = 1d0 
+
[…]

### 2014-04-07 non-ideal/cowling/sgmri.f90

@@ -1,0 +1,252 @@
+module global
+  logical :: refine, prin, use_old
+  character*3 :: eos, method
+  character*4 :: var, vbc 
+  integer, parameter :: nvar=3
+  integer :: nz, nk, big_nz, ntrials , kcount, nuse_old 
+  real*8, parameter :: omega=1.0, bigG = 1.0, mstar = 1.0, r0=1.0
+  real*8, parameter :: pi = 2d0*acos(0d0) 
+  real*8 :: bigQ, beta, shear, kmin, kmax, kx, dk, qmin, qmax, bmin, &
+       bmax, var_min, var_max 
+  real*8 :: zmax, dz    
+  real*8 :: kappa, omegak, omegaz2, fQ, Rm, amp, zstar, Rm_max, Rm_min
+  real*8 :: condition, min_growth, max_growth, sig_re, sig_im    
+  real*8, allocatable :: dnorm(:), drho(:), d2rho(:), valf(:), zaxis(:),&
+       kaxis(:), csq(:), eta(:), deta(:), d2eta(:) 
+  real*8, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)  
+  complex*16, parameter :: ii = (0d0, 1d0)      
+  complex*16 :: sig, sig_old
+  complex*16, allocatable :: vx(:), vy(:), bz(:)
+  complex*16, allocatable :: den(:), pot(:), wvector(:), sig_trials(:)
+end module global
+
+program sgmri
+  use global 
+  implicit none 
+  integer :: i, j, k  
+  integer :: haf_lmax, lmax  
+  character*3 :: str_mode 
+  real*8  :: zbar, T_l, dT_l, d2T_l, lmode, dummy
+  real*8, external :: density, dlogrho, alfven
+  namelist /params/ bigQ, beta, kx, var, shear, kmin, kmax, qmin,&
+       & qmax, bmin, bmax, eos 
+  namelist /resis/ Rm, amp, Rm_min, Rm_max
+  namelist /grid/ zmax, nz, nk, vbc, nuse_old   
+  namelist /eigen/ method, min_growth, max_growth, ntrials , refine, prin
+  
+  write(6,fmt='(A)') '--------------- welcome ---------------'
+  
+  !read input parameters.
+  open(7, file="input")
+  read(7, nml=params)
+  read(7, nml=resis)
+  read(7, nml=grid)
+  read(7, nml=eigen) 
+  close(7)
+
[…]

### 2014-04-07 tilted/functions.f90
added tilted

@@ -1,0 +1,674 @@
+SUBROUTINE F (NEQ, x, Y, YDOT, RPAR, IPAR)
+  use global
+  implicit none
+  integer :: NEQ, IPAR
+  real*8 :: x, Y(NEQ), YDOT(NEQ), RPAR
+  real*8 :: logrho, phidot
+  
+  logrho = Y(1)
+  phidot = Y(2)
+  
+  YDOT(1) = -omegaz2*x/fQ**2d0 - phidot
+  YDOT(2) = exp(logrho)/(bigQ*fQ**2d0)
+  
+  return
+end subroutine F
+
+subroutine jac
+  use global
+  implicit none
+  return
+end subroutine jac
+
+SUBROUTINE FCN_FQ_ISO(N,X,FVEC,IFLAG)
+  use global
+  implicit none
+  integer :: N, IFLAG
+  real*8 :: X(N),FVEC(N)
+  integer, parameter :: NEQ = 2, ITOL = 1 , MF = 10, LIW = 60, IOPT = 0
+  integer, parameter :: LRW = 2*(20 + 16*NEQ)
+  integer :: i, ITASK, ISTATE, IWORK(LIW), IPAR
+  real*8, parameter :: RTOL = 1d-15, ATOL=1d-15, eps=1d-2
+  real*8 :: Y(NEQ), xSTART, xEND, RWORK(LRW), RPAR
+  external :: F, JAC
+
+  Y(1) = 0d0 !log rho_hat = 0 at z=0
+  Y(2) = 0d0 !dphi/dz =0 at z=0
+
+  ITASK = 1
+  ISTATE= 1
+
+  xSTART  = 0d0 
+  xEND    = x(1)
+
+  call DVODE (F, NEQ, Y, xSTART, xEND, ITOL, RTOL, ATOL, ITASK, &
+          ISTATE, IOPT, RWORK, LRW, IWORK, LIW, JAC, MF, &
+          RPAR, IPAR)
+
+  FVEC(1) = exp(Y(1)) - eps 
+
+  return 
+END SUBROUTINE FCN_FQ_ISO
+
+
+subroutine get_fQ_iso
+   use global
+   implicit none
+   integer, parameter :: N = 1, LWA = N*(3*N+13)
+   integer :: INFO
+   real*8, parameter :: TOL = 1d-15
+   real*8 :: x(N) , fvec(N), wa(LWA)
+   external :: FCN_FQ_ISO 
+
+   fQ   = 1d0 
+
[…]

### 2014-04-07 tilted/sgmri.f90

@@ -1,0 +1,294 @@
+module global
+  logical :: refine, prin, use_old, beta_fix
+  character*3 :: eos, method
+  character*4 :: var, vbc 
+  integer, parameter :: nvar=4
+  integer :: nz, nk, big_nz, ntrials , kcount, nuse_old, Nzmid
+  real*8, parameter :: omega=1.0, bigG = 1.0, mstar = 1.0, r0=1.0
+  real*8, parameter :: pi = 2d0*acos(0d0) 
+  real*8 :: bigQ, beta, tilt, shear, kmin, kmax, kx, dk, qmin, qmax, bmin, &
+       bmax, var_min, var_max, ep_min, ep_max, beta_tot
+  real*8 :: zmax, dz    
+  real*8 :: kappa, omegak, omegaz2, fQ, Rm, amp, zstar, Rm_max, Rm_min
+  real*8 :: condition, min_growth, max_growth, sig_re, sig_im    
+  real*8, allocatable :: dnorm(:), drho(:), d2rho(:), valf(:), zaxis(:),&
+       kaxis(:), csq(:), eta(:), deta(:), d2eta(:) 
+  real*8, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)  
+  complex*16, parameter :: ii = (0d0, 1d0)      
+  complex*16 :: sig, sig_old
+  complex*16, allocatable :: vx(:), vy(:), bz(:)
+  complex*16, allocatable :: den(:), pot(:), wvector(:), sig_trials(:)
+end module global
+
+program sgmri
+  use global 
+  implicit none 
+  integer :: i, j, k  
+  integer :: haf_lmax, lmax  
+  character*3 :: str_mode 
+  real*8  :: zbar, T_l, dT_l, d2T_l, lmode, dummy
+  real*8, external :: density, dlogrho, alfven
+  namelist /params/ bigQ, beta, tilt, kx, var, shear, kmin, kmax, qmin,&
+       & qmax, bmin, bmax, ep_min, ep_max, eos, beta_fix  
+  namelist /resis/ Rm, amp, Rm_min, Rm_max
+  namelist /grid/ zmax, nz, nk, vbc, nuse_old   
+  namelist /eigen/ method, min_growth, max_growth, ntrials , refine, prin
+  
+  write(6,fmt='(A)') '--------------- welcome ---------------'
+  
+  !read input parameters.
+  open(7, file="input")
+  read(7, nml=params)
+  read(7, nml=resis)
+  read(7, nml=grid)
[…]

### 2014-04-07 tilted/cowling/functions.f90
added tilted/cowling

@@ -1,0 +1,667 @@
+SUBROUTINE F (NEQ, x, Y, YDOT, RPAR, IPAR)
+  use global
+  implicit none
+  integer :: NEQ, IPAR
+  real*8 :: x, Y(NEQ), YDOT(NEQ), RPAR
+  real*8 :: logrho, phidot
+  
+  logrho = Y(1)
+  phidot = Y(2)
+  
+  YDOT(1) = -omegaz2*x/fQ**2d0 - phidot
+  YDOT(2) = exp(logrho)/(bigQ*fQ**2d0)
+  
+  return
+end subroutine F
+
+subroutine jac
+  use global
+  implicit none
+  return
+end subroutine jac
+
+SUBROUTINE FCN_FQ_ISO(N,X,FVEC,IFLAG)
+  use global
+  implicit none
+  integer :: N, IFLAG
+  real*8 :: X(N),FVEC(N)
+  integer, parameter :: NEQ = 2, ITOL = 1 , MF = 10, LIW = 60, IOPT = 0
+  integer, parameter :: LRW = 2*(20 + 16*NEQ)
+  integer :: i, ITASK, ISTATE, IWORK(LIW), IPAR
+  real*8, parameter :: RTOL = 1d-15, ATOL=1d-15, eps=1d-2
+  real*8 :: Y(NEQ), xSTART, xEND, RWORK(LRW), RPAR
+  external :: F, JAC
+
+  Y(1) = 0d0 !log rho_hat = 0 at z=0
+  Y(2) = 0d0 !dphi/dz =0 at z=0
+
+  ITASK = 1
+  ISTATE= 1
+
+  xSTART  = 0d0 
+  xEND    = x(1)
+
+  call DVODE (F, NEQ, Y, xSTART, xEND, ITOL, RTOL, ATOL, ITASK, &
+          ISTATE, IOPT, RWORK, LRW, IWORK, LIW, JAC, MF, &
+          RPAR, IPAR)
+
+  FVEC(1) = exp(Y(1)) - eps 
+
+  return 
+END SUBROUTINE FCN_FQ_ISO
+
+
+subroutine get_fQ_iso
+   use global
+   implicit none
+   integer, parameter :: N = 1, LWA = N*(3*N+13)
+   integer :: INFO
+   real*8, parameter :: TOL = 1d-15
+   real*8 :: x(N) , fvec(N), wa(LWA)
+   external :: FCN_FQ_ISO 
+
+   fQ   = 1d0 
+
[…]

### 2014-04-07 tilted/cowling/sgmri.f90

@@ -1,0 +1,294 @@
+module global
+  logical :: refine, prin, use_old, beta_fix
+  character*3 :: eos, method
+  character*4 :: var, vbc 
+  integer, parameter :: nvar=3
+  integer :: nz, nk, big_nz, ntrials , kcount, nuse_old, Nzmid
+  real*8, parameter :: omega=1.0, bigG = 1.0, mstar = 1.0, r0=1.0
+  real*8, parameter :: pi = 2d0*acos(0d0) 
+  real*8 :: bigQ, beta, tilt, shear, kmin, kmax, kx, dk, qmin, qmax, bmin, &
+       bmax, var_min, var_max, ep_min, ep_max, beta_tot
+  real*8 :: zmax, dz    
+  real*8 :: kappa, omegak, omegaz2, fQ, Rm, amp, zstar, Rm_max, Rm_min
+  real*8 :: condition, min_growth, max_growth, sig_re, sig_im    
+  real*8, allocatable :: dnorm(:), drho(:), d2rho(:), valf(:), zaxis(:),&
+       kaxis(:), csq(:), eta(:), deta(:), d2eta(:) 
+  real*8, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)  
+  complex*16, parameter :: ii = (0d0, 1d0)      
+  complex*16 :: sig, sig_old
+  complex*16, allocatable :: vx(:), vy(:), bz(:)
+  complex*16, allocatable :: den(:), pot(:), wvector(:), sig_trials(:)
+end module global
+
+program sgmri
+  use global 
+  implicit none 
+  integer :: i, j, k  
+  integer :: haf_lmax, lmax  
+  character*3 :: str_mode 
+  real*8  :: zbar, T_l, dT_l, d2T_l, lmode, dummy
+  real*8, external :: density, dlogrho, alfven
+  namelist /params/ bigQ, beta, tilt, kx, var, shear, kmin, kmax, qmin,&
+       & qmax, bmin, bmax, ep_min, ep_max, eos, beta_fix  
+  namelist /resis/ Rm, amp, Rm_min, Rm_max
+  namelist /grid/ zmax, nz, nk, vbc, nuse_old   
+  namelist /eigen/ method, min_growth, max_growth, ntrials , refine, prin
+  
+  write(6,fmt='(A)') '--------------- welcome ---------------'
+  
+  !read input parameters.
+  open(7, file="input")
+  read(7, nml=params)
+  read(7, nml=resis)
+  read(7, nml=grid)
[…]

### 2014-04-07 paper/paper.tex
minor edits

--- 2014-04-03 paper/paper.tex
@@
   resistivity. It is found that the reduction in the disk thickness
-  by self-gravity (SG) can decrease MRI growth rates; the MRI
+  by self-gravity can decrease MRI growth rates; the MRI
   becomes global in the vertical direction, and MRI modes 
@@
   to regions where the Elsasser number exceeds unity. For MRI modes
-  with radial lengthscales on the order of the disk thickness, SG
+  with radial lengthscales on the order of the disk thickness, self-gravity 
   can enhance density perturbations, an effect that becomes 

### 2014-04-07 paper/results_azi.tex
minor edits

--- 2014-04-05 paper/results_azi.tex
@@
 \subsection{Ideal disks with MRI} 
-We consider disk with $Q=0.2$ ($Q_\mathrm{2D}=~0.72$) and
+We consider disks with $Q=0.2$ ($Q_\mathrm{2D}=~0.72$) and
 $\beta=10$ in the limit of ideal MHD ($\Lambda_0=100$). Gravitational
@@
     0.52cm]{figures/compare_growth3_tilted_odd.ps} 
-  \caption{Growth rates in massive isothermal disks with $Q=~0.2$ and
+  \caption{MRI growth rates in isothermal disks with $Q=~0.2$ ($Q_\mathrm{2D}=0.72$) and 
     $\beta=10$ for a range of azimuthal field strengths $B_y/B_z$. The
@@
     the solid curves are computed from the Cowling approximation. 
-    For $B_y=0$, modes in top and bottom panels 
+    For $B_y=0$, modes in top and bottom panels have
     $W^\prime(0)=0$ and $W(0)=0$, respectively.

### 2014-04-07 paper/paper.tex

--- 2014-04-05 paper/paper.tex
@@
   to regions where the Elsasser number exceeds unity. For MRI modes
-  with radial length scales on the order of the disk thickness, SG
+  with radial length scales on the order of the disk thickness, self-gravity 
   can enhance density perturbations, an effect that becomes 

### 2014-04-07 paper/summary.tex

@@
 the strength of self-gravity $Q$, the disk thickness $H=H(Q)$ 
-decreases with increasing SG, which can reducie MRI growth rates. 
-Thus, a sufficiently massive disk can potentially surpress the MRI. 
+decreases with increasing SG, which can reduce MRI growth rates. 
+Thus, a sufficiently massive disk can potentially suppress the MRI. 
 The MRI is also restricted to larger radial
@@
 in self-gravitating disks. Finally, the condition $\lambda < H$ 
-requries $f(Q)/\sqrt{\beta}$ to be sufficiently small, so weaker
+requires $f(Q)/\sqrt{\beta}$ to be sufficiently small, so weaker
 fields are required to permit the MRI with increasing self-gravity.   
@@
 This occurs when the MRI is weak, for example with a strong field or
-high resisitivty. We argue in that case $\lambda\sim H$, so the MRI is
+high resistivity. We argue in that case $\lambda\sim H$, so the MRI is
 compressible and the associated density
@@
 \citep{kim03,fromang04a,fromang04}. These authors find self-gravity
-did not enhance MRI density flucations significantly. However, they
+did not enhance MRI density fluctuations significantly. However, they
 employed ideal MHD simulations with gas-to-magnetic pressure ratios of
@@
 MRI turbulence is more coherent in self-gravitating disks. This may be related
-to our finding that small radial scale MRI is surpressed when
+to our finding that small radial scale MRI is suppressed when
 self-gravity is included in the background equilibria.
@@
 vertical direction, for it to affect the MRI, future non-linear
-simluations should consider parameter regimes in which the MRI is
+simulations should consider parameter regimes in which the MRI is
 vertically global. %this makes MRI more compressible
@@
 keeping the vertical field at fixed strength. In this case, we also
-relaxed the equitorial symmetry condition applied previously and
+relaxed the equatorial symmetry condition applied previously and
 considered the full disk column. Self-gravity affects the MRI
@@
 can be ignored. 
-Since compressibilty becomes important for strong
+Since compressibility becomes important for strong
 toroidal fields, the effect of self-gravity on modes with $k_xH\sim1$
 may become significant when super-thermal toroidal fields are
-considered. However, global disk models will be neccessary
+considered. However, global disk models will be necessary
 to self-consistently probe this regime.
@@
 \citep{elmegreen87,gammie96b,fan97,kim01}. However, two-dimensional models
-exclude the MRI. It will be neccessary to generalize these studies to
+exclude the MRI. It will be necessary to generalize these studies to
 3D in order to investigate the impact of the MRI on angular momentum
@@
 axisymmetric perturbations \citep{papaloizou89,papaloizou91}. Global
-non-axisymmetric linear models will be desirable to support non-linar
+non-axisymmetric linear models will be desirable to support non-linear
 simulations of this kind \citep{fromang04c,fromang05}.

### 2014-05-28 paper/linear.tex
first draft, spell-checked
revision to 1st ref report

@@
 the growth rate. We take $k_x>0$ without loss of generality. 
-Hereafter, we surpress the exponential factor and the real part notation.
+Hereafter, we suppress the exponential factor and the real part notation.
 
@@
 \subsection{Domain and boundary conditions}\label{domain}
+{\bf 
+For a vertical field, considered in \S\ref{result1} and
+\S\ref{result2}, we take $\bm{U}$ to be an even function of
+$z$. Odd modes are permitted in \S\ref{result3}, where an azimuthal field may
+be included. In both
+setups the gravitational potential boundary condition, given by
+\cite{goldreich65a}, is
+%The potential boundary
+%condition is 
+%\begin{align}
+%  \dphi^\prime(\zmax) + k_x\dphi(\zmax) = 0 \label{pot_bc1},
+%\end{align}
+%and is adapted from \cite{goldreich65a}. The general form of this
+%potential boundary condition is used for case 2 below.
+%and 
+%the potential boundary condition is 
+\begin{align}
+  \dphi^\prime(\pm \zmax) \pm k_x\dphi(\pm\zmax) =  - \left. \frac{\Omega^2
+    \rho \xi_z}{\rho_0Q}\right|_{\pm\zmax}, \label{pot_bc2}
+\end{align}
+where $\xi_z=\dvz/\imgi\sigma$ is the vertical Lagrangian
+displacement, and $z=\pm Z_s$ is the upper and lower disk surfaces,
+respectively. 
+}
+%In practice,  Eq. \ref{pot_bc1} and \ref{pot_bc2} are
+%very similar because $\rho(\pm\zmax)\ll\rho_0$. 
+%}
 
 \subsubsection{Case 1: vertical field}
-We impose $\bm{U}$ to be an even function of $z$, so that
-$d\bm{U}/dz=0$ at $z=0$. This permits higher numerical resolution by
-reducing the computational domain to $z\in~[0,Z_s]$.  
-At the upper disk boundary $z=Z_s$ we set 
+{\bf Here we impose} $d\bm{U}/dz=0$ at $z=0$. This permits higher
+numerical resolution by reducing the computational domain to
+$z\in~[0,Z_s]$. At the upper disk boundary $z=Z_s$ we set  
 \begin{align}
@@
 \end{align}
-so the field remains vertical. The potential boundary
-condition is 
-\begin{align}
-  \dphi^\prime(\zmax) + k_x\dphi(\zmax) = 0 \label{pot_bc1},
-\end{align}
-and is adapted from \cite{goldreich65a}. The general form of this
-potential boundary condition is used for case 2 below.
+so the field remains vertical. {\bf The derivation of the magnetic
+  field boundary conditions may be found in \cite{sano99}.}
 
@@
 In this more general setup the computational domain is
-$z\in[-Z_s, Z_s]$. At the disk surfaces we adopt the `halo' model of
+$z\in[-Z_s, Z_s]$ {\bf and no symmetry across the midplane is
+  enforced.} At the disk surfaces we adopt the `halo' model of
 \cite{gammie94}, so that
@@
 \end{align}
-and the potential boundary condition is 
-\begin{align}
-  \dphi^\prime(\pm \zmax) \pm k_x\dphi(\pm\zmax) =  - \left. \frac{\Omega^2
-    \rho \xi_z}{\rho_0Q}\right|_{\pm\zmax}, \label{pot_bc2}
-\end{align}
-where $\xi_z=\dvz/\imgi\sigma$ is the vertical Lagrangian
-displacement. In practice,  Eq. \ref{pot_bc1} and \ref{pot_bc2} are
-very similar because $\rho(\pm\zmax)\ll\rho_0$.
+    {\bf and this case permits $\dd v_z(\pm Z_s)\neq0$.}
+%      In practice, the gravitational potential boundary condition is
+%      similar to case 1 because $\rho(\pm Z_s)\ll\rho_0$.}
 
@@
 
-The above procedure discretizes the linear equations to a matrix equation,
+The above procedure discretize the linear equations to a matrix equation,
 \begin{align}\label{matrix_eqn}

### 2014-05-28 paper/paper.tex
revision to 1st ref report

@@
   becomes global in the vertical direction, and MRI modes 
-  with small radial length scales are stabilized. In massive
+  with small radial length scales are stabilized. {\bf The
+    upper limit to the vertical field strength to permit the MRI
+    in a strongly self-gravitating disk is estimated.} In massive
   disks with layered resistivity, the MRI is not well-localized
@@
   can enhance density perturbations, an effect that becomes 
-  significant in the presence of a strong toroidal field. In
+  significant in the presence of a strong toroidal field, {\bf and
+    which depends on the symmetry of the underlying MRI mode.} In
   gravitationally unstable disks where GI and MRI growth rates are
@@
 
+\acknowledgements
+{\bf I thank K. Menou, S. Fromang and A. Youdin for helpful discussions
+during the course of this project. 
+}
 \appendix

### 2014-05-28 paper/summary.tex
revision to 1st ref report

@@
 
-For a vertical field, the requirement for the MRI is that   
-its vertical wavelength $\lambda \lesssim H$. While $\lambda$ is independent of
-the strength of self-gravity $Q$, the disk thickness $H=H(Q)$ 
-decreases with increasing SG, which can reduce MRI growth rates. 
+For a vertical field, the requirement for the MRI {\bf to operate} is that   
+its vertical wavelength $\lambda \lesssim 2H$. 
+{\bf 
+%While $\lambda$ is independent of
+%the strength of self-gravity $Q$ {\bf for fixed $\beta$}, 
+The disk thickness $H=H(Q)$ decreases with increasing SG. This can
+reduce MRI growth rates when 
+$\beta$, and hence $\lambda$, is fixed.  
+}
 Thus, a sufficiently massive disk can potentially suppress the MRI. 
@@
 scales as $Q$ is lowered. This means that the MRI becomes more global
-in self-gravitating disks. Finally, the condition $\lambda < H$ 
-requires $f(Q)/\sqrt{\beta}$ to be sufficiently small, so weaker
-fields are required to permit the MRI with increasing self-gravity.   
-%since MRI is 3D, strong SG makes it 2D
+in self-gravitating disks.
+
+{\bf
+  The condition $\lambda < 2H$ 
+  %requires $f(Q)/\sqrt{\beta}$ to
+  %  be sufficiently small (see
+  %  Eq. \ref{lambda_ideal}---\ref{lambda_resis}), so larger values of
+  %  $\beta$ are required to  
+  %  permit the MRI with increasing self-gravity, since $f(Q)$ increases
+  %  with decreasing $Q$. 
+  may be written more precisely as 
+  \begin{align}\label{cond1}
+    \frac{n}{\mathrm{min}(\Lambda_0, 1)}\frac{f(Q)}{\sqrt{\beta}} \lesssim 1,
+  \end{align}
+  where $n\sim 3$ is a numerical factor and
+  $\mathrm{min}(\Lambda_0,1)$ accounts for the ideal and resistive
+  limits (see Eq. \ref{lambda_ideal}---\ref{lambda_resis}). Since $f$
+  increases with decreasing $Q$, Eq. \ref{cond1} implies
+  the MRI requires larger $\beta$ values with  
+  increasing self-gravity. 
+  For definiteness, consider the ideal polytropic disk. Then Eq. \ref{cond1} is
+  \begin{align}\label{cond2}
+    \beta^{-1/2} \lesssim
+    \frac{\sqrt{15}}{4\pi}\sqrt{Q}\arccos{\left(\frac{Q}{1+Q}\right)}. 
+  \end{align}
+  For a non-self-gravitating disk, $Q\to\infty$ and Eq. \ref{cond2} is
+  $\beta \gtrsim 16\pi^2/30\simeq 5$. For $Q\ll 1$, the condition is
+  $\beta \gtrsim 64/15Q$, giving $\beta\gtrsim 20$ for
+  $Q=0.2$. We confirmed this numerically, finding that the growth rate
+  $\gamma  \lesssim 0.1$ when $\beta \lesssim 3.3$ for $Q=20$ and
+  $\beta \lesssim 17$ for $Q=0.2$.  
+%  Inserting $Q\sim 0.2$ for an ideal disk marginally stable to
+%  self-gravity, we find $v_{A0}\lesssim 0.2\csmid$ for the MRI.
+
+  We can also place an upper bound on the field strength
+  $B_z$. Writing $v_{A0} = B_z\sqrt{4\pi G Q/\mu_0\Omega^2}$, we find 
+  Eq. \ref{cond2} is independent of $Q$ for $Q\ll1$, and 
+  \begin{align}\label{cond3}
+    \frac{B_z}{\csmid\Omega}\sqrt{\frac{\pi G}{\mu_0}}\lesssim
+    \frac{\sqrt{15}}{16}
+  \end{align}
+  is needed for the MRI to operate in an ideal polytropic disk with strong
+  self-gravity. Although both the MRI wavelength and disk thickness
+  vanish as $Q\to 0$, the MRI can still operate provided the field is
+  sufficiently weak according to Eq. \ref{cond3}.
+
+%Q->0, thickness -> 0
+
+%This condition translates to an upper
+%  limit on the actual field strength $B_z$. 
+%
+%  For definiteness, consider
+%  the polytropic disk with $Q \ll 1$. The disk thickness is then 
+%  $H\sim \csmid\pi\sqrt{Q}/2\Omega$. This applies at marginal
+%  gravitatational stability for which $Q\sim 0.2$. Next, expressing
+%  $\rho_0$ in terms of $Q$, we have $\lambda\sim 4\pi B_z \sqrt{\pi
+%    GQ/\mu_0}/\Omega^2$. Then 
+%  \begin{align}
+%    \frac{B_z}{\csmid\Omega}\sqrt{\frac{\pi G}{\mu_0}}\lesssim
+%    \frac{1}{4}
+%  \end{align}
+%  is needed for the MRI when self-gravity is strong. 
+  % We can quanitfy this by comparing
+  %$\lambda$ to the Jeans length $\lambda_J \equiv 2
+  %c_s^2/G\Sigma\sim 2\csmid^2/G\rho_0 H \sim 8\pi f^2Q H$. Then $\lambda
+  %< H $ translates to
+  %\begin{align}    
+  %\frac{\lambda}{\lambda_J}\lesssim \frac{1}{8\pi f^2(Q) Q}.
+  %\end{align}
+  %Given a mid-plane density $\rho_0$, and hence $Q$, the left-hand-side
+  %is a measure of the field strength.
+
+}
 

### 2014-05-28 paper/cover.tex

@@
 
-I am writing to submit the second revised manuscript for `Non-barotropic linear Rossby wave instability
-in three-dimensional disks'. I include a list of revisions made in light of the reviewer's comments.
+I am writing to submit the manuscript `Linear stability of magnetized massive protoplanetary disks'.
 

### 2014-05-28 paper/results.tex
minor edits

@@
 function of $Q$ and $\Lambda_0$. The resistivity is uniform ($A=1$). 
-For ideal MHD ($\Lambda_0>1$), there is negligible dependence on
-$Q$. In the resistive limit   
-($\Lambda_0<1$), however, growth rates decrease noticeably for $Q<0.5$
+For ideal MHD and {\bf a weak field ($\Lambda_0>1$, $\beta=100$)}, there is
+negligible dependence on $Q$. {\bf However, with $\beta=25$ or 
+in the resistive limit  ($\Lambda_0<1$)}, growth rates decrease noticeably for $Q<0.5$
 ($Q_\mathrm{2D}\lesssim 1.5$).  Since we find density and potential
@@
   \includegraphics[width=\linewidth]{figures/compare_growth_poly_uniresis2}
-  \caption{MRI growth rates as a function of $Q$ and midplane 
-    Elsasser numbers, in polytropic disks with $\beta=100$ and uniform
-    resistivity.  
+  \caption{{\bf MRI growth rates as a function of $Q$ and midplane 
+    Elsasser numbers $\Lambda_0$, in polytropic disks with $\beta=100$
+    (solid, dotted and dot-dashed lines). The dash-triple dot line is
+    the $\Lambda_0=10$ case with $\beta=25$. The resistivity is
+    uniform.}   
     \label{compare_growth_poly_uniresis}}
@@
 
-In the ideal limit, we find $\lambda < 2H$ throughout most of the disk
-for the values of $Q$ considered, so self-gravity does not affect
-growth rates significantly. However, the ratio $\lambda/2H$ does
-increase with stronger self-gravity. Consequently, the wavelength of
-the instability, in units of $H$, increases. This is shown in
+In the ideal limit {\bf with $\beta=100$}, we find $\lambda < 2H$
+throughout most of the disk for the values of $Q$ considered, so
+self-gravity does not affect growth rates significantly. However, the
+ratio $\lambda/2H$ does increase with stronger
+self-gravity. Consequently, the wavelength of the instability, in
+units of $H$, increases. This is shown in
 Fig. \ref{compare_result_lambda10} which plots the magnetic energies
@@
 because increasing vertical self-gravity makes it thinner.
+
+{\bf
+We repeated the $\Lambda_0=10$ case with a stronger field $\beta=25$,
+shown in Fig. \ref{compare_growth_poly_uniresis} as the dashed-triple
+dot line. Here, strong self-gravity is effective in reducing the
+growth rate, because decreasing $\beta$ enhances the dependence of
+$\lambda/H$ on $f(Q)$. 
+}
 
@@
 
-Self-gravity appreciably decreases the MRI growth rates in the
+Self-gravity {\bf also} appreciably decreases the MRI growth rates in the
 resistive limit. Fig. \ref{lambda_poly_resis} plots 
@@
 $f(Q)$ does not change significantly for the range of $Q$ considered,
-the dependence of $\lambda_\mathrm{resis}$ on $f(Q)$ is amplified by the
+the dependence of {\bf $\lambda/H$} on $f(Q)$ is amplified by the
 denominator $\Lambda_0<1$ in the resistive case. Modes in
@@
 $E_m$ except at $z\simeq0,\,H$, i.e. only the longest wavelength survives
-with large resistivity. 
+{\bf against} large resistivity. 
 %self-gravity tips it over
@@
 structure.
+
+{\bf 
+%We find qualitatively similar results for large $A$ with small
+%  $\Lambda_0$. 
+We also performed additional calculations with $A=10^3$ and $A=10^4$
+\citep[see Fig. 1 of ][]{gressel12}. For $\Lambda_0=0.1$, we find no
+significant increase in the magnetic energy in the resistive zones. 
+However, lowering $\Lambda_0$ gives similar
+results to Fig. \ref{poly_layer}, e.g. for $A=10^3$ and
+$\Lambda_0=10^{-2}$ or $A=10^4$ and $\Lambda_0=10^{-3}$ the magnetic
+energy penetrates into the resistive zone for strongly
+self-gravitating disks. In general, the magnetic energy density maximum moves
+toward the midplane with increasing self-gravity.    
+}
 
@@
 solve the above equations, starting from the Cowling approximation
-where $\delta\Phi=0$. This is the standard MRI and we denote the
+where $\delta\Phi\to0$ and $Q\to\infty$. This is the standard MRI and we denote the
 solution as $\dvx^{(0)}$, $\dvy^{(0)}$ and $W^{(0)}$. Eq. \ref{simp3} implies

### 2014-05-28 paper/setup.tex

@@
 We adopt constant resistivity or a 
-resistivity prescription such that $\eta(z)$ increases away from the
+resistivity prescription such that $\eta(z)$ increase {\bf towards} the
 midplane. In the latter case, we follow \cite{fleming03} and use 

### 2014-05-28 paper/paper.tex
first revision done

@@
   self-gravitating, uniformly magnetized, differentially rotating, 
-  three-dimensional disk subject to axisymmetric perturbations is
+  three-dimensional {\bf stratified} disk subject to axisymmetric perturbations is
   calculated numerically. The formulation includes 

### 2014-05-28 paper/results_azi.tex
first revision done

@@
     dots are solutions computed from the full problem, with the
-    colorbar measuring the gravitational potential perturbation, while
+    colorbar measuring the gravitational potential perturbation {\bf
+      via $\tau$}, while
     the solid curves are computed from the Cowling approximation. 
@@
 The modes in the bottom panel of Fig. \ref{compare_growth3_tilted} 
-display opposite behavior. For $B_y=0$ we find $W$ is odd, and
+display opposite behavior. For $B_y=0$ we find {\bf$W(z)$} is odd, and
 self-gravity has no effect. When $B_y>0$, $W$ deviates from an
@@
 is most effective at $k_xH =  O(1)$. 
-
 Fig. \ref{result_tilted} show eigenfunctions for $\epsilon=3$ and
@@
 magnetic modes between $0.8\lesssim k_xH\lesssim 1.5$ as $B_y$ is
-increased. For $B_y=2B_z$, MRI and GI are again independent, but
+increased. For $B_y=2B_z$, MRI and GI are again independent 
 because they operate at distinct radial scales. This implies that

### 2014-05-28 paper/results_gi.tex
revision to 1st ref report
first revision done

@@
   \caption{Growth rates for modes with $k_xH=1$ in isothermal ideal
-    disks with $Q=0.18$ (top), $Q=0.14$ (middle) and $Q=0.14$ (bottom). The
+    disks with $Q=0.18$ (top), $Q=0.14$ (middle) and $Q=0.12$ (bottom). The
     colorbar measures the importance of  
@@
 to the phenomenon of avoided crossing, as seen in stars
-\citep[e.g.][]{aizenman77} and accretion tori
-\citep[e.g.][]{christo93}, where physically distinct modes approach
+\citep[e.g.][]{aizenman77} and accretion tori{\bf /disks
+\citep[e.g.][]{christo93,ogilvie98}}, where physically distinct modes approach
 one another in frequency and exchange character. However, we cannot
@@
 the slight decrease in the most unstable $k_x$.) 
-This stabilization is due to magnetic pressure \citep{lizano10}, which
-is consistent with stabilizing the small wavelength GI only.
+This stabilization is due to magnetic pressure \citep{lizano10}, {\bf 
+consistent with pressure stabilizing small-wavelength GI only.}
 
@@
 be used as a code test for non-linear simulations. Note that this
-destabilization is through the linear response, rather 
+destabilization is through {\bf self-gravity in} the linear response, rather 
 than through the background stratification (which is stabilizing).

### 2014-05-28 paper/summary.tex
first revision done

@@
 %the strength of self-gravity $Q$ {\bf for fixed $\beta$}, 
-The disk thickness $H=H(Q)$ decreases with increasing SG. This can
-reduce MRI growth rates when 
+The disk thickness $H=H(Q)$ decreases with increasing SG. This 
+reduces MRI growth rates when 
 $\beta$, and hence $\lambda$, is fixed.  
@@
   increases with decreasing $Q$, Eq. \ref{cond1} implies
-  the MRI requires larger $\beta$ values with  
+  the MRI requires larger values of $\beta$ with  
   increasing self-gravity. 
@@
   $\beta \gtrsim 64/15Q$, giving $\beta\gtrsim 20$ for
-  $Q=0.2$. We confirmed this numerically, finding that the growth rate
+  $Q=0.2$. We confirm this numerically, finding the MRI growth rate
   $\gamma  \lesssim 0.1$ when $\beta \lesssim 3.3$ for $Q=20$ and 
@@
 
-  We can also place an upper bound on the field strength
+  We can also place an upper bound on the absolute field strength
   $B_z$. Writing $v_{A0} = B_z\sqrt{4\pi G Q/\mu_0\Omega^2}$, we find 
@@
   \end{align}
-  is needed for the MRI to operate in an ideal polytropic disk with strong
+  is needed for the MRI to operate in the ideal polytropic disk with strong
   self-gravity. Although both the MRI wavelength and disk thickness

### 2014-05-29 paper/apj/response.txt
final revised version

I thank the reviewer for comments and questions. I have performed
additional calculations and extended the discussion section to address
the issues raised below. Revisions to the manuscript are made in
bold.

**********************************************************************

1) For a fixed value of B_z, massive disks would have larger values of
plasma beta, compared to the less massive disks. In actual disks
around young stars or compact objects, how large is the required value
of B_z in order for self-gravity to affect MRI modes?  
**********************************************************************
I have expanded Section 7 to include a quantitative discussion on the
limiting field strength to permit the MRI. I show that a larger plasma
beta is required in strongly self-gravitating disks. I also estimate
the upper limit to the field strength, B_z, for Q<<1.

In addition, Fig. 2 in Section 4.1.1 has been updated to include a new
calculation, \beta=25 and \Lambda_0=10. This case demonstrates the
stabilizing effect of self-gravity in the ideal limit. 
**********************************************************************

2) The amplitude of the resistivity variation from the disk surface to
the midplane (A=eta(0) / eta(infinity) ) is fixed to 100. If you take
larger values of A, such as A=10^3 or 10^4, as seen in Figure 1 of
Gressel et al. (2012), is there a qualitative change in the result?
For example, in the Q=0.2 case in Figure 5, does the magnetic energy
still penetrate into the resistive zone? In a massive disk, the value
of A is expected to be large. Could you comment on the qualitative
tendency for high A?  
**********************************************************************
Additional calculations with A=10^3 and A=10^4 are described in
Section 4.1.2. Similar results were obtained for large A with small
Lambda_0. When the midplane resistivity (or Lambda_0) is fixed and A
is increased, the magnetic energy in the resistive zone lessens. For
A=10^3 or A=10^4 and Lambda_0=0.1, both self-gravitating and
non-self-gravitating disks show little magnetic energy in the
resistive zones. However, lowering Lambda_0 with these larger values
of A yields similar results to the case in Fig. 5.

**********************************************************************
3) Section 3.1.1 Case 1: vertical field
Please mention that the odd modes are considered in Section 6, and
cite references for explaining how the boundary conditions, Equation
(29), are derived, e.g., Sano and Miyama (1999). 
**********************************************************************
Section 3.1 has been re-structured. I describe where in the results
sections are the different field setups considered. I also make the
reference to Sano & Miyama (1999).

**********************************************************************
4) Caption of Figure 11
Q=0.14 (bottom) --> Q=0.12 (bottom) 
**********************************************************************
Corrected.
**********************************************************************

### 2014-05-30 paper/linear.tex
final edits

@@
 \subsection{Domain and boundary conditions}\label{domain}
-{\bf 
+
 For a vertical field, considered in \S\ref{result1} and
@@
 respectively.
-}
+
 %In practice,  Eq. \ref{pot_bc1} and \ref{pot_bc2} are
@@
 \subsubsection{Case 1: vertical field}
-{\bf Here we impose} $d\bm{U}/dz=0$ at $z=0$. This permits higher
+ Here we impose $d\bm{U}/dz=0$ at $z=0$. This permits higher
 numerical resolution by reducing the computational domain to
@@
 \end{align}
-so the field remains vertical. {\bf The derivation of the magnetic
-  field boundary conditions may be found in \cite{sano99}.}
+so the field remains vertical. The derivation of the magnetic
+  field boundary conditions may be found in \cite{sano99}.
 
@@
 In this more general setup the computational domain is
-$z\in[-Z_s, Z_s]$ {\bf and no symmetry across the midplane is
-  enforced.} At the disk surfaces we adopt the `halo' model of
+$z\in[-Z_s, Z_s]$ and no symmetry across the midplane is
+  enforced. At the disk surfaces we adopt the `halo' model of
 \cite{gammie94}, so that
@@
 \end{align}
-    {\bf and this case permits $\dd v_z(\pm Z_s)\neq0$.}
+     and this case permits $\dd v_z(\pm Z_s)\neq0$.
 %      In practice, the gravitational potential boundary condition is

### 2014-05-30 paper/paper.tex
final edits

--- 2014-04-07 paper/paper.tex
@@
   self-gravitating, uniformly magnetized, differentially rotating, 
-  three-dimensional disk subject to axisymmetric perturbations is
+  three-dimensional stratified disk subject to axisymmetric perturbations is
   calculated numerically. The formulation includes 
@@
   becomes global in the vertical direction, and MRI modes 
-  with small radial length scales are stabilized. In massive
+  with small radial length scales are stabilized. The
+    maximum vertical field strength that permits the MRI
+    in a strongly self-gravitating polytropic disk  with polytropic index $\Gamma=1$ 
+    is estimated to be $B_{z,\mathrm{max}} \simeq \csmid\Omega\sqrt{\mu_0/16\pi G} $, where $\csmid$ is the midplane sound speed and $\Omega$ is the angular velocity. 
+  In massive
   disks with layered resistivity, the MRI is not well-localized
@@
   can enhance density perturbations, an effect that becomes 
-  significant in the presence of a strong toroidal field. In
+  significant in the presence of a strong toroidal field, and
+    which depends on the symmetry of the underlying MRI mode. In
   gravitationally unstable disks where GI and MRI growth rates are
@@
 
+\acknowledgements
+I thank K. Menou, S. Fromang and A. Youdin for helpful discussions
+during the course of this project.
+
 \appendix

### 2014-05-30 paper/results_gi.tex
final edits

@@
 to the phenomenon of avoided crossing, as seen in stars
-\citep[e.g.][]{aizenman77} and accretion tori{\bf /disks
-\citep[e.g.][]{christo93,ogilvie98}}, where physically distinct modes approach
+\citep[e.g.][]{aizenman77} and accretion tori/disks
+\citep[e.g.][]{christo93,ogilvie98}, where physically distinct modes approach
 one another in frequency and exchange character. However, we cannot
@@
 the slight decrease in the most unstable $k_x$.) 
-This stabilization is due to magnetic pressure \citep{lizano10}, {\bf 
-consistent with pressure stabilizing small-wavelength GI only.}
+This stabilization is due to magnetic pressure \citep{lizano10},  
+consistent with pressure stabilizing small-wavelength GI only.
 
@@
 be used as a code test for non-linear simulations. Note that this
-destabilization is through {\bf self-gravity in} the linear response, rather 
+destabilization by self-gravity is through the linear response, rather 
 than through the background stratification (which is stabilizing).

### 2014-05-30 paper/setup.tex
first revision done
final revised version
final edits

--- 2014-04-05 paper/setup.tex
@@
 We adopt constant resistivity or a 
-resistivity prescription such that $\eta(z)$ increases away from the
+resistivity prescription such that $\eta(z)$ increase towards the
 midplane. In the latter case, we follow \cite{fleming03} and use 
-resistivity profile 
+the resistivity profile 
 \begin{align}
@@
 The plasma $\beta$ measures the inverse strength of the  
-magnetic field,  
+magnetic field 
 \begin{align}

### 2014-05-30 paper/summary.tex
final edits

@@
 
-For a vertical field, the requirement for the MRI {\bf to operate} is that   
+For a vertical field, the requirement for the MRI to operate is that   
 its vertical wavelength $\lambda \lesssim 2H$.  
-{\bf 
 %While $\lambda$ is independent of
@@
 $\beta$, and hence $\lambda$, is fixed.  
-}
 Thus, a sufficiently massive disk can potentially suppress the MRI. 
@@
 
-{\bf
   The condition $\lambda < 2H$ 
@@
   %is a measure of the field strength.
-
-}
 

### 2014-06-03 paper/cover.tex

@@
 
-I am writing to submit the manuscript `Linear stability of magnetized massive protoplanetary disks'.
+I am writing to submit the revised manuscript `Linear stability of
+magnetized massive protoplanetary disks', in response to the recent
+reviewer's report. Revisions are made in bold, and I attach my
+response to the reviewer.
 

### 2014-06-03 paper/results.tex
final revised version

@@
 function of $Q$ and $\Lambda_0$. The resistivity is uniform ($A=1$). 
-For ideal MHD and {\bf a weak field ($\Lambda_0>1$, $\beta=100$)}, there is
-negligible dependence on $Q$. {\bf However, with $\beta=25$ or 
-in the resistive limit  ($\Lambda_0<1$)}, growth rates decrease noticeably for $Q<0.5$
+For ideal MHD and a weak field ($\Lambda_0>1$, $\beta=100$), there is
+negligible dependence on $Q$. However, with $\beta=25$ or 
+in the resistive limit  ($\Lambda_0<1$), growth rates decrease noticeably for $Q<0.5$
 ($Q_\mathrm{2D}\lesssim 1.5$).  Since we find density and potential
@@
   \includegraphics[width=\linewidth]{figures/compare_growth_poly_uniresis2}
-  \caption{{\bf MRI growth rates as a function of $Q$ and midplane 
+  \caption{ MRI growth rates as a function of $Q$ and midplane 
     Elsasser numbers $\Lambda_0$, in polytropic disks with $\beta=100$
@@
     the $\Lambda_0=10$ case with $\beta=25$. The resistivity is
-    uniform.}   
+    uniform.   
     \label{compare_growth_poly_uniresis}}
@@
 
-In the ideal limit {\bf with $\beta=100$}, we find $\lambda < 2H$
+In the ideal limit with $\beta=100$, we find $\lambda < 2H$
 throughout most of the disk for the values of $Q$ considered, so
@@
 
-{\bf
 We repeated the $\Lambda_0=10$ case with a stronger field $\beta=25$,
@@
 growth rate, because decreasing $\beta$ enhances the dependence of
-$\lambda/H$ on $f(Q)$. 
-}
+$\lambda/H$ on $f(Q)$. For $Q=0.2$ and $\beta=25$ we find 
+$\lambda_\mathrm{ideal}/2H\sim 1$ at the midplane and the growth rate is reduced significantly.
 
@@
 
-Self-gravity {\bf also} appreciably decreases the MRI growth rates in the
+Self-gravity also appreciably decreases the MRI growth rates in the
 resistive limit. Fig. \ref{lambda_poly_resis} plots 
@@
 $f(Q)$ does not change significantly for the range of $Q$ considered,
-the dependence of {\bf $\lambda/H$} on $f(Q)$ is amplified by the
+the dependence of $\lambda/H$ on $f(Q)$ is amplified by the
 denominator $\Lambda_0<1$ in the resistive case. Modes in
@@
 $E_m$ except at $z\simeq0,\,H$, i.e. only the longest wavelength survives
-{\bf against} large resistivity. 
+ against large resistivity. 
 %self-gravity tips it over
@@
 
-{\bf 
 %We find qualitatively similar results for large $A$ with small
@@
 toward the midplane with increasing self-gravity.
-}
 

### 2014-06-03 paper/results_azi.tex

@@
     dots are solutions computed from the full problem, with the
-    colorbar measuring the gravitational potential perturbation {\bf
-      via $\tau$}, while
+    colorbar measuring the gravitational potential perturbation 
+      via $\tau$, while
     the solid curves are computed from the Cowling approximation. 
@@
 The modes in the bottom panel of Fig. \ref{compare_growth3_tilted} 
-display opposite behavior. For $B_y=0$ we find {\bf$W(z)$} is odd, and
+display opposite behavior. For $B_y=0$ we find $W(z)$ is odd, and
 self-gravity has no effect. When $B_y>0$, $W$ deviates from an
